---
title: "Finalize AHRP Metadata"
subtitle: "Compare SBF Bypass to Salmon Biological Fact + GOD (delete + pairings) + Stream Specimens + Riverdist"
author: "Kyle Shedd"
date: "Started: 2025-04-25, last opened: `r Sys.Date()`"
output:
  html_notebook:
    theme: united
    toc: yes
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

# Setup

```{r setup, message=FALSE, results='hide'}
rm(list = ls(all.names = TRUE))

if(!require("pacman")) install.packages("pacman"); library(pacman)

pacman::p_load(
  tidyverse,
  scales,
  janitor,
  GCLr
)

knitr::opts_chunk$set(fig.width = 10)

.username = readLines("~/usr_pw.txt", n = 1)  # LOKI username
.password = readLines("~/usr_pw.txt" , n = 2)[[2]]  # LOKI password
```

# Objective

The purpose of this notebook is to join together all the AHRP pedigree sample metadata (excluding genotypes). Salmon Biological Fact broke due to DFG migration to the cloud and Eric has been trying to rebuild a slimmer version in `AKFINADM.V_AHRP_SBF_BYPASS`. However, since building that new view has been an iterative process I want to double check it against the latest and greatest pre-GOD resolution data.

"V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\GIS\R\all_streams\stream_specimens_riverdist_all_streams_2013_2020.csv"
"V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\GitHub-PWS-Pink-Parentage\GOD\output\AHRP_GOD_wells_to_delete_otolith_read_data.csv"
"V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\GitHub-PWS-Pink-Parentage\GOD\output\PGOD21_assignments_for_LOKI_20250407.csv"
"V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\GitHub-PWS-Pink-Parentage\GOD\data\GOD_otolith_reads_all.csv"
"V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\OceanAK\AHRP Salmon Biological Data 20250303_092938.052746.csv"

# Background

# Import Data

## Stream Specimens + `riverdist`

This is from Stream Specimens from 2022-02-08.
```{r}
(
  stream_specimens <- readr::read_csv("../GIS/R/all_streams/stream_specimens_riverdist_all_streams_2013_2020.csv") %>%
    janitor::clean_names()
)
```

## AHRP Salmon Biological Fact

Need to read in the latest and greatest otolith data from the warehouse as of 2025-03-03. **NOTE** this does includes the 191 wells with otolith data to be deleted that were identified in `05_identify-wells-in-GOD`. Will remove those below.
```{r}
(
  AHRP_oceanAK <-
    readr::read_csv(file = "../OceanAK/AHRP Salmon Biological Data 20250303_092938.052746.csv") %>% 
    janitor::clean_names() %>% 
    dplyr::mutate(sample_date = lubridate::as_date(sample_date)) %>% 
    tidyr::unite(col = "heart_fish_id", silly_code:fish_id, sep = "_", remove = FALSE) %>%  # add heart SillySource as data key for GOD
    tidyr::unite(col = "sample", dna_tray_code:dna_tray_well_code, sep = "_", remove = FALSE)  # add DWP data key for stream specimens
)
```

## GOD Wells To Delete Otolith Data

Read in 191 wells with existing otolith data in Salmon Biological Fact to be deleted. This has occurred on the database end, but after the 2025-03-03 Salmon Biological Fact pull, so need to do it here.
```{r}
(
  AHRP_wells_to_delete_otolith_data <- readr::read_csv(file = "GOD/output/AHRP_GOD_wells_to_delete_otolith_read_data.csv") %>%
    tidyr::unite(
      col = "DWP_key",
      dna_tray_code:dna_tray_well_code,
      sep = "_",
      remove = FALSE
    )  # make key
)
```

## Assigned PGOD Matches

Read in final PGOD21 match assignments based on genotypes and logic.
```{r}
(PGOD21_48DWP_paired_genotype_logic_matches <- readr::read_csv(file = "GOD/output/PGOD21_48DWP_paired_genotype_logic_matches_final.csv"))
```

## GOD Otolith Reads

Got GOD otolith reads from Cordova Otolith lab (2025-02-05) with help from Eric L. (GCL AP) and Rachel E. (CDV Otolith Lab Manager).
```{r}
(GOD_otolith_reads <- readr::read_csv("GOD/data/GOD_otolith_reads_all.csv") %>% 
   janitor::clean_names())
```

Total of 10,816 GOD otoliths.

## Salmon Biological Fact Bypass

Read in Eric Lardizabal's AHRP view that bypasses SBF.
```{r}

```

# Compare Salmon Biological Fact to Bypass

See `oceanak_data_pull.R` for details. Already know that the Bypass contains duplicated rows for GOD samples & is missing samples that never had otoliths logged (e.g. `is.na(OTOLITH_MARK_STATUS_CODE)`).

TODO


# Modify/Join Data

## Delete 191 Otolith Reads

Wipe existing otolith reads from GOD wells identified in `AHRP_wells_to_delete_otolith_data`
```{r}
AHRP_oceanAK_clean <- AHRP_oceanAK %>% 
  tidyr::unite(col = "DWP_key", dna_tray_code:dna_tray_well_code, sep = "_", remove = FALSE) %>%  #  make key
  dplyr::mutate(otolith_mark_present = dplyr::case_when(DWP_key %in% AHRP_wells_to_delete_otolith_data$DWP_key ~ NA_character_,
                                                        TRUE ~ otolith_mark_present),
                otolith_mark_id = dplyr::case_when(DWP_key %in% AHRP_wells_to_delete_otolith_data$DWP_key ~ NA_character_,
                                                   TRUE ~ otolith_mark_id),
                otolith_mark_status_code = dplyr::case_when(DWP_key %in% AHRP_wells_to_delete_otolith_data$DWP_key ~ NA_character_,
                                                            TRUE ~ otolith_mark_status_code)) %>%   # wipe oto data
  dplyr::select(-DWP_key)
```

## Join GOD Otolith Reads

Join GOD otolith reads to the PGOD21 otolith transfer and genotype match assignments, filter for final assignments (5,520).
```{r}
(
  PGOD21_48DWP_paired_genotype_logic_matches_otolith_reads <- PGOD21_48DWP_paired_genotype_logic_matches %>% 
    dplyr::left_join(
      y = GOD_otolith_reads %>%
        dplyr::mutate(
          oto_tray = as.numeric(tray_id),
          oto_well = as.numeric(cell_number)
        ) %>%
        tidyr::unite(col = "oto_tray_and_well", c(oto_tray, oto_well), sep = "_") %>%
        dplyr::select(oto_tray_and_well, mark_present, mark_id, otolith_mark_status_code),
      by = "oto_tray_and_well"
    ) %>%  # join otolith reads
    dplyr::filter(!is.na(heart_fish_id) & non_GOD_assignment == FALSE) %>%  # filter for final GOD assignments
    dplyr::rename(otolith_mark_present = mark_present,
                  otolith_mark_id = mark_id)  # rename for joining
)
```

## Update GOD Otolith Reads

Replace otolith read data for 5,520 GOD assignments.
```{r}
(
  AHRP_oceanAK_GOD <- AHRP_oceanAK_clean %>%
    dplyr::mutate(
      GOD_oto_read = heart_fish_id %in% PGOD21_48DWP_paired_genotype_logic_matches_otolith_reads$heart_fish_id
    ) %>%  # record GOD oto assignments
    dplyr::select(-c(target_dna_tray_code:sw_age)) %>%  # toss irrelevant columns
    dplyr::left_join(
      y = PGOD21_48DWP_paired_genotype_logic_matches_otolith_reads %>%
        dplyr::select(
          heart_fish_id,
          otolith_mark_present:otolith_mark_status_code
        ),
      by = dplyr::join_by("heart_fish_id"),
      suffix = c("", "_GOD")
    ) %>%  # join assigned GOD otolith reads
    dplyr::mutate(
      otolith_mark_present = dplyr::case_when(
        GOD_oto_read ~ otolith_mark_present_GOD, 
        TRUE ~ otolith_mark_present),
      otolith_mark_id = dplyr::case_when(
        GOD_oto_read ~ otolith_mark_id_GOD, 
        TRUE ~ otolith_mark_id),
      otolith_mark_status_code = dplyr::case_when(
        GOD_oto_read ~ otolith_mark_status_code_GOD, 
        TRUE ~ otolith_mark_status_code)
    ) %>%  # update GOD otolith reads
    dplyr::select(-c(
      otolith_mark_present_GOD:otolith_mark_status_code_GOD
    ))  # toss extra GOD otolith read columns
)
```

## Join Stream Specimens + `riverdist`

First check and see if all `AHRP_oceanAK_GOD` in `stream_specimens`.
```{r}
all(AHRP_oceanAK_GOD$sample %in% stream_specimens$sample)
```
Hallelujah!

Join starting with stream_specimens, reformat columns to match `all_streams_paired_XX_YY_filter`
```{r}
(
  pws_pink_ped_dat <- stream_specimens %>%
    dplyr::left_join(y = AHRP_oceanAK_GOD, by = "sample") %>%  # join Salmon Bio Fact to Stream Specimens + riverdist
    dplyr::rename(date = sample_date,
                  stream_trib = stream_name,
                  length = length_mm,
                  silly = silly_code,
                  silly_source = heart_fish_id,
                  distance_mouth = mouthdist,
                  distance_tide = dist2tide,
                  high_tide = hitide,
                  partial_spawn = partial_spawner) %>% 
    dplyr::mutate(franz_id = stringr::str_c(stringr::str_sub(silly, 1, 2),  # 1st letter of species name, 1st letter of stream name
                                            stringr::str_sub(silly, -2, -1),  # 2 digit sample year
                                            "_",
                                            stringr::str_pad(fish_id, width = 5, pad = "0", side = "left")),  # 5 digit fishid using padded zeros up front
                  year = lubridate::year(date),
                  stream = stringr::str_remove(string = location_code, pattern = " Creek"),
                  DOY = lubridate::yday(date),  # day of year
                  sex = dplyr::case_when(
                    sex == "M" ~ "Male",
                    sex == "F" ~ "Female",
                    sex == "U" ~ NA_character_,
                    is.na(sex) ~ NA_character_),
                  origin = dplyr::case_when(
                    otolith_mark_present == "NO" ~ "Natural",
                    otolith_mark_present == "YES" ~ "Hatchery"
                  ),  # add origin variable
                  origin = base::factor(origin, levels = c("Natural", "Hatchery")),  # make factor to ensure hatchery != red
                  hatchery = dplyr::case_when(
                    stringr::str_detect(otolith_mark_id, "AFK") ~ "AFK",
                    stringr::str_detect(otolith_mark_id, "CCH") ~ "CC",
                    stringr::str_detect(otolith_mark_id, "ARMSTRONG") ~ "PA",
                    stringr::str_detect(otolith_mark_id, "GRAHAM") ~ "PG",
                    stringr::str_detect(otolith_mark_id, "SGH") ~ "SG",
                    stringr::str_detect(otolith_mark_id, "TUTKA") ~ "TUT",
                    stringr::str_detect(otolith_mark_id, "UNKNOWN") ~ NA_character_,
                    stringr::str_detect(otolith_mark_id, "WNH") ~ "WN",
                    stringr::str_detect(otolith_mark_id, "AFK") ~ "AFK",
                  ),  # add hatchery variable
                  intertidal = dplyr::case_when(
                    distance_tide <= 0 ~ "Intertidal",
                    distance_tide > 0 ~ "Upstream",
                    is.na(distance_tide) ~  NA_character_
                  ),  # create intertidal
                  spawning_state = dplyr::case_when(
                    spawning_state == "Unknown" ~ NA_character_,
                    TRUE ~ spawning_state
                  )  # force Unknown to NA
    ) %>%  
    dplyr::select(
      franz_id,  # unique identifier for FRANz (e.g. PE16_00001, species, stream, year, GCL fish_id)
      sample,  # ultimate data key for AHRP, DWP barcode + DWP position
      silly_source,  # unique identifier for LOKI (GCL database)
      stream,  # 5 pedigree streams
      year,  # sample year (aka death year for FRANz)
      origin,  # hatchery or wild
      hatchery,  # which hatchery
      sex,  # male, female, unknown
      date,  # sample date
      DOY,  # day of year (aka julian sample date)
      length,  # mideye to hypural plate length in mm
      intertidal,  # sampling location above or below intertidal
      distance_mouth,  # riverdist approximate sampling location from mouth in meters 
      spawning_state,  # alive, pink gill, grey gill, rotting (aka index of stream life)
      pre_spawn,  # TRUE/FALSE (added in 2016); if the eggs and milt do not easily release and carcass has completely full gonads
      partial_spawn,  # TRUE/FALSE (added in 2015); fish is sampled (dead or alive) with more than a little eggs or milt
      preyed_upon,  # TRUE/FALSE (added in 2015); gonads removed by predators
      stream_trib,  # stream tributary (Gilmour, Paddy, and Stockdale have multiple)
      latitude,  # sampling area (where fish were moved to prior to sampling, rough approximate of spawning location)
      longitude,  # sampling area (where fish were moved to prior to sampling, rough approximate of spawning location)
      otolith_mark_present,  # otolith mark present (yes = hatchery, no = natural)
      otolith_mark_id,  # otolith hatchery mark
      GOD_oto_read,  # TRUE/FALSE otolith read assigned via GOD match (aka great otolith debacle)
      silly,  # GCL collection code
      fish_id,  # GCL fish ID within a silly
      dna_tray_code,  # 10-digit DWP barcode
      dna_tray_well_code,  # 48 DWP position
      distance_tide,  # riverdist approximate sampling location from maximum high tide
      riverdist_seg,  # riverdist river segment
      riverdist_vert,  # riverdist vertex
      riverdist_snapdist,  # riverdist snapping distance (aka how far was sampling location lat/long from the stream polyline)
      high_tide  # highest extent of high tide in meters from mouth, unique to each stream
    )  
)
```

These are the columns in all_streams_paired_XX_YY_filter see 2_Parentage_XX R Notebook, line 258:

  * franz_id = PE16_00001, species, stream, year, GCL fish_id
  * stream = 5 pedigree streams
  * year
  * origin = factor of Hatchery vs Natural
  * sex = Female or Male
  * date = sample date
  * DOY = julian date
  * length
  * distance_mouth = mouthdist
  * intertidal = Intertidal vs. Upstream
  * distance_tide = dist2tide
  * otolith_mark_present
  * otolith_mark_id
  * *GOD_oto_read* (new)
  * *hatchery* (new)
  * *silly_source* (new)
  * silly
  * fish_id
  * sample = DWP barcode _ DWP position
  * dna_tray_code
  * dna_tray_well_code
  * stream_trib = stream_specimens$stream_name (includes separate tribs)
  * spawning_state
  * pre_spawn
  * partial_spawn
  * preyed_upon
  * riverdist_seg
  * riverdist_vert
  * riverdist_snapdist
  * latitude
  * longitude
  * high_tide

# Confirm

**Note** there are 151 more samples in `stream_specimens` than `AHRP_oceanAK` (aka Salmon Biological Fact). These 151 samples will likely be excluded from all final analyses.

## QA/QC

Confirm origin is correct.
```{r}
pws_pink_ped_dat %>% 
  dplyr::count(origin, otolith_mark_present)
```

Confirm hatchery origin is correct.
```{r}
pws_pink_ped_dat %>% 
  dplyr::count(origin, hatchery, otolith_mark_id)
```

The hatchery brood year thing is a bit weird, or am I missing something?
```{r}
pws_pink_ped_dat %>% 
  dplyr::filter(!is.na(otolith_mark_id)) %>% 
  dplyr::mutate(oto_mark_year = as.numeric(stringr::str_extract_all(otolith_mark_id, "\\d+")) + 2000,
                oto_mark_year_diff = year - oto_mark_year) %>% 
  dplyr::count(oto_mark_year_diff)
```

Okay, yes, there are some errors, but most are brood years.

Confirm sex is correct.
```{r}
pws_pink_ped_dat %>% 
  dplyr::count(sex)
```
What is date range within each year?
```{r}
pws_pink_ped_dat %>% 
  dplyr::group_by(year) %>% 
  dplyr::summarise(start = min(DOY),
                   end = max(DOY),
                   total_days = end - start)
```

Confirm intertidal.
```{r}
pws_pink_ped_dat %>% 
  dplyr::count(intertidal)
```

Confirm spawning state.
```{r}
pws_pink_ped_dat %>% 
  dplyr::count(spawning_state)
```

Confirm when data on pre_spawn, partial_spawn, and preyed_upon started being collected.
```{r}
pws_pink_ped_dat %>% 
  dplyr::filter(!is.na(pre_spawn)) %>% 
  dplyr::count(year)

pws_pink_ped_dat %>% 
  dplyr::filter(!is.na(partial_spawn)) %>% 
  dplyr::count(year)

pws_pink_ped_dat %>% 
  dplyr::filter(!is.na(preyed_upon)) %>% 
  dplyr::count(year)
```

Huh, `pre_spawn` started in 2016, but `partial_spawn` and `preyed_upon` started in 2015. Added definitions above based on the 2015 and 2016 AHRP contrator reports from PWSSC.

## Sample Sizes

```{r}
pws_pink_ped_dat %>% 
  dplyr::count(stream, year) %>% 
  dplyr::arrange(year) %>% 
  tidyr::pivot_wider(names_from = year, values_from = n)
```

# Save Final Dataset

TODO, currently this is based on Salmon Biological Fact!

Save as .rds (factors!) and .csv (last forever, transferrable)
```{r}
fs::dir_create("data")
GCLr::save_objects(objects = "pws_pink_ped_dat", path = "data", rds = TRUE)
readr::write_csv(x = pws_pink_ped_dat, path = "data/pws_pink_ped_dat_salmon_bio_fact_20250605.csv")
```

End