---
title: "PWS hydraulic distance"
output: html_notebook
---

# Introduction:    
The notebook calculates the river distance of each fish, from the river mouth. It relies on a river network shapefile and 'riverdist' package. Riverdist was created by ADFG biologist Matt Tyers for tracking fish movements through a river network. More details and vignette can be found here:        
https://cran.r-project.org/web/packages/riverdist/vignettes/riverdist_vignette.html#Importing%20a%20river%20network

The river network shapefiles are generally taken from the National Hydrography Database (NHD), however, in somce cases the polylines are not accurate or do not exist. Rivers not appearing in the NHD were delineated, in ArcMap, by simply creating polylines through the fish waypoints (We'll assume these datapoints represent fish in the stream...). The shapefiles for Stockdale Creek and Gilmore Creek were taken from the NHD. Paddy Creek, Erb Creek, and Hogan Creek, were maually delineated.


# Setup script
```{r setup}
while(!require(pacman)){ install.packages("pacman") }
  p_load(riverdist, rgdal, sp, tidyverse)  # use pacman to load or install + load necessary packages
```


```{r inputs}
stream = "stockdale"

# CAUTION - don't change this unless you want a different projection.
# For re-projecting in Alaska Albers Equal Area projection:
AKalbers <- "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154
    +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"
```



```{r prep}
# Provide shapefile (*.shp) of river network
shapefile = paste0("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GIS/R/",stream,"/",stream,"_stream_ckean.shp")

# CSV containing fish data, specifically, Lat/Longs
stream_data = "V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GIS/R/fish_2013to2018_chaseedits.csv"

# Where do you want output to go?
output_path <- paste0("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GIS/R/",stream,"/",stream,"_distances.txt")
```


# Begin Script

First, get the shapefiles and gps points ready for 'riverdist', then perform basic QC and visualization steps.    
    Note: Assumptions will be made!
    
    GPS points do not line up with streams so are snapped to the nearest (straight-line) stream segment
    Sometimes, points in the intertidal extend past stream network, so are snapped to the stream mouth (i.e., no negative distances)

```{r load_files}
# Import and make sure it's projected into AKAlbers
rivernetwork <- line2network(path = shapefile, reproject = AKalbers)
```

```{r plotnetwork}
plot(x= rivernetwork)
```

NOTE - Run this chunk in console as the notebook lags 1 graph behind, making it difficult to answer questions...    
    
    Cleanup steps taken for <i>STOCKDALE CREEK</i>:
    
    Insert verticies to increase snapping precision : y
    Minimum distance to use:  1 [this is in map units shown under the "load_files" chunk]
    Identify segment number of river mouth:  2
    Identify vertex number of river mouth:  1122
    Accept mouth assignment: y
    Remove additional segments:  n
    Build segment routes:  y
    
```{r CleanNetwork}
# run in console if you want / need visual guides
rivercleaned <- cleanup(rivernetwork)

# check topology so network distance can be calculated, make sure all confluences are green and endpoints are red
topologydots(rivers = rivercleaned)
```

# Import the raw fish data from the creek. 
```{r import_data}
raw_dat <- read_csv(file = stream_data)

xy_coords <- cbind(as.numeric(raw_dat$Longitude),
                  as.numeric(raw_dat$Latitude))
```

Now, convert the Lat/Long coordinates into AKalberts xy format. Make sure these are the same projection as the stream.
```{r xy2akalbers}
# Convert from Lat/Long dec. degree coordinates to the xy system.
# Note coords are long-lat, not lat-long.
xy2albers <- project(xy_coords, proj=AKalbers)

# Just xy-coords
head(xy2albers)

# now add xy-coords back to original fish data
fish_locs <- raw_dat %>% 
  select(SampleTray, SampleCell) %>% 
  unite_("Sample", c("SampleTray", "SampleCell")) %>% 
  add_column(x = xy2albers[,1], y = xy2albers[,2])

head(fish_locs)
```

Once the coordinates are in the correct format, the fish can be snapped to the stream network. 

```{r CalculateSnapDistance}
# Snap xy-coords to the nearest streamline. 
fish_locs_snapped <- xy2segvert(x = fish_locs$x, y = fish_locs$y, rivers = rivercleaned)

# Add fishID (Sample = TrayID_Cell)
fish_locs_snapped <- fish_locs_snapped %>% 
  add_column(Sample = fish_locs$Sample)

head(fish_locs_snapped)

# Visual of how far the waypoints were from the stream network (i.e., snapping distance)
ggplot(fish_locs_snapped, aes(x = snapdist)) +
  geom_histogram() +
  ggtitle(label = "Distance of GPS waypoint from river (m)")
```

View the waypoints on the stream network and compare snapped (blue) to raw (red) locations:

```{r RawVsSnapped}
plot(x = rivercleaned)

# Add points of raw data in red
points( x = fish_locs$x, y = fish_locs$y, pch =5, col = 2) 

# Add points from snapped data in blue
riverpoints(seg = fish_locs_snapped$seg, vert=fish_locs_snapped$vert, rivers = rivercleaned, pch=10, col=4)
```


Calculate the distance of every fish to the river mouth.
```{r CalculateDistance}
# Calculate distanc, using default algorithm. See riverdist docs for details of available algorithms.
distance <- mouthdist(seg = fish_locs_snapped$seg, vert = fish_locs_snapped$vert, rivers = rivercleaned)
```


Create dataframe containing fish information and distance calculation.
```{r FinalData}
final_df <- fish_locs_snapped %>%  
  add_column(distance)

head(final_df)
```


# Final QC / Cross-check

A good way to examine issues is to cross-check riverdist calculations with ArcMap (STARS v2.0.7) calculations.

```{r CompareArc2Riv}
# Read in csv from arcmap and select columns
arcmap <- read_csv(file = arc_updist) %>% 
  select(SampleTray, SampleCell, upDist, Longitude, Latitude, US_hi_tide) %>% 
  unite_("Sample", c("SampleTray", "SampleCell"))

# Remove unnecessary columns from dataframe with distance calculations
final_df_trim <- final_df %>% 
  select(Sample, distance, seg, vert)

joined <- left_join(x = arcmap, y = final_df_trim, "Sample")

# Visualize a 1:1 plot of distances
ggplot(joined, aes( x = upDist, y = distance)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0)
```


```{r DifferenceCalc}
# How different are the two distance calculations?? 
joined %>% 
  mutate(difference = upDist - distance) %>% 
  ggplot(aes(x = difference)) +
  geom_histogram()
```


Plot of distance data by Lat/Long - These should be a continuous color scale
```{r}
ggplot(joined, aes(x = Longitude, y= Latitude, color = upDist)) + 
  geom_point() +
  ggtitle("ArcMap calculations")

ggplot(joined, aes(x = Longitude, y= Latitude, color = distance)) + 
  geom_point() +
  ggtitle("Riverdist calculations")
```

# Export Data


If all looks good, write a csv containing:    
    
    Sample 
    Latitude
    Longitude
    Distance from river mouth
    Segment wpt was snapped to
    Vertex wpt was snapped to
    Snapping distance (how far wpt was from stream)
    Intertidal or not (1 = intertidal, 0 = upstream of tidal mark)
    


```{r exportCSV}
write_csv(x = final_df, path = output_path)
```


