---
title: "DWPs previously genotyped"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(DT)

source("~/../R/Functions.GCL.R")
```

# Purpose

The purpose of this R Notebook is to determine which DWPs Cordova should send to the GCL for otolith-DNA extraction testing (i.e. DWPs that have already been genotyped and have a large number of otoliths).

# Background

There was a major otolith DWP debacle that occurred when shipping PWS pink salmon otoliths to the MTA in Juneau for otolith reading. The DWPs had already been tissue separated (i.e. the hearts are safe in the correct location within a DWP). We were shipping these otoliths from 2015-2018 for Paddy, Erb, and Gilmour, because the Cordova otolith lab lacked the capacity to read them in a timely fashion and the MTA had excess capacity in the spring prior to fishing season starting in the summer. Many of the DWPs had been heat-sealed in the past, making it difficult to seal with the original impermamats. This difficulty in sealing the DPWs with impermamats and the overall expense of impermamats led us to order acetates to cover the DWPs (acetate + rubber band). Unfortunately for all involved, the boxes of DWPs were not paletized and clearly got some rough handling in shipment to Juneau. This meant that many DWPs had otoliths all over the place, losing the paired nature of the DNA (tissue) and origin (otolith) data.

In an attempt to rectify this situation, we are going to try to pair these otoliths back with the appropriate heart tissue by genotyping DNA from the otolith for a subset of SNPs from the GT-seq panel. The idea is, if we can extract DNA from otoliths (without harming the otolith), we can genotype that DNA and pair the otolith back with the heart tissue (because we know what DWP the otolith came from, but not the cell it came from). We should be able to do this with <24 SNPs. In order to do this cost-effectively, the idea is to order 96 *SNPtype* assays and select a subset of 24 SNPs that score well so that we can genotype the otolith-DNA with 192.24 *Fluidigm* chips.

Step 1 - send *Fluidigm* sequence information so they can design 96 *SNPtype* assays
Step 2 - extract DNA from right-side otoliths of known origin, for tissues that have already been genotyped with GT-seq
Step 3 - genotype this otolith-DNA for these 96 *SNPtype* assays
Step 4 - pair otoliths and tissues with the genotypes
Step 5 - send otoliths back to Cordova to verify they can still be read after going through DNA extraction
Step 6 - pick 24 assays to order in bulk
Step 7 - travel to Juneau to develop a protocol

# Read in *OceanAK* Data

```{r}
og_names <-
  suppressMessages(names(
    read_csv(file = "../OceanAK/PedigreeData_AHRP - Salmon Biological Data 2_PWS_2013-2018_no_otoliths.csv", progress = FALSE)
  ))

oceanak <-
  read_csv(file = "../OceanAK/AHRP Salmon Biological Data 20200129_1621.csv")

names(oceanak) <- og_names

oceanak_mod <- oceanak  %>%
  unite(SillySource,
        `Silly Code`,
        `Fish ID`,
        sep = "_",
        remove = FALSE) %>%
  unite(TrayCodeID,
        `DNA Tray Code`,
        `DNA Tray Well Code`,
        sep = "_",
        remove = FALSE) %>%
  mutate(otolith_read = !is.na(`Otolith Mark Status Code`) &
           `Otolith Mark Status Code` != "n") %>%
  mutate(stream = factor(
    x = `Location Code`,
    levels = c(
      "Gilmour Creek",
      "Paddy Creek",
      "Erb Creek",
      "Hogan Creek",
      "Stockdale Creek"
    )
  )) %>%
  rename(year = `Sample Year`) %>%
  mutate(
    origin = case_when(
      `Otolith Mark Present` == "NO" ~ "natural",
      `Otolith Mark Present` == "YES" ~ "hatchery"
    )
  ) %>%
  mutate(origin = factor(origin, levels = c("natural", "hatchery"))) %>%
  mutate(date = dmy(`Sample Date`))
```

How many samples per DWP?
```{r}
(
  stockdale_17 <- oceanak_mod %>%
    filter(stream == "Stockdale Creek" & year == 2017 & otolith_read == TRUE) %>%
    count(`DNA Tray Code`)
)
```

Histogram of sample size by DWP
```{r}
stockdale_17 %>% 
  ggplot(aes(x = n)) +
  geom_histogram()
```


# Read in Stockdale 2017 genotypes

```{r}
load_sillys("../Genotypes/Stockdale_15_17/OceanAK_Origin_PostQA/", sillyvec = "PSTOCK17")
```

How many samples genotyped per DWP?
```{r}
(
  stockdale_17_geno <- PSTOCK17.gcl$attributes %>% 
    as_tibble() %>% 
    count(DNA_TRAY_CODE) %>% 
    rename("DNA Tray Code" = DNA_TRAY_CODE)
)
```


Histogram of sample size by DWP
```{r}
stockdale_17_geno %>% 
  ggplot(aes(x = n)) +
  geom_histogram()
```

# Join

```{r}
(
  stockdale_17_join <-
    left_join(
      stockdale_17,
      stockdale_17_geno,
      by = "DNA Tray Code",
      suffix = c("_otolith", "_genotyped")
    ) %>%
    replace_na(list(n_genotyped = 0))
)
```

# Filter

These are the DWPs Stormy pulled
```{r}
stormy_DWPs <- c(22579, 22399, 22577, 22544, 22498, 22649, 22579, 22383, 22614, 22608, 22364, 22741, 22765, 22570, 22463, 22653, 22573, 22671, 22737, 22575, 22733)
```

Are they all in my joined table? Did I type them in correctly?
```{r}
all(stormy_DWPs %in% stockdale_17_join$`DNA Tray Code`)
```

Cool, filter and arrange by n_genotyped
```{r}
stockdale_17_join %>% 
  filter(`DNA Tray Code` %in% stormy_DWPs) %>% 
  arrange(desc(n_genotyped))
```

```{r}
stockdale_17_join %>% 
  filter(`DNA Tray Code` %in% stormy_DWPs) %>% 
  arrange(desc(n_genotyped)) %>% 
  write_csv("../Otolith Separation/DWP_otoliths_from_Cordova_DNA_test_20200318.csv")
```

## Round 2 10/1/20

These are the DWPs Crystal pulled
```{r}
crystal_DWPs <- as.numeric(readClipboard())
```

Are they all in my joined table? Did I type them in correctly?
```{r}
all(crystal_DWPs %in% stockdale_17_join$`DNA Tray Code`)
```

Cool, filter and arrange by n_genotyped
```{r}
stockdale_17_join %>% 
  filter(`DNA Tray Code` %in% crystal_DWPs) %>% 
  arrange(desc(n_genotyped))
```

```{r}
stockdale_17_join %>% 
  filter(`DNA Tray Code` %in% crystal_DWPs) %>% 
  arrange(desc(n_genotyped)) %>% 
  write_csv("../Otolith Separation/DWP_otoliths_from_Cordova_DNA_test_20201001.csv")
```