---
title: "Genotype Status for PWS Pink Salmon Fitness Study"
subtitle: "Which Samples Have NOT Been Genotyped?"
author: "Kyle Shedd"
date: "9 February 2022; update 14 April 2022; update 13 December 2022"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list=ls())

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(DT)
library(abind)
bbind <- function(...) { abind(..., along = 3) }

.username = readLines("~/../R/usr_pw.txt", n = 1)
.password = readLines("~/../R/usr_pw.txt" , n = 2)[[2]]
source("~/../R/Functions.GCL.R")  # develop branch!!!
```

# Objective

Figure out which hearts need are remaning and need to be genotyped.

# Introduction

This is a quick exploratory analysis of the 2014/2016 genotype data from all five fitness streams (Hogan, Stockdale, Gilmour, Paddy, and Erb). Here we will:

  * Read in paired field/otolith data from [OceanAK](http://www.oceanak.adfg.alaska.gov/)
  * Read in *LocusControl*
  * Read in all project genotypes
  * Join genetic data with paired data from OceanAK
    + Will need to create data key for join (SillySource, Silly_FishID)
  * Determine which streams/years/origin have not been genotyped

**NOTE** this script operates off of the `develop` branch of *GCL-R-Scripts*

# OceanAK

## Read and format

Read in paired field and otolith data from *OceanAK* and join that data to the attributes table of each silly object
```{r read_oceanak, message=FALSE}
(
  oceanak <-
    readr::read_csv(file = "../OceanAK/AHRP Salmon Biological Data 20221212_162324_names.csv") %>%
    tidyr::unite(
      col = "silly_source",
      c("Silly Code", "Fish ID"),
      sep = "_",
      remove = FALSE
    )
)  # read in renamed OceanAK data; recommend recent data pull prior to running
```

## Visualize and subset

How many samples per stream/year?
```{r}
oceanak %>%
  dplyr::count(`Sample Year`, `Location Code`) %>%
  tidyr::pivot_wider(names_from = `Sample Year`,
                     values_from = n,
                     values_fill =  0)
```

# Genotypes

Create *LocusControl* and read in genotypes by *silly*

First grab the unique `Silly Code` values from `oceanak` and `loci298`
```{r sillyvec}
(sillyvec <- sort(unique(oceanak$`Silly Code`)))  # derived from oceanak

loci298 <- dget("../Objects/loci298.txt")  # saved list of loci
```

## LocusControl

Create `LocusControl` but only for 1 locus, just need to know what got genotyped or not, not saving anything
```{r LocusControl}
CreateLocusControl.GCL(locusnames = loci298[1],
                       username = .username,
                       password = .password)

loci <- LocusControl$locusnames
```

## Genotypes

Read in genotypes
```{r}
LOKI2R.GCL(sillyvec = sillyvec,
           username = .username,
           password = .password,
           test_type = "GTSNP")
```

Not planning to save anything, since we are just looking

# What has not been genotyped?

## Tidy things up

Verify that all genotypes are in OceanAK data
```{r}
all.gcl <- lapply(objects(pattern = "\\.gcl"), get) %>%
  dplyr::bind_rows()

all(all.gcl$SillySource %in% oceanak$silly_source)

table(all.gcl$SillySource %in% oceanak$silly_source)
```

Fabulous, all in there. That is the huge upside of working with the data warehouse that Tim and Eric put together.

Subset for samples that we **HAVE** genotyped using `inner_join` of `oceanak` and `.gcl` objects.
```{r}
(
  genotyped <- all.gcl %>%
    dplyr::select(SillySource) %>%
    dplyr::inner_join(
      x = oceanak,
      y = .,
      by = c("silly_source" = "SillySource")
    )
)
```

Subset for samples that we have **NOT** genotyped using `anti_join` of `oceanak` and `.gcl` objects.
```{r}
(
  not_genotyped <- all.gcl %>%
    dplyr::select(SillySource) %>%
    dplyr::anti_join(
      x = oceanak,
      y = .,
      by = c("silly_source" = "SillySource")
    )
)
```

Make a tidy dataset showing what **has** and **has not** been genotyped.
Bind together genotyped and non-genotyped samples, rename, and reformat to new standard
```{r}
(
  oceanak_geno <- dplyr::bind_rows(
    genotyped %>%
      dplyr::mutate(genotyped = TRUE)
    ,
    not_genotyped %>%
      dplyr::mutate(genotyped = FALSE)
  ) %>%
    dplyr::select(
      `Silly Code`,
      `Fish ID`,
      `DNA Tray Code`,
      `DNA Tray Well Code`,
      `Location Code`,
      `Sample Year`,
      `Sample Date`,
      Sex,
      `Length Mm`,
      `Otolith Mark Present`,
      `Otolith Mark ID`,
      `Otolith Mark Status Code`,
      `Is Missing Paired Data Exists`,
      `Well Has More Than One Sample`,
      `Tissue Type`,
      genotyped
    ) %>%
    dplyr::rename(
      silly = "Silly Code",
      fish_id = "Fish ID",
      dna_tray_code = "DNA Tray Code",
      dna_tray_well_code = "DNA Tray Well Code",
      year = "Sample Year",
      length = "Length Mm",
      otolith_mark_present = "Otolith Mark Present",
      otolith_mark_id = "Otolith Mark ID",
      otolith_mark_status_code = "Otolith Mark Status Code",
      is_missing = "Is Missing Paired Data Exists",
      more_than_one = "Well Has More Than One Sample",
      tissue_type = "Tissue Type"
    ) %>%
    dplyr::mutate(
      stream = stringr::str_remove(string = `Location Code`, pattern = " Creek"),
      date = lubridate::as_date(`Sample Date`),
      DOY = lubridate::yday(date),
      # day of year
      sex = dplyr::case_when(Sex == "M" ~ "Male",
                             Sex == "F" ~ "Female"),
      origin = dplyr::case_when(
        otolith_mark_present == "NO" ~ "Natural",
        otolith_mark_present == "YES" ~ "Hatchery"
      ),
      # add origin variable
      otolith_read = dplyr::case_when(
        !is.na(otolith_mark_status_code) ~ TRUE,
        is.na(otolith_mark_status_code) ~ FALSE
      )
    ) %>%
    dplyr::mutate(origin = base::factor(origin, c(
      "Natural", "Hatchery"
    ))) %>%  # make factor to ensure hatchery != red
    tidyr::unite(
      col = "sample",
      c("dna_tray_code", "dna_tray_well_code"),
      sep = "_",
      remove = FALSE
    ) %>%
    dplyr::select(
      genotyped,
      stream,
      year,
      origin,
      sex,
      date,
      DOY,
      length,
      otolith_mark_present,
      otolith_mark_id,
      otolith_mark_status_code,
      otolith_read,
      silly,
      fish_id,
      sample,
      dna_tray_code,
      dna_tray_well_code,
      is_missing,
      more_than_one,
      tissue_type
    )
)
```

**NOTE** my naive attempt to create a boolean (TRUE/FALSE) `otolith_read` does **NOT** appear to be working as intended. Long story short, GOD screwed it all up. Since we have DWPs with *GOD-affected* otoliths being removed, but then the *unaffected* otoliths remaining in the DWPs and subsequently being logged and read by Cordova or MTAL, those empty wells get coded as `3N66` - missing. There is no way to tell which 3N66 otoliths are *true missing* vs. *GOD-affected*, so that sucks. Also, I have no earthly idea what `n` means (probably `null`) or how those otoliths got coded as such. Please, please, please, treat `otolith_read` with a dumptruck of salt.
```{r}
oceanak_geno %>% 
  dplyr::count(otolith_read)
```


## Table samples by genotyped vs. not

### Genotyped

Make a table of what we **HAVE** genotyped
```{r}
oceanak_geno %>%
  dplyr::filter(genotyped == TRUE) %>%
  dplyr::count(stream, year, otolith_mark_present) %>%
  tidyr::pivot_wider(names_from = otolith_mark_present,
                     values_from = n,
                     values_fill =  0)
```

### Not Genotyped

How about the other way around? What have we **NOT** genotyped?
```{r}
oceanak_geno %>%
  dplyr::filter(genotyped == FALSE) %>%
  dplyr::count(stream, year, otolith_mark_present) %>%
  tidyr::pivot_wider(names_from = otolith_mark_present,
                     values_from = n,
                     values_fill =  0)
```

## Plot samples by genotyped vs. not

```{r fig.width = 12}
oceanak_geno %>%
  ggplot2::ggplot(aes(x = DOY, fill = genotyped)) +
  ggplot2::geom_bar(colour = "black") +  # group = interaction(DOY, origin))
  ggplot2::scale_x_continuous(breaks = seq(180, 320, by = 5),
                              labels = format(x = (
                                as.Date("2013-12-31") + seq(180, 320, by = 5)
                              ), "%b %d")) +
  ggplot2::facet_grid(year ~ stream) +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::ylab("Number of Samples") +
  ggplot2::xlab("") +
  ggplot2::labs(title = "AHRP Samples Collected by Day - Genotype Status 12/13/22", fill = "Genotyped") +
  ggplot2::theme_bw() +
  ggplot2::theme(text = element_text(size = 15),
                 axis.text.x = element_text(angle = 90)) +
  ggplot2::scale_fill_manual(values = c("grey20", "grey80"))
```

## Entire collections not genotyped

  * **PERB13** - previously decided to deep 6 2013 samples due to low offspring assignment rates in Stockdale/Hogan and non-representative sampling throughout the season
  * **PERB15** - previously decided that there were more parents than necessary, subsampled post-GOD
  * **PERB17** - previously decided to wait for post-GOD to filter out new hatchery strays
  * **PERB20** - previously waiting on otolith reads, but we have them now!
  * **PPADDY13** - previously decided to deep 6 2013 samples due to low offspring assignment rates in Stockdale/Hogan and non-representative sampling throughout the season
  * **PPADDY15** - previously decided that there were more parents than necessary, subsampled post-GOD
  * **PPADDY17** - previously decided to wait for post-GOD to filter out new hatchery strays
  * **PPADDY20** - previously waiting on otolith reads, but we have them now!

## Collections partially genotyped

### Hatchery strays excluded in final year (lineage)

  * **PGILMOUR19** - last year for odd-lineage, excluded hatchery strays
  * **PHOGAN18** - last year for even-lineage, excluded hatchery strays
  * **PHOGAN19** - last year for odd-lineage, excluded hatchery strays
  * **PSTOCK19** - last year for odd-lineage, excluded hatchery strays

### Revisit?

  * **PHOGAN14** - previously decided that we didn't need more hatchery-origin parents, already have 500-ish, there are **1,820** remaining, but they could help with F2 analyses
  * **PHOGAN17** - previously extracted all natural-origin available(offspring of 2015) + 1K hatchery-origin parents, looks like there are more natural-origin now because at the time, only every other DWP got otolith read (there were so many, so they skipped every other one at the time), there are 3,145 non-genotyped natural-origin fish and 10,193 non-genotyped hatchery-origin fish
  * **PSTOCK14** - previously decided that we didn't need more hatchery-origin parents, already have have 500-ish, there are **557** remaning

# Samples we should genotype

## eP013 - 2014 Stockdale and Hogan

Extract these left-over 2014 hatchery-origin parents to aid in finding more offspring in 2016 for subsequent F2 analyses in 2018. Unfortunately this will involve going back through previously cherry picked DPWs.
```{r}
(
  eP013 <- oceanak_geno %>%
    dplyr::filter(
      genotyped == FALSE,
      stream %in% c("Stockdale", "Hogan"),
      year == 2014,
      origin == "Hatchery",
      is_missing != 1,
      more_than_one != 1,
      sex %in% c("Male", "Female")
    )
)
```

## eP014 - 2020 Paddy and Erb

Extract these natural-origin offspring now that we have all of the otolith reads back. This is the final year of even-lineage analyses, so no need to grab hatchery-origin or `NA` otolith fish.
```{r}
(
  eP014 <- oceanak_geno %>%
    dplyr::filter(
      genotyped == FALSE,
      stream %in% c("Paddy", "Erb"),
      year == 2020,
      origin == "Natural",
      is_missing != 1,
      more_than_one != 1,
      sex %in% c("Male", "Female")
    )
)
```

Verify that all of the `NA` otolith fish have been read

**WAIT** on this until Tim Frawley confirms what these different codes mean...we want to make sure we have "all the reads" before we go picking samples to extract.
```{r}
oceanak_geno %>%
  dplyr::filter(genotyped == FALSE,
                stream %in% c("Paddy", "Erb"),
                year == 2020,
                is.na(origin)) %>%
  dplyr::count(stream, otolith_mark_status_code) %>%
  tidyr::pivot_wider(names_from = stream, values_from = n)
```

Hrmm, I should follow up on `3N66` - missing (could be *true missing* or *GOD-affected*) and `n` - null (dunno how an otolith gets this code)

## eP015 - 2017 Hogan

At the time this collection was originally extracted, we only had otolith reads from every-other DWP, so we grabbed all of the natural-origin fish and 1K hatchery-origin fish. Now that all of the DWPs have been read, I propose that we genotype all fish from that 2nd round of otolith reads (i.e. all the "new" natural-origin fish + all the "new" hatchery-origin fish). This will get us more natural-origin fish as offspring for the 2015/2017 F~1~ analyses and 2015/2017/2019 F~2~ analyses, and also give us a lot more hatchery-origin parents for the 2017/2019 F~1~ analyses. I do not think it is worth going back to genotype all of the non-1K hatchery-origin fish that were originally passed over.

How many previously genotyped
```{r}
oceanak_geno %>% 
  dplyr::filter(stream == "Hogan",
                year == 2017,
                genotyped == TRUE)
```

What should we genotype now?
```{r}
hogan_2017_DWPs_to_genotype <- oceanak_geno %>%
  dplyr::filter(stream == "Hogan",
                year == 2017) %>%
  dplyr::group_by(dna_tray_code) %>%
  dplyr::summarise(genotyped = sum(genotyped)) %>%
  dplyr::filter(genotyped == 0) %>%
  dplyr::pull(dna_tray_code)

(
  eP015 <- oceanak_geno %>%
    dplyr::filter(
      dna_tray_code %in% hogan_2017_DWPs_to_genotype,
      is.na(is_missing) | is_missing != 1,
      is.na(more_than_one) | more_than_one != 1,
      sex %in% c("Male", "Female")
    )
)
```

Created this extraction list in `ExtractionSelection.R` on 4/18/22.

## eP016 - 2015 Paddy and Erb

Extract all, despite GOD. Originally we were going to wait until GOD is done, but it’s gonna take too long. These will be parents for the odd-lineage 2015/2017 F~1~ analyses. Ideally we could filter out `NA` origin fish, if they have been otolith read, but this is likely fairly few samples.
```{r}
(
  eP016 <- oceanak_geno %>%
    dplyr::filter(
      genotyped == FALSE,
      stream %in% c("Paddy", "Erb"),
      year == 2015,
      is.na(is_missing) | is_missing != 1,
      is.na(more_than_one) | more_than_one != 1,
      sex %in% c("Male", "Female")
    )
)
```

Not done yet!!! Need to remove GOD-affected DWPs that will be in eP018.

Updating to remove eP018 fish, should remove 705 fish from PERB15, leaving us with at least 20,950 fish.
```{r}
(
  eP016 <- oceanak_geno %>%
    dplyr::filter(
      genotyped == FALSE,
      stream %in% c("Paddy", "Erb"),
      year == 2015,
      is.na(is_missing) | is_missing != 1,
      is.na(more_than_one) | more_than_one != 1,
      sex %in% c("Male", "Female"),
      !dna_tray_code %in% GOD_affected_DWPs_consensus
    )
)
```

Created this extraction list in `ExtractionSelection.R` on 4/19/22.

## eP017 - 2017 Paddy and Erb

Extract all natural-origin and `NA` origin, despite GOD. Originally we were going to wait until GOD is done, but it’s gonna take too long. These will be offspring for the odd-lineage 2015/2017 F~1~ analyses. Ideally we could filter out `NA` origin fish, if they have been otolith read, but this is likely fairly few samples. We do **NOT** need any hatchery-origin individuals.
```{r}
(
  eP017 <- oceanak_geno %>%
    dplyr::filter(
      genotyped == FALSE,
      stream %in% c("Paddy", "Erb"),
      year == 2017,
      is.na(origin) | origin == "Natural",
      is.na(is_missing) | is_missing != 1,
      is.na(more_than_one) | more_than_one != 1,
      sex %in% c("Male", "Female")
    )
)
```

Not done yet!!! Need to remove GOD-affected DWPs that will be in eP018.

Updating to remove eP018 fish, should remove 1,053 fish from PERB17, leaving us with at least 17,648 fish.
```{r}
(
  eP017 <- oceanak_geno %>%
    dplyr::filter(
      genotyped == FALSE,
      stream %in% c("Paddy", "Erb"),
      year == 2017,
      is.na(origin) | origin == "Natural",
      is.na(is_missing) | is_missing != 1,
      is.na(more_than_one) | more_than_one != 1,
      sex %in% c("Male", "Female"),
      !dna_tray_code %in% GOD_affected_DWPs_consensus
    )
)
```

Created this extraction list in `ExtractionSelection.R` on 4/19/22.

## eP018 - GOD Hearts 2015/2017 Paddy and Erb

In order to try to resolve GOD prior to Chris' retirement in February 2023, we are going to prioritize genotyping of all remaining hearts involved in GOD. This means that we want to grab all of the 2015 and 2017 Paddy and Erb DWPs **with affected otoliths** and remove them from eP016 and eP017 to get them in front of the line for extraction and genotyping.

### Otolith Transfer 11/17/21

First read in the GOD data from the MTAL transfer sheet to figure out which DWPs were affected.
```{r}
(otolith_transfer <- readr::read_csv(file = "GOD/otolith_transfer_111721.csv"))
```

Find unique, affected DWPs
```{r}
(GOD_affected_DWPs <- otolith_transfer %>% 
   dplyr::mutate(DWP = stringr::str_pad(string = `48_DWP_tray_barcode`, width = 10, side = "left", pad = "0")) %>% 
   dplyr::filter(!is.na(DWP)) %>% 
   dplyr::distinct(DWP) %>% 
   dplyr::pull(DWP)
)
```

Are all of these in *OceanAK*?
```{r}
table(GOD_affected_DWPs %in% oceanak_geno$dna_tray_code)
```

Ugh, one of these isn't even in *OceanAK*

Which one is not in *OceanAK*?
```{r}
setdiff(GOD_affected_DWPs, oceanak_geno$dna_tray_code)
```

How many otos?
```{r}
otolith_transfer %>% 
   dplyr::mutate(DWP = stringr::str_pad(string = `48_DWP_tray_barcode`, width = 10, side = "left", pad = "0")) %>% 
   dplyr::filter(DWP %in% setdiff(GOD_affected_DWPs, oceanak_geno$dna_tray_code))
```

#### Change 14037 to 14307

Looks like we found a typo, *14037* is wrong and should be *14307*.
```{r}
"0000014307" %in% oceanak_geno$dna_tray_code
```

One 4/19/22, Kyle corrected this error in 3 files:
  * "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\GOD\otolith_transfer_96_template_LEFT_all trays_17Nov21.xlsx"
  * "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\GitHub-PWS-Pink-Parentage\GOD\otolith_transfer_111721.csv"
  * "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\GitHub-PWS-Pink-Parentage\GOD\otolith_transfer_111721.xlsx"

#### Re-read otolith_transfer_111721.csv

Re-read in the *corrected* GOD data from the MTAL transfer sheet to figure out which DWPs were affected.
```{r}
(otolith_transfer <- readr::read_csv(file = "GOD/otolith_transfer_111721.csv"))
```

Find unique, affected DWPs
```{r}
(GOD_affected_DWPs <- otolith_transfer %>% 
   dplyr::mutate(DWP = stringr::str_pad(string = `48_DWP_tray_barcode`, width = 10, side = "left", pad = "0")) %>% 
   dplyr::filter(!is.na(DWP)) %>% 
   dplyr::distinct(DWP) %>% 
   dplyr::pull(DWP)
)
```

Are all of these in *OceanAK*?
```{r}
table(GOD_affected_DWPs %in% oceanak_geno$dna_tray_code)
```

Great, all are in *OceanAK*

### working_oto_sheet_combined

Reading in the GOD data from the DWP investigations to cross check the transfer sheet.
```{r}
(working_oto_sheet_combined <- readxl::read_xlsx(path = "../GOD/working_oto_sheet_combined.xlsm", sheet = "oto unpacking data"))
```

Find unique, affected DWPs
```{r}
(GOD_affected_DWPs_v2 <- working_oto_sheet_combined %>% 
   dplyr::mutate(DWP = stringr::str_pad(string = DWPbarcode, width = 10, side = "left", pad = "0")) %>% 
   dplyr::filter(genotype_otot == "Y") %>% 
   dplyr::distinct(DWP) %>% 
   dplyr::pull(DWP)
)
```

Are all of these in *OceanAK*?
```{r}
table(GOD_affected_DWPs_v2 %in% oceanak_geno$dna_tray_code)
```

Great, all are in *OceanAK* (425 DWPs)

### Cross Reference

Which DWPs are in the `otolith_transfer_111721.csv` spreadsheet, but were not originally flagged in the `working_oto_sheet_combined.xlsm`?
```{r}
setdiff(GOD_affected_DWPs, GOD_affected_DWPs_v2)
```

Looks like there are 2 DWPs in `otolith_transfer_111721.csv` that were **NOT** flagged in `working_oto_sheet_combined.xlsm`
```{r}
working_oto_sheet_combined %>% 
   dplyr::mutate(DWP = stringr::str_pad(string = DWPbarcode, width = 10, side = "left", pad = "0")) %>% 
   dplyr::filter(DWP %in% setdiff(GOD_affected_DWPs, GOD_affected_DWPs_v2))
```

Huh, 22770 is not in `working_oto_sheet_combined.xlsm`

Upon review of `otolith_transfer_111721.csv`, there is a typo, 22770 should be 22700
```{r}
oceanak_geno %>% 
  dplyr::filter(dna_tray_code %in% c("0000022770", "0000022700")) %>% 
  dplyr::distinct(silly, dna_tray_code)
```

Ah, so there was a typo, because none of the PSTOCK17 DWPs should have been in GOD.

#### Change 22770 to 22700

Looks like we found a typo, *22770* is wrong and should be *22700*.
```{r}
"0000014307" %in% oceanak_geno$dna_tray_code
```

One 4/19/22, Kyle corrected this error in 3 files:
  * "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\GOD\otolith_transfer_96_template_LEFT_all trays_17Nov21.xlsx"
  * "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\GitHub-PWS-Pink-Parentage\GOD\otolith_transfer_111721.csv"
  * "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\GitHub-PWS-Pink-Parentage\GOD\otolith_transfer_111721.xlsx"


#### Re-read otolith_transfer_111721.csv

Re-read in the *corrected* GOD data from the MTAL transfer sheet to figure out which DWPs were affected.
```{r}
(otolith_transfer <- readr::read_csv(file = "GOD/otolith_transfer_111721_updated_save_me.csv"))
```

Find unique, affected DWPs
```{r}
(GOD_affected_DWPs <- otolith_transfer %>% 
   dplyr::mutate(DWP = stringr::str_pad(string = `48_DWP_tray_barcode`, width = 10, side = "left", pad = "0")) %>% 
   dplyr::filter(!is.na(DWP)) %>% 
   dplyr::distinct(DWP) %>% 
   dplyr::pull(DWP)
)
```

Are all of these in *OceanAK*?
```{r}
table(GOD_affected_DWPs %in% oceanak_geno$dna_tray_code)
```

Great, all are in *OceanAK* (now 421)

Okay, so now redo the cross reference

Which DWPs are in the `otolith_transfer_111721.csv` spreadsheet, but were not originally flagged in the `working_oto_sheet_combined.xlsm`?
```{r}
setdiff(GOD_affected_DWPs, GOD_affected_DWPs_v2)
```

Okay, just 2164 which is in `working_oto_sheet_combined.xlsm`, but not flagged for genotyping. We'll just want to include it in eP018 (if it is from Paddy or Erb 2015/2017).

Okay, howe abotu the other way now....

Which DWPs are flagged in the `working_oto_sheet_combined.xlsm`, but not in the `otolith_transfer_111721.csv` spreadsheet?
```{r}
setdiff(GOD_affected_DWPs_v2, GOD_affected_DWPs)
```

We know that all of these are in *OceanAK* (see above), but why are they not in `otolith_transfer_111721.csv`?
```{r}
working_oto_sheet_combined %>% 
   dplyr::mutate(DWP = stringr::str_pad(string = DWPbarcode, width = 10, side = "left", pad = "0")) %>% 
   dplyr::filter(DWP %in% setdiff(GOD_affected_DWPs_v2, GOD_affected_DWPs))
```

Hrmm, which silly codes?
```{r}
oceanak_geno %>% 
  dplyr::filter(dna_tray_code %in% setdiff(GOD_affected_DWPs_v2, GOD_affected_DWPs)) %>% 
  dplyr::distinct(silly, dna_tray_code)
```

Interesting, well none of these are in Paddy/Erb 2015/2017, so kinda moot re: eP018

### Consensus GOD affected DWPs

Venn diagram
```{r}
ggVennDiagram::ggVennDiagram(x = list("otolith_transfer" = GOD_affected_DWPs,
                                      "working_oto_sheet" = GOD_affected_DWPs_v2))
```

Generate a *consensus* vector with all unique DWPs (426)
```{r}
(GOD_affected_DWPs_consensus <- sort(unique(c(GOD_affected_DWPs, GOD_affected_DWPs_v2))))
```

Save this vector so we can use it in `ExtractionSelection.R`!!!
```{r}
save_objects(objects = "GOD_affected_DWPs_consensus", path = "GOD")
```

What is the silly code breakdown of GOD?
```{r}
oceanak_geno %>% 
  dplyr::filter(dna_tray_code %in% GOD_affected_DWPs_consensus) %>% 
  dplyr::distinct(silly, dna_tray_code) %>% 
  dplyr::count(silly)
```

Oh, interesting, that's not as bad as we thought! Unfortunately, most of them are Gilmour. However, for eP018, there are only 37 DWPs total, all Erb!

### Final

How many of these DWPs are Paddy/Erb 2015/2017?
```{r}
oceanak_geno %>% 
  dplyr::filter(dna_tray_code %in% GOD_affected_DWPs_consensus,
                silly %in% c("PPADDY15", "PPADDY17", "PERB15", "PERB17")) %>% 
  dplyr::distinct(silly, dna_tray_code) %>% 
  dplyr::count(silly)
```

How many of these fish are Paddy/Erb 2015/2017?
```{r}
oceanak_geno %>% 
  dplyr::filter(dna_tray_code %in% GOD_affected_DWPs_consensus,
                silly %in% c("PPADDY15", "PPADDY17", "PERB15", "PERB17")) %>% 
  dplyr::count(silly)
```

Damn, not bad, only 1,776 total fish!

How many would we actually want to extract (with all of our filters)?
```{r}
(
  eP018 <- oceanak_geno %>%
    dplyr::filter(
      genotyped == FALSE,
      stream %in% c("Paddy", "Erb"),
      year %in% c(2015, 2017),
      is.na(is_missing) | is_missing != 1,
      is.na(more_than_one) | more_than_one != 1,
      sex %in% c("Male", "Female"),
      dna_tray_code %in% GOD_affected_DWPs_consensus
    )
)

eP018 %>% 
  dplyr::count(silly)
```

Extract these 1,758 fish!

# Summary of genotype status 2/9/22

```{r}
oceanak_geno %>% 
  dplyr::count(genotyped)
```


We have genotyped 133,369 fish.

There are 102,269 fish that have not been genotyped. Some should be, some (mainly hatchery-origin strays in the final offspring year) do not need to be genotyped.

# Magnitude of GOD

How many samples do **not** have otolith reads per stream/year (based on `NA` for `origin`, will include otoliths that were unreadable/lost)?
```{r}
oceanak_geno %>%
  dplyr::filter(is.na(origin)) %>% 
  dplyr::count(year, stream) %>%
  tidyr::pivot_wider(names_from = year,
                     values_from = n,
                     values_fill =  0)
```

How many samples do **not** have otolith reads per stream/year (based on `NA` for `otolith_mark_status_code`)?
```{r}
oceanak_geno %>%
  dplyr::filter(otolith_read == FALSE) %>% 
  dplyr::count(year, stream) %>%
  tidyr::pivot_wider(names_from = year,
                     values_from = n,
                     values_fill =  0) %>% 
  dplyr::arrange(stream)
```

Why are there so few for Gilmour?
```{r}
oceanak_geno %>% 
  dplyr::filter(stream == "Gilmour",
                is.na(origin)) %>% 
  dplyr::count(year, otolith_mark_status_code) %>% 
  tidyr::pivot_wider(names_from = year, values_from = n)
```

Okay, this is a global issue, need to check with Tim to see what these mean...(update, see above, unresolved, but I don't think we'll be able to figure out which otoliths have truly not been read and which are *GOD-affected*)
```{r rows.print = 22}
oceanak_geno %>% 
  dplyr::filter(is.na(otolith_mark_present)) %>% 
  dplyr::count(otolith_mark_status_code) %>% 
  dplyr::arrange(dplyr::desc(n))
```

Okay, something is not right. I have no idea what "n" is, but I think 3N66 is "missing"? Evidently my attempt to get a TRUE/FALSE for `otolith_read` based on whether `otolith_mark_status_code` = `NA` is not the way to go about this...

## Plot samples by stream/year/origin

```{r fig.width = 12}
oceanak_geno %>%
  ggplot2::ggplot(aes(x = DOY, fill = origin)) +
  ggplot2::geom_bar(colour = "black") +  # group = interaction(DOY, origin))
  ggplot2::scale_x_continuous(breaks = seq(180, 320, by = 5),
                              labels = format(x = (
                                as.Date("2013-12-31") + seq(180, 320, by = 5)
                              ), "%b %d")) +
  ggplot2::facet_grid(year ~ stream) +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::ylab("Number of Samples") +
  ggplot2::xlab("") +
  ggplot2::labs(title = "AHRP Samples Collected by Day - Otolith Status", fill = "Origin") +
  ggplot2::theme_bw() +
  ggplot2::theme(text = element_text(size = 15),
                 axis.text.x = element_text(angle = 90))
```

### Erb

Which samples were affected by GOD (`NA` origin)?

**NOTE** `NA` for origin could be due to a number of non-GOD issues (overground, true missing, etc.), this is just a coarse estimate.
```{r fig.width = 12}
oceanak_geno %>%
  dplyr::filter(stream == "Erb",
                year %in% 2015:2018) %>% 
  ggplot2::ggplot(aes(x = DOY, fill = origin)) +
  ggplot2::geom_bar(colour = "black") +  # group = interaction(DOY, origin))
  ggplot2::scale_x_continuous(breaks = seq(180, 320, by = 5),
                              labels = format(x = (
                                as.Date("2013-12-31") + seq(180, 320, by = 5)
                              ), "%b %d")) +
  ggplot2::facet_grid(year ~ stream) +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::ylab("Number of Samples") +
  ggplot2::xlab("") +
  ggplot2::labs(title = "Erb Samples Collected by Day - Otolith Status", fill = "Origin") +
  ggplot2::theme_bw() +
  ggplot2::theme(text = element_text(size = 15),
                 axis.text.x = element_text(angle = 90))
```

Compare density plot of otoliths that have been read vs. those that haven't (i.e. GOD). If no difference in distributions, then GOD has not biased our field collections.
```{r fig.width = 12}
oceanak_geno %>%
  dplyr::filter(stream == "Erb",
                year %in% 2015:2018) %>%
  dplyr::mutate(otolith_read = dplyr::case_when(is.na(origin) ~ FALSE, TRUE ~ TRUE)) %>%
  ggplot2::ggplot(aes(x = DOY, fill = otolith_read)) +
  ggplot2::geom_density(alpha = 0.5) +
  ggplot2::scale_x_continuous(breaks = seq(180, 320, by = 5),
                              labels = format(x = (
                                as.Date("2013-12-31") + seq(180, 320, by = 5)
                              ), "%b %d")) +
  ggplot2::facet_grid(year ~ stream) +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::ylab("Number of Samples") +
  ggplot2::xlab("") +
  ggplot2::labs(title = "Erb Samples Collected by Day - Otolith Status", fill = "Otolith Read") +
  ggplot2::theme_bw() +
  ggplot2::theme(text = element_text(size = 15),
                 axis.text.x = element_text(angle = 90)) +
  ggplot2::scale_fill_manual(values = c("grey20", "grey80"))
```

### Paddy

Which samples were affected by GOD (`NA` origin)?

**NOTE** `NA` for origin could be due to a number of non-GOD issues (overground, true missing, etc.), this is just a coarse estimate.
```{r fig.width = 12}
oceanak_geno %>%
  dplyr::filter(stream == "Paddy",
                year %in% 2015:2018) %>% 
  ggplot2::ggplot(aes(x = DOY, fill = origin)) +
  ggplot2::geom_bar(colour = "black") +  # group = interaction(DOY, origin))
  ggplot2::scale_x_continuous(breaks = seq(180, 320, by = 5),
                              labels = format(x = (
                                as.Date("2013-12-31") + seq(180, 320, by = 5)
                              ), "%b %d")) +
  ggplot2::facet_grid(year ~ stream) +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::ylab("Number of Samples") +
  ggplot2::xlab("") +
  ggplot2::labs(title = "Paddy Samples Collected by Day - Otolith Status", fill = "Origin") +
  ggplot2::theme_bw() +
  ggplot2::theme(text = element_text(size = 15),
                 axis.text.x = element_text(angle = 90))
```

Compare density plot of otoliths that have been read vs. those that haven't (i.e. GOD). If no difference in distributions, then GOD has not biased our field collections.
```{r fig.width = 12}
oceanak_geno %>%
  dplyr::filter(stream == "Paddy",
                year %in% 2015:2018) %>%
  dplyr::mutate(otolith_read = dplyr::case_when(is.na(origin) ~ FALSE, TRUE ~ TRUE)) %>%
  ggplot2::ggplot(aes(x = DOY, fill = otolith_read)) +
  ggplot2::geom_density(alpha = 0.5) +
  ggplot2::scale_x_continuous(breaks = seq(180, 320, by = 5),
                              labels = format(x = (
                                as.Date("2013-12-31") + seq(180, 320, by = 5)
                              ), "%b %d")) +
  ggplot2::facet_grid(year ~ stream) +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::ylab("Number of Samples") +
  ggplot2::xlab("") +
  ggplot2::labs(title = "Paddy Samples Collected by Day - Otolith Status", fill = "Otolith Read") +
  ggplot2::theme_bw() +
  ggplot2::theme(text = element_text(size = 15),
                 axis.text.x = element_text(angle = 90)) +
  ggplot2::scale_fill_manual(values = c("grey20", "grey80"))
```

### Gilmour

Which samples were affected by GOD (`NA` origin)?

**NOTE** `NA` for origin could be due to a number of non-GOD issues (overground, true missing, etc.), this is just a coarse estimate.
```{r fig.width = 12}
oceanak_geno %>%
  dplyr::filter(stream == "Gilmour",
                year %in% 2015:2018) %>% 
  ggplot2::ggplot(aes(x = DOY, fill = origin)) +
  ggplot2::geom_bar(colour = "black") +  # group = interaction(DOY, origin))
  ggplot2::scale_x_continuous(breaks = seq(180, 320, by = 5),
                              labels = format(x = (
                                as.Date("2013-12-31") + seq(180, 320, by = 5)
                              ), "%b %d")) +
  ggplot2::facet_grid(year ~ stream) +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::ylab("Number of Samples") +
  ggplot2::xlab("") +
  ggplot2::labs(title = "Gilmour Samples Collected by Day - Otolith Status", fill = "Origin") +
  ggplot2::theme_bw() +
  ggplot2::theme(text = element_text(size = 15),
                 axis.text.x = element_text(angle = 90))
```

Compare density plot of otoliths that have been read vs. those that haven't (i.e. GOD). If no difference in distributions, then GOD has not biased our field collections.
```{r fig.width = 12}
oceanak_geno %>%
  dplyr::filter(stream == "Gilmour",
                year %in% 2015:2018) %>%
  dplyr::mutate(otolith_read = dplyr::case_when(is.na(origin) ~ FALSE, TRUE ~ TRUE)) %>%
  ggplot2::ggplot(aes(x = DOY, fill = otolith_read)) +
  ggplot2::geom_density(alpha = 0.5) +
  ggplot2::scale_x_continuous(breaks = seq(180, 320, by = 5),
                              labels = format(x = (
                                as.Date("2013-12-31") + seq(180, 320, by = 5)
                              ), "%b %d")) +
  ggplot2::facet_grid(year ~ stream) +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::ylab("Number of Samples") +
  ggplot2::xlab("") +
  ggplot2::labs(title = "Gilmour Samples Collected by Day - Otolith Status", fill = "Otolith Read") +
  ggplot2::theme_bw() +
  ggplot2::theme(text = element_text(size = 15),
                 axis.text.x = element_text(angle = 90)) +
  ggplot2::scale_fill_manual(values = c("grey20", "grey80"))
```

# Summary of GOD

While the total proportion of samples affected by GOD in a given stream/year varies quite a bit (e.g. hardly anything for Paddy 2015 vs. significant loss for all Gilmour), the loss of samples to GOD was non-random, and appears to bias the density distribution of sampling date relative to what was collected in the field (and is assumed to be representative).

End