---
title: "Hogan_popgen"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(DT)
library(abind)
bbind <- function(...) { abind(..., along = 3) }

source("~/../R/Functions.GCL.R")

load_objects(path = "../Objects/")
load_sillys(path = "../Genotypes/OceanAK_Origin_PostQA/")
```

# Introduction

This is a quick exploratory analysis of the population structure of the NPRB Hogan Bay 2013-2016 genotype data, post-QA. Here we will:

  * Read in all post-QA project genotypes
  * Create a *genepop* file to read back in with *adegenet*
  * Explore genetic structure
    + Among brood lines (even/odd)
    + Among years
    + Among hatchery/wild (note different hatcheries have different brood sources)
  * Export genotypes as a *Genepop* file to feed in to *genepop2franz.GCL.r*
  
# Create genepop

Create a *genepop* using `gcl2Genepop.GCL`, individuals are named by *SillySource*.
```{r genepop}
gcl2Genepop.GCL(sillyvec = c("PHOGAN13n", "PHOGAN13h", "PHOGAN14n", "PHOGAN14h", "PHOGAN15", "PHOGAN16"), 
                loci = loci, 
                path = "../Genepop/hogan_postQA.gen", 
                VialNums = TRUE, 
                usat = FALSE)
```

```{r combine attributes tables}
#When we pooled PHOGAN 13 and 14, the FK_FISH_ID column was converted from numeric to character. However, it remained numeric for PHOGAN 15 and 16. Its structure needs to be uniform across tables to allow for combining. 

PHOGAN15.gcl$attributes <- PHOGAN15.gcl$attributes %>% 
  mutate(FK_FISH_ID = as.character(FK_FISH_ID))
PHOGAN16.gcl$attributes <- PHOGAN16.gcl$attributes %>% 
  mutate(FK_FISH_ID = as.character(FK_FISH_ID))
HOGAN_attributes_tibble <- dplyr::bind_rows(PHOGAN13n.gcl$attributes, PHOGAN13h.gcl$attributes, PHOGAN14n.gcl$attributes, PHOGAN14h.gcl$attributes, PHOGAN15.gcl$attributes, PHOGAN16.gcl$attributes)
```


# Analysis in *adegenet*

```{r adegenet}
install.packages("adegenet")
library("adegenet")

Hogan <- read.genepop(file="v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Genepop/hogan_postQA.gen")
Hogan 

Hoganpop <- genind2genpop(Hogan)

Hoganpop@tab[1:6, 1:7] #Shows population names.  


```

##Inbreeding Estimates

Here we will:
1. Create separate genind files for each year and origin.
2. Estimate and plot inbreeding (F).
3. Calculate and plot mean F across individuals.
4. Estimate and plot probability density functions. 
5. Estimate and plot MLE.
6. Compare MLE with observed values. 


```{r estimate inbreeding}
#From adgenet manual - see p 106

#isolate populations
PHOGAN13n_genind <- seppop(Hogan)$PHOGAN13_813
PHOGAN13h_genind <- seppop(Hogan)$PHOGAN13_829
PHOGAN14n_genind <- seppop(Hogan)$PHOGAN14_2649
PHOGAN14h_genind <- seppop(Hogan)$PHOGAN14_2647
PHOGAN15_genind <- seppop(Hogan)$PHOGAN15_9409
PHOGAN16_genind <- seppop(Hogan)$PHOGAN16_13054

#Estimate inbreeding and return sample of F values
Fs_13n <- inbreeding(PHOGAN13n_genind, N=30)
Fs_13h <- inbreeding(PHOGAN13h_genind, N=30)
Fs_14n <- inbreeding(PHOGAN14n_genind, N=30)
Fs_14h <- inbreeding(PHOGAN14h_genind, N=30)
Fs_15 <- inbreeding(PHOGAN15_genind, N=30)
Fs_16 <- inbreeding (PHOGAN16_genind, N=30)

#plot first 10 results
invisible(sapply(Fs_13n[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of Sampled F Values"))) 
invisible(sapply(Fs_13h[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of sampled F Values")))
invisible(sapply(Fs_14n[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of sampled F Values")))
invisible(sapply(Fs_14h[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of sampled F Values")))
invisible(sapply(Fs_15[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of sampled F Values")))
invisible(sapply(Fs_16[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of sampled F Values")))

```

```{r compute mean F}
#compute means for all individuals
Fmean_13n <- sapply(Fs_13n, mean) 
Fmean_13h <- sapply(Fs_13h, mean)
Fmean_14n <- sapply(Fs_14n, mean)
Fmean_14h <- sapply(Fs_14h, mean)
Fmean_15 <- sapply(Fs_15, mean)
Fmean_16 <- sapply(Fs_16, mean)

par(mfrow=c(3,3))
hist(Fmean_13n, col="orange", xlab="Mean value of F", main="2013 natural")
hist(Fmean_13h, col="orange", xlab="Mean value of F", main="2013 hatchery")
hist(Fmean_14n, col="Orange", xlab="Mean value of F", main="2014 natural")
hist(Fmean_14h, col="orange", xlab="Mean value of F", main="2014 hatchery")
hist(Fmean_15, col="orange", xlab="Mean value of F", main="2015")
hist(Fmean_16, col="orange", xlab="Mean value of F", main="2016")
```

There appear to be potential qualitative differences in F between lineages, but not origins. 

```{r estimate inbreeding and probability density functions}
#estimate inbreeding and return probability density functions
Fdens_13n <- inbreeding(PHOGAN13n_genind, res.type = "function")
Fdens_13h <- inbreeding(PHOGAN13h_genind, res.type = "function")
Fdens_14n <- inbreeding(PHOGAN14n_genind, res.type = "function")
Fdens_14h <- inbreeding(PHOGAN14h_genind, res.type = "function")
Fdens_15 <- inbreeding(PHOGAN15_genind, res.type = "function")
Fdens_16 <- inbreeding(PHOGAN16_genind, res.type = "function")

#plot the first 10 functions
invisible(sapply(Fdens_13n[1:10], plot, ylab="Density", main="Density of probability of F values")) 
invisible(sapply(Fdens_13h[1:10], plot, ylab="Density", main="Density of probability of F values"))
invisible(sapply(Fdens_14n[1:10], plot, ylab="Density", main="Density of probability of F values"))
invisible(sapply(Fdens_14h[1:10], plot, ylab="Density", main="Density of probability of F values"))
invisible(sapply(Fdens_15[1:10], plot, ylab="Density", main="Density of probability of F values"))
invisible(sapply(Fdens_16[1:10], plot, ylab="Density", main="Density of probability of F values"))
```

```{r estimate inbreeding and maximum likelihood estimates}
#estimate inbreeding and return maximum likelihood estimates
Fest_13n <- inbreeding(PHOGAN13n_genind, res.type = "estimate")
Fest_13h <- inbreeding(PHOGAN13h_genind, res.type = "estimate")
Fest_14n <- inbreeding(PHOGAN14n_genind, res.type = "estimate")
Fest_14h <- inbreeding(PHOGAN14h_genind, res.type = "estimate")
Fest_15 <- inbreeding(PHOGAN15_genind, res.type = "estimate")
Fest_16 <- inbreeding(PHOGAN16_genind, res.type = "estimate")

mostinbred_13n <- which.max(Fest_13n)
mostinbred_13h <- which.max(Fest_13h)
mostinbred_14n <- which.max(Fest_14n)
mostinbred_14h <- which.max(Fest_14h)
mostinbred_15 <- which.max(Fest_15)
mostinbred_16 <- which.max(Fest_16)

par(mfrow=c(3,3))
plot(Fdens_13n[[mostinbred_13n]], ylab="Density", xlab="F", main = paste("Probability density of F values 2013 natural", names(mostinbred_13n)))
abline(v=Fest_13n[mostinbred_13n], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
plot(Fdens_13h[[mostinbred_13h]], ylab="Density", xlab="F", main = paste("Probability density of F values 2013 hatchery", names(mostinbred_13h)))
abline(v=Fest_13h[mostinbred_13h], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
plot(Fdens_14n[[mostinbred_14n]], ylab="Density", xlab="F", main = paste("Probability density of F values 2014 natural", names(mostinbred_14n)))
abline(v=Fest_14n[mostinbred_14n], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
plot(Fdens_14h[[mostinbred_14h]], ylab="Density", xlab="F", main = paste("Probability density of F values 2014 hatchery", names(mostinbred_14h)))
abline(v=Fest_14h[mostinbred_14h], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
plot(Fdens_15[[mostinbred_15]], ylab="Density", xlab="F", main = paste("Probability density of F values 2015", names(mostinbred_15)))
abline(v=Fest_15[mostinbred_15], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
plot(Fdens_16[[mostinbred_16]], ylab="Density", xlab="F", main = paste("Probability density of F values 2016", names(mostinbred_16)))
abline(v=Fest_16[mostinbred_16], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
```

```{r compare MLE and means}
#compare MLE and means
par(mfrow=c(3,3))
plot(Fest_13n, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_13n, pch=2, col="red")
arrows(x0=1:length(Fest_13n), y0=Fest_13n, y1=Fmean_13n, x1=1:length(Fest_13n), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")
plot(Fest_13h, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_13h, pch=2, col="red")
arrows(x0=1:length(Fest_13h), y0=Fest_13h, y1=Fmean_13h, x1=1:length(Fest_13h), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")
plot(Fest_14n, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_14n, pch=2, col="red")
arrows(x0=1:length(Fest_14n), y0=Fest_14n, y1=Fmean_14n, x1=1:length(Fest_14n), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")
plot(Fest_14h, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_14h, pch=2, col="red")
arrows(x0=1:length(Fest_14h), y0=Fest_14h, y1=Fmean_14h, x1=1:length(Fest_14h), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")
plot(Fest_15, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_15, pch=2, col="red")
arrows(x0=1:length(Fest_15), y0=Fest_15, y1=Fmean_15, x1=1:length(Fest_15), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")
plot(Fest_16, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_16, pch=2, col="red")
arrows(x0=1:length(Fest_16), y0=Fest_16, y1=Fmean_16, x1=1:length(Fest_16), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")

```

##Fst

Here we will:
1. Calculate Fst, Fit, and Fis.
2. Calculate PW Fst across year*origin. 
3. Calculate Fst for each locus. 

```{r Calculate F statistics}
install.packages("hierfstat")
library(hierfstat)
fstat(Hogan)
```

```{r Calculate PW Fst}
PW_Fst_Hogan <- pairwise.fst(Hogan)
PW_Fst_Hogan
```

Note that I've turned this into an Excel heatmap located here: ("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Genepop"). Differentiation is evident across, but not within, lineages. 

```{r Calculate Fst across loci}
install.packages("pegas")
library(pegas)
Fst(as.loci(Hogan))
Gtest_Hogan <- gstat.randtest(Hogan, nsim=99)
Gtest_Hogan
plot(Gtest_Hogan)
```

##Principal Components Analysis

```{r Calculate the PCA}
#Count instances of missing data. 
sum(is.na(Hogan$tab))

#Replace missing data.
Hogan_imputed <- scaleGen(Hogan, NA.method="mean")
class(Hogan_imputed)
dim(Hogan_imputed)

pca1 <- dudi.pca(Hogan_imputed, cent=FALSE, scale=FALSE, scannf=FALSE, nf=3)
barplot(pca1$eig[1:50], main="PCA Eigenvalues", col=heat.colors(50))
```

Wow, take a look at PC 1. 

```{r Examine the output object}
pca1
```

The output object is a list containing: 1) $eig (eigenvalues), 2) $li (PC's), 3) $c1 (allele loadings).

```{r scatterplot of axes 1 and 2}
s.class(pca1$li, pop(Hogan))
title("PCA\naxes 1-2")
add.scatter.eig(pca1$eig[1:20], 3, 1, 2)
```

PC 1 clearly separates the two lineages. 

```{r scatterplot of axes 1 and 3}
s.class(pca1$li, pop(Hogan), xax=1, yax=3, sub="PCA 1-3", csub=2)
title("PCA\naxes 1-3")
add.scatter.eig(pca1$eig[1:20], nf=3, xax=1, yax=3)
```

Hmm... I wonder what's separating out in the odd lineage? 

```{r PCs 1 and 3}
col <- funky(15)
s.class(pca1$li, pop(Hogan), xax=1, yax=3, col=transp(col, .6), axesell=FALSE, cstar=0, cpoint=3, grid=FALSE)

```

```{Plot of PCs 1 and 2}
colorplot(pca1$li, pca1$li, transp = TRUE, cex=3, xlab="PC 1", ylab="PC 2")
title("PCA of axes 1-2")
```

```{r Plot of PCs 1 and 3}
colorplot(pca1$li[c(1,3)], pca1$li, transp=TRUE, cex=3, xlab="PC 1", ylab="PC 3")
title("PCA of axes 1-3")
```

