---
title: "Figures for Hogan-Stockdale 13-16 MS"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

The purpose of this notebook is to create figures for the manuscript focused on RRS from Hogan and Stockdale for years 2013-2016.

```{r setup, include=FALSE}
library(coin)
library(MASS)
library(scales)
library(tidyverse)
# library(leaflet)
library(lubridate)
library(PBSmapping)
library(grid)
library(ggthemes)
library(mapproj) 
library(gridExtra)
# Load
```

# Figure 1: Map of study area

Try PBSmapping, clunky, but great basemap for AK coastlines. Got the idea from Ben Williams GitHub [link](https://ben-williams.github.io/updated_ggplot_figures.html). Still used the idea from Quantitative Palaeoecology [link](https://quantpalaeo.wordpress.com/2016/06/05/ggplot2-maps-with-inset/)

```{r map of study area, message=FALSE, warning=FALSE}
# Load data
dat <- read_csv("map_data_2.csv")

dat <- dat %>%
  mutate(Legend = case_when(type == "Hogan" ~ "Hogan Bay",
                            type == "Stockdale" ~ "Stockdale Creek",
                            type == "Fitness Stream" ~ "Fitness Streams",
                            type == "Hatchery" & name != "VFDA" ~ "PWSAC",
                            name == "VFDA" ~ "VFDA")) %>% 
  mutate(Legend = factor(x = Legend, levels = c("Hogan Bay", "Stockdale Creek", "Fitness Streams", "VFDA", "PWSAC"))) %>% 
  mutate(abbr = case_when(name == "AFK" ~ "AFK",
                          name == "WHN" ~ "WNH",
                          name == "Cannery Creek" ~ "CCH",
                          name == "VFDA" ~ "SG"))

x_lim <- range(dat$long) + c(-0.25, 0.25)
y_lim <- range(dat$lat) + c(-0.25, 0.25)

dat <- dat %>%
  filter(type != "City")

# Plot map with PBS mapping
data("nepacLLhigh")
ak <- nepacLLhigh %>% 
  dplyr::select(group = PID, POS = POS,long = X, lat = Y)

maptheme <- theme(
  panel.grid = element_blank(),
  panel.border = element_rect(fill = NA, colour = "black"),
  panel.background = element_blank()
)

# PWS map
pws_map <- ggplot() + 
  geom_polygon(data = ak, aes(long, lat, group = group), fill = "grey90", color = "black", lwd = 0.3) +
  geom_point(data = dat, aes(x = long, y = lat, fill = Legend), shape = 21, colour = "black", stroke = 2, size = 5) +
  geom_label(data = filter(dat, type == "Hatchery"), aes(long, lat, label = abbr), hjust = 0, nudge_x = 0.1) +
  geom_segment(aes(x = -148.7, xend = -148.7, y = 59.85, yend = 60.02), lineend = "butt", linejoin = "round", size = 2, arrow = arrow(length = unit(0.2, "inches"))) +
  annotate("text", label = "Prince\nWilliam\nSound", x = -147, y = 60.6, size = 5, colour = "black") +
  theme(panel.background = element_rect(fill = 'white')) +
  scale_y_continuous(breaks = c(60.0, 60.5, 61.0)) +
  xlab(expression(paste(Longitude^o, ~'W'))) +
  ylab(expression(paste(Latitude^o, ~'N'))) +
  coord_map(xlim = x_lim, ylim = y_lim) +
  maptheme

# north2(pws_map, symbol = 12, 0.18, 0.20)  # couldn't get this to work with the inset

# Inset map
ak_map <- ggplot() + 
  geom_polygon(data = ak, aes(long, lat, group = group), fill = "grey90", color = "black", lwd = 0.3) +
  borders("world", xlim = c(-170, -130, ylim = c(20, 50)), lwd = 1) +
  geom_rect(data = data.frame(), aes(xmin = x_lim[1], xmax = x_lim[2], ymin = y_lim[1], ymax = y_lim[2]), colour = "red", fill = NA, lwd = 1.3) +
  annotate("text", label = "Alaska", x = -150.5, y = 63.5, size = 5, colour = "black") +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  coord_map(xlim = c(-170, -130), ylim = c(53, 65))

# Create map with inset and North arrow
grid.newpage()
vp_b <- viewport(width = 1, height = 1, x = 0.5, y = 0.5)  # the larger map
vp_a <- viewport(width = 0.23, height = 0.35, x = 0.605, y = 0.24)  # the inset in lower left
print(pws_map, vp = vp_b)
print(ak_map + maptheme, vp = vp_a)
```

This is the code used to actually save the map as a .png file
```{r save_map}
png(filename = "V:/Documents/5_Coastwide/Multispecies/AHRP/Manuscripts/02.Hogan_Stockdale_EA/0002.Figures&Tables/Figures/Figure1_map.png", width = 6.5, height = 6.5, units = "in", res = 300)
grid.newpage()
vp_b <- viewport(width = 1, height = 1, x = 0.5, y = 0.5)  # the larger map
vp_a <- viewport(width = 0.23, height = 0.35, x = 0.63, y = 0.31)  # the inset in lower left
print(pws_map, vp = vp_b)
print(ak_map + maptheme, vp = vp_a)
dev.off()
```

# Figure 2: Sample sizes throughout run
```{r sample sizes throughout run}
paired_13_15_filter <- read_csv("Hogan/paired_13_15_filter.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

paired_14_16_filter <- read.csv("Hogan/paired_14_16_filter.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

stock_paired_13_15_filter <- read_csv("Stockdale/stock_paired_13_15_filter.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

stock_paired_14_16_filter <- read_csv("Stockdale/stock_paired_14_16_filter.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

#Stacked
Hogan_Odd <- paired_13_15_filter %>% 
  ggplot(aes(x= Julian)) +
  geom_bar(aes(fill=origin)) +
  theme_bw(base_size = 14) +
  theme(legend.position = c(0.85, 0.75),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.key.size = unit(1, "line"),
        panel.grid = element_blank(),
        axis.title.x = element_blank()) +
  scale_y_continuous(limits = c(0, 500), breaks = seq(0, 500, by = 100)) +
  scale_x_continuous(limits = c(220, 270), breaks = seq(220, 270, by = 10), labels = format(x = (as.Date("2012-12-31") + seq(220, 270, by = 10)), "%b %d")) +
  facet_grid(`Sample Year` ~ .) +
  labs(title = "Hogan", y = "Number of Samples", fill = "Origin") +
  ylab(expression(atop("Odd", paste("Number of Samples"))))

Hogan_Even <- paired_14_16_filter %>% 
  ggplot(aes(x= Julian)) +
  geom_bar(aes(fill=origin)) +
  theme_bw(base_size = 14) +
  theme(legend.position = 'none',
        panel.grid = element_blank()) +
  scale_y_continuous(limits = c(0, 500), breaks = seq(0, 500, by = 100)) +
  scale_x_continuous(limits = c(220, 270), breaks = seq(220, 270, by = 10), labels = format(x = (as.Date("2012-12-31") + seq(220, 270, by = 10)), "%b %d")) +
  facet_grid(Sample.Year ~ .) +
  labs(y = "Number of Samples", x = "Sample Date") +
  ylab(expression(atop("Even", paste("Number of Samples"))))

Stock_Odd <- stock_paired_13_15_filter %>% 
  ggplot(aes(x=Julian)) +
  geom_bar(aes(fill=origin)) +
  theme_bw(base_size = 14) +
  scale_y_continuous(limits = c(0, 500), breaks = seq(0, 500, by = 100)) +
  scale_x_continuous(limits = c(220, 270), breaks = seq(220, 270, by = 10), labels = format(x = (as.Date("2012-12-31") + seq(220, 270, by = 10)), "%b %d")) +
  theme(legend.position = 'none',
        panel.grid = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank()) +
  facet_grid(`Sample Year` ~ .) +
  labs(title = "Stockdale")

Stock_Even <- stock_paired_14_16_filter %>% 
  ggplot(aes(x=Julian)) +
  geom_bar(aes(fill=origin)) +
  theme_bw(base_size = 14) +
  scale_y_continuous(limits = c(0, 500), breaks = seq(0, 500, by = 100)) +
  scale_x_continuous(limits = c(220, 270), breaks = seq(220, 270, by = 10), labels = format(x = (as.Date("2012-12-31") + seq(220, 270, by = 10)), "%b %d")) +
  theme(legend.position = 'none',
        panel.grid = element_blank(),
        axis.title.y = element_blank()) +
  facet_grid(`Sample Year` ~ .) +
  labs(x = "Sample Date")

Figure_2 <- 
  grid.arrange(Hogan_Odd, Stock_Odd, Hogan_Even, Stock_Even, nrow = 2, widths = c(1.10, 1))  # added width argument to get plot area equal


ggsave("Figure_2.pdf", plot = Figure_2, path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Manuscripts/02.Hogan_Stockdale_EA/0002.Figures&Tables/Figures", height = 6, width = 12)
```

# Figure 3: Reproductive success

```{r Plot RS}
paired_13_15_filter_parents <- read_csv("Hogan/paired_13_15_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(SEX = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

paired_14_16_filter_parents <- read_csv("Hogan/paired_14_16_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(SEX = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

stock_paired_13_15_filter_parents <- read_csv("Stockdale/stock_paired_13_15_filter_parents.csv")%>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(SEX = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

stock_paired_14_16_filter_parents <- read_csv("Stockdale/stock_paired_14_16_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(SEX = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

Hogan_Odd_RS <- paired_13_15_filter_parents %>% 
  count(SEX, origin, n) %>% 
  group_by(SEX, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  theme_bw(base_size = 14) +
  theme(legend.position = c(0.85, 0.75),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.key.size = unit(1, "line"),
        panel.grid = element_blank()) +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1, by = 0.1)) +
  scale_x_continuous(limits = c(-0.5,20), breaks = seq(0,20, by = 2)) +
  facet_grid(~ SEX, labeller = labeller(F = "Female", M = "Male")) +
  labs(title="Hogan",
       fill = "Parent origin") +
  ylab(expression(atop("Odd", paste("Proportion of Parents")))) +
  xlab(" ")

Hogan_Even_RS <- paired_14_16_filter_parents %>% 
  count(SEX, origin, n) %>% 
  group_by(SEX, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  theme_bw(base_size = 14) +
  theme(legend.position = 'none',
        panel.grid = element_blank()) +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1, by = 0.1)) +
  scale_x_continuous(limits = c(-0.5,20), breaks = seq(0,20, by = 2)) +
  facet_grid(~ SEX) +
  labs(fill = "Parent origin") +
  xlab("Number of Offspring (RS)")+
  ylab(expression(atop("Even", paste("Proportion of Parents")))) 

Stockdale_Odd_RS <- stock_paired_13_15_filter_parents %>% 
  count(SEX, origin, n) %>% 
  group_by(SEX, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  theme_bw(base_size = 14) +
  theme(legend.position = 'none',
        panel.grid = element_blank()) +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1, by = 0.1)) +
  scale_x_continuous(limits = c(-0.5,20), breaks = seq(0,20, by = 2)) +
  facet_grid(~ SEX, labeller = labeller(F = "Female", M = "Male")) +
  labs(title="Stockdale",
       fill = "Parent origin") +
  ylab(" ") +
  xlab (" ")

Stockdale_Even_RS <- stock_paired_14_16_filter_parents %>% 
  count(SEX, origin, n) %>% 
  group_by(SEX, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  theme_bw(base_size = 14) +
  theme(legend.position = 'none',
        panel.grid = element_blank()) +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1, by = 0.1)) +
  scale_x_continuous(limits = c(-0.5,20), breaks = seq(0,20, by = 2)) +
  facet_grid(~ SEX, labeller = labeller(F = "Female", M = "Male")) +
  labs(fill = "Parent origin") +
  xlab("Number of Offspring (RS)") +
  ylab(" ")

Figure_3 <- 
  grid.arrange(Hogan_Odd_RS, Stockdale_Odd_RS, Hogan_Even_RS, Stockdale_Even_RS, nrow = 2, widths = c(1.05, 1))


ggsave("Figure_3.pdf", plot = Figure_3, path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Manuscripts/02.Hogan_Stockdale_EA/0002.Figures&Tables/Figures", height = 6, width = 12)

```

# Figure 4: RS by date
```{r Plot RS by Date}
Hogan_Even_Date <- paired_14_16_filter_parents %>% 
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(height = 0, alpha = 0.5, size = 3) +
  geom_smooth(method = "loess", se = TRUE, level = 0.95) +
  theme_bw(base_size = 14) +
  theme(legend.position = c(0.85, 0.90),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.key.size = unit(1, "line"),
        panel.grid = element_blank()) +
  scale_y_continuous(limits = c(0,20), breaks = seq(0,20, by = 5)) +
  facet_grid(SEX ~ .)+
  labs(title = "Hogan",  colour = "Parent origin") +
  xlab("Parent Sample Date") +
  ylab("Number of Offspring (RS)")

Stockdale_Even_Date <- stock_paired_14_16_filter_parents %>% 
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(height = 0, alpha = 0.5, size = 3) +
  geom_smooth(method = "loess", se = TRUE, level = 0.95) +
  theme_bw(base_size = 14) +
  theme(legend.position = 'none',
        panel.grid = element_blank()) +
  scale_y_continuous(limits = c(0,20), breaks = seq(0,20, by = 5)) +
  facet_grid(SEX ~ .)+
  labs(title = "Stockdale",  colour = "Parent origin") +
  xlab("Parent Sample Date") +
  ylab(" ")

Figure_4 <- grid.arrange(Hogan_Even_Date, Stockdale_Even_Date, nrow = 1, widths = c(1, 1))

ggsave("Figure_4.pdf", plot = Figure_4, path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Manuscripts/02.Hogan_Stockdale_EA/0002.Figures&Tables/Figures", height = 6, width = 12)

```

# Figure 5: RS by location
``` {r Plot RS by Location}
riverdist_parentage_hogan <-
  read_csv("Hogan/riverdist_parentage_hogan.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>%
  mutate(SEX = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))
riverdist_parentage_stock <-
  read_csv("Stockdale/riverdist_parentage_stock.csv") %>%
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>%
  mutate(SEX = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

Hogan_Location <- riverdist_parentage_hogan %>% 
  ggplot(aes(x = Distance, y = n, color = origin)) +
  geom_vline(aes(xintercept = hitide), color = "black", lwd = 1.5) +
  geom_jitter(height = 0, alpha = 0.5, size = 3) +
  geom_smooth(method = "loess", se = TRUE, level = 0.95) +
  theme_bw(base_size = 14) +
  theme(legend.position = c(0.85, 0.90),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.key.size = unit(1, "line"),
        panel.grid = element_blank()) +
  scale_x_continuous(limits = c(0, 1500), breaks = seq(0, 1500, by = 250)) +
  scale_y_continuous(limits = c(0,20), breaks = seq(0,20, by = 5)) +
  facet_grid(SEX ~ .)+
  labs(title = "Hogan",  colour = "Parent origin") +
  xlab("Distance from Stream Mouth (m)") +
  ylab("Number of Offspring (RS)")


Stockdale_Location <- riverdist_parentage_stock %>% 
  ggplot(aes(x = Distance, y = n, color = origin)) +
  geom_vline(aes(xintercept = tidedist), color = "black", lwd = 1.5) +
  geom_jitter(height = 0, alpha = 0.5, size = 3) +
  geom_smooth(method = "loess", se = TRUE, level = 0.95) +
  theme_bw(base_size = 14) +
  theme(legend.position = "none",
        panel.grid = element_blank()) +
  scale_x_continuous(limits = c(0, 1500), breaks = seq(0, 1500, by = 250)) +
  scale_y_continuous(limits = c(0,20), breaks = seq(0,20, by = 5)) +
  facet_grid(SEX ~ .)+
  labs(title = "Stockdale") +
  xlab("Distance from Stream Mouth (m)") +
  ylab("")


Figure_5 <- grid.arrange(Hogan_Location, Stockdale_Location, nrow = 1, widths = c(1, 1))

ggsave("Figure_5.pdf", plot = Figure_5, path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Manuscripts/02.Hogan_Stockdale_EA/0002.Figures&Tables/Figures", height = 6, width = 12)

```

# Figure 6: RS by length
```{r Plot RS by Length}
Hogan_Even_Length <- paired_14_16_filter_parents %>% 
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(height = 0, alpha = 0.5, size = 3) +
  geom_smooth(method = "loess", se = TRUE, level = 0.95) +
  theme_bw(base_size = 14) +
  theme(legend.position = c(0.85, 0.90),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.key.size = unit(1, "line"),
        panel.grid = element_blank()) +
  scale_x_continuous(limits = c(325, 500), breaks = seq(325, 500, by = 25)) +
  scale_y_continuous(limits = c(0,20), breaks = seq(0,20, by = 5)) +
  facet_grid(SEX ~ .)+
  labs(title = "Hogan",  colour = "Parent origin") +
  xlab("Parent Length (mm)") +
  ylab("Number of Offspring (RS)")

Stockdale_Even_Length <- stock_paired_14_16_filter_parents %>% 
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(height = 0, alpha = 0.5, size = 3) +
  geom_smooth(method = "loess", se = TRUE, level = 0.95) +
  theme_bw(base_size = 14) +
  theme(legend.position = 'none',
        panel.grid = element_blank()) +
  scale_x_continuous(limits = c(325, 500), breaks = seq(325, 500, by = 25)) +
  scale_y_continuous(limits = c(0,20), breaks = seq(0,20, by = 5)) +
  facet_grid(SEX ~ .)+
  labs(title = "Stockdale",  colour = "Parent origin") +
  xlab("Parent Length (mm)") +
  ylab(" ")


Figure_6 <- grid.arrange(Hogan_Even_Length, Stockdale_Even_Length, nrow = 1, widths = c(1, 1))

ggsave("Figure_6.pdf", plot = Figure_6, path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Manuscripts/02.Hogan_Stockdale_EA/0002.Figures&Tables/Figures", height = 6, width = 12)
```

# Figure S1: Heterozygosity distribution
```{r Plot Het}
Hogan_het_tibble <- dget("../Objects/Hogan_13_14_15_16/het_tibble.txt")
Stockdale_het_tibble <- dget("../Objects/Stockdale_13_14_15_16//het_tibble.txt")

# IQR
Hogan_low_box_het <- Hogan_het_tibble %>% 
  dplyr::group_by(lineage) %>% 
  dplyr::summarise(het = quantile(het, probs = 0.25) - 1.5 * diff(quantile(het, probs = c(0.25, 0.75))))

Hogan_high_box_het <- Hogan_het_tibble %>% 
  dplyr::group_by(lineage) %>% 
  dplyr::summarise(het = quantile(het, probs = 0.75) + 1.5 * diff(quantile(het, probs = c(0.25, 0.75))))

Stockdale_low_box_het <- Stockdale_het_tibble %>% 
  dplyr::group_by(lineage) %>% 
  dplyr::summarise(het = quantile(het, probs = 0.25) - 1.5 * diff(quantile(het, probs = c(0.25, 0.75))))

Stockdale_high_box_het <- Stockdale_het_tibble %>% 
  dplyr::group_by(lineage) %>% 
  dplyr::summarise(het = quantile(het, probs = 0.75) + 1.5 * diff(quantile(het, probs = c(0.25, 0.75))))

# 3SD
Hogan_low_sd_het <- Hogan_het_tibble %>% 
  dplyr::group_by(lineage) %>% 
  dplyr::summarise(het = mean(het) - (3 * sd(het)))

Hogan_high_sd_het <- Hogan_het_tibble %>% 
  dplyr::group_by(lineage) %>% 
  dplyr::summarise(het = mean(het) + (3 * sd(het)))

Stockdale_low_sd_het <- Stockdale_het_tibble %>% 
  dplyr::group_by(lineage) %>% 
  dplyr::summarise(het = mean(het) - (3 * sd(het)))

Stockdale_high_sd_het <- Stockdale_het_tibble %>% 
  dplyr::group_by(lineage) %>% 
  dplyr::summarise(het = mean(het) + (3 * sd(het)))



Hogan_odd_het <- Hogan_het_tibble %>% 
  filter(lineage == "odd") %>% 
  ggplot(aes(x = het)) +
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 1/303) +
  geom_vline(aes(xintercept = het, color = "1.5 IQR"), filter(Hogan_low_box_het, lineage == "odd")) +
  geom_vline(aes(xintercept = het, color = "1.5 IQR"), filter(Hogan_high_box_het, lineage == "odd")) +
  geom_vline(aes(xintercept = het, color = "3 SD"), filter(Hogan_low_sd_het, lineage == "odd")) +
  geom_vline(aes(xintercept = het, color = "3 SD"), filter(Hogan_high_sd_het, lineage == "odd")) +
  theme_bw(base_size = 14) +
  theme(legend.position = c(0.85, 0.75),
        legend.title = element_text(size = 10),
        legend.text = element_text(size = 10),
        legend.key.size = unit(1, "line"),
        panel.grid = element_blank()) +
  xlim(0.15, 0.85) +
  ylim(0, 0.05) +
  xlab("") +
  labs(title = "Hogan", y = "Frequency", color = "Cutoff")

Hogan_even_het <- Hogan_het_tibble %>% 
  filter(lineage == "even") %>% 
  ggplot(aes(x = het)) +
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 1/303) +
  geom_vline(aes(xintercept = het, color = "1.5 IQR"), filter(Hogan_low_box_het, lineage == "even")) +
  geom_vline(aes(xintercept = het, color = "1.5 IQR"), filter(Hogan_high_box_het, lineage == "even")) +
  geom_vline(aes(xintercept = het, color = "3 SD"), filter(Hogan_low_sd_het, lineage == "even")) +
  geom_vline(aes(xintercept = het, color = "3 SD"), filter(Hogan_high_sd_het, lineage == "even")) +
  theme_bw(base_size = 14) +
  theme(legend.position = 'none',
        panel.grid = element_blank()) +
  xlim(0.15, 0.85) +
  ylim(0, 0.05) +
  xlab("Individual Heterozygosity") +
  ylab("Frequency")

Stockdale_odd_het <- Stockdale_het_tibble %>% 
  filter(lineage == "odd") %>% 
  ggplot(aes(x = het)) +
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 1/303) +
  geom_vline(aes(xintercept = het, color = "1.5 IQR"), filter(Stockdale_low_box_het, lineage == "odd")) +
  geom_vline(aes(xintercept = het, color = "1.5 IQR"), filter(Stockdale_high_box_het, lineage == "odd")) +
  geom_vline(aes(xintercept = het, color = "3 SD"), filter(Stockdale_low_sd_het, lineage == "odd")) +
  geom_vline(aes(xintercept = het, color = "3 SD"), filter(Stockdale_high_sd_het, lineage == "odd")) +
  theme_bw(base_size = 14) +
  theme(legend.position = 'none',
        panel.grid = element_blank()) +
  xlim(0.15, 0.85) +
  ylim(0, 0.05) +
  xlab("") +
  ylab("") +
  ggtitle("Stockdale")

Stockdale_even_het <- Stockdale_het_tibble %>% 
  filter(lineage == "even") %>% 
  ggplot(aes(x = het)) +
  geom_histogram(aes(y = ..count../sum(..count..)), binwidth = 1/303) +
  geom_vline(aes(xintercept = het, color = "1.5 IQR"), filter(Stockdale_low_box_het, lineage == "even")) +
  geom_vline(aes(xintercept = het, color = "1.5 IQR"), filter(Stockdale_high_box_het, lineage == "even")) +
  geom_vline(aes(xintercept = het, color = "3 SD"), filter(Stockdale_low_sd_het, lineage == "even")) +
  geom_vline(aes(xintercept = het, color = "3 SD"), filter(Stockdale_high_sd_het, lineage == "even")) +
  theme_bw(base_size = 14) +
  theme(legend.position = 'none',
        panel.grid = element_blank()) +
  xlim(0.15, 0.85) +
  ylim(0, 0.05) +
  xlab("Individual Heterozygosity") +
  ylab("")


Figure_S1 <- 
  grid.arrange(Hogan_odd_het, Stockdale_odd_het, Hogan_even_het, Stockdale_even_het, nrow = 2, widths = c(1.10, 1))  # added width argument to get plot area equal


ggsave("Figure_S1.pdf", plot = Figure_S1, path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Manuscripts/02.Hogan_Stockdale_EA/0002.Figures&Tables/Figures", height = 6, width = 12)
```

# Figure S2
``` {r Plot Cross Type}
Hogan_parents_paired_14_16_cross <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Hogan/parents_paired_14_16_cross.csv") %>% 
  mutate(cross = factor(x = cross, levels = c("NN", "NH", "HN", "HH"))) 

Stockdale_parents_paired_14_16_cross <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Stockdale/stock_parents_paired_14_16_cross.csv") %>% 
  mutate(cross = factor(x = cross, levels = c("NN", "NH", "HN", "HH"))) 

Hogan_Cross_Type <- Hogan_parents_paired_14_16_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  count(cross, n) %>% 
  complete(cross, n, fill = list(nn = 0)) %>% 
  group_by(cross) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = cross))+
  geom_col(position = position_dodge2(preserve = "single"), width = 0.5) +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1, by = 0.1)) +
  scale_x_continuous(limits = c(0,5), breaks = seq(0, 5, by = 1)) +
  labs(title = "Hogan",
       fill = "Cross Type") +
  xlab("Number of Offspring (RS)")+
  ylab("Proportion of Families of Each Cross Type") +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(legend.position = c(0.7, 0.75)) 

Stockdale_Cross_Type <- Stockdale_parents_paired_14_16_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  count(cross, n) %>% 
  complete(cross, n, fill = list(nn = 0)) %>% 
  group_by(cross) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = cross))+
  geom_col(position = position_dodge2(preserve = "single"), width = 0.5) +
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1, by = 0.1)) +
  scale_x_continuous(limits = c(0,5), breaks = seq(0, 5, by = 1)) +
  labs(title = "Stockdale",
       fill = "Cross Type") +
  xlab("Number of Offspring (RS)")+
  ylab(" ") +
  theme_bw() +
  theme(text = element_text(size = 20)) +
  theme(legend.position = 'none') 

Figure_S2 <- grid.arrange(Hogan_Cross_Type, Stockdale_Cross_Type, nrow = 1)

ggsave("Figure_S2.pdf", plot = Figure_S2, path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Manuscripts/02.Hogan_Stockdale_EA/0002.Figures&Tables/Figures", height = 6, width = 12)

```
