---
title: "Comparison of Posterior Probabilities for the 3 Even-Year Franz Runs"
output: html_notebook
---

Parameters:

Run 1: FRANz.exe --Nmmax 4500 --Nfmax 4500 --femrepro 1:2 --malerepro 1:2 --typingerror 0.005 --updatefreqs --poutformat 2 --fullsibtest "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\hogan_postQA_2014_2016_HOGAN.dat"
Run 2: FRANz.exe --Nmmax 3300 --Nfmax 3300 --femrepro 1:2 --malerepro 1:2 --typingerror 0.005 --updatefreqs --poutformat 2 --fullsibtest "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\hogan_postQA_2014_2016_HOGAN.dat"
Run 3: FRANz.exe --Nmmax 4500 --Nfmax 4500 --femrepro 1:2 --malerepro 1:2 --typingerror 0.01 --updatefreqs --poutformat 2 --fullsibtest "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\hogan_postQA_2014_2016_HOGAN.dat"

```{r Setup , message=FALSE, warning=FALSE}
library(tidyverse)
library(scales)
parentage_14_16_r1 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Franz/hogan_postQA_2014_2016_HOGAN_Run 1/parentage.csv")
parentage_14_16_r2 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Franz/hogan_postQA_2014_2016_HOGAN_Run2/parentage.csv")
parentage_14_16_r3 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Franz/hogan_postQA_2014_2016_HOGAN_Run3/parentage.csv")
```

```{r Pair data}
paired_14_16 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Franz/hogan_postQA_OceanAK_paired_2014_2016_HOGAN.csv")
```

Plot a histogram of parentage posterior probabilities to show robustness of assignments
```{r FRANz_posterior_plot}
parentage_14_16_r1  %>% 
  filter(!is.na(`Parent 1`)) %>% 
  ggplot(aes(x = Posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  labs(title = "Run 1")

parentage_14_16_r2  %>% 
  filter(!is.na(`Parent 1`)) %>% 
  ggplot(aes(x = Posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  labs(title = "Run 2")

parentage_14_16_r3  %>% 
  filter(!is.na(`Parent 1`)) %>% 
  ggplot(aes(x = Posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  labs(title = "Run 3")
```

Filter Data
```{r filter}
# Filter for non-genotype columns
paired_14_16_filter <- paired_14_16 %>% 
  dplyr::select(franz_id, SILLY, `Fish ID`, `DNA Tray Code`, `DNA Tray Well Code`, `Sample Year`, `Sample Date`, SEX, `Length Mm`, `Otolith Mark Present`, `Otolith Mark ID`)

# Filter for posterior > 0.9 + make tidy relative to assigment (1 row per parentage assignment)
parentage_14_16_filter_r1 <- parentage_14_16_r1 %>% 
  filter(Posterior > 0.9) %>% 
  dplyr::select(Offspring, `Parent 1`, `Parent 2`) %>% 
  gather(Parent, Parent_ID, - Offspring) %>% 
  filter(!is.na(Parent_ID))

parentage_14_16_filter_r2 <- parentage_14_16_r2 %>% 
  filter(Posterior > 0.9) %>% 
  dplyr::select(Offspring, `Parent 1`, `Parent 2`) %>% 
  gather(Parent, Parent_ID, - Offspring) %>% 
  filter(!is.na(Parent_ID))

parentage_14_16_filter_r3 <- parentage_14_16_r3 %>% 
  filter(Posterior > 0.9) %>% 
  dplyr::select(Offspring, `Parent 1`, `Parent 2`) %>% 
  gather(Parent, Parent_ID, - Offspring) %>% 
  filter(!is.na(Parent_ID))
```


```{r Join parentage and paired objects Even}
# Creating a single, "tidy" object where each row is a parent-offspring relationship
# Offspring with 2 parents have 2 rows
# .x data is offspring, .y data is parents
(parents_paired_14_16_r1 <- parentage_14_16_filter_r1 %>% 
   left_join(paired_14_16_filter, by=c("Offspring"="franz_id")) %>% 
   left_join(paired_14_16_filter, by=c("Parent_ID"="franz_id")))

(parents_paired_14_16_r2 <- parentage_14_16_filter_r2 %>% 
   left_join(paired_14_16_filter, by=c("Offspring"="franz_id")) %>% 
   left_join(paired_14_16_filter, by=c("Parent_ID"="franz_id")))

(parents_paired_14_16_r3 <- parentage_14_16_filter_r3 %>% 
   left_join(paired_14_16_filter, by=c("Offspring"="franz_id")) %>% 
   left_join(paired_14_16_filter, by=c("Parent_ID"="franz_id")))

parents_paired_14_16_r1 %>% 
  count(Parent)

parents_paired_14_16_r2 %>% 
  count(Parent)

parents_paired_14_16_r3 %>% 
  count(Parent)

```

```{r count number of distinct parents for each run}
n_distinct(parents_paired_14_16_r1$Parent_ID)
n_distinct(parents_paired_14_16_r2$Parent_ID)
n_distinct(parents_paired_14_16_r3$Parent_ID)
```

```{r summarize numbers of hatchery and natural origin fish assigned as parents}
parents_paired_14_16_r1 %>% 
  summarise(count_h = n_distinct(Parent_ID[`Otolith Mark Present.y`=="YES"]), count_n = n_distinct(Parent_ID[`Otolith Mark Present.y`=="NO"]))

parents_paired_14_16_r2 %>% 
  summarise(count_h = n_distinct(Parent_ID[`Otolith Mark Present.y`=="YES"]), count_n = n_distinct(Parent_ID[`Otolith Mark Present.y`=="NO"]))

parents_paired_14_16_r3 %>% 
  summarise(count_h = n_distinct(Parent_ID[`Otolith Mark Present.y`=="YES"]), count_n = n_distinct(Parent_ID[`Otolith Mark Present.y`=="NO"]))
```

```{r find out where hatchery origin parents originate from}
parents_paired_14_16_r1 %>% 
  count(`Otolith Mark Present.y`, `Otolith Mark ID.y`) %>% 
  spread(`Otolith Mark Present.y`, n)

parents_paired_14_16_r2 %>% 
  count(`Otolith Mark Present.y`, `Otolith Mark ID.y`) %>% 
  spread(`Otolith Mark Present.y`, n)

parents_paired_14_16_r3 %>% 
  count(`Otolith Mark Present.y`, `Otolith Mark ID.y`) %>% 
  spread(`Otolith Mark Present.y`, n)
```

# Summary
For runs 2 and 3, 458 individuals were assigned to parents, whereas in run 1, 456 offspring were assigned. There were 187 unique parents in runs 2 and 3 as opposed to 186 unique parents in run 1. The additional parent assigned in runs 1 and 2 was of hatchery origin (CCH). 

