---
title: "Hogan_popgen"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(DT)
library(abind)
bbind <- function(...) { abind(..., along = 3) }

source("~/../R/Functions.GCL.R")

load_objects(path = "../Objects/")
load_sillys(path = "../Genotypes/OceanAK_Origin_PostQA/")
```

# Introduction

This is an exploratory analysis of the population structure of the NPRB Hogan Bay 2013-2016 genotype data, post-QA. Here we will:

  * Read in all post-QA project genotypes
  * Create a *genepop* file to read back in with *adegenet*
  * Explore genetic structure
    + Among brood lines (even/odd)
    + Among years
    + Among hatchery/wild (note different hatcheries have different brood sources) 
    + Between sexes
  * Export genotypes as a *Genepop* file to feed in to *genepop2franz.GCL.r*
  
# Create genepop

Create a *genepop* using `gcl2Genepop.GCL`, individuals are named by *SillySource*.
```{r genepop}
gcl2Genepop.GCL(sillyvec = c("PHOGAN13n", "PHOGAN13h", "PHOGAN14n", "PHOGAN14h", "PHOGAN15", "PHOGAN16"), 
                loci = loci, 
                path = "../Genepop/hogan_postQA.gen", 
                VialNums = TRUE, 
                usat = FALSE)

gcl2Genepop.GCL(sillyvec = c("PHOGAN13n", "PHOGAN13h", "PHOGAN15"), 
                loci = loci, 
                path = "../Genepop/hogan_odd_postQA.gen", 
                VialNums = TRUE, 
                usat = FALSE)

gcl2Genepop.GCL(sillyvec = c("PHOGAN14n", "PHOGAN14h", "PHOGAN16"), 
                loci = loci, 
                path = "../Genepop/hogan_even_postQA.gen", 
                VialNums = TRUE, 
                usat = FALSE)
```

Note that I created separate genepop files for each lineage in addition to one for the complete dataset so I could run separate PCAs by lineage.

```{r combine attributes tables}
#When we pooled PHOGAN 13 and 14, the FK_FISH_ID column was converted from numeric to character. However, it remained numeric for PHOGAN 15 and 16. Its structure needs to be uniform across tables to allow for combining. 

PHOGAN15.gcl$attributes <- PHOGAN15.gcl$attributes %>% 
  mutate(FK_FISH_ID = as.character(FK_FISH_ID))
PHOGAN16.gcl$attributes <- PHOGAN16.gcl$attributes %>% 
  mutate(FK_FISH_ID = as.character(FK_FISH_ID))
HOGAN_attributes_tibble <- dplyr::bind_rows(PHOGAN13n.gcl$attributes, PHOGAN13h.gcl$attributes, PHOGAN14n.gcl$attributes, PHOGAN14h.gcl$attributes, PHOGAN15.gcl$attributes, PHOGAN16.gcl$attributes) %>% 
  mutate(Lineage = case_when(SILLY_CODE %in% c("PHOGAN13", "PHOGAN15") ~ "ODD", 
                             SILLY_CODE %in% c("PHOGAN14", "PHOGAN16") ~ "EVEN"))
write_csv(HOGAN_attributes_tibble, "../Attributes table/HOGAN_attributes_tibble.csv")

HOGAN_attributes_odd_tibble <- dplyr::bind_rows(PHOGAN13n.gcl$attributes, PHOGAN13h.gcl$attributes, PHOGAN15.gcl$attributes)

HOGAN_attributes_even_tibble <- dplyr::bind_rows(PHOGAN14n.gcl$attributes, PHOGAN14h.gcl$attributes, PHOGAN16.gcl$attributes)

```


# Analysis in *adegenet*

```{r adegenet}
install.packages("adegenet")
library("adegenet")

Hogan <- read.genepop(file="v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Genepop/hogan_postQA.gen")
Hogan 

Hoganpop <- genind2genpop(Hogan)

Hoganpop@tab[1:6, 1:7] #Shows population names.  

Hogan_odd <- read.genepop(file="v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Genepop/hogan_odd_postQA.gen")

Hogan_even <- read.genepop(file="v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Genepop/hogan_even_postQA.gen")

```

##Inbreeding Estimates

Here we will:
1. Create separate genind files for each year and origin.
2. Estimate and plot inbreeding (F).
3. Calculate and plot mean F across individuals.
 


```{r estimate inbreeding, include=FALSE}
#From adgenet manual - see p 106

#isolate populations
PHOGAN13n_genind <- seppop(Hogan)$PHOGAN13_813
PHOGAN13h_genind <- seppop(Hogan)$PHOGAN13_829
PHOGAN14n_genind <- seppop(Hogan)$PHOGAN14_2649
PHOGAN14h_genind <- seppop(Hogan)$PHOGAN14_2647
PHOGAN15_genind <- seppop(Hogan)$PHOGAN15_9409
PHOGAN16_genind <- seppop(Hogan)$PHOGAN16_13054

#Estimate inbreeding and return sample of F values
Fs_13n <- inbreeding(PHOGAN13n_genind, N=30)
Fs_13h <- inbreeding(PHOGAN13h_genind, N=30)
Fs_14n <- inbreeding(PHOGAN14n_genind, N=30)
Fs_14h <- inbreeding(PHOGAN14h_genind, N=30)
Fs_15 <- inbreeding(PHOGAN15_genind, N=30)
Fs_16 <- inbreeding (PHOGAN16_genind, N=30)

Fs_13n_1 <- inbreeding(PHOGAN13n_genind, N=1)
Fs_13h_1 <- inbreeding(PHOGAN13h_genind, N=1)
Fs_14n_1 <- inbreeding(PHOGAN14n_genind, N=1)
Fs_14h_1 <- inbreeding(PHOGAN14h_genind, N=1)
Fs_15_1 <- inbreeding(PHOGAN15_genind, N=1)
Fs_16_1 <- inbreeding(PHOGAN16_genind, N=1)

```

```{r F density plots, include=FALSE}
#plot first 10 results
invisible(sapply(Fs_13n[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of Sampled F Values"))) 
invisible(sapply(Fs_13h[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of sampled F Values")))
invisible(sapply(Fs_14n[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of sampled F Values")))
invisible(sapply(Fs_14h[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of sampled F Values")))
invisible(sapply(Fs_15[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of sampled F Values")))
invisible(sapply(Fs_16[1:10], function(e) plot(density(e), xlab="F", xlim=c(0,1), main="Density of sampled F Values")))

```


```{r compute mean F}
#compute means for all individuals
Fmean_13n <- sapply(Fs_13n, mean) 
Fmean_13h <- sapply(Fs_13h, mean)
Fmean_14n <- sapply(Fs_14n, mean)
Fmean_14h <- sapply(Fs_14h, mean)
Fmean_15 <- sapply(Fs_15, mean)
Fmean_16 <- sapply(Fs_16, mean)

par(mfrow=c(3,3))
hist(Fmean_13n, col="orange", xlab="Mean value of F", main="2013 natural")
hist(Fmean_13h, col="orange", xlab="Mean value of F", main="2013 hatchery")
hist(Fmean_14n, col="Orange", xlab="Mean value of F", main="2014 natural")
hist(Fmean_14h, col="orange", xlab="Mean value of F", main="2014 hatchery")
hist(Fmean_15, col="orange", xlab="Mean value of F", main="2015")
hist(Fmean_16, col="orange", xlab="Mean value of F", main="2016")
```

These histograms show the mean value of F (inbreeding coefficient) across years and origin types. There appear to be potential qualitative differences in F between lineages, but not origins. 

```{r estimate inbreeding and probability density functions, include=FALSE}
#estimate inbreeding and return probability density functions
Fdens_13n <- inbreeding(PHOGAN13n_genind, res.type = "function")
Fdens_13h <- inbreeding(PHOGAN13h_genind, res.type = "function")
Fdens_14n <- inbreeding(PHOGAN14n_genind, res.type = "function")
Fdens_14h <- inbreeding(PHOGAN14h_genind, res.type = "function")
Fdens_15 <- inbreeding(PHOGAN15_genind, res.type = "function")
Fdens_16 <- inbreeding(PHOGAN16_genind, res.type = "function")

#plot the first 10 functions
invisible(sapply(Fdens_13n[1:10], plot, ylab="Density", main="Density of probability of F values")) 
invisible(sapply(Fdens_13h[1:10], plot, ylab="Density", main="Density of probability of F values"))
invisible(sapply(Fdens_14n[1:10], plot, ylab="Density", main="Density of probability of F values"))
invisible(sapply(Fdens_14h[1:10], plot, ylab="Density", main="Density of probability of F values"))
invisible(sapply(Fdens_15[1:10], plot, ylab="Density", main="Density of probability of F values"))
invisible(sapply(Fdens_16[1:10], plot, ylab="Density", main="Density of probability of F values"))
```

```{r estimate inbreeding and maximum likelihood estimates, include=FALSE}
#estimate inbreeding and return maximum likelihood estimates
Fest_13n <- inbreeding(PHOGAN13n_genind, res.type = "estimate")
Fest_13h <- inbreeding(PHOGAN13h_genind, res.type = "estimate")
Fest_14n <- inbreeding(PHOGAN14n_genind, res.type = "estimate")
Fest_14h <- inbreeding(PHOGAN14h_genind, res.type = "estimate")
Fest_15 <- inbreeding(PHOGAN15_genind, res.type = "estimate")
Fest_16 <- inbreeding(PHOGAN16_genind, res.type = "estimate")

mostinbred_13n <- which.max(Fest_13n)
mostinbred_13h <- which.max(Fest_13h)
mostinbred_14n <- which.max(Fest_14n)
mostinbred_14h <- which.max(Fest_14h)
mostinbred_15 <- which.max(Fest_15)
mostinbred_16 <- which.max(Fest_16)

par(mfrow=c(3,3))
plot(Fdens_13n[[mostinbred_13n]], ylab="Density", xlab="F", main = paste("Probability density of F values 2013 natural", names(mostinbred_13n)))
abline(v=Fest_13n[mostinbred_13n], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
plot(Fdens_13h[[mostinbred_13h]], ylab="Density", xlab="F", main = paste("Probability density of F values 2013 hatchery", names(mostinbred_13h)))
abline(v=Fest_13h[mostinbred_13h], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
plot(Fdens_14n[[mostinbred_14n]], ylab="Density", xlab="F", main = paste("Probability density of F values 2014 natural", names(mostinbred_14n)))
abline(v=Fest_14n[mostinbred_14n], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
plot(Fdens_14h[[mostinbred_14h]], ylab="Density", xlab="F", main = paste("Probability density of F values 2014 hatchery", names(mostinbred_14h)))
abline(v=Fest_14h[mostinbred_14h], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
plot(Fdens_15[[mostinbred_15]], ylab="Density", xlab="F", main = paste("Probability density of F values 2015", names(mostinbred_15)))
abline(v=Fest_15[mostinbred_15], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
plot(Fdens_16[[mostinbred_16]], ylab="Density", xlab="F", main = paste("Probability density of F values 2016", names(mostinbred_16)))
abline(v=Fest_16[mostinbred_16], col="red", lty=2)
legend("topright", legend="MLE", col="red", lty=2)
```

```{r compare MLE and means, include=FALSE}
#compare MLE and means
par(mfrow=c(3,3))
plot(Fest_13n, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_13n, pch=2, col="red")
arrows(x0=1:length(Fest_13n), y0=Fest_13n, y1=Fmean_13n, x1=1:length(Fest_13n), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")
plot(Fest_13h, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_13h, pch=2, col="red")
arrows(x0=1:length(Fest_13h), y0=Fest_13h, y1=Fmean_13h, x1=1:length(Fest_13h), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")
plot(Fest_14n, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_14n, pch=2, col="red")
arrows(x0=1:length(Fest_14n), y0=Fest_14n, y1=Fmean_14n, x1=1:length(Fest_14n), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")
plot(Fest_14h, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_14h, pch=2, col="red")
arrows(x0=1:length(Fest_14h), y0=Fest_14h, y1=Fmean_14h, x1=1:length(Fest_14h), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")
plot(Fest_15, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_15, pch=2, col="red")
arrows(x0=1:length(Fest_15), y0=Fest_15, y1=Fmean_15, x1=1:length(Fest_15), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")
plot(Fest_16, ylim=c(0,0.8), ylab="F", col="blue", main = "Comparison of MLE and average sample estimates of F")
points(Fmean_16, pch=2, col="red")
arrows(x0=1:length(Fest_16), y0=Fest_16, y1=Fmean_16, x1=1:length(Fest_16), length=0.125)
legend("topleft", legend=c("estimate", "sample"), col=c("blue", "red"), pch=c(1,2), title="res.type")

```

##Fst

Here we will:
1. Calculate Fst, Fit, and Fis.
2. Calculate PW Fst across year*origin. 
 

```{r Calculate F statistics}
install.packages("hierfstat")
library(hierfstat)
fstat(Hogan)
```

This is a table of global F statistics (across all individuals in the dataset). 
Fst=pop/total=0.05
Fit=ind/total=0.04
Fis=ind/pop=0

```{r Calculate PW Fst}
PW_Fst_Hogan <- pairwise.fst(Hogan)
PW_Fst_Hogan
```

These are pairwise Fst values among populations and origins. 
Note that I've turned this into an Excel heatmap ("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Genepop"), which is much easier to interpret because it includes population names. Differentiation is evident across, but not within, lineages. 

```{r Calculate Fst across loci, include=FALSE}
install.packages("pegas")
library(pegas)
Fst(as.loci(Hogan))
Gtest_Hogan <- gstat.randtest(Hogan, nsim=99)
Gtest_Hogan
plot(Gtest_Hogan)
```

##Principal Components Analysis
The purpose of the PCAs is to visualize the distribution of variation across the dataset (and subsets of the data) and calculate the extent to which the following variables contribute to that variation: year, lineage, origin, hatchery, and sex. 

```{r Calculate the PCA}
#Count instances of missing data. 
sum(is.na(Hogan$tab))

#Replace missing data.
Hogan_imputed <- scaleGen(Hogan, NA.method="mean")
class(Hogan_imputed)
dim(Hogan_imputed)

pca1 <- dudi.pca(Hogan_imputed, cent=FALSE, scale=FALSE, scannf=FALSE, nf=3)
barplot(pca1$eig[1:50], main="PCA Eigenvalues", col=heat.colors(50))
pca1$eig[1:3]
summary(pca1)
sum_pca1 <- sum(pca1$eig)
sum_pca1
```

These are the eigenvalues for the PCA that includes the entire dataset. PC 1 explains the vast majority of the variation. 

```{r Examine the output object}
pca1
```

The output object is a list containing: 1) $eig (eigenvalues), 2) $li (PC's), 3) $c1 (allele loadings).


```{r scatterplot of axes 1 and 2}
s.class(pca1$li, pop(Hogan))
title("PCA\naxes 1-2")
add.scatter.eig(pca1$eig[1:20], 3, 1, 2)
```

Axis 1 clearly separates the two lineages. Odd is on the left and even is on the right. 

```{r scatterplot of axes 1 and 3, include=FALSE}
s.class(pca1$li, pop(Hogan), xax=1, yax=3, sub="PCA 1-3", csub=2)
title("PCA\naxes 1-3")
add.scatter.eig(pca1$eig[1:20], nf=3, xax=1, yax=3)
```



```{r PCs 1 and 3, include = FALSE}
col <- funky(15)
s.class(pca1$li, pop(Hogan), xax=1, yax=3, col=transp(col, .6), axesell=FALSE, cstar=0, cpoint=3, grid=FALSE)

```

```{Plot of PCs 1 and 2, include=FALSE}
colorplot(pca1$li, pca1$li, transp = TRUE, cex=3, xlab="PC 1", ylab="PC 2")
title("PCA of axes 1-2")
```

```{r Plot of PCs 1 and 3, include=FALSE}
colorplot(pca1$li[c(1,3)], pca1$li, transp=TRUE, cex=3, xlab="PC 1", ylab="PC 3")
title("PCA of axes 1-3")
```

```{r combine attributes table with PC scores, include=FALSE}
Scores_Attributes <- pca1$li %>% 
  as_tibble(rownames="silly_source") %>% 
  left_join(HOGAN_attributes_tibble, by=c("silly_source"= "SillySource"))
```

```{r Exploratory Plots for Year and Origin}
par(mfrow=c(1,3))
ggplot(Scores_Attributes, aes(x=Axis1, y=Axis2, shape=`Otolith Mark Present`, color=SILLY_CODE)) + 
  geom_point(alpha=0.2) +
  labs(title="PC Scores By Year and Origin", x="Axis 1 (~5.6%)", y="Axis 2 (~0.6%)")
ggplot(Scores_Attributes, aes(x=Axis1, y=Axis3, shape=`Otolith Mark Present`, color=SILLY_CODE)) + 
  geom_point(alpha=0.2) +
  labs(title="PC Scores by Year and Origin", x="Axis 1 (~5.6%)", y="Axis 3 (0.5%)")
ggplot(Scores_Attributes, aes(x=Axis2, y=Axis3, shape=`Otolith Mark Present`, color=SILLY_CODE)) + 
  geom_point(alpha=0.2) +
  labs(title="PC Scores by Year and Origin", x= "Axis 2 (~0.6%)", y="Axis 3 (0.5%)")
```

These plots visualize variation in the full dataset by year and origin. Axis 1 clearly separates lineages. We do see some separation within lineages along axis 2, but it does not appear to be due to origin or year. However, note that some natural-origin fish from years 2015 and 2016 separate out. Are these natural strays?

Axis labels provide the percent variation explained. Note that the axes do not explain much of the overall variation. This supports that there's little overall population structure. 

```{r Plots of PCs Showing Origin and Lineage}

ggplot(Scores_Attributes, aes(x=Axis1, y=Axis2, shape=`Otolith Mark Present`, color=Lineage)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Lineage", x="Axis 1 (~5.6%)", y="Axis 2 (~0.6%)")
ggplot(Scores_Attributes, aes(x=Axis1, y=Axis3, shape=`Otolith Mark Present`, color=Lineage)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Lineage", x="Axis 1 (~5.6%)", y="Axis 3 (0.5%)")
ggplot(Scores_Attributes, aes(x=Axis2, y=Axis3, shape=`Otolith Mark Present`, color=Lineage)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Lineage",  x= "Axis 2 (~0.6%)", y="Axis 3 (0.5%)")



```

These are PC plots of the full dataset with individuals categorized by origin and lineage. Axis 1 separates the lineages. Axis 2 shows separation within lineages. Axis 3 shows separation of some natural origin fish from the odd lineage (natural strays?). The plot of Axis2 versus Axis3 shows evidence of multimodality within lineages, suggesting that the data are not normally distributed. We had considered plotting mean scores by lineage, year, origin, and sex, but given this non-normality, the approach seemed inappropriate and potentially misleading. 

Update: After speaking with Chris and Wei, we determined that these clusters are likely artefacts and represent genotype frequencies. This is a limitation of using categorical data as input in a PCA. This is corroborated by (https://privefl.github.io/bigsnpr/articles/how-to-PCA.html) - Scroll down to the explanation of plots of axes 3 and 4. 

Let's examine our loadings.

```{r Plot distribution of loadings }
plot(pca1$c1)
max_loading <- max(pca1$c1)
max_loading
min_loading <- min(pca1$c1)
min_loading
range_loading <- range(pca1$c1)
range_loading
```

Similar to the example cited above, it appears that the banding we observed in axes 2 and 3 may be representing variation at just a couple of loci. 

```{r Plots of PCs Showing Origin and Sex}
ggplot(Scores_Attributes, aes(x=Axis1, y=Axis2, shape=`Otolith Mark Present`, color=Sex)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Sex", x="Axis 1 (~5.6%)", y="Axis 2 (~0.6%)")
ggplot(Scores_Attributes, aes(x=Axis1, y=Axis3, shape=`Otolith Mark Present`, color=Sex)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Sex", x="Axis 1 (~5.6%)", y="Axis 3 (0.5%)")
ggplot(Scores_Attributes, aes(x=Axis2, y=Axis3, shape=`Otolith Mark Present`, color=Sex)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Sex", x= "Axis 2 (~0.6%)", y="Axis 3 (0.5%)")

```

These are plots of the full dataset with individuals categorized by origin and sex. Overall, hatchery and natural origin fish and males and females appear admixed. The clustering we see along axis two does not seem to be due to sex or origin differences. The individuals that separate out along axis 3 appear to be a combination of males and females, but qualitatively may be male-biased. 

These plots are consistently showing that axis 1 separates the two lineages, which is what we expect. However, we do see some apparent clustering along axes 2 and 3. We used anovas to determine which variables are driving the variation in these scores. 
 
```{r anova on PC 2 scores}
aov_PC2 <- aov(Scores_Attributes$Axis2 ~ Scores_Attributes$SILLY_CODE*Scores_Attributes$Sex*Scores_Attributes$`Otolith Mark Present`)
aov_PC2
summary(aov_PC2)

```

This is an anova for Axis 2 scores. There is a significant interaction between sex and origin (p=0.003). 

```{r anova on PC 3 scores}
aov_PC3 <- aov(Scores_Attributes$Axis3 ~ Scores_Attributes$SILLY_CODE*Scores_Attributes$Sex*Scores_Attributes$`Otolith Mark Present`)
aov_PC3
summary(aov_PC3)
```

This is an anova of Axis 3 scores. SILLY_CODE (year and origin) has a significant effect (p<0.001), as does sex (p=0.03). 

Let's examine the distributions of Axis 2 and 3 scores by sex to see if there truly are sex effects. 

```{r Violin plots of scores by sex}
#par(mfrow=c(1,2))
#boxplot(Scores_Attributes$Axis2~Scores_Attributes$Sex, ylab="PC 2 Scores")
#boxplot(Scores_Attributes$Axis3~Scores_Attributes$Sex, ylab="PC 3 Scores")

ggplot(Scores_Attributes, aes(x=Sex, y=Axis2)) +
  geom_violin()
ggplot(Scores_Attributes, aes(x=Sex, y=Axis3)) +
  geom_violin()
```

The distributions of Axis 2 and 3 scores between sexes do not appear to differ. The distributions seen here along Axis 2 are reflective of the artefactual clustering that we saw in the PC plots. 


```{r Biplots of F and PC 1 Scores}
Fs_all <- c(Fs_13n_1, Fs_13h_1, Fs_14n_1, Fs_14h_1, Fs_15_1, Fs_16_1)
head(Fs_all) #this is a list, need to convert to DF
Fs_all_df <- data.frame(matrix(unlist(Fs_all), nrow=7367, byrow = T))
head(Fs_all_df)

Fs_odd <- c(Fs_13n_1, Fs_13h_1, Fs_15)

Fs_even <- c(Fs_14n_1, Fs_14h_1, Fs_16)

par(mfrow=c(1,3))
plot(Fs_all, Scores_Attributes$Axis1)
plot(Fs_all, Scores_Attributes$Axis2)
plot(Fs_all, Scores_Attributes$Axis3)

plot(Fs_odd, Scores_Attributes_odd$Axis1)
plot(Fs_even, Scores_Attributes_even$Axis1)
head(Fs_odd)
Scores_all <- pca1$li

Fs_Scores <- left_join(Fs_all, Scores_all)
```

These are plots of F versus Scores for Axes 1-3. There are no apparent patterns between the two.  

Now that we've examined the full dataset, let's focus on only the hatchery origin fish to see if there is differentiation among the hatcheries. 

```{r Plots of PCs by Hatchery}
par(mfrow=c(3,1))

Scores_Attributes %>% 
  filter(`Otolith Mark Present`=="YES") %>% 
  ggplot(aes(x=Axis1, y=Axis2, color=`Otolith Mark ID`)) +
  geom_point(alpha=0.2) +
  facet_grid(rows = vars(Lineage)) +
  labs(title="PC Scores by Hatchery", x="Axis 1 (~5.6%)", y="Axis 2 (~0.6%)")

Scores_Attributes %>% 
  filter(`Otolith Mark Present`=="YES") %>% 
  ggplot(aes(x=Axis2, y=Axis3, color=`Otolith Mark ID`)) +
  geom_point(alpha=0.2) +
  facet_grid(rows = vars(Lineage)) +
  labs(title="PC Scores by Hatchery", x="Axis 2 (~0.6%)", y="Axis 3 (0.5%)")
  
Scores_Attributes %>% 
  filter(`Otolith Mark Present`=="YES") %>% 
  ggplot(aes(x=Axis1, y=Axis3, color=`Otolith Mark ID`)) +
  geom_point(alpha=0.2) +
  facet_grid(rows = vars(Lineage)) +
  labs(title="PC Scores by Hatchery", x="Axis 1 (~5.6%)", y="Axis 3 (0.5%)")


```

These are plots of only hatchery fish. We do see some clustering along axes 2 and 3, but particular hatcheries do not seem to be grouping together. Overall, hatchery fish appear to be mixed.    

```{r Hatchery anova}
aov_PC1_hatchery <- aov(Scores_Attributes$Axis1 ~ Scores_Attributes$`Otolith Mark ID`)
summary(aov_PC1_hatchery)
aov_PC2_hatchery <- aov(Scores_Attributes$Axis2 ~ Scores_Attributes$`Otolith Mark ID`)
summary(aov_PC2_hatchery)
aov_PC3_hatchery <- aov(Scores_Attributes$Axis3 ~ Scores_Attributes$`Otolith Mark ID`)
summary(aov_PC3_hatchery)
```

There are significant differences in Axis 1 and 3 scores among the different hatcheries. However, this significance is confounded by year. 

```{r Violin plots of PC Scores for fish from each hatchery}
#par(mfrow=c(3,1))

#boxplot(Scores_Attributes$Axis1~Scores_Attributes$`Otolith Mark ID`, ylab="PC 1 Scores", par(cex.axis=0.5))
#boxplot(Scores_Attributes$Axis2~Scores_Attributes$`Otolith Mark ID`, ylab="PC 2 Scores", par(cex.axis=0.5))
#boxplot(Scores_Attributes$Axis3~Scores_Attributes$`Otolith Mark ID`, ylab="PC 3 Scores", par(cex.axis=0.5))

ggplot(Scores_Attributes, aes(x=`Otolith Mark ID`, y=Axis1)) +
  geom_violin() +
  theme(axis.text=element_text(size=6))
ggplot(Scores_Attributes, aes(x=`Otolith Mark ID`, y=Axis2)) +
  geom_violin() +
  theme(axis.text=element_text(size=6))
ggplot(Scores_Attributes, aes(x=`Otolith Mark ID`, y=Axis3)) +
  geom_violin() +
  theme(axis.text=element_text(size=6))
```

We can visualize differences in Axis scores among hatcheries using violin plots. The strongest differentiation is seen in Axis 1, with hatcheries splitting by lineage. Distributions appear more similar in scores from axes 2 and 3. Note that NA denotes natural fish.  

```{r Sample sizes by hatchery}
Scores_Attributes %>% 
  count(SILLY_CODE, `Otolith Mark ID`) %>% 
  spread(`Otolith Mark ID`, n, fill=0)
```

Here are the counts of hatchery fish by year. Note that there is only 1 fish from Solomon Gulch, which explains the missing data in the plots above.

Since the PCA for the full dataset was so strongly driven by differences between lineages, we decided to perform separate PCA's by lineage to better examine differentiation. 

```{r Separate PCA for Odd Lineage}
#Replace missing data.
Hogan_imputed_odd <- scaleGen(Hogan_odd, NA.method="mean")
class(Hogan_imputed_odd)
dim(Hogan_imputed_odd)

pca_odd <- dudi.pca(Hogan_imputed_odd, cent=FALSE, scale=FALSE, scannf=FALSE, nf=3)
barplot(pca_odd$eig[1:50], main="PCA Eigenvalues", col=heat.colors(50))
summary(pca_odd)
```

These are eigenvalues for a PCA that only includes odd-lineage fish. Note that the eigenvalues are much more similar. 

```{r Explore PC Scores by origin and sex}
Scores_Attributes_odd <- pca_odd$li %>% 
  as_tibble(rownames="silly_source") %>% 
  left_join(HOGAN_attributes_odd_tibble, by=c("silly_source"= "SillySource"))

ggplot(Scores_Attributes_odd, aes(x=Axis1, y=Axis2, color=`Otolith Mark Present`, shape=Sex)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Sex", x="Axis 1 (~0.9%)", y="Axis 2 (~0.8%)")

ggplot(Scores_Attributes_odd, aes(x=Axis1, y=Axis3, color=`Otolith Mark Present`, shape=Sex)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Sex", x="Axis 1 (~0.9%)", y="Axis 3 (~0.7%)")

ggplot(Scores_Attributes_odd, aes(x=Axis2, y=Axis3, color=`Otolith Mark Present`, shape=Sex)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Sex", x="Axis 2 (~0.8%)", y="Axis 3 (~0.7%)")
```

For odd-year fish, axis 1 separates a small cluster of natural origin fish, which may be natural strays. Axes 2 and 3 also separate a small cluster of natural origin fish that appear to be a mix of males and females. 

```{r Explore PC Scores by hatchery}
Scores_Attributes_odd %>% 
  filter(`Otolith Mark Present`=="YES") %>% 
  ggplot(aes(x=Axis1, y=Axis2, color=`Otolith Mark ID`)) +
  geom_point(alpha=0.2) +
  labs(title="PC Scores by Hatchery", x="Axis 1 (~0.9%)", y="Axis 2 (~0.8%)")

Scores_Attributes_odd %>% 
  filter(`Otolith Mark Present`=="YES") %>% 
  ggplot(aes(x=Axis1, y=Axis3, color=`Otolith Mark ID`)) +
  geom_point(alpha=0.2) +
  labs(title="PC Scores by Hatchery", x="Axis 1 (~0.9%)", y="Axis 3 (~0.7%)")

Scores_Attributes_odd %>% 
  filter(`Otolith Mark Present`=="YES") %>% 
  ggplot(aes(x=Axis2, y=Axis3, color=`Otolith Mark ID`)) +
  geom_point(alpha=0.2) +
  labs(title="PC Scores by Hatchery", x="Axis 2 (~0.8%)", y="Axis 3 (~0.7%)")
```

Fish from different hatcheries appear intermixed.

```{r Separate PCA for even lineage}
#Replace missing data.
Hogan_imputed_even <- scaleGen(Hogan_even, NA.method="mean")
class(Hogan_imputed_even)
dim(Hogan_imputed_even)

pca_even <- dudi.pca(Hogan_imputed_even, cent=FALSE, scale=FALSE, scannf=FALSE, nf=3)
barplot(pca_even$eig[1:50], main="PCA Eigenvalues", col=heat.colors(50))
summary(pca_even)
```

These are eigenvalues for a PCA including only even-year fish. These values are similar to what we observed for the odd-year lineage. 

```{r Explore PCs by origin and sex}
Scores_Attributes_even <- pca_even$li %>% 
  as_tibble(rownames="silly_source") %>% 
  left_join(HOGAN_attributes_even_tibble, by=c("silly_source"= "SillySource"))

ggplot(Scores_Attributes_even, aes(x=Axis1, y=Axis2, color=`Otolith Mark Present`, shape=Sex)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Sex", x="Axis 1 (~0.7%)", y="Axis 2 (~0.5%)")

ggplot(Scores_Attributes_even, aes(x=Axis1, y=Axis3, color=`Otolith Mark Present`, shape=Sex)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Sex", x="Axis 1 (~0.7%)", y="Axis 3 (~0.5%)")

ggplot(Scores_Attributes_even, aes(x=Axis2, y=Axis3, color=`Otolith Mark Present`, shape=Sex)) + 
  geom_point(alpha=0.2) +
  labs(title = "PC Scores by Origin and Sex", x="Axis 2 (~0.5%)", y="Axis 3 (~0.5%)")
```

Axis 1 separates fish into three clusters. Outliers appear to be mostly natural fish. 

Let's examine loadings again. 

```{r Loadings for even year PCA}
plot(pca_even$c1)

```

It appears that the banding is again driven by a small number of outlier loci. 

```{r PC Scores by hatchery}
Scores_Attributes_even %>% 
  filter(`Otolith Mark Present`=="YES") %>% 
  ggplot(aes(x=Axis1, y=Axis2, color=`Otolith Mark ID`)) +
  geom_point(alpha=0.2) +
  labs(title="PC Scores by Hatchery", x="Axis 1 (~0.7%)", y="Axis 2 (~0.5%)")

Scores_Attributes_even %>% 
  filter(`Otolith Mark Present`=="YES") %>% 
  ggplot(aes(x=Axis1, y=Axis3, color=`Otolith Mark ID`)) +
  geom_point(alpha=0.2) +
  labs(title="PC Scores by Hatchery", x="Axis 1(~0.7%)", y="Axis 3 (~0.5%)")

Scores_Attributes_even %>% 
  filter(`Otolith Mark Present`=="YES") %>% 
  ggplot(aes(x=Axis2, y=Axis3, color=`Otolith Mark ID`)) +
  geom_point(alpha=0.2) +
  labs(title="PC Scores by Hatchery", x="Axis 2 (~0.5%)", y="Axis 3 (~0.5%)")
```

Again, axis 1 seems to separate the fish into clusters, although fish do not cluster by hatchery. 

```{r ANOVA for PC 1 scores}
aov_even <- aov(Scores_Attributes_even$Axis1~Scores_Attributes_even$SILLY_CODE*Scores_Attributes_even$Sex*Scores_Attributes_even$`Otolith Mark Present`)
summary(aov_even)

```

Despite apparent clustering along axis 1, there are no significant differences in Axis 1 scores due to year, sex, origin, or interactions. 


# Main take-aways:
1. The two lineages are strongly genetically differentiated. 
2. Fish from different hatcheries are differentiated in scores from Axes 1 and 3 in the full dataset, due to strong associations between hatchery and year. These differences go away when we examine the lineages separately. 
3. There may be some slight partitioning of variation by sex and origin, but overall fish appear to be admixed. 
4. PCAs are limiting when categorical data are used as input. The clustering that we observe along Axis 2 appears to be due to loadings from a small number of loci, rather than biological effects. 

The remaining mystery seems to be explaining the outgroup in Axes 2 and 3. They appear to be natural-origin strays.  
