---
title: "Stockdale + Gilmour F2 Analysis"
subtitle: "How many grandoffspring produced from cross matings?"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list = ls())

library(kinship2)
library(coin)
library(MASS)
library(scales)
library(tidyverse)
library(leaflet)
library(lubridate)
library(ggExtra)

knitr::opts_chunk$set(out.width = "100%", fig.width = 10)
```

# Stockdale

## 14/16 Cross Data

Read offspring-centric cross type data in.
```{r}
(stock_parents_paired_14_16_cross <- read_csv("Stockdale/stock_parents_paired_14_16_cross.csv"))
```

How many offspring by cross type?
```{r}
stock_parents_paired_14_16_cross %>% 
  count(cross)
```

## 16/18 Parentage Data

Get parent-centric 2016/2018 parentage data
```{r}
(stock_paired_16_18_filter_parents <- read_csv("Stockdale/stock_paired_16_18_filter_parents.csv"))
```

Filter for parents that came from known crosses in 2014
```{r}
stock_paired_16_18_filter_parents %>% 
  right_join(stock_parents_paired_14_16_cross, by = c("franz_id" = "Offspring")) %>% 
  dplyr::select(cross, n) %>% 
  rename(n_off = n) %>% 
  group_by(cross) %>% 
  summarise(n_par = length(cross), n_off = sum(n_off))
```

# Gilmour

## 14/16 Cross Data

Read offspring-centric cross type data in.
```{r}
(gilmour_parents_paired_14_16_cross <- read_csv("Gilmour/gilmour_parents_paired_14_16_cross.csv"))
```

How many offspring by cross type?
```{r}
gilmour_parents_paired_14_16_cross %>% 
  count(cross)
```

How many unique cross matings?
```{r}
gilmour_parents_paired_14_16_cross %>% 
  dplyr::select(`Parent 1`, `Parent 2`, cross) %>% 
  distinct() %>% 
  count(cross)
```


## 16/18 Parentage Data

Get parent-centric 2016/2018 parentage data
```{r}
(gilmour_paired_16_18_filter_parents <- read_csv("Gilmour/gilmour_paired_16_18_filter_parents.csv"))
```

Filter for parents that came from known crosses in 2014
```{r}
gilmour_paired_16_18_filter_parents %>% 
  right_join(gilmour_parents_paired_14_16_cross, by = c("franz_id" = "Offspring")) %>% 
  dplyr::select(cross, n) %>% 
  rename(n_off = n) %>% 
  group_by(cross) %>% 
  summarise(n_par = length(cross), n_off = sum(n_off))
```