---
title: "Publication Maps"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

 

```{r setup, include=FALSE}
library(tidyverse)
library(leaflet)
library(ggmap)
library(maps)
library(mapdata)
library(PBSmapping)
library(grid)
```

# Read in, format, and filter data

```{r read_map_data}
(dat <- read_csv("map_data.csv"))
```

# Make Map

## ggplot + map_data

The inspiration for this approach came from a wordpress blog on Quantitative Paleoecology [link](https://quantpalaeo.wordpress.com/2016/06/05/ggplot2-maps-with-inset/)

Try "world" and "worldHires" maps...but they suck because their coastline is terrible. No Knight Island?!?!?!
```{r map_prep}
x_lim <- range(dat$long) + c(-0.5, 0.5)
y_lim <- range(dat$lat) + c(-0.5, 0.5)

ak_map <- map_data("worldHires", "USA:Alaska")
ak_map <- subset(ak_map, long < 0)  # drop the end of the Aleutian Islands, or use world2Hires 

ggplot() +
  geom_map(data = ak_map, aes(x = long, y = lat, map_id = region), map = ak_map, colour = NA, fill = "grey60")

ggplot() + 
  geom_polygon(data = ak_map, aes(long, lat, group = group), fill = "grey60", color = "black", lwd = 0.5) +
  theme(panel.background = element_rect(fill = 'white')) +
  xlab(expression(paste(Longitude^o, ~'W'))) +
  ylab(expression(paste(Latitude^o, ~'N'))) +
  coord_map(xlim = x_lim, ylim = y_lim)
```

## ggplot + PBSmapping

Try PBSmapping, clunky, but great basemap for AK coastlines. Got the idea from Ben Williams GitHub [link](https://ben-williams.github.io/updated_ggplot_figures.html). Still used the idea from the previous link on Quantitative Palaeoecology

```{r pbs_mapping}
x_lim <- range(dat$long) + c(-0.25, 0.25)
y_lim <- range(dat$lat) + c(-0.25, 0.25)

data("nepacLLhigh")
ak <- nepacLLhigh %>% 
  dplyr::select(group = PID, POS = POS,long = X, lat = Y)

pws_map <- ggplot() + 
  geom_polygon(data = ak, aes(long, lat, group = group), fill = "grey90", color = "black", lwd = 0.3) +
  geom_point(data = dat, aes(x = long, y = lat, colour = type), cex = 5) +
  theme(panel.background = element_rect(fill = 'white')) +
  xlab(expression(paste(Longitude^o, ~'W'))) +
  ylab(expression(paste(Latitude^o, ~'N'))) +
  coord_map(xlim = x_lim, ylim = y_lim)

ak_map <- ggplot() + 
  geom_polygon(data = ak, aes(long, lat, group = group), fill = "grey90", color = "black", lwd = 0.3) +
  geom_rect(data = data.frame(), aes(xmin = x_lim[1], xmax = x_lim[2], ymin = y_lim[1], ymax = y_lim[2]), colour = "red", fill = NA, lwd = 1.3) +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  coord_map(xlim = c(-170, -130), ylim = c(53, 65))

maptheme <- theme(
  panel.grid = element_blank(),
  panel.border = element_rect(fill = NA, colour = "black"),
  panel.background = element_blank()
)

grid.newpage()
vp_b <- viewport(width = 1, height = 1, x = 0.5, y = 0.5)  # the larger map
vp_a <- viewport(width = 0.23, height = 0.35, x = 0.605, y = 0.24)  # the inset in lower left
print(pws_map + maptheme, vp = vp_b)
print(ak_map + maptheme, vp = vp_a)
```

## ggmp

This fails and is super frustraiting. Google changed their API...

Attempted to follow a tutorial on Eric Anderson's reproducible research [link](http://eriqande.github.io/rep-res-web/lectures/making-maps-with-R.html)

```{r ggmap}
pws_box <- make_bbox(lon = dat$long, lat = dat$lat, f = 0.1)

# pws_map <- get_map(location = pws_box, source = "osm")
pws_map <- get_openstreetmap(bbox = pws_box, urlonly = FALSE)
ggmap(pws_map)

get_map(location = pws_box, maptype = "toner-lite", source = "osm")

```

