---
title: "RRS Sensitivity Analysis: How many parents to genotype?"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(coin)
library(MASS)
library(scales)
library(tidyverse)
library(leaflet)
library(lubridate)
library(ggExtra)

knitr::opts_chunk$set(out.width = "100%", fig.width = 10)
```

# Background

For many of the PWS pink salmon fitness streams, we have thousands of tissue samples from each origin and sex. The number of parents that we genotype by origin and sex is the number of draws we get from a given distribution of RS for each origin/sex (i.e. sample size). In some cases, we may have more parent samples for some origin/sex combinations than strictly necessary to estimate RRS. In this notebook, I'm going to go back and subsample fewer parents (*bootstrap*) to see how that would affect our estimates of RRS. The goal of this exercise is to determine if we can scale back some of the genotyping for parent years where we have several thousand samples for some origin/sex combinations. For example, do we really need to genotype 4,383 hatchery-origin parents from Hogan 2015 to get an accurate estimate of RRS?

## Things to keep in mind

Parent sample size is just **one** factor that affects power to detect RRS < 1. Others include:

  * True RRS
  * Sampled proportion of offspring
  * Productivity (true R/S in a given BY)

# Objectives

  1) Estimate RRS with full sample size for Hogan 2015/2017 and Stockdale 2015/2017  
  2) Bootstrap RRS estimates with subsamples of parents, equal sample size by sex and origin
  3) Compare bootstrapped estimates to see what parent sample sizes were necessary to detect different levels of RRS

# Read in parent data

This is parent-centric data, one row = one potential parent (fish sampled in 2015), `n` is the number of returning adult offspring sampled in 2017 that were assigned to that potential parent via parentage analysis in `FRANz`.

```{r}
(
  hogan_paired_15_17_filter_parents <-
    read_csv("Hogan/hogan_paired_15_17_filter_parents.csv") %>%
    filter(SEX != "?") %>%
    mutate(origin = factor(
      x = origin, levels = c("Natural", "Hatchery")
    ))
)

(
  stock_paired_15_17_filter_parents <-
    read_csv("Stockdale/stock_paired_15_17_filter_parents.csv") %>%
    filter(SEX != "?") %>%
    mutate(origin = factor(
      x = origin, levels = c("Natural", "Hatchery")
    ))
)
```

# Sample Sizes

What were the sample sizes of parents by sex and origin? How many draws from the distribution of reproductive success were we getting?

## Hogan 2015

We purposely chose to genotype more potential parents than necessary for Hogan 2015 in order to do this type of sensitivity analysis. Since natural-origin fish represent parents for one generation and offspring from the previous generation, we chose to genotype all available natural-origin fish since we want to maximize the proportion of offspring genotyped for the previous generation. The main question for the sensitivity analysis is how many hatchery-origin parents do we need to genotype to:

  1) determine if RRS < 1, and
  2) get a precise estimate of RRS

```{r}
hogan_paired_15_17_filter_parents %>% 
  count(origin, Sex) %>% 
  spread(origin, n)
```

## Stockdale 2015

```{r}
stock_paired_15_17_filter_parents %>% 
  count(origin, Sex) %>% 
  spread(origin, n)
```

# Calculate RS and RRS with all samples

Average reproductive success (RS) by sex and origin, as well as the relative reproductive success (RRS).

## Hogan 2015

RRS for Hogan 2015 is very low, ~16%. With this large of an effect size, it likely will not take very many parents genotyped to be able to conclude that RRS < 1.

```{r}
hogan_rs_15_17_0 <- hogan_paired_15_17_filter_parents %>%
  group_by(origin, Sex) %>% 
  summarise(RS = mean(n, na.rm = TRUE))

hogan_rs_15_17_0 %>%
  mutate(RS = round(RS, 3)) %>% 
  spread(origin, RS) %>% 
  mutate(RRS = Hatchery / Natural)
```

### Plot distribution of RS

Most potential parents did not have any offspring assigned (>95%). Family sizes were very small.

```{r, fig.width=10}
hogan_paired_15_17_filter_parents %>% 
  filter(SEX != "?") %>% 
  count(Sex, origin, n) %>% 
  group_by(Sex, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  facet_grid(~ Sex) +
  labs(title="Distribution of Family Size",
       fill = "Parent Origin") +
  xlab("Number of Offspring")+
  ylab("Proportion of Parents")  +
  theme_bw() +
  theme(text = element_text(size = 20))
```

## Stockdale 2015

RRS for Stockdale 2015 is higher than Hogan 2015. Given this smaller effect size, it will take more parents genotyped than Hogan 2015 to be able to conclude that RRS < 1.

```{r}
stock_rs_15_17_0 <- stock_paired_15_17_filter_parents %>%
  group_by(origin, Sex) %>% 
  summarise(RS = mean(n, na.rm = TRUE))

stock_rs_15_17_0 %>%
  mutate(RS = round(RS, 3)) %>% 
  spread(origin, RS) %>% 
  mutate(RRS = Hatchery / Natural)
```

### Plot distribution of RS

Most potential parents did not have any offspring assigned (>95%). Family sizes were very small.

```{r fig.width=10}
stock_paired_15_17_filter_parents %>% 
  filter(SEX != "?") %>% 
  count(Sex, origin, n) %>% 
  group_by(Sex, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  facet_grid(~ Sex) +
  labs(title="Distribution of Family Size",
       fill = "Parent Origin") +
  xlab("Number of Offspring")+
  ylab("Proportion of Parents")  +
  theme_bw() +
  theme(text = element_text(size = 20))
```

# Bootstrapped RRS with a reduced sample size of parents

**Note** all of this is done with equal sample size among sexes and origins. The reason for this is that in any subsampling scheme we design, we want to maximize power by gettiNg equal sample sizes among all 4 distributions we are trying to estimate (hatchery females, natural females, hatchery males, and natural males). On the other hand, a random subsample would represent a decrease in field sampling (i.e. what if the crews did not sample as intensively). What we are trying to do here is simulate what would happen if we had explicitly chosen to genotype a subsample of the available field samples. All sampling is **with** replacement.

## Hogan 2015

This simulates what would happen if we **chose** to genotype a subset of samples. **Note** here we are assuming that we would take a random subsample of parents **within** each origin and sex (i.e. equal sample size by sex and origin).

```{r}
hogan_rs_15_17_0_sub <-
  sapply(seq(from = 50, to = 2000, by = 50), function(samp_n) {  # loop over number of subsamples
    sapply(1:1000, function(rep_n) {  # loop over repetitions for a given number of subsamples
      
      # stratified subsample by origin and sex for `samp_n` individuals
      my_subsample <- hogan_paired_15_17_filter_parents %>%
        dplyr::group_by(origin, Sex) %>%  # this gives us a stratified sample, random within origin and Sex
        dplyr::sample_n(samp_n, replace = TRUE) %>%
        dplyr::ungroup()
      
      # p-value for permutation test by sex for RRS < 1 (coin package)
      if (sum(subset(x = my_subsample, Sex == "Female")$n) == 0 |
          sum(subset(x = my_subsample, Sex == "Male")$n) == 0)  # deal with n = 0 for all
      {
        my_p_values <-
          tibble::tibble(Sex = c("Female", "Male"),
                         p_value = c(1, 1))
      } else {
        my_p_values <- my_subsample %>%
          base::split(.$Sex) %>%
          purrr::map(~ coin::oneway_test(
            n ~ origin,
            data = .,
            distribution = approximate(nresample = 10000)
          )) %>%
          purrr::map(~ coin::pvalue(., alternative = "greater")) %>%
          purrr::map(as.numeric) %>%
          tibble::as_tibble() %>%
          tidyr::gather(Sex, p_value)
      }
      
      # calculate RS and RRS, join with p-value
      my_subsample %>%
        dplyr::group_by(origin, Sex) %>%
        dplyr::summarise(RS = mean(n, na.rm = TRUE)) %>%
        tidyr::spread(origin, RS) %>%
        dplyr::group_by(Sex) %>%
        dplyr::summarise(RRS = Hatchery / Natural) %>%
        dplyr::ungroup() %>%
        dplyr::mutate(samp_n = samp_n,
                      rep_n = rep_n) %>%
        dplyr::left_join(my_p_values, by = "Sex")
      
    }, simplify = FALSE) %>%
      dplyr::bind_rows()
  }, simplify = FALSE) %>%
  dplyr::bind_rows()
```

### Power plot

How often did we detect that RRS < 1 via a permutation test (*alpha* = 0.05)? **Note** our best estimate of RRS is ~ 16% for both males and females, thus we should expect that we need a relatively low sample size to detect RRS < 1.

```{r}
hogan_rs_15_17_0_sub %>% 
  group_by(Sex, samp_n) %>% 
  summarise(power = sum(p_value <= 0.05) / 1000) %>% 
  ggplot(aes(x = samp_n, y = power)) +
  geom_line() +
  geom_hline(yintercept = 0.8, colour = "red") +
  facet_grid(. ~ Sex) +
  xlab("Sample Size of Parents for Each Origin") +
  ylab("Power") +
  ggtitle("Power to Detect RRS < 1 with Equal Sample Size by Origin") +
  theme_bw()
```

### Bootstrap plot

Find 95% range of RRS (2.5-97.5% interval).

```{r}
hogan_rs_15_17_0_sub %>% 
  filter(!RRS %in% c("NaN", "Inf")) %>% 
  group_by(Sex, samp_n) %>% 
  summarise(high = quantile(RRS, probs = 0.975),
            low = quantile(RRS, probs = 0.025)) %>% 
  ggplot(aes(x = samp_n, ymin = low, ymax = high)) +
  geom_ribbon(fill = "grey30") +
  geom_hline(yintercept = 1, colour = "red") +
  facet_grid(. ~ Sex) +
  xlab("Sample Size of Parents for Each Origin") +
  ylab("RRS Estimate") +
  ggtitle("Bootstrapped 95% Range of RRS Estimates with Equal Sample Size by Origin") +
  theme_bw()
```

## Stockdale 2015

This simulates what would happen if we **chose** to genotype a subset of samples. **Note** here we are assuming that we would take a random subsample of parents **within** each origin and sex (i.e. equal sample size by sex and origin).

```{r}
stock_rs_15_17_0_sub <-
  sapply(seq(from = 50, to = 1000, by = 50), function(samp_n) {  # loop over number of subsamples
    sapply(1:1000, function(rep_n) {  # loop over repetitions for a given number of subsamples
      
      # stratified subsample by origin and sex for `samp_n` individuals
      my_subsample <- stock_paired_15_17_filter_parents %>%
        dplyr::group_by(origin, Sex) %>%  # this gives us a stratified sample, random within origin and Sex
        dplyr::sample_n(samp_n, replace = TRUE) %>%
        dplyr::ungroup()
      
      # p-value for permutation test by sex for RRS < 1 (coin package)
      if (sum(subset(x = my_subsample, Sex == "Female")$n) == 0 |
          sum(subset(x = my_subsample, Sex == "Male")$n) == 0)  # deal with n = 0 for all
      {
        my_p_values <-
          tibble::tibble(Sex = c("Female", "Male"),
                         p_value = c(1, 1))
      } else {
        my_p_values <- my_subsample %>%
          base::split(.$Sex) %>%
          purrr::map(~ coin::oneway_test(
            n ~ origin,
            data = .,
            distribution = approximate(nresample = 10000)
          )) %>%
          purrr::map(~ coin::pvalue(., alternative = "greater")) %>%
          purrr::map(as.numeric) %>%
          tibble::as_tibble() %>%
          tidyr::gather(Sex, p_value)
      }
      
      # calculate RS and RRS, join with p-value
      my_subsample %>%
        dplyr::group_by(origin, Sex) %>%
        dplyr::summarise(RS = mean(n, na.rm = TRUE)) %>%
        tidyr::spread(origin, RS) %>%
        dplyr::group_by(Sex) %>%
        dplyr::summarise(RRS = Hatchery / Natural) %>%
        dplyr::ungroup() %>%
        dplyr::mutate(samp_n = samp_n,
                      rep_n = rep_n) %>%
        dplyr::left_join(my_p_values, by = "Sex")
      
    }, simplify = FALSE) %>%
      dplyr::bind_rows()
  }, simplify = FALSE) %>%
  dplyr::bind_rows()
```

### Power plot

How often did we detect that RRS < 1 via a permutation test (*alpha* = 0.05)? **Note** our best estimate of RRS is ~ 16% for both males and females, thus we should expect that we need a relatively low sample size to detect RRS < 1.

```{r}
stock_rs_15_17_0_sub %>% 
  group_by(Sex, samp_n) %>% 
  summarise(power = sum(p_value <= 0.05) / 1000) %>% 
  ggplot(aes(x = samp_n, y = power)) +
  geom_line() +
  geom_hline(yintercept = 0.8, colour = "red") +
  facet_grid(. ~ Sex) +
  xlab("Sample Size of Parents for Each Origin") +
  ylab("Power") +
  ggtitle("Power to Detect RRS < 1 with Equal Sample Size by Origin") +
  theme_bw()
```

### Bootstrap plot

Find 95% range of RRS (2.5-97.5% interval). These stop at 1700 fish, because that is the lowest number of parents for any given origin/sex combination (i.e. only XXX natural-origin females in Stockdale 2015)
```{r}
stock_rs_15_17_0_sub %>% 
  filter(!RRS %in% c("NaN", "Inf")) %>% 
  group_by(Sex, samp_n) %>% 
  summarise(high = quantile(RRS, probs = 0.975),
            low = quantile(RRS, probs = 0.025)) %>% 
  ggplot(aes(x = samp_n, ymin = low, ymax = high)) +
  geom_ribbon(fill = "grey30") +
  geom_hline(yintercept = 1, colour = "red") +
  facet_grid(. ~ Sex) +
  xlab("Sample Size of Parents for Each Origin") +
  ylab("RRS Estimate") +
  ggtitle("Bootstrapped 95% Range of RRS Estimates with Equal Sample Size by Origin") +
  theme_bw()
```