---
title: "Stream Surveys"
subtitle: "Investigating Live/Dead Counts"
author: "Kyle Shedd"
date: "15 February 2022"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list = ls())
library(tidyverse)
library(leaflet)
library(lubridate)
library(ggExtra)
library(gganimate)

knitr::opts_chunk$set(out.width = "100%", fig.width = 10)
```

# Background

There are concerns that field crews quit sampling pre-maturely in 2017 and 2019. The contractor reports basically confess to quiting early in 2017 and hint that they might have in 2019. Here I look at the live count data to determine the magnitude of this issue.

# Load Stream Survey Data

All years
```{r}
(stream_surveys_2013_2020 <- readr::read_csv("~/../Desktop/Local_PWS_pinks/StreamSpecimens/stream_surveys_2013-2020.csv"))
```

## Function

```{r}
live_count_by_fitness_stream <-
  function(indata = stream_surveys_2013_2020,
           sample_year = 2013) {
    
    indata_filter_fitness <- indata %>%
      dplyr::mutate(date = lubridate::mdy(SurveyDate)) %>%
      dplyr::mutate(year = lubridate::year(date)) %>%
      dplyr::filter(
        CommonName == "Pink Salmon",!is.na(AliveCount),
        year == sample_year,!is.na(DistanceMeters)
      ) %>%
      dplyr::filter(
        base::grepl(pattern = "Erb", x = StreamName) |
          base::grepl(pattern = "Paddy", x = StreamName) |
          base::grepl(pattern = "Hogan", x = StreamName) |
          base::grepl(pattern = "Gilmour", x = StreamName) |
          base::grepl(pattern = "Stockdale", x = StreamName)
      ) %>%
      dplyr::select(year,
                    date,
                    StreamName,
                    CommonName,
                    AliveCount,
                    DistanceMeters)
    
    date_min_xlim <- indata_filter_fitness %>%
      dplyr::summarise(min(date)) %>%
      dplyr::pull()
    
    date_max_xlim <- indata_filter_fitness %>% 
      dplyr::summarise(max(date)) %>%
      dplyr::pull()
    
    indata_filter_fitness %>%
      ggplot2::ggplot(ggplot2::aes(x = date, y = AliveCount)) +
      ggplot2::geom_hline(yintercept = 0) +
      ggplot2::geom_line(colour = "pink2", lwd = 2) +
      ggplot2::geom_point(data = . %>% group_by(StreamName) %>% filter(date == max(date))) +
      ggplot2::geom_text(
        data = . %>% group_by(StreamName) %>% filter(date == max(date)),
        ggplot2::aes(label = AliveCount),
        hjust = 0,
        nudge_x = 1
      ) +
      ggplot2::labs(x = "Sample Date",
                    y = "Live Count",
                    title = sample_year) +
      ggplot2::scale_x_date(limits = c(date_min_xlim, date_max_xlim + 10)) +
      ggplot2::theme_bw() +
      ggplot2::facet_wrap(ggplot2::vars(StreamName), scales = "free_y")
  }
```

## 2013

Plot live counts by stream
**NOTE** this is naively ignoring the start/end locations of the surveys!!!
```{r}
live_count_by_fitness_stream(sample_year = 2013)
```

## 2014

Plot live counts by stream
**NOTE** this is naively ignoring the start/end locations of the surveys!!!
```{r}
live_count_by_fitness_stream(sample_year = 2014)
```

## 2015

Plot live counts by stream
**NOTE** this is naively ignoring the start/end locations of the surveys!!!
```{r}
live_count_by_fitness_stream(sample_year = 2015)
```

## 2016

Plot live counts by stream
**NOTE** this is naively ignoring the start/end locations of the surveys!!!
```{r}
live_count_by_fitness_stream(sample_year = 2016)
```

## 2017

Plot live counts by stream
**NOTE** this is naively ignoring the start/end locations of the surveys!!!
```{r}
live_count_by_fitness_stream(sample_year = 2017)
```

## 2018

Plot live counts by stream
**NOTE** this is naively ignoring the start/end locations of the surveys!!!
```{r}
live_count_by_fitness_stream(sample_year = 2018)
```

## 2019

Plot live counts by stream
**NOTE** this is naively ignoring the start/end locations of the surveys!!!
```{r}
live_count_by_fitness_stream(sample_year = 2019)
```

## 2020

Plot live counts by stream
**NOTE** this is naively ignoring the start/end locations of the surveys!!!
```{r}
live_count_by_fitness_stream(sample_year = 2020)
```

# Explore Density vs. RS


Read in live/dead count data
```{r}
(survey_data <- readr::read_csv(file = "~/../Desktop/Local_PWS_pinks/StreamSpecimens/stream_surveys_2013-2020.csv"))
```

Reformat
```{r}
(
  survey_data <- survey_data %>%
    dplyr::mutate(date = lubridate::mdy(SurveyDate)) %>%
    dplyr::mutate(year = lubridate::year(date)) %>%
    dplyr::filter(
      CommonName == "Pink Salmon",
      !is.na(AliveCount),
      year == 2014,
      !is.na(DistanceMeters)
    ) %>%
    dplyr::filter(
      base::grepl(pattern = "Erb", x = StreamName) |
        base::grepl(pattern = "Paddy", x = StreamName) |
        base::grepl(pattern = "Hogan", x = StreamName) |
        base::grepl(pattern = "Gilmour", x = StreamName) |
        base::grepl(pattern = "Stockdale", x = StreamName)
    ) %>%
    dplyr::mutate(
      stream = dplyr::case_when(
        base::grepl(pattern = "Erb", x = StreamName) ~ "Erb",
        base::grepl(pattern = "Paddy", x = StreamName) ~ "Paddy",
        base::grepl(pattern = "Hogan", x = StreamName) ~ "Hogan",
        base::grepl(pattern = "Gilmour", x = StreamName) ~ "Gilmour",
        base::grepl(pattern = "Stockdale", x = StreamName) ~ "Stockdale"
      )
    ) %>%
    dplyr::select(
      year,
      date,
      stream,
      StreamName,
      CommonName,
      AliveCount,
      DeadCount,
      DistanceMeters
    )
)
```

```{r}
survey_data %>% 
  dplyr::count(stream, StreamName)
```

## Erb 2014

```{r}
reduced_Erb <- all_streams_paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Erb",
                length > 300) %>% 
  dplyr::mutate(success = dplyr::case_when(n > 0 ~ "Successful",
                                           n == 0 ~ "Unsuccessful")) %>%
  dplyr::select(stream, origin, n, success, sex, date, length, intertidal, distance_mouth) %>% 
  tidyr::drop_na()


reduced_Erb %>% 
  dplyr::count(stream, date) %>% 
  dplyr::left_join(survey_data, by = c("stream", "date"))
```

```{r}
reduced_Erb %>% 
  dplyr::left_join(survey_data, by = c("stream", "date")) %>% 
  ggplot2::ggplot(ggplot2::aes(x = DeadCount, y = n, colour = origin)) +
  ggplot2::geom_jitter(alpha = 0.5) +
  ggplot2::geom_smooth(method = loess) +
  ggplot2::theme_bw()
```

End...for now