---
title: "Stream Surveys"
subtitle: "Investigating Live/Dead Counts"
author: "Kyle Shedd"
date: "15 February 2022"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list = ls())
library(tidyverse)
library(leaflet)
library(lubridate)
library(ggExtra)
library(gganimate)

knitr::opts_chunk$set(out.width = "100%", fig.width = 10)
```

# Background

There are concerns that field crews quit sampling pre-maturely in 2017 and 2019. The contractor reports basically confess to quiting early in 2017 and hint that they might have in 2019. Here I look at the live count data to determine the magnitude of this issue.

# Load Stream Survey Data

## 2017

```{r}
(stream_surveys_2017 <- readr::read_csv("~/../Desktop/Local_PWS_pinks/StreamSpecimens/stream_surveys_2017.csv"))
```

Plot live counts by stream
**NOTE** this is naively ignoring the start/end locations of the surveys!!!
```{r}
date_min_xlim <- stream_surveys_2017 %>%
  dplyr::mutate(date = lubridate::mdy(SurveyDate)) %>%
  dplyr::filter(CommonName == "Pink Salmon", !is.na(AliveCount)) %>% 
  dplyr::summarise(min(date)) %>% 
  dplyr::pull()

date_max_xlim <- stream_surveys_2017 %>%
  dplyr::mutate(date = lubridate::mdy(SurveyDate)) %>%
  dplyr::filter(CommonName == "Pink Salmon", !is.na(AliveCount)) %>% 
  dplyr::summarise(max(date)) %>% 
  dplyr::pull()

stream_surveys_2017 %>%
  dplyr::mutate(date = lubridate::mdy(SurveyDate)) %>%
  dplyr::filter(CommonName == "Pink Salmon", !is.na(AliveCount)) %>%
  ggplot2::ggplot(ggplot2::aes(x = date, y = AliveCount)) +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_line(colour = "pink2", lwd = 2) +
  ggplot2::geom_point(data = . %>% group_by(StreamName) %>% filter(date == max(date))) +
  ggplot2::geom_text(
    data = . %>% group_by(StreamName) %>% filter(date == max(date)),
    ggplot2::aes(label = AliveCount),
    hjust = 0,
    nudge_x = 1
  ) +
  ggplot2::labs(x = "Sample Date",
                y = "Live Count",
                title = "2017") +
  ggplot2::scale_x_date(limits = c(date_min_xlim, date_max_xlim + 10)) +
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(ggplot2::vars(StreamName), scales = "free_y")
```

## 2019

```{r}
(stream_surveys_2019 <- readr::read_csv("~/../Desktop/Local_PWS_pinks/StreamSpecimens/stream_surveys_2019.csv"))
```

Plot live counts by stream
**NOTE** this is naively ignoring the start/end locations of the surveys!!!
```{r}
date_min_xlim <- stream_surveys_2019 %>%
  dplyr::mutate(date = lubridate::mdy(SurveyDate)) %>%
  dplyr::filter(CommonName == "Pink Salmon", !is.na(AliveCount)) %>% 
  dplyr::summarise(min(date)) %>% 
  dplyr::pull()

date_max_xlim <- stream_surveys_2019 %>%
  dplyr::mutate(date = lubridate::mdy(SurveyDate)) %>%
  dplyr::filter(CommonName == "Pink Salmon", !is.na(AliveCount)) %>% 
  dplyr::summarise(max(date)) %>% 
  dplyr::pull()

stream_surveys_2019 %>%
  dplyr::mutate(date = lubridate::mdy(SurveyDate)) %>%
  dplyr::filter(CommonName == "Pink Salmon", !is.na(AliveCount)) %>%
  ggplot2::ggplot(ggplot2::aes(x = date, y = AliveCount)) +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_line(colour = "pink2", lwd = 2) +
  ggplot2::geom_point(data = . %>% group_by(StreamName) %>% filter(date == max(date))) +
  ggplot2::geom_text(
    data = . %>% group_by(StreamName) %>% filter(date == max(date)),
    ggplot2::aes(label = AliveCount),
    hjust = 0,
    nudge_x = 1
  ) +
  ggplot2::labs(x = "Sample Date",
                y = "Live Count",
                title = "2019") +
  ggplot2::scale_x_date(limits = c(date_min_xlim, date_max_xlim + 10)) +
  ggplot2::theme_bw() +
  ggplot2::facet_wrap(ggplot2::vars(StreamName), scales = "free_y")
```