---
title: "Data Checking for Sam May Tidal Gradients Manuscript Resubmission"
subtitle: "Make sure all of our numbers are correct!"
author: "Kyle Shedd"
date: "21 July 2023"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list=ls())

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DT)

.username = readLines("~/R/usr_pw.txt", n = 1)
.password = readLines("~/R/usr_pw.txt" , n = 2)[[2]]
source("~/R/Functions.GCL.R")  # develop branch!!!
```

# Objective

Check Sam May's numbers in our "Tidal gradients, fine-scale homing, and a potential cryptic ecotype of wild spawning pink salmon (Oncorhynchus gorbuscha)" Manuscript

# Read in Data

```{r}
load_objects(path = "../Objects/All_Streams_14_16_develop/", pattern = "sample_size_qa")

(paired_14_16_filter <- readr::read_csv(file = "Stockdale_Hogan_Gilmour_Paddy_Erb/all_streams_paired_14_16_filter.csv"))

(parents_paired_14_16 <- readr::read_csv(file = "Stockdale_Hogan_Gilmour_Paddy_Erb/all_streams_parents_paired_14_16.csv"))
```

# QA of Results

## Samples Sizes from Genotype QA

```{r}
sample_size_qa %>% 
  dplyr::filter(!silly %in% c("PPADDY14", "PPADDY16")) %>% 
  dplyr::mutate(sample_year = as.numeric(paste0("20", stringr::str_sub(string = silly, start = -2, end = -1)))) %>% 
  dplyr::group_by(sample_year) %>%
  dplyr::summarise(sum(genotyped), sum(missing), sum(duplicate), sum(heterozygosity), sum(final))
```

## How Many Unique Parents?

This is the mistake I think Sam made...
```{r}
parents_paired_14_16 %>% 
  dplyr::filter(stream_off != "Paddy") %>% 
  dplyr::pull(parent_id) %>% 
  dplyr::n_distinct()
```

He forgot to exclude both parents and offspring from Paddy!
```{r}
parents_paired_14_16 %>% 
  dplyr::filter(stream_off != "Paddy" & stream_par!= "Paddy") %>% 
  dplyr::pull(parent_id) %>% 
  dplyr::n_distinct()
```

## How Many Offspring with at Least 1 Parent?

This is the mistake I think Sam made...
```{r}
parents_paired_14_16 %>% 
  dplyr::filter(stream_off != "Paddy") %>% 
  dplyr::pull(offspring_id) %>% 
  dplyr::n_distinct()
```

He forgot to exclude both parents and offspring from Paddy!
```{r}
parents_paired_14_16 %>% 
  dplyr::filter(stream_off != "Paddy" & stream_par!= "Paddy") %>% 
  dplyr::pull(offspring_id) %>% 
  dplyr::n_distinct()
```

## How Many Unique Parents WITHOUT Dispersers?

No mistakes here since `stream_off == stream_par` and we are already filtering out Paddy
```{r}
parents_paired_14_16 %>% 
  dplyr::filter(stream_off != "Paddy" & stream_par!= "Paddy",
                stream_off == stream_par) %>%
  dplyr::pull(parent_id) %>% 
  dplyr::n_distinct()
```

## How Many Unique Offspring WITHOUT Dispersers?

No mistakes here since `stream_off == stream_par` and we are already filtering out Paddy
```{r}
parents_paired_14_16 %>% 
  dplyr::filter(stream_off != "Paddy" & stream_par!= "Paddy",
                stream_off == stream_par) %>%
  dplyr::pull(offspring_id) %>% 
  dplyr::n_distinct()
```

## How Many Dyads vs. Triads?

```{r}
parents_paired_14_16 %>% 
  dplyr::filter(stream_off != "Paddy" & stream_par!= "Paddy",
                stream_off == stream_par) %>% 
  dplyr::count(parent)
```

## How Many Offspring Left After Removing `NA` for GLM Variables?

```{r}
parents_paired_14_16 %>% 
  dplyr::filter(stream_off != "Paddy" & stream_par!= "Paddy",
                stream_off == stream_par,
                !is.na(distance_mouth_off),
                !is.na(distance_mouth_par),
                !is.na(sex_off),
                !is.na(sex_par),
                !is.na(length_off),
                !is.na(DOY_off)) %>%
  dplyr::pull(offspring_id) %>% 
  dplyr::n_distinct()
```

## How Many Dyads vs. Triads Left After Removing `NA` for GLM Variables?

```{r}
parents_paired_14_16 %>% 
  dplyr::filter(stream_off != "Paddy" & stream_par!= "Paddy",
                stream_off == stream_par,
                !is.na(distance_mouth_off),
                !is.na(distance_mouth_par),
                !is.na(sex_off),
                !is.na(sex_par),
                !is.na(length_off),
                !is.na(DOY_off)) %>%
  dplyr::count(parent)
```

What about just Dyads?
```{r}
5459-1045
```

Otherwise the numbers check out!

End