---
title: "Gilmour 2016/2018 Parentage Analysis with 298 Markers"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list=ls())

library(kinship2)
library(coin)
library(MASS)
library(scales)
library(tidyverse)
library(leaflet)
library(lubridate)
library(ggExtra)

knitr::opts_chunk$set(out.width = "100%", fig.width = 10)
```

Below are analyses of parentage results for Even (2016-2018) using *FRANz*. 

# Even lineage (2016/2018)

## *FRANz* parameters
*FRANz* was run at Mon Oct 26 13:45 2020. We used the following parameters:
FRANz.exe --Nmmax 5000 --Nfmax 5000 --femrepro 1:2 --malerepro 1:2 --typingerror 0.006 --updatefreqs --poutformat 2 "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\Gilmour_2016_2018\Gilmour_16_18_postQA_2016_2018_GILMOUR.dat"

Do not use --fullsibtest if running on Kyle's Genetics Domain Computer, it won't save output!!!

The parameters are defined as follows:
--Nmmax and --Nfmax are the maximum numbers of candidate mothers and fathers. To obtain our values, we used an estimated escapement of 10,000 and divided by 2 (*note: no aerial survey for 2016 and stream walk was 4,274*). This escapement number came from the stream walk AUC number, as I did not have an aerial survey AUC. This estimate is listed under 'pop est walk' in the Excel file found here: "V:\Documents\5_Coastwide\Multispecies\AHRP\Field data\Sample Summary Extract and Genotype.xlsx"
--femrepro and --malerepro specify the age range in which an individual can reproduce  
--typingerror refers to the overall genotyping error rate. Ours was ~0.006 (for Gilmour was actually ???, from QC)
--updatefreqs specifies that *FRANz* should update allele frequencies using MCMC sampling  
--poutformat specifies that all potential parents should be listed, not just the most likely  
--fullsibtest tests for full siblings among offspring   

All output files can be found here: "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\Gilmour_2016_2018"  

The summary file provides information about the power of our marker suite:   
Cumulative exclusion probability when 1 to 7 fullsibs are genotyped  
  First Parent              : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Second Parent             : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Parent Pair               : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000
  
According to the FRANz manual, marker sets are not considered powerful if these cumulative exclusion probabilities are less than 0.95, which indicates that the probability that a random pair of individuals in the population has a 5% chance of having a genotype pair compatible to an offspring genotype. Since all of our probabilities are 1, we can be **confident** in the power of our 298 amplicons to make parent assignments.  

## Import data

The first step is to read in .csv files for parentage assignments produced by *FRANz* as well as paired genotype and OceanAK data. 
```{r Setup Even, message=FALSE, warning=FALSE}
parentage_16_18 <- read_csv("../../Franz/Gilmour_2016_2018/parentage.csv")
paired_16_18 <- read_csv("../Franz/Gilmour_16_18_postQA_OceanAK_paired_2016_2018_GILMOUR.csv")
```

Plot a histogram of *FRANz* parentage posterior probabilities to show robustness of assignments
```{r FRANz_posterior_plot_even}
parentage_16_18  %>% 
  filter(!is.na(`Parent 1`)) %>% 
  ggplot(aes(x = Posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  ggtitle("Histogram of FRANz posterior probabilities for parentage assignments")
```

Almost all parentage assingments have a posterior probability of 1, which is very robust.  

## Filter paired data
The file containing paired genotype and OceanAK data has a lot of information that we do not need at this time (e.g. genotypes for each marker). We therefore separate out only the columns that contain identifying information for each individual.

```{r Filter paired data Even}
# What are the non-genotype columns
grep(pattern = "RAD", x = colnames(paired_16_18), value = TRUE, invert = TRUE)

# Filter for non-genotype columns
paired_16_18_filter <- paired_16_18 %>% 
  dplyr::select(franz_id, SILLY, `Fish ID`, `DNA Tray Code`, `DNA Tray Well Code`, `Sample Year`, `Sample Date`, SEX, `Length Mm`, `Otolith Mark Present`, `Otolith Mark ID`) %>% 
  mutate(origin = case_when(`Otolith Mark Present` == "NO" ~ "Natural",
                            `Otolith Mark Present` == "YES" ~ "Hatchery")) %>%  # add origin variable
  mutate(origin = factor(origin, c("Natural", "Hatchery"))) %>%  # make factor to ensure hatchery != red
  mutate(Sex = case_when(SEX == "M" ~ "Male", SEX == "F" ~ "Female")) %>% 
  mutate(date = dmy(`Sample Date`)) %>% 
  mutate(DOY = yday(date)) 

write_csv(paired_16_18_filter, "gilmour_paired_16_18_filter.csv")
```

```{r Filter parentage data Even}
# Filter for posterior > 0.9 + make tidy relative to assigment (1 row per parentage assignment)
parentage_16_18_filter <- parentage_16_18 %>% 
  filter(Posterior > 0.9) %>% 
  dplyr::select(Offspring, `Parent 1`, `Parent 2`) %>% 
  gather(Parent, Parent_ID, - Offspring) %>% 
  filter(!is.na(Parent_ID))
```

## Sample sizes
Quick look at what our sample sizes are for potential parents (2016) and potential offspring (2018).

```{r even_sample_sizes}
paired_16_18_filter %>% 
  count(`Sample Year`, SEX, origin) %>% 
  spread(origin, n, fill = 0)
```


```{r even_sample_sizes_plot}
paired_16_18_filter %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(Sex = case_when(SEX == "M" ~ "Male", SEX == "F" ~ "Female")) %>% 
  mutate(date = dmy(`Sample Date`)) %>% 
  mutate(DOY = yday(date)) %>% 
  ggplot(aes(x = DOY)) +
  theme_bw() +
  geom_bar(aes(fill = origin), colour = "black") +
  facet_grid(`Sample Year` ~ .) +
  geom_hline(yintercept = 0, colour = "black") +
  ylab("Number of Samples") +
  xlab("Day of Year") +
  labs(title= "Gilmour 2016/2018",
       fill = "Origin") +
  theme(text = element_text(size = 20))
```

## Join parentage and individual data
Here, we joined parentage data from *FRANz* with individual data so that we can match data to each offspring and parent. **Note** that any column ending in `.off` refers to offspring and `.par` refers to parent.  

```{r Join parentage and paired objects Even}
# Creating a single, "tidy" object where each row is a parent-offspring relationship
# Offspring with 2 parents have 2 rows
# .off data is offspring, .par data is parents
(parents_paired_16_18 <- parentage_16_18_filter %>% 
   left_join(paired_16_18_filter, by = c("Offspring" = "franz_id")) %>% 
   left_join(paired_16_18_filter, by = c("Parent_ID" = "franz_id"), suffix = c(".off", ".par")) %>% 
   mutate(origin = case_when(`Otolith Mark Present.par` == "NO" ~ "Natural",
                             `Otolith Mark Present.par` == "YES" ~ "Hatchery")) %>% 
   mutate(origin = factor(origin, c("Natural", "Hatchery"))))  # added hatchery/natural variable

write_csv(parents_paired_16_18, "gilmour_parents_paired_16_18.csv")

# How many single-parent offspring pair assignments?
parents_paired_16_18 %>% 
  count(Parent)

# How many unique parents had offspring assigned?
n_distinct(parents_paired_16_18$Parent_ID)

# How many unique parents by origin had offspring assigned?
parents_paired_16_18 %>% 
  group_by(origin) %>% 
  summarise(n = n_distinct(Parent_ID))

# For each offspring assigment, what was the parent's origin?
parents_paired_16_18 %>% 
  count(origin, `Otolith Mark ID.par`) %>% 
  spread(origin, n)
```

**Note** that *FRANz* assigned ???? offspring to ???? parents. ???? of these offspring have natural-origin parents and ???? offspring have hatchery-origin parents. ???? of the hatchery origin fish are from AFK14B, ???? are from CCH12, and ???? are from WNH12PINKB.

## Proportion test
Let's use a Chi-Square test to determine whether the proportions of offspring assigned to hatchery- and natural-origin parents significantly differ from the proportions of hatchery- and natural-origin parents sampled. 

```{r proportion_even_data}
# How many spawners from each origin?
paired_16_18_filter %>% 
  filter(`Sample Year` == 2016) %>% 
  count(origin)

# How many offspring from each origin?
parents_paired_16_18 %>% 
  count(origin)
```

```{r Chi square test even}
O.even <- c("H", "N")
Spawners.even <- c(28, 278) #  number of hatchery- and natural-origin parents
Offspring.even <- c(5, 80)  # number of offspring assigned to hatchery- and natural-origin parents

df.even <- data.frame(Spawners.even, Offspring.even)  # made a data.frame because tibbles don't like rownames
row.names(df.even) <- O.even

df.even
chisq.test(df.even)
```

There is a highly significant difference in the proportions of offspring assigned to hatchery- and natural-origin parents relative to the proportions of potential parents sampled, indicating an under-representation of offspring assigning to hatchery-origin parents. We see that although **`r round(df.even["H", "Spawners.even"] / sum(df.even[, "Spawners.even"]), 2) * 100`%** of parents genotyped were of hatchery-origin, only **`r round(df.even["H", "Offspring.even"] / sum(df.even[, "Offspring.even"]), 2) * 100`%** of offspring assigned to hatchery-origin parents.

## Family size

Here, we calculate and plot family size for both natural and hatchery parents.  Family size ranges from 1-????. Most assignments were to 1 parent (*single parent-offspring duos*) - there were ???? 2-parent assignments (*parent pair-offspring trios*). 

# Of the 2,710 offspring successfully genotyped (of 2,833 collected) in 2017, ???? were assigned to parents, for an overall offspring assignment rate of **????**. 

### Single parent-offspring pairs

#### Successful parents (no 0's)
This plot shows the distribution of family size by sex (number of offspring assigned to a female parent or a male parent) for both hatchery and natural parents.  **Note** this plot doesn't include the 0's.
```{r Family size by sex Even}
#parents_paired_16_18 %>% 
  #count(`Fish ID.par`, origin, SEX.par) %>% 
  #count(origin, SEX.par, n) %>% 
  #complete(origin, SEX.par, n, fill = list(nn = 0)) %>% 
  #ggplot(aes(x=n, y = nn, fill=origin)) +
  #geom_col(position = position_dodge2(preserve = "single")) +
  #scale_y_continuous(breaks = pretty_breaks()) +
  #scale_x_continuous(breaks = pretty_breaks()) +
  #facet_grid(~ SEX.par) +
  #labs(title="Distribution of Family Size by Females (F) and Males (M)\nNumber of Families",
  #     fill = "Parent: Origin") +
  #xlab("Number of Offspring")+
  #ylab("Number of Parents") 

# Plot histogram of proportion of parents with a given family size
parents_paired_16_18 %>% 
  count(`Fish ID.par`, origin, SEX.par) %>% 
  count(SEX.par, origin, n) %>% 
  group_by(SEX.par, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x=n, y=p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  facet_grid(~ SEX.par) +
  labs(title="Distribution of Family Size by Females (F) and Males (M)\nProportion of Families",
       fill = "Parent: Origin") +
  xlab("Number of Offspring")+
  ylab("Proportion of Parents") 
```

Female parents produced ???? offspring, while male parents produced ????.  

Below, we'll calculate reproductive success of natural and hatchery females and males. 

```{r even_rrs}
(rs_16_18 <- parents_paired_16_18 %>%
  count(`Fish ID.par`, origin, SEX.par) %>%
  group_by(origin, SEX.par) %>% 
  summarise(RS = mean(n))) %>% 
  spread(SEX.par, RS)
  

F_h_even <- rs_16_18 %>% 
  filter(origin == "Hatchery" & SEX.par == "F") %>% 
  pull(RS)
M_h_even <- rs_16_18 %>% 
  filter(origin == "Hatchery" & SEX.par == "M") %>% 
  pull(RS)
F_n_even <- rs_16_18 %>% 
  filter(origin == "Natural" & SEX.par == "F") %>% 
  pull(RS)
M_n_even <- rs_16_18 %>% 
  filter(origin == "Natural" & SEX.par == "M") %>% 
  pull(RS)
```

#### Calculate RRS again including 0's 
We'll calculate RRS again including sampled fish from the parental generation to which offspring were not assigned.

```{r Re-calculate RRS even}
# Make one tibble with all assignment data
# RS_combined <- bind_rows(parents_paired_16_18, parents_paired_16_18) 

# Transform to one row per parent with n = number of offspring produced
parents_paired_16_18_count <- parents_paired_16_18 %>% 
  count(Parent_ID, origin, SEX.par, `Sample Year.par`, `Length Mm.par`, `Sample Date.par`) %>% 
  mutate(lineage = dplyr::case_when(`Sample Year.par` %in% c("2015", "2017") ~ "odd",
                                    `Sample Year.par` %in% c("2014", "2016", "2018") ~ "even"))

paired_16_18_filter_parents <- paired_16_18_filter %>% 
  filter(`Sample Year` == "2016")

# Join with all parents that produced no progeny
paired_16_18_filter_parents <- left_join(paired_16_18_filter_parents, parents_paired_16_18_count, by = c("franz_id" = "Parent_ID")) %>% 
  mutate(n = replace_na(n, 0)) %>% 
  rename(origin = origin.x) %>% 
  dplyr::select(-origin.y, -SEX.par, -`Sample Year.par`, -`Length Mm.par`, -`Sample Date.par`)

write_csv(paired_16_18_filter_parents, "gilmour_paired_16_18_filter_parents.csv")

# Mean RS including 0's
rs_16_18_0 <- paired_16_18_filter_parents %>%
  group_by(origin, SEX) %>% 
  summarise(RS = mean(n, na.rm = TRUE))

rs_16_18_0 %>%
  mutate(RS = round(RS, 2)) %>% 
  spread(SEX, RS)

# Plot histogram of number of parents with a given family size
#paired_16_18_filter_parents %>% 
#  ggplot(aes(x = n, fill = origin)) +
#  geom_bar(position = position_dodge2(preserve = "single")) +
#  facet_grid( ~ SEX) +
#  xlab("Number of Offspring") +
#  ylab("Number of Parents") +
#  labs(title = "Distribution of Family Size by Females (F) and Males (M)\nNumber of Families",
#       fill = "Parent: Origin")

# Plot histogram of proportion of parents with a given family size
paired_16_18_filter_parents %>% 
  filter(SEX != "?") %>% 
  mutate(Sex = case_when(SEX == "M" ~ "Male", SEX == "F" ~ "Female")) %>% 
  count(Sex, origin, n) %>% 
  group_by(Sex, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  facet_grid(~ Sex) +
  labs(title="Distribution of Family Size",
       fill = "Parent Origin") +
  xlab("Number of Offspring")+
  ylab("Proportion of Parents")  +
  theme_bw() +
  theme(text = element_text(size = 20))

# Save RS values
F_h_0_16_18 <- rs_16_18_0 %>% 
  filter(origin == "Hatchery" & SEX == "F") %>% 
  pull(RS)
M_h_0_16_18 <- rs_16_18_0 %>% 
  filter(origin == "Hatchery" & SEX == "M") %>% 
  pull(RS)
F_n_0_16_18 <- rs_16_18_0 %>% 
  filter(origin == "Natural" & SEX == "F") %>% 
  pull(RS)
M_n_0_16_18 <- rs_16_18_0 %>% 
  filter(origin == "Natural" & SEX == "M") %>% 
  pull(RS)
```

Here are the plots again with lines showing mean reproductive success. 

```{r Plots of RS by sex with mean lines even}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

#paired_16_18_filter_parents %>% 
#  filter(SEX == "F") %>% 
#  ggplot(aes(x = n, fill = origin)) +
#  geom_bar(position = position_dodge2(preserve = "single")) +
#  scale_y_continuous(breaks = pretty_breaks()) +
#  scale_x_continuous(breaks = pretty_breaks()) +
#  xlab("Number of Offspring") +
#  ylab("Number of Females") +
#  labs(title="Number of Offspring for Females",
#      fill = "Parent: Origin") +
#  geom_vline(xintercept=F_n_0_16_18, size=2) +
#  geom_vline(xintercept=F_n_0_16_18, color=gg_color_hue(2)[1], linetype = "dashed", size=2) +
#  geom_vline(xintercept=F_h_0_16_18, size=2) +
#  geom_vline(xintercept=F_h_0_16_18, color=gg_color_hue(2)[2], linetype = "dashed", size=2) +
#  annotate("text", x = F_n_0_16_18 + 0.5, y = 250, label = paste("F_n_0 =", round(F_n_0_16_18, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[1]) +
#  annotate("text", x = F_h_0_16_18 + 0.5, y = 300, label = paste("F_h_0 =", round(F_h_0_16_18, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[2])

#paired_16_18_filter_parents %>% 
#  filter(SEX == "M") %>% 
#  ggplot(aes(x = n, fill = origin)) +
#  geom_bar(position = position_dodge2(preserve = "single")) +
#  scale_y_continuous(breaks = pretty_breaks()) +
#  scale_x_continuous(breaks = pretty_breaks()) +
#  xlab("Number of Offspring") +
#  ylab("Number of Males") +
#  labs(title="Number of Offspring for Males",
#       fill = "Parent: Origin") +
#  geom_vline(xintercept=M_n_0_16_18, size=2) +
#  geom_vline(xintercept=M_n_0_16_18, color=gg_color_hue(2)[1], linetype = "dashed", size=2) +
#  geom_vline(xintercept=M_h_0_16_18, size=2) +
#  geom_vline(xintercept=M_h_0_16_18, color=gg_color_hue(2)[2], linetype = "dashed", size=2) +
#  annotate("text", x = M_n_0_16_18 + 0.5, y = 150, label = paste("M_n_0 =", round(M_n_0_16_18, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[1]) +
#  annotate("text", x = M_h_0_16_18 + 0.5, y = 100, label = paste("M_h_0 =", round(M_h_0_16_18, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[2])

paired_16_18_filter_parents %>% 
  filter(SEX == "F") %>% 
  count(origin, n) %>% 
  group_by(origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x=n, y=p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +xlab("Number of Offspring") +
  ylab("Proportion of Females") +
  labs(title="Number of Offspring for Females",
       fill = "Parent: Origin") +
  geom_vline(xintercept=F_n_0_16_18, size=2) +
  geom_vline(xintercept=F_n_0_16_18, color=gg_color_hue(2)[1], linetype = "dashed", size=2) +
  geom_vline(xintercept=F_h_0_16_18, size=2) +
  geom_vline(xintercept=F_h_0_16_18, color=gg_color_hue(2)[2], linetype = "dashed", size=2) +
  annotate("text", x = F_n_0_16_18 + 0.5, y = .250, label = paste("F_n_0 =", round(F_n_0_16_18, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[1]) +
  annotate("text", x = F_h_0_16_18 + 0.5, y = .300, label = paste("F_h_0 =", round(F_h_0_16_18, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[2]) 

paired_16_18_filter_parents %>% 
  filter(SEX == "M") %>% 
  count(origin, n) %>% 
  group_by(origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x=n, y=p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  xlab("Number of Offspring") +
  ylab("Proportion of Males") +
  labs(title="Number of Offspring for Males",
       fill = "Parent: Origin") +
  geom_vline(xintercept=M_n_0_16_18, size=2) +
  geom_vline(xintercept=M_n_0_16_18, color=gg_color_hue(2)[1], linetype = "dashed", size=2) +
  geom_vline(xintercept=M_h_0_16_18, size=2) +
  geom_vline(xintercept=M_h_0_16_18, color=gg_color_hue(2)[2], linetype = "dashed", size=2) +
  annotate("text", x = M_n_0_16_18 + 0.5, y = .250, label = paste("M_n_0 =", round(M_n_0_16_18, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[1]) +
  annotate("text", x = M_h_0_16_18 + 0.5, y = .300, label = paste("M_h_0 =", round(M_h_0_16_18, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[2])

```

```{r Calculate relative reproductive success including 0s even}
RRS_m_0_16_18 <- M_h_0_16_18/M_n_0_16_18
RRS_f_0_16_18 <- F_h_0_16_18/F_n_0_16_18

RRS_tibble_0_16_18 <- tibble(RRS_m_0_16_18, RRS_f_0_16_18)
round(RRS_tibble_0_16_18, 2)
```

**RRS for males = `r round(RRS_m_0_16_18, 2)`. RRS for females = `r round(RRS_f_0_16_18, 2)`.**

Let's test for significant differences in RRS between male and female hatchery and natural origin fish. We'll perform two different statistical tests:  

  1. Nonparametric permutation test
  2. Parametric negative binomial GLM  
  
**Note** we are only doing this for the odd lineage because the sample sizes are too low in the odd lineage. 

```{r RRS statistical testing including 0s even}
# The format of the data to test was very simple, it was a data.frame called “mydata” with 2 columns
# 1.	nOff = number of offspring per family
# 2.	Origin = “H” or “W” for if the parent was hatchery or natural (wild)
# 

mydata_M_0 <- paired_16_18_filter_parents %>%
  filter(SEX == "M")

mydata_F_0 <- paired_16_18_filter_parents %>%
  filter(SEX == "F")

perm_1tail_pvalue_M_0 <- round(x = as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_M_0, distribution = approximate(B = 10000), alternative = "greater"))), digits = 4)

perm_1tail_pvalue_F_0 <- round(x = as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_F_0, distribution = approximate(B = 10000), alternative = "greater"))), digits = 4)

perm_1tail_pvalue_both_0 <- round(x = as.numeric(coin::pvalue(oneway_test(n ~ origin, data = paired_16_18_filter_parents, distribution = approximate(B = 10000), alternative = "greater"))), digits = 4)
#And to do a 1-tailed negative binomial GLM

fit_M_0 <- glm.nb(n ~ origin, data = mydata_M_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_M_0 <- round(summary(fit_M_0)$coefficients[2, 4], 4)

fit_F_0 <- glm.nb(n ~ origin, data = mydata_F_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_F_0 <- round(summary(fit_F_0)$coefficients[2, 4], 4)

fit_both_0 <- glm.nb(n ~ origin, data = paired_16_18_filter_parents, init.theta = 1, link = log)
nbGLM_1tail_pvalue_both_0 <- round(summary(fit_both_0)$coefficients[2, 4], 4)

RRS_0 <- tibble(
  Sex = c("M", "F", "Both", "M", "F", "Both"),
  Test = c(rep("Permutation", 3), rep("Negative Binomial GLM", 3)),
  p_value = c(
    perm_1tail_pvalue_M_0,
    perm_1tail_pvalue_F_0,
    perm_1tail_pvalue_both_0,
    nbGLM_1tail_pvalue_M_0,
    nbGLM_1tail_pvalue_F_0,
    nbGLM_1tail_pvalue_both_0
  )
)
RRS_0 %>%
  spread(Sex, p_value)

# (RRS_0 <- tibble(perm_1tail_pvalue_F_0, perm_1tail_pvalue_M_0, nbGLM_1tail_pvalue_F_0, nbGLM_1tail_pvalue_M_0))
```

There is a significant difference in RRS for both males and females. 

#### Calculate CI's for RRS

We will calculate 95% CI's for RRS based on Kalinowski and Taper 2005 [link] (http://www.nrcresearchpress.com/doi/10.1139/f04-239). This is a maximum likelihood estimate based off of Hinrichsen 2003. This method has been used in other peer reviewed papers such as Ford et al. 2016 [link] (https://www.researchgate.net/publication/309147116_Broodstock_History_Strongly_Influences_Natural_Spawning_Success_in_Hatchery_Steelhead_Oncorhynchus_mykiss#pf14).

```{r RRS_confidence_intervals_function_even}
rrs_ci_kalinowski <- function(n_h_off, n_w_off, n_h_par, n_w_par, alpha){
  chi_alpha <- qchisq(p = (1 - alpha), df = 1)
  n_off <- sum(c(n_h_off, n_w_off))
  n_par <- sum(c(n_h_par, n_w_par))
  
  rs_h <- n_h_off / n_h_par
  rs_w <- n_w_off / n_w_par
  
  p_h_par <- n_h_par / n_par
  p_w_par <- n_w_par / n_par
  
  rrs_h <- rs_h / rs_w
  rrs_w <- rs_w / rs_w
  rrs_avg <- (rrs_h * p_h_par) + (rrs_w * p_w_par)
  
  rrs_ml <- (n_h_off * log(p_h_par * rrs_h / rrs_avg)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg))
  
  xi_dist <- bind_rows(
    lapply(seq(0.01, 5, by = 0.01), function(rrs_h_xi) {
      rrs_avg_xi <- (rrs_h_xi * p_h_par) + (rrs_w * p_w_par)
      tibble(rrs_crit = rrs_h_xi,
             logl = (n_h_off * log(p_h_par * rrs_h_xi / rrs_avg_xi)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg_xi)) - (rrs_ml - chi_alpha / 2)
      )
    } )
  )
  
  rrs_min <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit < rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  rrs_max <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit > rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  xi_plot <- xi_dist %>% 
    ggplot(aes(x = rrs_crit, y = logl)) +
    geom_line() +
    geom_hline(yintercept = 0, colour = "red", lwd = 2) +
    geom_vline(xintercept = c(rrs_h, rrs_min, rrs_max), colour = "blue") +
    ylim(c(-5, 5)) +
    xlim(c(0, 2)) +
    ylab("Log Likelihood - Chi Sq Value") +
    annotate("text", x = rrs_h + 0.1, y = xi_dist %>% filter(rrs_crit == xi_dist$rrs_crit[which.min(abs(xi_dist$rrs_crit  - rrs_h))]) %>% pull(logl) + 0.4, label = round(rrs_h, 2)) +
    annotate("text", x = rrs_min - 0.1, y = xi_dist %>% filter(rrs_crit == rrs_min) %>% pull(logl) + 0.4, label = rrs_min) +
    annotate("text", x = rrs_max + 0.1, y = xi_dist %>% filter(rrs_crit == rrs_max) %>% pull(logl) + 0.4, label = rrs_max)
  
  print(xi_plot)
  return(c(rrs_min, rrs_h, rrs_max))
}
```

##### Even 2016/2018

Calculate 95% CIs for Gilmour 16-18 females. RRS = 0.?? (0.??-0.??)

```{r RRS_confidence_intervals_female_16_18}
# Females
# Get parent sample sizes
paired_16_18_filter_parents %>% 
  select(-n) %>% 
  filter(SEX == "F") %>% 
  count(SEX, origin) %>% 
  spread(origin, n)

# Get offspring sample sizes
paired_16_18_filter_parents %>% 
  filter(SEX == "F") %>% 
  group_by(SEX, origin) %>% 
  summarise(n_off = sum(n)) %>% 
  spread(origin, n_off)

rrs_ci_kalinowski(n_h_off = 3, n_w_off = 48, n_h_par = 12, n_w_par = 119, alpha = 0.05)
```

Calculate 95% CIs for Gilmour 16-18 males. RRS = 0.?? (0.??-0.??)

```{r RRS_confidence_intervals_male_16_18}
# Males
# Get parent sample sizes
paired_16_18_filter_parents %>% 
  select(-n) %>% 
  filter(SEX == "M") %>% 
  count(SEX, origin) %>% 
  spread(origin, n)

# Get offspring sample sizes
paired_16_18_filter_parents %>% 
  filter(SEX == "M") %>% 
  group_by(SEX, origin) %>% 
  summarise(n_off = sum(n)) %>% 
  spread(origin, n_off)

rrs_ci_kalinowski(n_h_off = 2, n_w_off = 32, n_h_par = 16, n_w_par = 159, alpha = 0.05)
```

#### RS vs. field data plots

```{r paired_16_18_filter_parents_sex_even}
paired_16_18_filter_parents <- paired_16_18_filter_parents %>% 
  filter(SEX != "?") %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female")) %>% 
  mutate(date = dmy(`Sample Date`))
```

Natural-origin males and females both do better early in the season (density-dependence or increased fishing pressure later in the season?).
```{r Examine reproductive success by sample date even with zeros}
paired_16_18_filter_parents %>% 
  ggplot(aes(x = date, y = n, color = origin)) +
  geom_jitter(aes(shape = sex), height = 0) +
  geom_smooth() +
  labs(title = "Reproductive Success by Parental Sample Date",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Sample Date") +
  ylab("Reproductive Success")

paired_16_18_filter_parents %>% 
  ggplot(aes(x = date, y = n, color = origin)) +
  geom_jitter(height = 0) +
  geom_smooth() +
  facet_grid(sex ~ .)+
  labs(title = "Reproductive Success by Parental Sample Date",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Sample Date") +
  ylab("Reproductive Success")
```

Larger males are much more successful (but only natural-origin males).
```{r Examine reproductive success by parent length even with zeros}
paired_16_18_filter_parents %>% 
  filter(`Length Mm` > 300) %>% 
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(aes(shape = sex), height = 0) +
  geom_smooth() +
  labs(title = "Reproductive Success by Parental Length (mm)",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Reproductive Success")

paired_16_18_filter_parents %>% 
  filter(`Length Mm` > 300) %>% 
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(height = 0) +
  geom_smooth() +
  facet_grid(sex ~ .) +
  labs(title = "Reproductive Success by Parental Length (mm)",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Reproductive Success")
```

### Parent pair-offspring trios

Because we have ???? individuals who were assigned to 2 parents, we are able to examine instances of crosses between hatchery fish, between natural fish, and between natural and hatchery fish. **Note** that we have divided natural-hatchery crosses into "HN", which indicates a hatchery origin mother and natural origin father and "NH", which indicates a natural origin mother and hatchery origin father. 

```{r Create cross as new grouping variable even, message=FALSE}
# Which offspring have 2 parents?
offspring_trio_16_18 <- parents_paired_16_18 %>% 
  filter(Parent == "Parent 2") %>% 
  pull(Offspring)

parents_paired_16_18_cross <- parentage_16_18 %>% 
  filter(Posterior > 0.9) %>% 
  dplyr::select(Offspring, `Parent 1`, `Parent 2`) %>% 
  filter(Offspring %in% offspring_trio_16_18) %>% 
  left_join(paired_16_18_filter, by = c("Offspring" = "franz_id")) %>% 
  left_join(paired_16_18_filter, by = c("Parent 1" = "franz_id"), suffix = c(".off", ".par1")) %>% 
  left_join(paired_16_18_filter, by = c("Parent 2" = "franz_id")) %>% 
  dplyr::mutate(cross = dplyr::case_when(
                `Otolith Mark Present.par1` == "NO" & `Otolith Mark Present` == "NO" ~ "NN",
                `Otolith Mark Present.par1`== "YES" & `Otolith Mark Present` == "YES" ~ "HH",
                `Otolith Mark Present.par1` == "YES" & SEX.par1 == "F" & `Otolith Mark Present` == "NO" & SEX == "M" ~ "HN",
                `Otolith Mark Present.par1` == "NO" & SEX.par1 == "F" & `Otolith Mark Present` == "YES" & SEX == "M" ~ "NH",
                `Otolith Mark Present.par1` == "YES" & SEX.par1 == "M" & `Otolith Mark Present` == "NO" & SEX == "F" ~ "NH",
                `Otolith Mark Present.par1` == "NO" & SEX.par1 == "M" & `Otolith Mark Present` == "YES" & SEX == "F" ~ "HN" ))

write_csv(parents_paired_16_18_cross, "gilmour_parents_paired_16_18_cross.csv")

parents_paired_16_18_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  count(cross, n) %>% 
  complete(cross, n, fill = list(nn = 0)) %>% 
  ggplot(aes(x = n, y = nn, fill = cross))+
  geom_col(position = position_dodge2(preserve = "single"), width = 0.5) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = 1:6) +
  theme_bw() +
  # ylim(0, 40) +
  # xlim(0, 5) +
  labs(title = "Distribution of Family Size by Cross") +
  xlab("Number of Offspring")+
  ylab("Number of Families") 
```
Offspring from two types of crosses are represented in our dataset.

Calculate RS of cross types
```{r cross_rs_even}
parents_paired_16_18_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  group_by(cross) %>% 
  summarise(rs = mean(n))
```

Is this significant?
```{r RRS_confidence_intervals_16_18_setup}
# How many matings by cross type?
parents_paired_16_18_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  count(cross)

# Get offspring sample sizes
parents_paired_16_18_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  group_by(cross) %>% 
  summarise(nn = sum(n))
```

```{r RRS_confidence_intervals_16_18}
# HH vs. NN
# rrs_ci_kalinowski(n_h_off = 16, n_w_off = 109, n_h_par = 15, n_w_par = 62, alpha = 0.05) # Significant!

# HN vs. NN
#rrs_ci_kalinowski(n_h_off = 5, n_w_off = 61, n_h_par = 5, n_w_par = 55, alpha = 0.05) # NS

# NH vs. NN
#rrs_ci_kalinowski(n_h_off = 2, n_w_off = 61, n_h_par = 2, n_w_par = 55, alpha = 0.05) # NS
```

Investigate the potential for multiple matings
```{r multiple_matings_even}
parents_paired_16_18_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  group_by(`Fish ID`) %>% 
  summarise(n_mates = n_distinct(`Fish ID.par1`)) %>% 
  count(n_mates)

parents_paired_16_18_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  group_by(`Fish ID.par1`) %>% 
  summarise(n_mates = n_distinct(`Fish ID`)) %>% 
  count(n_mates)
```

Multiple matings clearly occur, however, the way we have our data set up it is hard to ask whether it is males that have multiple partners or females.
```{r cross_dam_sire_even}
(parents_paired_16_18_cross_dam_sire <- parents_paired_16_18_cross %>% 
  dplyr::select(Offspring, `Parent 1`, `Parent 2`, cross, SEX.par1, origin.par1, SEX, origin) %>% 
  mutate(Sire = case_when(SEX.par1 == "M" ~ `Parent 1`,
                          SEX == "M" ~ `Parent 2`)) %>% 
  mutate(Dam = case_when(SEX.par1 == "F" ~ `Parent 1`,
                         SEX == "F" ~ `Parent 2`)) %>% 
  dplyr::select(Offspring, Dam, Sire, cross) %>% 
  left_join(paired_16_18_filter, by = c("Dam" = "franz_id")) %>% 
  left_join(paired_16_18_filter, by = c("Sire" = "franz_id"), suffix = c("_dam", "_sire")))
```

There were 62 unique matings.
```{r unique_matings_even}
parents_paired_16_18_cross_dam_sire %>% 
  count(Dam, Sire)
```

Do our sample dates for cross type make sense?
```{r cross_sample_date_even}
parents_paired_16_18_cross_dam_sire %>% 
  mutate(`Sample Date_dam` = dmy(`Sample Date_dam`),
         `Sample Date_sire` = dmy(`Sample Date_sire`)) %>% 
  group_by(Dam, Sire, `Sample Date_dam`, `Sample Date_sire`, cross) %>% 
  summarise(n = n_distinct(Offspring)) %>% 
  ggplot(aes(x = `Sample Date_dam`, y = `Sample Date_sire`, colour = cross)) +
  geom_jitter(size = 2)
```

How many Dams had multiple Sires?
```{r dams_with_mult_sires_even}
# How many Dams had multiple Sires
parents_paired_16_18_cross_dam_sire %>% 
  group_by(Dam) %>% 
  summarise(n_sires = n_distinct(Sire)) %>% 
  count(n_sires)
```

How many Sires had multiple Dams?
```{r sires_with_mult_dams_even}
# How many Dams had multiple Sires
parents_paired_16_18_cross_dam_sire %>% 
  group_by(Sire) %>% 
  summarise(n_dams = n_distinct(Dam)) %>% 
  count(n_dams)
```

How many Dams had multiple Sires of differing origin? What was their reproductive success?
```{r rs_dams_mult_sire_even}
# How many Dams had multiple Sires of differing origin
# dams_with_mult_sire <- parents_paired_16_18_cross_dam_sire %>% 
#   group_by(Dam) %>% 
#   summarise(n_cross = n_distinct(cross)) %>% 
#   filter(n_cross > 1) %>% 
#   pull(Dam)
# 
# parents_paired_16_18_cross_dam_sire %>% 
#   filter(Dam %in% dams_with_mult_sire) %>% 
#   group_by(Dam, Sire, cross) %>% 
#   summarise(n = n_distinct(Offspring)) %>% 
#   spread(cross, n)
# 
# parents_paired_16_18_cross_dam_sire %>% 
#   filter(Dam %in% dams_with_mult_sire) %>% 
#   group_by(Dam, Sire, cross) %>% 
#   summarise(n = n_distinct(Offspring)) %>% 
#   ggplot(aes(x = Dam, y = n, colour = cross)) +
#   geom_jitter(size = 2, width = 0.2, height = 0.2) +
#   ylab("Number of offspring") +
#   ggtitle("Reproductive success of Dams with Sires\nof a different origin")
```

How many Sires had multiple Dams of differing origin? What was their reproductive success?
```{r rs_sires_mult_dam_even}
# How many Sires had multiple Dams of differing origin
# sires_with_mult_dam <- parents_paired_16_18_cross_dam_sire %>% 
#   group_by(Sire) %>% 
#   summarise(n_cross = n_distinct(cross)) %>% 
#   filter(n_cross > 1) %>% 
#   pull(Sire)
# 
# parents_paired_16_18_cross_dam_sire %>% 
#   filter(Sire %in% sires_with_mult_dam) %>% 
#   group_by(Dam, Sire, cross) %>% 
#   summarise(n = n_distinct(Offspring)) %>% 
#   spread(cross, n)
# 
# parents_paired_16_18_cross_dam_sire %>% 
#   filter(Sire %in% sires_with_mult_dam) %>% 
#   group_by(Dam, Sire, cross) %>% 
#   summarise(n = n_distinct(Offspring)) %>% 
#   ggplot(aes(x = Sire, y = n, colour = cross)) +
#   geom_jitter(size = 2, width = 0.2, height = 0.2) +
#   ylab("Number of offspring") +
#   ggtitle("Reproductive success of Sires with Dams\nof a different origin")
```

## Grandparentage

Find the 2016 fish that were offspring from 2014 and see how many offspring those had returning in 2018!
```{r}
parents_paired_14_16 <- read_csv("gilmour_parents_paired_14_16.csv") %>% 
  dplyr::select(-origin)

paired_16_18_filter_parents <- read_csv("gilmour_paired_16_18_filter_parents.csv") 

paired_16_18_filter_parents %>% 
  right_join(parents_paired_14_16, by = c("franz_id" = "Offspring")) %>% 
  count(origin.par, n)
```

???? of the hatchery fish from 2014 produced grandoffspring. In 2014, fish produced ???? offspring in 2016 and ???? grandoffspring in 2018. 

### RRS of fish with hatchery vs. natural parents

Origin of 2014 fish, their F1 offspring, and their F2 offspring.
```{r}
paired_16_18_filter_parents %>% 
  right_join(parents_paired_14_16, by = c("franz_id" = "Offspring")) %>% 
  group_by(origin.par) %>% 
  summarise(n_par = length(franz_id),
            n_off = sum(n, na.rm = TRUE))
```

```{r}
rrs_ci_kalinowski(n_h_off = 176, n_w_off = 197, n_h_par = 645, n_w_par = 604, alpha = 0.05)
```

### How many grandparents?
```{r}
(parents_paired_14_16 <- read_csv("gilmour_parents_paired_14_16.csv") %>% 
  dplyr::select(-origin))

(parents_paired_16_18 <- read_csv("gilmour_parents_paired_16_18.csv") %>% 
  dplyr::select(-origin))

parents_paired_16_18 %>% 
  inner_join(parents_paired_14_16, by = c("Parent_ID" = "Offspring"), suffix = c("_16", "_14"))
```

How many grand offspring (n) are the result of a given fish from 2014. How often that occurred is (nn).
```{r}
parents_paired_16_18 %>% 
  inner_join(parents_paired_14_16, by = c("Parent_ID" = "Offspring"), suffix = c("_16", "_14")) %>% 
  count(Parent_ID_14) %>% 
  count(n)
```

Get dam and sire for each offspring 2014/2016
```{r}
(
  sires_14 <- parents_paired_14_16 %>%
    filter(Sex.par == "Male") %>%
    rename(Sire = Parent_ID, Sire_origin = origin.par) %>%
    dplyr::select(Offspring, Sire, Sire_origin)
)

(
  dams_14 <- parents_paired_14_16 %>%
    filter(Sex.par == "Female") %>%
    rename(Dam = Parent_ID, Dam_origin = origin.par) %>%
    dplyr::select(Offspring, Dam, Dam_origin)
)

(parents_14_16 <- full_join(sires_14, dams_14, by = "Offspring") %>%
    arrange(Offspring))
```

Get dam and sire for each offspring 2016/2018
```{r}
(
  sires_16 <- parents_paired_16_18 %>%
    filter(Sex.par == "Male") %>%
    rename(Sire = Parent_ID, Sire_origin = origin.par) %>%
    dplyr::select(Offspring, Sire, Sire_origin)
)

(
  dams_16 <- parents_paired_16_18 %>%
    filter(Sex.par == "Female") %>%
    rename(Dam = Parent_ID, Dam_origin = origin.par) %>%
    dplyr::select(Offspring, Dam, Dam_origin)
)

(
  gilmour_grandparentage <- full_join(sires_16, dams_16, by = "Offspring") %>% 
    arrange(Offspring) %>% 
    left_join(parents_14_16, by = c("Sire" = "Offspring"), suffix = c("", "_paternal")) %>% 
    left_join(parents_14_16, by = c("Dam" = "Offspring"), suffix = c("", "_maternal")) %>%
    group_by(Offspring) %>% 
    mutate(n_par = sum(!is.na(Sire), !is.na(Dam)),
           n_g_par = sum(!is.na(Sire_paternal), !is.na(Sire_maternal), !is.na(Dam_paternal), !is.na(Dam_maternal))) %>% 
    ungroup() %>% 
    filter(n_g_par > 0)
)
```

```{r}
gilmour_grandparentage %>% 
  count(n_g_par)
```


### Plot pedigrees with `kinship2`

This failed because it won't plot for single parents...

```{r}
pedigree_df <- bind_rows(
  gilmour_grandparentage %>%
    mutate(Origin = "Natural") %>%
    dplyr::select(Offspring, Sire, Dam, Origin) %>%
    left_join(dplyr::select(parents_paired_16_18, Offspring, Sex.off), by = "Offspring") %>%
    dplyr::rename(Sex = Sex.off) %>%
    distinct(),
  gilmour_grandparentage %>%
    dplyr::select(Sire, Sire_paternal, Dam_paternal, Sire_origin) %>%
    drop_na(Sire) %>%
    dplyr::rename(
      Offspring = Sire,
      Sire = Sire_paternal,
      Dam = Dam_paternal,
      Origin = Sire_origin
    ) %>%
    mutate(Sex = "Male") %>% 
    distinct(),
  gilmour_grandparentage %>%
    dplyr::select(Dam, Sire_maternal, Dam_maternal, Dam_origin) %>%
    drop_na(Dam) %>%
    dplyr::rename(
      Offspring = Dam,
      Sire = Sire_maternal,
      Dam = Dam_maternal,
      Origin = Dam_origin
    ) %>%
    mutate(Sex = "Female") %>% 
    distinct(),
  gilmour_grandparentage %>% 
    dplyr::select(Sire_paternal, Sire_origin_paternal) %>% 
    drop_na(Sire_paternal) %>% 
    dplyr::rename(
      Offspring = Sire_paternal,
      Origin = Sire_origin_paternal
    ) %>% 
    mutate(Sex = "Male",
           Sire = NA,
           Dam = NA) %>% 
    distinct(),
  gilmour_grandparentage %>% 
    dplyr::select(Dam_paternal, Dam_origin_paternal) %>% 
    drop_na(Dam_paternal) %>% 
    dplyr::rename(
      Offspring = Dam_paternal,
      Origin = Dam_origin_paternal
    ) %>% 
    mutate(Sex = "Female",
           Sire = NA,
           Dam = NA) %>% 
    distinct(),
  gilmour_grandparentage %>% 
    dplyr::select(Sire_maternal, Sire_origin_maternal) %>% 
    drop_na(Sire_maternal) %>% 
    dplyr::rename(
      Offspring = Sire_maternal,
      Origin = Sire_origin_maternal
    ) %>% 
    mutate(Sex = "Male",
           Sire = NA,
           Dam = NA) %>% 
    distinct(),
  gilmour_grandparentage %>% 
    dplyr::select(Dam_maternal, Dam_origin_maternal) %>% 
    drop_na(Dam_maternal) %>% 
    dplyr::rename(
      Offspring = Dam_maternal,
      Origin = Dam_origin_maternal
    ) %>% 
    mutate(Sex = "Female",
           Sire = NA,
           Dam = NA) %>% 
    distinct()
    
)

pedigree_df %>% 
  replace_na(list(Sire = "0", Dam = "0")) %>% 
  with(.,
       fixParents(id = Offspring, dadid = Sire, momid = Dam, sex = Sex, missid = "0")) %>% 
  left_join(pedigree_df, by = c("id" = "Offspring")) %>% 
  dplyr::select(-Sire, -Dam, -Sex) %>% 
  with(.,
     pedigree(id, dadid, momid, sex, missid = "0", affected = as.factor(Origin))) %>% 
  plot
```

### Plot pedigrees with `visPedigree`

```{r}
(
  pedigree_vis <- pedigree_df %>%
    dplyr::rename(ID = Offspring) %>%
    dplyr::select(ID, Sire, Dam, Sex)
)
```

```{r}
(pedigree_tdy <- tidyped(pedigree_vis))
```

```{r}
visped(pedigree_tdy)
```

## Summary of even lineage results
1. Almost all exclusion probabilities were equal to 1, which means we can be confident in the ability of our marker set to correctly assign parents to offspring.  
2. ???? offspring were assigned to ???? parents, for an assignment rate of **????%**.  
3. ???? offspring assigned to 2 parents (triads).  
4. RRS (with 0's) was significant for females (0.??), but not for males (0.??).  
5. Family size varied from 0-?? for females and 0-?? for males.  
