---
title: "Heritability of body length, sample date, location, and RS according to Cross Type and Parental/Offspring Sex for Gilmour 2014-2016"
author: "Kristen Gruenthal"
date: "30 July 2021"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

# **INTRODUCTION**

The purpose of these analyses are to more fully explore the parent-pair offspring trio/triad data (i.e. cross data) for Gilmour Creek 2014-2016 in terms of spatiotemporal patterning and generate preliminary estimates of narrow-sense heritability using regression slope analysis. **NOTE: The bulk of 2016 origin data for Gilmour is missing as a result of the otolith shipping issue.**

```{r setup, include=FALSE}
rm(list=ls())

library(coin)
library(MASS)
library(scales)
library(tidyverse)
library(leaflet)
library(lubridate)
library(ggExtra)
library(gridExtra)
library(MuMIn)
library(GGally)
library(gganimate)
library(broom)
library(tidyr)
library(purrr)
library(MCMCglmm)
library(lme4)
library(lmerTest)

knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.width = 10)
```

# **WHO**

Who are the two-parents, both dam and sire, for each cross type offspring and where were they sampled? Note: River kilometer not determined yet for Gilmour.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# gilmour_riverdist_14_16 <- read_csv("../../GIS/R/Gilmour/gilmour_distances.csv") %>% 
#   mutate(Intertidal = case_when(dist2tide > 0 ~ "Upstream",
#                                 dist2tide <= 0 ~ "Intertidal")) %>% 
#   rename(Distance = mouthdist, Segment = seg)
# 
# gilmour_parents_paired_14_16_cross <- read_csv("../Gilmour/gilmour_parents_paired_14_16_cross.csv")
# 
# # organize by sire and dam, join riverdist data
# parents_paired_cross_riverdist_spacetime <- function(indata_spacetime = gilmour_parents_paired_14_16_cross, indata_riverdist = gilmour_riverdist_14_16) {
#   indata_spacetime %>% 
#   count(cross, `Fish ID`, `Fish ID.par1`, `Sample Date`, `Sample Date.par1`, SEX, SEX.par1, `DNA Tray Code`, `DNA Tray Code.par1`, `DNA Tray Well Code`, `DNA Tray Well Code.par1`, origin.par1, origin, `Length Mm.par1`, `Length Mm`) %>% 
#   rename(`Fish ID.par2` = `Fish ID`, 
#          `Sample Date.par2` = `Sample Date`,
#          SEX.par2 = SEX,
#          `DNA Tray Code.par2` = `DNA Tray Code`,
#          `DNA Tray Well Code.par2` = `DNA Tray Well Code`,
#          origin.par2 = origin,
#          `Length Mm.par2` = `Length Mm`) %>% 
#   unite(Sample.par1, c("DNA Tray Code.par1", "DNA Tray Well Code.par1"), sep = "_", remove = TRUE) %>% 
#   unite(Sample.par2, c("DNA Tray Code.par2", "DNA Tray Well Code.par2"), sep = "_", remove = TRUE) %>% 
#   left_join(indata_riverdist, by = c("Sample.par1" = "Sample")) %>% 
#   left_join(indata_riverdist, by = c("Sample.par2" = "Sample"), suffix = c(".par1", ".par2")) %>% 
#   mutate(fish_id.sire = case_when(SEX.par1 == "M" ~ `Fish ID.par1`,
#                                   TRUE ~ `Fish ID.par2`)) %>% 
#   mutate(fish_id.dam = case_when(SEX.par1 == "F" ~ `Fish ID.par1`,
#                                   TRUE ~ `Fish ID.par2`)) %>% 
#   mutate(sample_date.sire = case_when(SEX.par1 == "M" ~ `Sample Date.par1`,
#                                   TRUE ~ `Sample Date.par2`)) %>% 
#   mutate(sample_date.dam = case_when(SEX.par1 == "F" ~ `Sample Date.par1`,
#                                   TRUE ~ `Sample Date.par2`)) %>% 
#   mutate(sex.sire = case_when(SEX.par1 == "M" ~ SEX.par1,
#                                   TRUE ~ SEX.par2)) %>% 
#   mutate(sex.dam = case_when(SEX.par1 == "F" ~ SEX.par1,
#                                   TRUE ~ SEX.par2)) %>%
#   mutate(origin.sire = case_when(SEX.par1 == "M" ~ origin.par1,
#                                   TRUE ~ origin.par2)) %>% 
#   mutate(origin.dam = case_when(SEX.par1 == "F" ~ origin.par1,
#                                   TRUE ~ origin.par2)) %>%
#   mutate(distance.sire = case_when(SEX.par1 == "M" ~ Distance.par1,
#                                   TRUE ~ Distance.par2)) %>% 
#   mutate(distance.dam = case_when(SEX.par1 == "F" ~ Distance.par1,
#                                   TRUE ~ Distance.par2)) %>%
#   mutate(intertidal.sire = case_when(SEX.par1 == "M" ~ Intertidal.par1,
#                                   TRUE ~ Intertidal.par2)) %>% 
#   mutate(intertidal.dam = case_when(SEX.par1 == "F" ~ Intertidal.par1,
#                                   TRUE ~ Intertidal.par2)) %>%
#   mutate(length.sire = case_when(SEX.par1 == "M" ~ `Length Mm.par1`,
#                                   TRUE ~ `Length Mm.par2`)) %>% 
#   mutate(length.dam = case_when(SEX.par1 == "F" ~ `Length Mm.par1`,
#                                   TRUE ~ `Length Mm.par2`)) %>%
#   select(cross, n, ends_with(".sire"), ends_with(".dam"))
# }
# 
# (gilmour_parents_paired_14_16_cross_riverdist_spacetime <- parents_paired_cross_riverdist_spacetime())
```

## Offspring, dam, and sire relationships - dyads and triads

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# dyad relationships - one row per offspring
gilmour_parents_paired_14_16 <- read_csv("../Gilmour/gilmour_parents_paired_14_16.csv") %>% 
     mutate(origin.par = factor(x = origin.par, levels = c("Natural", "Hatchery")))

# Read in paired data
gilmour_postQA_OceanAK_paired_14_16 <- read_csv("../Franz/Gilmour_14_16_postQA_OceanAK_paired_2014_2016_GILMOUR.csv")

OceanAK_paired <- function(indata_OceanAK_paired = gilmour_postQA_OceanAK_paired_14_16) {
   indata_OceanAK_paired %>% 
   select(franz_id, starts_with("Sample"), SILLY, `Fish ID`, starts_with("DNA"), SEX, `Length Mm`, `Otolith Mark Present`, `Otolith Mark ID`) %>% 
   rename(sample_id = "Sample ID", 
          Year = "Sample Year", 
          silly = "SILLY", 
          fish_id = "Fish ID", 
          dna_tray = "DNA Tray Code", 
          dna_well = "DNA Tray Well Code",
          Sex = "SEX",
          Length = "Length Mm") %>% 
   mutate(Origin = case_when(`Otolith Mark Present` == "YES" ~ "Hatchery",
                             `Otolith Mark Present` == "NO" ~ "Natural")) %>% 
   mutate(Origin = factor(x = Origin, levels = c("Natural", "Hatchery")))
}         

gilmour_OceanAK_paired_14_16 <- OceanAK_paired()

# Get offspring, dam, sire relationships and join
ods_relationships <- function(indata_ods = gilmour_parents_paired_14_16) {
  offspring_sire <- indata_ods %>% 
    filter(SEX.par == "M") %>% 
    select(Offspring, Parent_ID)

  offspring_dam <- indata_ods %>% 
    filter(SEX.par == "F") %>% 
    select(Offspring, Parent_ID)

  offspring_dam_sire <- offspring_dam %>% 
    full_join(offspring_sire, by = "Offspring", suffix = c(".dam", ".sire"))
}

gilmour_offspring_dam_sire_14_16 <- ods_relationships()

# Join back paired data to offspring, dam, and sire
(gilmour_parents_paired_14_16_dam_sire <- gilmour_offspring_dam_sire_14_16 %>% 
  left_join(gilmour_OceanAK_paired_14_16, by = c("Offspring" = "franz_id")) %>% 
  left_join(gilmour_OceanAK_paired_14_16, by = c("Parent_ID.dam" = "franz_id"), suffix = c(".off", "")) %>% 
  left_join(gilmour_OceanAK_paired_14_16, by = c("Parent_ID.sire" = "franz_id"), suffix = c(".dam", ".sire")) %>% 
  mutate(date.off = dmy(`Sample Date.off`),
         date.dam = dmy(`Sample Date.dam`),
         date.sire = dmy(`Sample Date.sire`)) %>% 
  mutate(DOY.off = yday(date.off),
         DOY.dam = yday(date.dam),
         DOY.sire = yday(date.sire)))
```

## Triads - used for remaining analyses

```{r, echo=FALSE, warning=FALSE, message=FALSE}
triads <- function(indata_triads = gilmour_parents_paired_14_16_dam_sire) {
  indata_triads %>% 
  filter(!is.na(Parent_ID.dam) & !is.na(Parent_ID.sire)) 
}

(gilmour_triads_14_16 <- triads())
```

# **HERITABILITY**

For traits associated with fitness in natural populations, heritability is typically 0.1–0.2 (Wray and Visscher 2008 [https://www.nature.com/scitable/topicpage/estimating-trait-heritability-46889/] per Visscher et al. 2008). 

## Body Length

```{r}
# mid-parent trait values and cross info
triads_length <- function(indata_triads_length = gilmour_triads_14_16) {
  indata_triads_length %>%
  mutate(Length.par = (Length.dam + Length.sire) / 2) %>% 
  mutate(Cross = case_when(Origin.dam == "Natural" & Origin.sire == "Natural" ~ "NN",
                           Origin.dam == "Natural" & Origin.sire == "Hatchery" ~ "NH",
                           Origin.dam == "Hatchery" & Origin.sire == "Natural" ~ "HN",
                           Origin.dam == "Hatchery" & Origin.sire == "Hatchery" ~ "HH"))
}

gilmour_triads_length_14_16 <- triads_length()
```

### Scatterplots

```{r}
parents_paired_length_graph <- function(indata_parents_paired_length_graph = gilmour_parents_paired_14_16, title = "Gilmour_14_16\n\nRegression of Body Length by Sex") {
  indata_parents_paired_length_graph %>% 
  rename(length.par = "Length Mm.par", 
         length.off = "Length Mm.off") %>% 
  mutate(sex.par = case_when(SEX.par == "F" ~ "Parent: Female",
                             SEX.par == "M" ~ "Parent: Male")) %>% 
  mutate(sex.off = case_when(SEX.off == "F" ~ "Offspring: Female",
                             SEX.off == "M" ~ "Offspring: Male")) %>% 
  group_by(sex.par, sex.off) %>%
  ggplot(aes(x = length.par, y = length.off, colour = origin.par)) +
  geom_abline(slope = 1) +
  geom_jitter(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_grid(sex.off ~ sex.par) +
  labs(title = title,  colour = "Parent: Origin") +
  xlab("Parent Length (mm)") +
  ylab("Offspring Length (mm)") +
  theme_bw() +
  theme(text = element_text(size = 20))
}

parents_paired_length_graph()

h2_length_graph <- function(indata_h2_length_graph = gilmour_triads_length_14_16, title = "Gilmour_14_16\n\nRegression of Body Length by Cross") {
  indata_h2_length_graph %>% 
  ggplot(aes(x = Length.par, y = Length.off, colour = Cross)) +
  geom_abline(slope = 1) +
  geom_jitter(size = 2) +
  geom_smooth(method = "lm") +
  labs(title = title, colour = "Parent: Origin") +
  xlab("Mid-parent Length (mm)") +
  ylab("Offspring Length (mm)") +
  theme_bw() +
  theme(text = element_text(size = 20))
}

h2_length_graph()
```

### Correlation

We estimate narrow-sense heritability here, as well as for location and sample date, from r = √(½)h^2.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
parents_paired_length <- function(indata_triads_length = gilmour_triads_length_14_16) {
  indata_triads_length %>% 
  drop_na(Length.par, Length.off) %>% 
  mutate(sex.dam = case_when(Sex.dam == "F" ~ "Female")) %>%
  mutate(sex.sire = case_when(Sex.sire == "M" ~ "Male")) %>%
  mutate(sex.off = case_when(Sex.off == "F" ~ "Female",
                             Sex.off == "M" ~ "Male")) 
}

(gilmour_parents_paired_14_16_length1 <- parents_paired_length() %>% 
  group_by(Cross) %>%
  summarise(correlation = round(cor(Length.off, Length.par), 3), r2 = round(cor(Length.off, Length.par)^2, 3)))
(gilmour_parents_paired_14_16_length2 <- parents_paired_length() %>% 
  group_by(sex.dam, sex.off) %>%
  summarise(correlation = round(cor(Length.off, Length.par), 3), r2 = round(cor(Length.off, Length.par)^2, 3)))
(gilmour_parents_paired_14_16_length3 <- parents_paired_length() %>% 
  group_by(sex.sire, sex.off) %>%
  summarise(correlation = round(cor(Length.off, Length.par), 3), r2 = round(cor(Length.off, Length.par)^2, 3)))

gilmour_parents_paired_14_16_length1_NN <- as.data.frame(c(gilmour_parents_paired_14_16_length1[4,2], gilmour_parents_paired_14_16_length1[4,3]*2))
gilmour_parents_paired_14_16_length1_NH <- as.data.frame(c(gilmour_parents_paired_14_16_length1[3,2], gilmour_parents_paired_14_16_length1[3,3]*2))
gilmour_parents_paired_14_16_length1_HN <- as.data.frame(c(gilmour_parents_paired_14_16_length1[2,2], gilmour_parents_paired_14_16_length1[2,3]*2))
gilmour_parents_paired_14_16_length1_HH <- as.data.frame(c(gilmour_parents_paired_14_16_length1[1,2], gilmour_parents_paired_14_16_length1[1,3]*2))
gilmour_parents_paired_14_16_length2_dF <- as.data.frame(c(gilmour_parents_paired_14_16_length2[1,3], gilmour_parents_paired_14_16_length2[1,4]*2))
gilmour_parents_paired_14_16_length2_dM <- as.data.frame(c(gilmour_parents_paired_14_16_length2[2,3], gilmour_parents_paired_14_16_length2[2,4]*2))
gilmour_parents_paired_14_16_length3_sF <- as.data.frame(c(gilmour_parents_paired_14_16_length3[1,3], gilmour_parents_paired_14_16_length3[1,4]*2))
gilmour_parents_paired_14_16_length3_sM <- as.data.frame(c(gilmour_parents_paired_14_16_length3[2,3], gilmour_parents_paired_14_16_length3[2,4]*2))

gilmour_r_and_h2_length_14_16 <- rbind(gilmour_parents_paired_14_16_length1_NN, gilmour_parents_paired_14_16_length1_NH, gilmour_parents_paired_14_16_length1_HN, gilmour_parents_paired_14_16_length1_HH, gilmour_parents_paired_14_16_length2_dF, gilmour_parents_paired_14_16_length2_dM, gilmour_parents_paired_14_16_length3_sF, gilmour_parents_paired_14_16_length3_sM) %>% round(3)
colnames(gilmour_r_and_h2_length_14_16) <- c("Correlation (r)", "Heritability (h^2)")
gilmour_r_and_h2_length_14_16$Category <- c("NN", "NH", "HN", "HH", "Dam/Female offspring", "Dam/Male offspring", "Sire/Female offspring", "Sire/Male offspring")
(gilmour_r_and_h2_length_14_16 <- gilmour_r_and_h2_length_14_16[,c(3,1,2)])
```

### Slope

If we use the parents' phenotypic variance as a measure of VP, we can estimate h^2 by solving for:
trait_value_offspring = Cov_parent_offspring / Var_parent x trait_value_parent + c. The slope of the line (number under "estimate" for "Length.*" in the tibbles below) is mathematically equivalent to h^2.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# by cross
gilmour_h2_length_14_16_all_tbl <- tidy(lm(Length.off ~ Length.par, data = gilmour_triads_length_14_16))
gilmour_h2_length_14_16_NN_tbl <- tidy(lm(Length.off ~ Length.par, data = gilmour_triads_length_14_16 %>% filter(Cross == "NN")))
gilmour_h2_length_14_16_NH_tbl <- tidy(lm(Length.off ~ Length.par, data = gilmour_triads_length_14_16 %>% filter(Cross == "NH")))
gilmour_h2_length_14_16_HN_tbl <- tidy(lm(Length.off ~ Length.par, data = gilmour_triads_length_14_16 %>% filter(Cross == "HN")))
gilmour_h2_length_14_16_HH_tbl <- tidy(lm(Length.off ~ Length.par, data = gilmour_triads_length_14_16 %>% filter(Cross == "HH")))

# by sex
gilmour_h2_length_14_16_dam_tbl <- tidy(lm(Length.off ~ Length.dam, data = gilmour_triads_length_14_16))
gilmour_h2_length_14_16_sire_tbl <- tidy(lm(Length.off ~ Length.sire, data = gilmour_triads_length_14_16))
gilmour_h2_length_14_16_dF_tbl <- tidy(lm(Length.off ~ Length.dam, data = gilmour_triads_length_14_16 %>% filter(Sex.off == "F")))
gilmour_h2_length_14_16_dM_tbl <- tidy(lm(Length.off ~ Length.dam, data = gilmour_triads_length_14_16 %>% filter(Sex.off == "M")))
gilmour_h2_length_14_16_sF_tbl <- tidy(lm(Length.off ~ Length.sire, data = gilmour_triads_length_14_16 %>% filter(Sex.off == "F")))
gilmour_h2_length_14_16_sM_tbl <- tidy(lm(Length.off ~ Length.sire, data = gilmour_triads_length_14_16 %>% filter(Sex.off == "M")))

gilmour_slope_as_h2_length_14_16 <- as.data.frame(rbind(gilmour_h2_length_14_16_all_tbl[2,2], gilmour_h2_length_14_16_NN_tbl[2,2], gilmour_h2_length_14_16_NH_tbl[2,2], gilmour_h2_length_14_16_HN_tbl[2,2], gilmour_h2_length_14_16_HH_tbl[2,2], gilmour_h2_length_14_16_dam_tbl[2,2], gilmour_h2_length_14_16_sire_tbl[2,2], gilmour_h2_length_14_16_dF_tbl[2,2], gilmour_h2_length_14_16_dM_tbl[2,2], gilmour_h2_length_14_16_sF_tbl[2,2], gilmour_h2_length_14_16_sM_tbl[2,2]) %>% round(3))
colnames(gilmour_slope_as_h2_length_14_16) <- c("Heritability (h^2)")
gilmour_slope_as_h2_length_14_16$Category <- c("Overall", "NN", "NH", "HN", "HH", "Dam/All offspring", "Sire/All offspring", "Dam/Female offspring", "Dam/Male offspring", "Sire/Female offspring", "Sire/Male offspring")
(gilmour_slope_as_h2_length_14_16 <- gilmour_slope_as_h2_length_14_16[,c(2,1)])
```

### Animal model

Narrow-sense heritabilities (h^2) of were estimated from the proportion of the total phenotypic variance explained by the additive genetic variance component (h^2 = VA/VP) per Evans et al. (2019) [https://link.springer.com/article/10.1007/s10592-019-01174-4]. Maternal (M) versus paternal (S) effects were estimated as the proportion of total phenotypic variance explained by each parental variance component (M = VM/VP; S = VS/VP).

```{r}
gilmour_h2_length_14_16_dam_sire <- lmer(Length.off ~ (1|Length.sire)+(1|Length.dam), data = gilmour_triads_length_14_16, na.action = na.omit)
(summary(gilmour_h2_length_14_16_dam_sire))
plot(gilmour_h2_length_14_16_dam_sire)
hist(resid(gilmour_h2_length_14_16_dam_sire))

gilmour_h2_length_14_16_NN <- gilmour_triads_length_14_16 %>% filter(Cross == "NN")
gilmour_h2_length_14_16_dam_sire_NN <- lmer(Length.off ~ (1|Length.sire)+(1|Length.dam), data = gilmour_h2_length_14_16_NN, na.action = na.omit)
(summary(gilmour_h2_length_14_16_dam_sire_NN))

gilmour_h2_length_14_16_NH <- gilmour_triads_length_14_16 %>% filter(Cross == "NH")
gilmour_h2_length_14_16_dam_sire_NH <- lmer(Length.off ~ (1|Length.sire)+(1|Length.dam), data = gilmour_h2_length_14_16_NH, na.action = na.omit)
(summary(gilmour_h2_length_14_16_dam_sire_NH))

gilmour_h2_length_14_16_HN <- gilmour_triads_length_14_16 %>% filter(Cross == "HN")
gilmour_h2_length_14_16_dam_sire_HN <- lmer(Length.off ~ (1|Length.sire)+(1|Length.dam), data = gilmour_h2_length_14_16_HN, na.action = na.omit)
(summary(gilmour_h2_length_14_16_dam_sire_HN))

gilmour_h2_length_14_16_HH <- gilmour_triads_length_14_16 %>% filter(Cross == "HH")
gilmour_h2_length_14_16_dam_sire_HH <- lmer(Length.off ~ (1|Length.sire)+(1|Length.dam), data = gilmour_h2_length_14_16_HH, na.action = na.omit)
(summary(gilmour_h2_length_14_16_dam_sire_HH))

gilmour_h2_length_14_16_F <- gilmour_triads_length_14_16 %>% filter(Sex.off == "F")
gilmour_h2_length_14_16_dam_sire_F <- lmer(Length.off ~ (1|Length.sire)+(1|Length.dam), data = gilmour_h2_length_14_16_F, na.action = na.omit)
(summary(gilmour_h2_length_14_16_dam_sire_F))

gilmour_h2_length_14_16_M <- gilmour_triads_length_14_16 %>% filter(Sex.off == "M")
gilmour_h2_length_14_16_dam_sire_M <- lmer(Length.off ~ (1|Length.sire)+(1|Length.dam), data = gilmour_h2_length_14_16_M, na.action = na.omit)
(summary(gilmour_h2_length_14_16_dam_sire_M))

gilmour_h2_length_14_16_all <- 17.64 / (17.64 + 0.00 + 474.84)
gilmour_h2_length_14_16_NN <- (47.25 + 70.31) / (47.25 + 70.31 + 441.91)
gilmour_h2_length_14_16_NH <- 0.00 / (0.00 + 0.00 + 326.5)
gilmour_h2_length_14_16_HN <- (196.536 + 9.784) / (196.536 + 9.784 + 539.361)
gilmour_h2_length_14_16_HH <- 37.41 / (0.00 + 37.41 + 406.01)
gilmour_h2_length_14_16_dam_F <- 39.85 / (80.32 + 39.85 + 280.48)
gilmour_h2_length_14_16_sire_F <- 80.32 / (80.32 + 39.85 + 280.48)
gilmour_h2_length_14_16_dam_M <- 0.00 / (0.00 + 0.00 + 592.7)
gilmour_h2_length_14_16_sire_M <- 0.00 / (0.00 + 0.00 + 592.7)

gilmour_am_for_h2_length_14_16 <- as.data.frame(c(gilmour_h2_length_14_16_all, gilmour_h2_length_14_16_NN, gilmour_h2_length_14_16_NH, gilmour_h2_length_14_16_HN, gilmour_h2_length_14_16_HH, gilmour_h2_length_14_16_dam_F, gilmour_h2_length_14_16_dam_M, gilmour_h2_length_14_16_sire_F, gilmour_h2_length_14_16_sire_M)) %>% round(3)
colnames(gilmour_am_for_h2_length_14_16) <- c("Heritability (h^2)")
gilmour_am_for_h2_length_14_16$Category <- c("Overall", "NN", "NH", "HN", "HH", "Dam/Female offspring", "Dam/Male offspring", "Sire/Female offspring", "Sire/Male offspring")
(gilmour_am_for_h2_length_14_16 <- gilmour_am_for_h2_length_14_16[,c(2,1)])
```

### *Results*

```{r}
gilmour_r_and_h2_length_14_16 <- gilmour_r_and_h2_length_14_16[,c(1,3)]

gilmour_h2_length_14_16 <- merge(gilmour_r_and_h2_length_14_16, gilmour_slope_as_h2_length_14_16,  by="Category") %>% 
    merge(gilmour_am_for_h2_length_14_16, by="Category") %>% 
    rename(gilmour_14_16 = Category, "h^2.correlation" = "Heritability (h^2).x", "h^2.slope" = "Heritability (h^2).y", "h^2.variance" = "Heritability (h^2)")
(gilmour_h2_length_14_16 <- gilmour_h2_length_14_16[c(6,5,4,3,1,2,7,8),])
```

There is some evidence for heritability of body size for certain crosses, although cross data is confounded by missing origin info in 2018.


## Sample Date

```{r}
# mid-parent trait values and cross info
triads_date <- function(indata_triads_date = gilmour_triads_14_16) {
  indata_triads_date %>%
  mutate(DOY.par = (DOY.dam + DOY.sire) / 2) %>% 
  mutate(Cross = case_when(Origin.dam == "Natural" & Origin.sire == "Natural" ~ "NN",
                           Origin.dam == "Natural" & Origin.sire == "Hatchery" ~ "NH",
                           Origin.dam == "Hatchery" & Origin.sire == "Natural" ~ "HN",
                           Origin.dam == "Hatchery" & Origin.sire == "Hatchery" ~ "HH"))
}

gilmour_triads_date_14_16 <- triads_date()
```

### Scatterplots

```{r}
parents_paired_date_graph <- function(indata_parents_paired_date_graph = gilmour_parents_paired_14_16, title = "Gilmour_14_16\n\nRegression of Sample Date by Sex") {
  indata_parents_paired_date_graph %>% 
  rename(DOY.par = "DOY.par", 
         DOY.off = "DOY.off") %>% 
  mutate(sex.par = case_when(SEX.par == "F" ~ "Parent: Female",
                             SEX.par == "M" ~ "Parent: Male")) %>% 
  mutate(sex.off = case_when(SEX.off == "F" ~ "Offspring: Female",
                             SEX.off == "M" ~ "Offspring: Male")) %>% 
  group_by(sex.par, sex.off) %>%
  ggplot(aes(x = DOY.par, y = DOY.off, colour = origin.par)) +
  geom_abline(slope = 1) +
  geom_jitter(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_grid(sex.off ~ sex.par) +
  labs(title = title,  colour = "Parent: Origin") +
  xlab("Parent DOY (mm)") +
  ylab("Offspring DOY (mm)") +
  theme_bw() +
  theme(text = element_text(size = 20))
}

parents_paired_date_graph()

h2_date_graph <- function(indata_h2_date_graph = gilmour_triads_date_14_16, title = "Gilmour_14_16\n\nRegression of Sample Date by Cross") {
  indata_h2_date_graph %>% 
  ggplot(aes(x = DOY.par, y = DOY.off, colour = Cross)) +
  geom_abline(slope = 1) +
  geom_jitter(size = 2) +
  geom_smooth(method = "lm") +
  labs(title = title, colour = "Parent: Origin") +
  xlab("Mid-parent DOY (mm)") +
  ylab("Offspring DOY (mm)") +
  theme_bw() +
  theme(text = element_text(size = 20))
}

h2_date_graph()
```

### Correlation

We estimate narrow-sense heritability here, as well as for location and sample date, from r = √(½)h^2.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
parents_paired_date <- function(indata_triads_date = gilmour_triads_date_14_16) {
  indata_triads_date %>% 
  drop_na(DOY.par, DOY.off) %>% 
  mutate(sex.dam = case_when(Sex.dam == "F" ~ "Female")) %>%
  mutate(sex.sire = case_when(Sex.sire == "M" ~ "Male")) %>%
  mutate(sex.off = case_when(Sex.off == "F" ~ "Female",
                             Sex.off == "M" ~ "Male")) 
}

(gilmour_parents_paired_14_16_date1 <- parents_paired_date() %>% 
  group_by(Cross) %>%
  summarise(correlation = round(cor(DOY.off, DOY.par), 3), r2 = round(cor(DOY.off, DOY.par)^2, 3)))
(gilmour_parents_paired_14_16_date2 <- parents_paired_date() %>% 
  group_by(sex.dam, sex.off) %>%
  summarise(correlation = round(cor(DOY.off, DOY.par), 3), r2 = round(cor(DOY.off, DOY.par)^2, 3)))
(gilmour_parents_paired_14_16_date3 <- parents_paired_date() %>% 
  group_by(sex.sire, sex.off) %>%
  summarise(correlation = round(cor(DOY.off, DOY.par), 3), r2 = round(cor(DOY.off, DOY.par)^2, 3)))

gilmour_parents_paired_14_16_date1_NN <- as.data.frame(c(gilmour_parents_paired_14_16_date1[4,2], gilmour_parents_paired_14_16_date1[4,3]*2))
gilmour_parents_paired_14_16_date1_NH <- as.data.frame(c(gilmour_parents_paired_14_16_date1[3,2], gilmour_parents_paired_14_16_date1[3,3]*2))
gilmour_parents_paired_14_16_date1_HN <- as.data.frame(c(gilmour_parents_paired_14_16_date1[2,2], gilmour_parents_paired_14_16_date1[2,3]*2))
gilmour_parents_paired_14_16_date1_HH <- as.data.frame(c(gilmour_parents_paired_14_16_date1[1,2], gilmour_parents_paired_14_16_date1[1,3]*2))
gilmour_parents_paired_14_16_date2_dF <- as.data.frame(c(gilmour_parents_paired_14_16_date2[1,3], gilmour_parents_paired_14_16_date2[1,4]*2))
gilmour_parents_paired_14_16_date2_dM <- as.data.frame(c(gilmour_parents_paired_14_16_date2[2,3], gilmour_parents_paired_14_16_date2[2,4]*2))
gilmour_parents_paired_14_16_date3_sF <- as.data.frame(c(gilmour_parents_paired_14_16_date3[1,3], gilmour_parents_paired_14_16_date3[1,4]*2))
gilmour_parents_paired_14_16_date3_sM <- as.data.frame(c(gilmour_parents_paired_14_16_date3[2,3], gilmour_parents_paired_14_16_date3[2,4]*2))

gilmour_r_and_h2_date_14_16 <- rbind(gilmour_parents_paired_14_16_date1_NN, gilmour_parents_paired_14_16_date1_NH, gilmour_parents_paired_14_16_date1_HN, gilmour_parents_paired_14_16_date1_HH, gilmour_parents_paired_14_16_date2_dF, gilmour_parents_paired_14_16_date2_dM, gilmour_parents_paired_14_16_date3_sF, gilmour_parents_paired_14_16_date3_sM) %>% round(3)
colnames(gilmour_r_and_h2_date_14_16) <- c("Correlation (r)", "Heritability (h^2)")
gilmour_r_and_h2_date_14_16$Category <- c("NN", "NH", "HN", "HH", "Dam/Female offspring", "Dam/Male offspring", "Sire/Female offspring", "Sire/Male offspring")
(gilmour_r_and_h2_date_14_16 <- gilmour_r_and_h2_date_14_16[,c(3,1,2)])
```

### Slope

If we use the parents' phenotypic variance as a measure of VP, we can estimate h^2 by solving for:
trait_value_offspring = Cov_parent_offspring / Var_parent x trait_value_parent + c. The slope of the line is mathematically equivalent to h^2.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# by cross
gilmour_h2_date_14_16_all_tbl <- tidy(lm(DOY.off ~ DOY.par, data = gilmour_triads_date_14_16))
gilmour_h2_date_14_16_NN_tbl <- tidy(lm(DOY.off ~ DOY.par, data = gilmour_triads_date_14_16 %>% filter(Cross == "NN")))
gilmour_h2_date_14_16_NH_tbl <- tidy(lm(DOY.off ~ DOY.par, data = gilmour_triads_date_14_16 %>% filter(Cross == "NH")))
gilmour_h2_date_14_16_HN_tbl <- tidy(lm(DOY.off ~ DOY.par, data = gilmour_triads_date_14_16 %>% filter(Cross == "HN")))
gilmour_h2_date_14_16_HH_tbl <- tidy(lm(DOY.off ~ DOY.par, data = gilmour_triads_date_14_16 %>% filter(Cross == "HH")))

# by sex
gilmour_h2_date_14_16_dam_tbl <- tidy(lm(DOY.off ~ DOY.dam, data = gilmour_triads_date_14_16))
gilmour_h2_date_14_16_sire_tbl <- tidy(lm(DOY.off ~ DOY.sire, data = gilmour_triads_date_14_16))
gilmour_h2_date_14_16_dF_tbl <- tidy(lm(DOY.off ~ DOY.dam, data = gilmour_triads_date_14_16 %>% filter(Sex.off == "F")))
gilmour_h2_date_14_16_dM_tbl <- tidy(lm(DOY.off ~ DOY.dam, data = gilmour_triads_date_14_16 %>% filter(Sex.off == "M")))
gilmour_h2_date_14_16_sF_tbl <- tidy(lm(DOY.off ~ DOY.sire, data = gilmour_triads_date_14_16 %>% filter(Sex.off == "F")))
gilmour_h2_date_14_16_sM_tbl <- tidy(lm(DOY.off ~ DOY.sire, data = gilmour_triads_date_14_16 %>% filter(Sex.off == "M")))

gilmour_slope_as_h2_date_14_16 <- rbind(gilmour_h2_date_14_16_all_tbl[2,2], gilmour_h2_date_14_16_NN_tbl[2,2], gilmour_h2_date_14_16_NH_tbl[2,2], gilmour_h2_date_14_16_HN_tbl[2,2], gilmour_h2_date_14_16_HH_tbl[2,2], gilmour_h2_date_14_16_dam_tbl[2,2], gilmour_h2_date_14_16_sire_tbl[2,2], gilmour_h2_date_14_16_dF_tbl[2,2], gilmour_h2_date_14_16_dM_tbl[2,2], gilmour_h2_date_14_16_sF_tbl[2,2], gilmour_h2_date_14_16_sM_tbl[2,2]) %>% round(3)
colnames(gilmour_slope_as_h2_date_14_16) <- c("Heritability (h^2)")
gilmour_slope_as_h2_date_14_16$Category <- c("Overall", "NN", "NH", "HN", "HH", "Dam/All offspring", "Sire/All offspring", "Dam/Female offspring", "Dam/Male offspring", "Sire/Female offspring", "Sire/Male offspring")
(gilmour_slope_as_h2_date_14_16 <- gilmour_slope_as_h2_date_14_16[,c(2,1)])
```

### Animal model

Narrow-sense heritabilities (h^2) of were estimated from the proportion of the total phenotypic variance explained by the additive genetic variance component (h^2 = VA/VP) per Evans et al. (2019) [https://link.springer.com/article/10.1007/s10592-019-01174-4]. Maternal (M) versus paternal (S) effects were estimated as the proportion of total phenotypic variance explained by each parental variance component (M = VM/VP; S = VS/VP).

```{r}
gilmour_h2_date_14_16_dam_sire <- lmer(DOY.off ~ (1|DOY.sire)+(1|DOY.dam), data = gilmour_triads_date_14_16, na.action = na.omit)
(summary(gilmour_h2_date_14_16_dam_sire))
plot(gilmour_h2_date_14_16_dam_sire)
hist(resid(gilmour_h2_date_14_16_dam_sire))

gilmour_h2_date_14_16_NN <- gilmour_triads_date_14_16 %>% filter(Cross == "NN")
gilmour_h2_date_14_16_dam_sire_NN <- lmer(DOY.off ~ (1|DOY.sire)+(1|DOY.dam), data = gilmour_h2_date_14_16_NN, na.action = na.omit)
(summary(gilmour_h2_date_14_16_dam_sire_NN))

gilmour_h2_date_14_16_NH <- gilmour_triads_date_14_16 %>% filter(Cross == "NH")
gilmour_h2_date_14_16_dam_sire_NH <- lmer(DOY.off ~ (1|DOY.sire)+(1|DOY.dam), data = gilmour_h2_date_14_16_NH, na.action = na.omit)
(summary(gilmour_h2_date_14_16_dam_sire_NH))

gilmour_h2_date_14_16_HN <- gilmour_triads_date_14_16 %>% filter(Cross == "HN")
gilmour_h2_date_14_16_dam_sire_HN <- lmer(DOY.off ~ (1|DOY.sire)+(1|DOY.dam), data = gilmour_h2_date_14_16_HN, na.action = na.omit)
(summary(gilmour_h2_date_14_16_dam_sire_HN))

gilmour_h2_date_14_16_HH <- gilmour_triads_date_14_16 %>% filter(Cross == "HH")
gilmour_h2_date_14_16_dam_sire_HH <- lmer(DOY.off ~ (1|DOY.sire)+(1|DOY.dam), data = gilmour_h2_date_14_16_HH, na.action = na.omit)
(summary(gilmour_h2_date_14_16_dam_sire_HH))

gilmour_h2_date_14_16_F <- gilmour_triads_date_14_16 %>% filter(Sex.off == "F")
gilmour_h2_date_14_16_dam_sire_F <- lmer(DOY.off ~ (1|DOY.sire)+(1|DOY.dam), data = gilmour_h2_date_14_16_F, na.action = na.omit)
(summary(gilmour_h2_date_14_16_dam_sire_F))

gilmour_h2_date_14_16_M <- gilmour_triads_date_14_16 %>% filter(Sex.off == "M")
gilmour_h2_date_14_16_dam_sire_M <- lmer(DOY.off ~ (1|DOY.sire)+(1|DOY.dam), data = gilmour_h2_date_14_16_M, na.action = na.omit)
(summary(gilmour_h2_date_14_16_dam_sire_M))

gilmour_h2_date_14_16_all <- (3.985 + 1.571) / (3.985 + 1.571 + 17.646)
gilmour_h2_date_14_16_NN <- (6.3645 + 0.7492) / (6.3645 + 0.7492 + 21.3311)
gilmour_h2_date_14_16_NH <- (0.4543 + 5.8177) / (0.4543 + 5.8177 + 14.0447)
gilmour_h2_date_14_16_HN <- (15.212 + 3.167) / (15.212 + 3.167 + 12.954)
gilmour_h2_date_14_16_HH <- (5.556 + 1.187) / (5.556 + 1.187 + 19.191)
gilmour_h2_date_14_16_dam_F <- 5.063 / (5.063 + 0.00 + 19.949)
gilmour_h2_date_14_16_sire_F <- 0.00 / (5.063 + 0.00 + 19.949)
gilmour_h2_date_14_16_dam_M <- 0.5972 / (0.5972 + 6.2724 + 15.6431)
gilmour_h2_date_14_16_sire_M <- 6.2724 / (0.5972 + 6.2724 + 15.6431)

gilmour_am_for_h2_date_14_16 <- as.data.frame(c(gilmour_h2_date_14_16_all, gilmour_h2_date_14_16_NN, gilmour_h2_date_14_16_NH, gilmour_h2_date_14_16_HN, gilmour_h2_date_14_16_HH, gilmour_h2_date_14_16_dam_F, gilmour_h2_date_14_16_dam_M, gilmour_h2_date_14_16_sire_F, gilmour_h2_date_14_16_sire_M)) %>% round(3)
colnames(gilmour_am_for_h2_date_14_16) <- c("Heritability (h^2)")
gilmour_am_for_h2_date_14_16$Category <- c("Overall", "NN", "NH", "HN", "HH", "Dam/Female offspring", "Dam/Male offspring", "Sire/Female offspring", "Sire/Male offspring")
(gilmour_am_for_h2_date_14_16 <- gilmour_am_for_h2_date_14_16[,c(2,1)])
```

### *Results*

```{r}
gilmour_r_and_h2_date_14_16 <- gilmour_r_and_h2_date_14_16[,c(1,3)]

gilmour_h2_date_14_16 <- merge(gilmour_r_and_h2_date_14_16, gilmour_slope_as_h2_date_14_16,  by="Category") %>% 
    merge(gilmour_am_for_h2_date_14_16, by="Category") %>% 
    rename(gilmour_14_16 = Category, "h^2.correlation" = "Heritability (h^2).x", "h^2.slope" = "Heritability (h^2).y", "h^2.variance" = "Heritability (h^2)")
(gilmour_h2_date_14_16 <- gilmour_h2_date_14_16[c(6,5,4,3,1,2,7,8),])
```

There is good evidence of heritability of sample date as a proxy for return time, although cross data is confounded by missing origin info in 2018.


## Sample Location

**Note:** Correlation between parent-offspring sample location and narrow-sense heritability of sample location *not yet available* for parental years beyond 2014. GIS-based stream distance mapping is *in progress* by Chase Jalbert.

```{r}

```

### Scatterplots

```{r}

```

### Correlation

Location is reported as distance from stream mouth. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}

```

### Slope

```{r, echo=FALSE, warning=FALSE, message=FALSE}

```

### Animal model

```{r}

```

### *Results*

Stream location/distance data not yet available for all sample sets.


## Reproductive success

```{r}
# Read in data
gilmour_paired_14_16_filter_parents_n <- read_csv("../Gilmour/gilmour_paired_14_16_filter_parents.csv") %>% 
  select(franz_id, n) %>% 
  rename(n_14 = n)
gilmour_paired_16_18_filter_parents_n <- read_csv("../Gilmour/gilmour_paired_16_18_filter_parents.csv") %>% 
  select(franz_id, n) %>% 
  rename(n.off = n)

# join by n (i.e. individual RS)
triads_n <- function(indata_triads = gilmour_triads_14_16, indata_paired_filter_parents_n = gilmour_paired_14_16_filter_parents_n, indata_paired_filter_offspring_n = gilmour_paired_16_18_filter_parents_n) {
  inner_join(indata_triads, indata_paired_filter_parents_n, by = c("Parent_ID.dam" = "franz_id")) %>% 
  rename(n.dam = n_14) %>% 
  inner_join(indata_paired_filter_parents_n, by = c("Parent_ID.sire" = "franz_id")) %>% 
  rename(n.sire = n_14) %>% 
  inner_join(indata_paired_filter_offspring_n, by = c("Offspring" = "franz_id"))
}

gilmour_triads_n_14_16 <- triads_n()

# mid-parent trait values and cross info
triads_rs <- function(indata_triads_rs = gilmour_triads_n_14_16) {
  indata_triads_rs %>%
  mutate(rs.par = (n.dam + n.sire) / 2) %>% 
  mutate(Cross = case_when(Origin.dam == "Natural" & Origin.sire == "Natural" ~ "NN",
                           Origin.dam == "Natural" & Origin.sire == "Hatchery" ~ "NH",
                           Origin.dam == "Hatchery" & Origin.sire == "Natural" ~ "HN",
                           Origin.dam == "Hatchery" & Origin.sire == "Hatchery" ~ "HH"))
}

gilmour_triads_rs_14_16 <- triads_rs()
```

### Scatterplots

```{r}
h2_rs_graph <- function(indata_h2_rs_graph = gilmour_triads_rs_14_16, title = "Gilmour_14_16\n\nHeritability of RS by Cross") {
  indata_h2_rs_graph %>% 
  ggplot(aes(x = rs.par, y = n.off, colour = Cross)) +
  geom_abline(slope = 1) +
  geom_jitter(size = 2) +
  geom_smooth(method = "lm") +
  labs(title = title, colour = "Parent: Origin") +
  xlab("Mid-parent RS") +
  ylab("Offspring RS") +
  theme_bw() +
  theme(text = element_text(size = 20))
}

h2_rs_graph()
```

### Correlation

```{r, echo=FALSE, warning=FALSE, message=FALSE}
parents_paired_rs <- function(indata_triads_rs = gilmour_triads_rs_14_16) {
  indata_triads_rs %>% 
  drop_na(rs.par, n.off) %>% 
  mutate(sex.dam = case_when(Sex.dam == "F" ~ "Female")) %>%
  mutate(sex.sire = case_when(Sex.sire == "M" ~ "Male")) %>%
  mutate(sex.off = case_when(Sex.off == "F" ~ "Female",
                             Sex.off == "M" ~ "Male")) 
}

(gilmour_parents_paired_14_16_rs1 <- parents_paired_rs() %>% 
  group_by(Cross) %>%
  summarise(correlation = round(cor(n.off, rs.par), 3), r2 = round(cor(n.off, rs.par)^2, 3)))
(gilmour_parents_paired_14_16_rs2 <- parents_paired_rs() %>% 
  group_by(sex.dam, sex.off) %>%
  summarise(correlation = round(cor(n.off, rs.par), 3), r2 = round(cor(n.off, rs.par)^2, 3)))
(gilmour_parents_paired_14_16_rs3 <- parents_paired_rs() %>% 
  group_by(sex.sire, sex.off) %>%
  summarise(correlation = round(cor(n.off, rs.par), 3), r2 = round(cor(n.off, rs.par)^2, 3)))

gilmour_parents_paired_14_16_rs1_NN <- as.data.frame(c(gilmour_parents_paired_14_16_rs1[4,2], gilmour_parents_paired_14_16_rs1[4,3]*2))
gilmour_parents_paired_14_16_rs1_NH <- as.data.frame(c(gilmour_parents_paired_14_16_rs1[3,2], gilmour_parents_paired_14_16_rs1[3,3]*2))
gilmour_parents_paired_14_16_rs1_HN <- as.data.frame(c(gilmour_parents_paired_14_16_rs1[2,2], gilmour_parents_paired_14_16_rs1[2,3]*2))
gilmour_parents_paired_14_16_rs1_HH <- as.data.frame(c(gilmour_parents_paired_14_16_rs1[1,2], gilmour_parents_paired_14_16_rs1[1,3]*2))
gilmour_parents_paired_14_16_rs2_dF <- as.data.frame(c(gilmour_parents_paired_14_16_rs2[1,3], gilmour_parents_paired_14_16_rs2[1,4]*2))
gilmour_parents_paired_14_16_rs2_dM <- as.data.frame(c(gilmour_parents_paired_14_16_rs2[2,3], gilmour_parents_paired_14_16_rs2[2,4]*2))
gilmour_parents_paired_14_16_rs3_sF <- as.data.frame(c(gilmour_parents_paired_14_16_rs3[1,3], gilmour_parents_paired_14_16_rs3[1,4]*2))
gilmour_parents_paired_14_16_rs3_sM <- as.data.frame(c(gilmour_parents_paired_14_16_rs3[2,3], gilmour_parents_paired_14_16_rs3[2,4]*2))

gilmour_r_and_h2_rs_14_16 <- rbind(gilmour_parents_paired_14_16_rs1_NN, gilmour_parents_paired_14_16_rs1_NH, gilmour_parents_paired_14_16_rs1_HN, gilmour_parents_paired_14_16_rs1_HH, gilmour_parents_paired_14_16_rs2_dF, gilmour_parents_paired_14_16_rs2_dM, gilmour_parents_paired_14_16_rs3_sF, gilmour_parents_paired_14_16_rs3_sM) %>% round(3)
colnames(gilmour_r_and_h2_rs_14_16) <- c("Correlation (r)", "Heritability (h^2)")
gilmour_r_and_h2_rs_14_16$Category <- c("NN", "NH", "HN", "HH", "Dam/Female offspring", "Dam/Male offspring", "Sire/Female offspring", "Sire/Male offspring")
(gilmour_r_and_h2_rs_14_16 <- gilmour_r_and_h2_rs_14_16[,c(3,1,2)])
```

### Slope

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# by cross
gilmour_h2_rs_14_16_all_tbl <- tidy(lm(n.off ~ rs.par, data = gilmour_triads_rs_14_16))
gilmour_h2_rs_14_16_NN_tbl <- tidy(lm(n.off ~ rs.par, data = gilmour_triads_rs_14_16 %>% filter(Cross == "NN")))
gilmour_h2_rs_14_16_NH_tbl <- tidy(lm(n.off ~ rs.par, data = gilmour_triads_rs_14_16 %>% filter(Cross == "NH")))
gilmour_h2_rs_14_16_HN_tbl <- tidy(lm(n.off ~ rs.par, data = gilmour_triads_rs_14_16 %>% filter(Cross == "HN")))
gilmour_h2_rs_14_16_HH_tbl <- tidy(lm(n.off ~ rs.par, data = gilmour_triads_rs_14_16 %>% filter(Cross == "HH")))

# by sex
gilmour_h2_rs_14_16_dam_tbl <- tidy(lm(n.off ~ n.dam, data = gilmour_triads_rs_14_16))
gilmour_h2_rs_14_16_sire_tbl <- tidy(lm(n.off ~ n.sire, data = gilmour_triads_rs_14_16))
gilmour_h2_rs_14_16_dF_tbl <- tidy(lm(n.off ~ n.dam, data = gilmour_triads_rs_14_16 %>% filter(Sex.off == "F")))
gilmour_h2_rs_14_16_sM_tbl <- tidy(lm(n.off ~ n.sire, data = gilmour_triads_rs_14_16 %>% filter(Sex.off == "M")))
gilmour_h2_rs_14_16_dM_tbl <- tidy(lm(n.off ~ n.dam, data = gilmour_triads_rs_14_16 %>% filter(Sex.off == "M")))
gilmour_h2_rs_14_16_sF_tbl <- tidy(lm(n.off ~ n.sire, data = gilmour_triads_rs_14_16 %>% filter(Sex.off == "F")))

gilmour_slope_as_h2_rs_14_16 <- rbind(gilmour_h2_rs_14_16_all_tbl[2,2], gilmour_h2_rs_14_16_NN_tbl[2,2], gilmour_h2_rs_14_16_NH_tbl[2,2], gilmour_h2_rs_14_16_HN_tbl[2,2], gilmour_h2_rs_14_16_HH_tbl[2,2], gilmour_h2_rs_14_16_dam_tbl[2,2], gilmour_h2_rs_14_16_sire_tbl[2,2], gilmour_h2_rs_14_16_dF_tbl[2,2], gilmour_h2_rs_14_16_dM_tbl[2,2], gilmour_h2_rs_14_16_sF_tbl[2,2], gilmour_h2_rs_14_16_sM_tbl[2,2]) %>% round(3)
colnames(gilmour_slope_as_h2_rs_14_16) <- c("Heritability (h^2)")
gilmour_slope_as_h2_rs_14_16$Category <- c("Overall", "NN", "NH", "HN", "HH", "Dam/All offspring", "Sire/All offspring", "Dam/Female offspring", "Dam/Male offspring", "Sire/Female offspring", "Sire/Male offspring")
(gilmour_slope_as_h2_rs_14_16 <- gilmour_slope_as_h2_rs_14_16[,c(2,1)])
```

### Animal model

```{r}
gilmour_h2_rs_14_16_dam_sire <- lmer(n.off ~ (1|n.sire)+(1|n.dam), data = gilmour_triads_rs_14_16, na.action = na.omit)
(summary(gilmour_h2_rs_14_16_dam_sire))
plot(gilmour_h2_rs_14_16_dam_sire)
hist(resid(gilmour_h2_rs_14_16_dam_sire))

gilmour_h2_rs_14_16_NN <- gilmour_triads_rs_14_16 %>% filter(Cross == "NN")
gilmour_h2_rs_14_16_dam_sire_NN <- lmer(n.off ~ (1|n.sire)+(1|n.dam), data = gilmour_h2_rs_14_16_NN, na.action = na.omit)
(summary(gilmour_h2_rs_14_16_dam_sire_NN))

gilmour_h2_rs_14_16_NH <- gilmour_triads_rs_14_16 %>% filter(Cross == "NH")
gilmour_h2_rs_14_16_dam_sire_NH <- lmer(n.off ~ (1|n.sire)+(1|n.dam), data = gilmour_h2_rs_14_16_NH, na.action = na.omit)
(summary(gilmour_h2_rs_14_16_dam_sire_NH))

gilmour_h2_rs_14_16_HN <- gilmour_triads_rs_14_16 %>% filter(Cross == "HN")
gilmour_h2_rs_14_16_dam_sire_HN <- lmer(n.off ~ (1|n.sire)+(1|n.dam), data = gilmour_h2_rs_14_16_HN, na.action = na.omit)
(summary(gilmour_h2_rs_14_16_dam_sire_HN))

gilmour_h2_rs_14_16_HH <- gilmour_triads_rs_14_16 %>% filter(Cross == "HH")
gilmour_h2_rs_14_16_dam_sire_HH <- lmer(n.off ~ (1|n.sire)+(1|n.dam), data = gilmour_h2_rs_14_16_HH, na.action = na.omit)
(summary(gilmour_h2_rs_14_16_dam_sire_HH))

gilmour_h2_rs_14_16_F <- gilmour_triads_rs_14_16 %>% filter(Sex.off == "F")
gilmour_h2_rs_14_16_dam_sire_F <- lmer(n.off ~ (1|n.sire)+(1|n.dam), data = gilmour_h2_rs_14_16_F, na.action = na.omit)
(summary(gilmour_h2_rs_14_16_dam_sire_F))

gilmour_h2_rs_14_16_M <- gilmour_triads_rs_14_16 %>% filter(Sex.off == "M")
gilmour_h2_rs_14_16_dam_sire_M <- lmer(n.off ~ (1|n.sire)+(1|n.dam), data = gilmour_h2_rs_14_16_M, na.action = na.omit)
(summary(gilmour_h2_rs_14_16_dam_sire_M))

gilmour_h2_rs_14_16_all <- 0.005904 / (0.005904 + 0.00 + 0.806979)
gilmour_h2_rs_14_16_NN <- 0.00 / (0.00 + 0.00 + 0.2564)
gilmour_h2_rs_14_16_NH <- 0.08873 / (0.00 + 0.08873 + 0.83867)
gilmour_h2_rs_14_16_HN <- 0.02971 / (0.02971 + 0.00 + 1.32096)
gilmour_h2_rs_14_16_HH <- (0.2591) / (0.2591 + 0.00 + 0.6904)
gilmour_h2_rs_14_16_dam_F <- 0.00 / (0.00 + 0.00 + 0.1211)
gilmour_h2_rs_14_16_sire_F <- 0.05681 / (0.05681 + 0.00 + 0.85955)
gilmour_h2_rs_14_16_dam_M <- 0.00 / (0.00 + 0.00 + 0.7266)
gilmour_h2_rs_14_16_sire_M <- 0.00 / (0.00 + 0.00 + 0.7266)

gilmour_am_for_h2_rs_14_16 <- as.data.frame(c(gilmour_h2_rs_14_16_all, gilmour_h2_rs_14_16_NN, gilmour_h2_rs_14_16_NH, gilmour_h2_rs_14_16_HN, gilmour_h2_rs_14_16_HH, gilmour_h2_rs_14_16_dam_F, gilmour_h2_rs_14_16_dam_M, gilmour_h2_rs_14_16_sire_F, gilmour_h2_rs_14_16_sire_M)) %>% round(3)
colnames(gilmour_am_for_h2_rs_14_16) <- c("Heritability (h^2)")
gilmour_am_for_h2_rs_14_16$Category <- c("Overall", "NN", "NH", "HN", "HH", "Dam/Female offspring", "Dam/Male offspring", "Sire/Female offspring", "Sire/Male offspring")
(gilmour_am_for_h2_rs_14_16 <- gilmour_am_for_h2_rs_14_16[,c(2,1)])
```

### *Results*

```{r}
gilmour_r_and_h2_rs_14_16 <- gilmour_r_and_h2_rs_14_16[,c(1,3)]

gilmour_h2_rs_14_16 <- merge(gilmour_r_and_h2_rs_14_16, gilmour_slope_as_h2_rs_14_16,  by="Category") %>% 
    merge(gilmour_am_for_h2_rs_14_16, by="Category") %>% 
    rename(gilmour_14_16 = Category, "h^2.correlation" = "Heritability (h^2).x", "h^2.slope" = "Heritability (h^2).y", "h^2.variance" = "Heritability (h^2)")
(gilmour_h2_rs_14_16 <- gilmour_h2_rs_14_16[c(6,5,4,3,1,2,7,8),])
```

Nothing to see here. I wouldn't bank on the HH result being real; it's probably an artifact of small sample size due to a lack of origin data for 2016.