---
title: "05 - Investigate Otoliths With Reads But No Heart Match"
subtitle: "What To Do With Wells With 2 Otoliths, But No Matching Data?"
author: "Kyle Shedd"
date: "Started: 2025-02-06, last opened: `r Sys.Date()`"
output:
  html_notebook:
    theme: united
    toc: yes
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

# Setup

Load all necessary packages, as of 2023-08-11 Kyle switched to [GCLr](https://github.com/commfish/GCLr).
```{r setup, message=FALSE, results='hide'}
rm(list = ls(all.names = TRUE))

if(!require("pacman")) install.packages("pacman"); library(pacman)

pacman::p_load(
  tidyverse,
  lubridate,
  scales,
  janitor,
  GCLr
)

knitr::opts_chunk$set(fig.width = 10)

.username = readLines("~/R/usr_pw.txt", n = 1)  # LOKI username
.password = readLines("~/R/usr_pw.txt" , n = 2)[[2]]  # LOKI password
```

# Objective

The purpose of this notebook is to dig into the otolith read data for each DWP to determine if we can salvage any fish that didn't have a heart-otolith match above our thresholds. For example, some DWPs may not have any GOD otoliths with hatchery marks, in which case heart-otolith pairing is moot. Alternatively, we could have some wells with 2 otoliths and all surrounding wells have been resolved, but no valid heart-otolith match for that well.

The next step after this will be to *compare heart matches to existing AHRP data*. 

# Import Data

## Otolith Transfer Data

Read in the otolith transfer data that matches up 96 well otolith tray barcodes and wells to the original 48 DWP barcodes and wells. This was generated in `01_import-genotypes-join-otlith-transfer-records`.

Recall that only left and unknown otoliths were transferred. No right-side otoliths were transferred.
```{r}
PGOD21_48DWP_paired <- readr::read_csv(file = "../output/PGOD21_48DWP_paired.csv")
```

## Otolith Reads

Got GOD otolith reads from Cordova Otolith lab (2025-02-05) with help from Eric L. (GCL AP) and Rachel E. (CDV Otolith Lab Manager).
```{r}
(otolith_reads <- readr::read_csv("../data/GOD_otolith_reads_all.csv") %>% 
   janitor::clean_names())
```

Total of 10,816 otoliths.

# Join

```{r}
(
  PGOD21_48DWP_paired_otolith_reads <- PGOD21_48DWP_paired %>% dplyr::left_join(
    y = otolith_reads %>%
      dplyr::mutate(
        oto_tray = as.numeric(tray_id),
        oto_well = as.numeric(cell_number)
      ) %>%
      tidyr::unite(col = "oto_tray_and_well", c(oto_tray, oto_well), sep = "_") %>%
      dplyr::select(oto_tray_and_well, mark_present, mark_id),
    by = "oto_tray_and_well"
  ) # join otolith reads
)
```

# Investigate

## DWP 14326

Looking at `GOD/data/raw_working_oto_sheets.xlsm` and `GOD/output/otolith_shift_DWP_visualization_dup-rate-0.8_diff-0.25.pdf`, this was a pretty clean DWP. Otoliths appear to have shifted from just a few wells.

According to the DWP map, there could be otoliths from up to 26 different fish in GOD.
```{r}
PGOD21_48DWP_paired_otolith_reads %>% 
  dplyr::filter(`48_DWP_tray_barcode` == 14326) %>% 
  dplyr::distinct(`48_DWP_tray_position`) %>% 
  dplyr::arrange(`48_DWP_tray_position`)
```

D5, E3, E5, and F4 must have been a right otoliths and thus **not** part of GOD.

```{r}
PGOD21_48DWP_paired_otolith_reads %>% 
  dplyr::filter(`48_DWP_tray_barcode` == 14326) %>% 
  dplyr::count(oto_side, mark_present)
```

Of the GOD otoliths, there is only 1 hatchery-origin fish in this DWP.

```{r}
PGOD21_48DWP_paired_otolith_reads %>% 
  dplyr::filter(`48_DWP_tray_barcode` == 14326,
                is.na(mark_present) | mark_present != "NO") %>% 
  dplyr::select(`48_DWP_tray_barcode`, `48_DWP_tray_position`, mark_present, mark_id)
```

