---
title: "Stockdale Parentage"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(coin)
library(MASS)
library(scales)
library(tidyverse)
library(leaflet)
library(lubridate)
```

Below are analyses of parentage results calculated separately for Odd (2013-2015) and Even (2014-2016) lineages using *FRANz*. 

# Odd lineage (2013/2015)

## *FRANz* parameters
*FRANz* was run at 8:20AM on November 6, 2018. We used the following parameters:
FRANz.exe --Nmmax 2900 --Nfmax 2900 --femrepro 1:2 --malerepro 1:2 --typingerror 0.005 --updatefreqs --poutformat 2 --fullsibtest "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\stock_postQA_2013_2015_STOCK.dat"

The parameters are defined as follows:
--Nmmax and --Nfmax are the maximum numbers of candidate mothers and fathers. To obtain our values, we used an estimated escapement of 47000 and divided by 2. This escapement number came from the larger of the two estimates (aerial survey AUC and stream walk AUC). This estimate is listed under 'pop est aerial' in the Excel file found here: "V:\Documents\5_Coastwide\Multispecies\AHRP\Field data\Sample Summary Extract and Genotype.xlsx"   
--femrepro and --malerepro specify the age range in which an individual can reproduce  
--typingerror refers to the overall genotyping error rate. Ours was ~0.005  
--updatefreqs specifies that *FRANz* should update allele frequencies using MCMC sampling  
--poutformat specifies that all potential parents should be listed, not just the most likely  
--fullsibtest tests for full siblings among offspring   

All output files can be found here: "v:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\stock_postQA_2013_2015_STOCK_Run1"  

The summary file provides information about the power of our marker suite:   
Cumulative exclusion probability when 1 to 7 fullsibs are genotyped  
  First Parent              : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Second Parent             : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Parent Pair               : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000
  
According to the FRANz manual, marker sets are not considered powerful if these cumulative exclusion probabilities are less than 0.95, which indicates that the probability that a random pair of individuals in the population has a 5% chance of having a genotype pair compatible to an offspring genotype. Since all of our probabilities are 1, we can be **confident** in the power of our 304 amplicons to make parent assignments.  

## Import data
The first step is to read in .csv files for parentage assignments produced by *FRANz* as well as paired genotype and OceanAK data. 

```{r Setup Odd, message=FALSE, warning=FALSE}
parentage_13_15 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Franz/Stockdale/Odd_Run 1/parentage.csv")
paired_13_15 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Franz/stockdale_postQA_OceanAK_paired_2013_2015_STOCK.csv")
```

Plot a histogram of *FRANz* parentage posterior probabilities to show robustness of assignments
```{r FRANz_posterior_plot_odd}
parentage_13_15  %>% 
  filter(!is.na(`Parent 1`)) %>% 
  ggplot(aes(x = Posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  ggtitle("Histogram of FRANz posterior probabilities for parentage assignments")
```

All parentage assigments have a posterior probability of 1, which is very robust.  

## Filter paired data
The file containing paired genotype and OceanAK data has a lot of information that we do not need at this time (e.g. genotypes for each marker). We therefore separate out only the columns that contain identifying information for each individual.

```{r Filter paired data Odd}
# What are the non-genotype columns
grep(pattern = "RAD", x = colnames(paired_13_15), value = TRUE, invert = TRUE)

# Filter for non-genotype columns
paired_13_15_filter <- paired_13_15 %>% 
  dplyr::select(franz_id, SILLY, `Fish ID`, `DNA Tray Code`, `DNA Tray Well Code`, `Sample Year`, `Sample Date`, SEX, `Length Mm`, `Otolith Mark Present`, `Otolith Mark ID`) %>% 
  mutate(origin = case_when(`Otolith Mark Present` == "NO" ~ "Natural",
                            `Otolith Mark Present` == "YES" ~ "Hatchery")) %>%  # add origin variable
  mutate(origin = factor(origin, c("Natural", "Hatchery")))  # make factor to ensure hatchery != red
```

```{r Filter parentage data Odd}
# Filter for posterior > 0.9 + make tidy relative to assigment (1 row per parentage assignment)
parentage_13_15_filter <- parentage_13_15 %>% 
  filter(Posterior > 0.9) %>% 
  dplyr::select(Offspring, `Parent 1`, `Parent 2`) %>% 
  gather(Parent, Parent_ID, - Offspring) %>% 
  filter(!is.na(Parent_ID))
```

## Sample sizes
Quick look at what our sample sizes are for potential parents (2013) and potential offspring (2015).

```{r odd_sample_sizes}
paired_13_15_filter %>% 
  count(`Sample Year`, SEX, origin) %>% 
  spread(origin, n, fill = 0)
```


## Join parentage and individual data
Here, we joined parentage data from *FRANz* with individual data so that we can match data to each offspring and parent. **Note** that any column ending in `.off` refers to offspring and `.par` refers to parent.  

```{r Join parentage and paired objects Odd}
# Creating a single, "tidy" object where each row is a parent-offspring relationship
# Offspring with 2 parents have 2 rows
# .off data is offspring, .par data is parents
(parents_paired_13_15 <- parentage_13_15_filter %>% 
   left_join(paired_13_15_filter, by = c("Offspring" = "franz_id")) %>% 
   left_join(paired_13_15_filter, by = c("Parent_ID" = "franz_id"), suffix = c(".off", ".par")) %>% 
   mutate(origin = case_when(`Otolith Mark Present.par` == "NO" ~ "Natural",
                             `Otolith Mark Present.par` == "YES" ~ "Hatchery")) %>% 
   mutate(origin = factor(origin, c("Natural", "Hatchery"))))  # added hatchery/natural variable

# How many single-parent offspring pair assignments?
parents_paired_13_15 %>% 
  count(Parent)

# How many unique parents had offspring assigned?
n_distinct(parents_paired_13_15$Parent_ID)

# For each offspring assigment, what was the parent's origin?
parents_paired_13_15 %>% 
  count(origin, `Otolith Mark ID.par`) %>% 
  spread(origin, n)
```

**Note** that *FRANz* assigned 72 offspring to 54 parents. 69 of these parents are natural origin and 3 are hatchery origin. The hatchery origin fish are from AFK11A and AFK11B.

## Proportion test
Let's use a Chi-Square test to determine whether the proportions of offspring assigned to hatchery- and natural-origin parents significantly differ from the proportions of hatchery- and natural-origin parents sampled. 

```{r proportion_odd_data}
# How many spawners from each origin?
paired_13_15_filter %>% 
  filter(`Sample Year` == 2013) %>% 
  count(origin)

# How many offspring from each origin?
parents_paired_13_15 %>% 
  count(origin)
```

```{r Chi square test odd}
O.odd <- c("H", "N")
Spawners.odd <- c(163, 811) #  number of hatchery- and natural-origin parents
Offspring.odd <- c(3, 69)  # number of offspring assigned to hatchery- and natural-origin parents

df.odd <- data.frame(Spawners.odd, Offspring.odd)  # made a data.frame because tibbles don't like rownames
row.names(df.odd) <- O.odd

df.odd
chisq.test(df.odd)
```

There is a highly significant difference in the proportions of offspring assigned to hatchery- and natural-origin parents relative to the proportions of potential parents sampled, indicating an under-representation of offspring assigning to hatchery-origin parents. We see that although **`r round(df.odd["H", "Spawners.odd"] / sum(df.odd[, "Spawners.odd"]), 2) * 100`%** of parents genotyped were of hatchery-origin, only **`r round(df.odd["H", "Offspring.odd"] / sum(df.odd[, "Offspring.odd"]), 2) * 100`%** of offspring assigned to hatchery-origin parents.

## Family size
Here, we calculate and plot family size for both natural and hatchery parents.  Family size ranges from 1-4. All assignments were to 1 parent (*single parent-offspring duos*) - there were no 2 parent assignments (*parent pair-offspring trios*). 

Of the 3,093 offspring collected in 2015, 72 were assigned to parents, for an overall offspring assignment rate of **2.3%**. 

### Single parent-offspring pairs
This plot shows the distribution of family size by sex (number of offspring assigned to a female parent or a male parent) for both hatchery and natural parents.  

```{r Family size by sex Odd}
#parents_paired_13_15 %>% 
  #count(`Fish ID.par`, origin, SEX.par) %>% 
  #count(origin, SEX.par, n) %>% 
  #complete(origin, SEX.par, n, fill = list(nn = 0)) %>% 
  #ggplot(aes(x=n, y = nn, fill=origin)) +
  #geom_col(position = position_dodge2(preserve = "single")) +
  #scale_y_continuous(breaks = pretty_breaks()) +
  #scale_x_continuous(breaks = pretty_breaks()) +
  #facet_grid(~ SEX.par) +
  #labs(title="Distribution of Family Size by Females (F) and Males (M)\nNumber of Families",
  #     fill = "Parent: Origin") +
  #xlab("Number of Offspring")+
  #ylab("Number of Parents") 



# Plot histogram of proportion of parents with a given family size
parents_paired_13_15 %>% 
  count(`Fish ID.par`, origin, SEX.par) %>% 
  count(SEX.par, origin, n) %>% 
  group_by(SEX.par, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x=n, y=p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  facet_grid(~ SEX.par) +
  labs(title="Distribution of Family Size by Females (F) and Males (M)\nProportion of Families",
       fill = "Parent: Origin") +
  xlab("Number of Offspring")+
  ylab("Proportion of Parents") 
```

When we plot the sexes separately, we can see that most of the assigned parents are males. Female parents produced 1-2 offspring, while male parents produced 1-4.  

Below, we'll calculate reproductive success of natural and hatchery females and males. 

```{r odd_rrs}
(rs_13_15 <- parents_paired_13_15 %>%
  count(`Fish ID.par`, origin, SEX.par) %>%
  group_by(origin, SEX.par) %>% 
  summarise(RS = mean(n))) %>% 
  spread(SEX.par, RS)
  

F_h_odd <- rs_13_15 %>% 
  filter(origin == "Hatchery" & SEX.par == "F") %>% 
  pull(RS)
M_h_odd <- rs_13_15 %>% 
  filter(origin == "Hatchery" & SEX.par == "M") %>% 
  pull(RS)
F_n_odd <- rs_13_15 %>% 
  filter(origin == "Natural" & SEX.par == "F") %>% 
  pull(RS)
M_n_odd <- rs_13_15 %>% 
  filter(origin == "Natural" & SEX.par == "M") %>% 
  pull(RS)
```

#### Calculate RRS again including 0's 
We'll calculate RRS again including sampled fish from the parental generation to which offspring were not assigned.

```{r Re-calculate RRS odd}
# Make one tibble with all assignment data
# RS_combined <- bind_rows(parents_paired_13_15, parents_paired_14_16) 

# Transform to one row per parent with n = number of offspring produced
parents_paired_13_15_count <- parents_paired_13_15 %>% 
  count(Parent_ID, origin, SEX.par, `Sample Year.par`, `Length Mm.par`, `Sample Date.par`) %>% 
  mutate(lineage = dplyr::case_when(`Sample Year.par` %in% c("2013", "2015") ~ "odd",
                                    `Sample Year.par` %in% c("2014", "2016") ~ "even"))

paired_13_15_filter_parents <- paired_13_15_filter %>% 
  filter(`Sample Year` == "2013")

# Join with all parents that produced no progeny
paired_13_15_filter_parents <- left_join(paired_13_15_filter_parents, parents_paired_13_15_count, by = c("franz_id" = "Parent_ID")) %>% 
  mutate(n = replace_na(n, 0)) %>% 
  rename(origin = origin.x) %>% 
  select(-origin.y, -SEX.par, -`Sample Year.par`, -`Length Mm.par`, -`Sample Date.par`)

write_csv(paired_13_15_filter_parents, "stock_paired_13_15_filter_parents.csv")

# Mean RS including 0's
rs_13_15_0 <- paired_13_15_filter_parents %>%
  group_by(origin, SEX) %>% 
  summarise(RS = mean(n, na.rm = TRUE))

rs_13_15_0 %>%
  mutate(RS = round(RS, 2)) %>% 
  spread(SEX, RS)

# Plot histogram of number of parents with a given family size
#paired_13_15_filter_parents %>% 
#  ggplot(aes(x = n, fill = origin)) +
#  geom_bar(position = position_dodge2(preserve = "single")) +
#  facet_grid( ~ SEX) +
#  xlab("Number of Offspring") +
#  ylab("Number of Parents") +
#  labs(title = "Distribution of Family Size by Females (F) and Males (M)\nNumber of Families",
#       fill = "Parent: Origin")

# Plot histogram of proportion of parents with a given family size
paired_13_15_filter_parents %>% 
  count(SEX, origin, n) %>% 
  group_by(SEX, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  facet_grid(~ SEX) +
  labs(title="Distribution of Family Size by Females (F) and Males (M)\nProportion of Families",
       fill = "Parent: Origin") +
  xlab("Number of Offspring")+
  ylab("Proportion of Parents") 

# Save RS values
F_h_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Hatchery" & SEX == "F") %>% 
  pull(RS)
M_h_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Hatchery" & SEX == "M") %>% 
  pull(RS)
F_n_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Natural" & SEX == "F") %>% 
  pull(RS)
M_n_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Natural" & SEX == "M") %>% 
  pull(RS)
```

Here are the plots again with lines showing mean reproductive success. 

```{r Plots of RS by sex with mean lines odd}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

#paired_13_15_filter_parents %>% 
#  filter(SEX == "F") %>% 
#  ggplot(aes(x = n, fill = origin)) +
#  geom_bar(position = position_dodge2(preserve = "single")) +
#  scale_y_continuous(breaks = pretty_breaks()) +
#  scale_x_continuous(breaks = pretty_breaks()) +
#  xlab("Number of Offspring") +
#  ylab("Number of Females") +
#  labs(title="Number of Offspring for Females",
#      fill = "Parent: Origin") +
#  geom_vline(xintercept=F_n_0_13_15, size=2) +
#  geom_vline(xintercept=F_n_0_13_15, color=gg_color_hue(2)[1], linetype = "dashed", size=2) +
#  geom_vline(xintercept=F_h_0_13_15, size=2) +
#  geom_vline(xintercept=F_h_0_13_15, color=gg_color_hue(2)[2], linetype = "dashed", size=2) +
#  annotate("text", x = F_n_0_13_15 + 0.5, y = 250, label = paste("F_n_0 =", round(F_n_0_13_15, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[1]) +
#  annotate("text", x = F_h_0_13_15 + 0.5, y = 300, label = paste("F_h_0 =", round(F_h_0_13_15, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[2])

#paired_13_15_filter_parents %>% 
#  filter(SEX == "M") %>% 
#  ggplot(aes(x = n, fill = origin)) +
#  geom_bar(position = position_dodge2(preserve = "single")) +
#  scale_y_continuous(breaks = pretty_breaks()) +
#  scale_x_continuous(breaks = pretty_breaks()) +
#  xlab("Number of Offspring") +
#  ylab("Number of Males") +
#  labs(title="Number of Offspring for Males",
#       fill = "Parent: Origin") +
#  geom_vline(xintercept=M_n_0_13_15, size=2) +
#  geom_vline(xintercept=M_n_0_13_15, color=gg_color_hue(2)[1], linetype = "dashed", size=2) +
#  geom_vline(xintercept=M_h_0_13_15, size=2) +
#  geom_vline(xintercept=M_h_0_13_15, color=gg_color_hue(2)[2], linetype = "dashed", size=2) +
#  annotate("text", x = M_n_0_13_15 + 0.5, y = 150, label = paste("M_n_0 =", round(M_n_0_13_15, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[1]) +
#  annotate("text", x = M_h_0_13_15 + 0.5, y = 100, label = paste("M_h_0 =", round(M_h_0_13_15, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[2])

paired_13_15_filter_parents %>% 
  filter(SEX == "F") %>% 
  count(origin, n) %>% 
  group_by(origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x=n, y=p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +xlab("Number of Offspring") +
  ylab("Proportion of Females") +
  labs(title="Number of Offspring for Females",
       fill = "Parent: Origin") +
  geom_vline(xintercept=F_n_0_13_15, size=2) +
  geom_vline(xintercept=F_n_0_13_15, color=gg_color_hue(2)[1], linetype = "dashed", size=2) +
  geom_vline(xintercept=F_h_0_13_15, size=2) +
  geom_vline(xintercept=F_h_0_13_15, color=gg_color_hue(2)[2], linetype = "dashed", size=2) +
  annotate("text", x = F_n_0_13_15 + 0.5, y = .250, label = paste("F_n_0 =", round(F_n_0_13_15, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[1]) +
  annotate("text", x = F_h_0_13_15 + 0.5, y = .300, label = paste("F_h_0 =", round(F_h_0_13_15, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[2]) 

paired_13_15_filter_parents %>% 
  filter(SEX == "M") %>% 
  count(origin, n) %>% 
  group_by(origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x=n, y=p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  xlab("Number of Offspring") +
  ylab("Proportion of Males") +
  labs(title="Number of Offspring for Males",
       fill = "Parent: Origin") +
  geom_vline(xintercept=M_n_0_13_15, size=2) +
  geom_vline(xintercept=M_n_0_13_15, color=gg_color_hue(2)[1], linetype = "dashed", size=2) +
  geom_vline(xintercept=M_h_0_13_15, size=2) +
  geom_vline(xintercept=M_h_0_13_15, color=gg_color_hue(2)[2], linetype = "dashed", size=2) +
  annotate("text", x = M_n_0_13_15 + 0.5, y = .250, label = paste("M_n_0 =", round(M_n_0_13_15, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[1]) +
  annotate("text", x = M_h_0_13_15 + 0.5, y = .300, label = paste("M_h_0 =", round(M_h_0_13_15, 2)), hjust = 0, size = 6, color = gg_color_hue(2)[2])

```

```{r Calculate relative reproductive success including 0s odd}
RRS_m_0_13_15 <- M_h_0_13_15/M_n_0_13_15
RRS_f_0_13_15 <- F_h_0_13_15/F_n_0_13_15

RRS_tibble_0_13_15 <- tibble(RRS_m_0_13_15, RRS_f_0_13_15)
round(RRS_tibble_0_13_15, 2)
```

**RRS for males = `r round(RRS_m_0_13_15, 2)`. RRS for females = `r round(RRS_f_0_13_15, 2)`.**

Let's test for significant differences in RRS between male and female hatchery and natural origin fish. We'll perform two different statistical tests:  

  1. Nonparametric permutation test
  2. Parametric negative binomial GLM  
  
**Note** we are only doing this for the even lineage because the sample sizes are too low in the odd lineage. 

```{r RRS statistical testing including 0s odd}
# The format of the data to test was very simple, it was a data.frame called “mydata” with 2 columns
# 1.	nOff = number of offspring per family
# 2.	Origin = “H” or “W” for if the parent was hatchery or natural (wild)
# 

mydata_M_0 <- paired_13_15_filter_parents %>%
  filter(SEX == "M")

mydata_F_0 <- paired_13_15_filter_parents %>%
  filter(SEX == "F")

perm_1tail_pvalue_M_0 <- round(x = as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_M_0, distribution = approximate(B = 10000), alternative = "greater"))), digits = 4)

perm_1tail_pvalue_F_0 <- round(x = as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_F_0, distribution = approximate(B = 10000), alternative = "greater"))), digits = 4)

#And to do a 1-tailed negative binomial GLM

fit_M_0 <- glm.nb(n ~ origin, data = mydata_M_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_M_0 <- round(summary(fit_M_0)$coefficients[2, 4], 4)

fit_F_0 <- glm.nb(n ~ origin, data = mydata_F_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_F_0 <- round(summary(fit_F_0)$coefficients[2, 4], 4)

RRS_0 <- tibble(Sex = c("M", "F", "M", "F"),
                Test = c(rep("Permutation", 2), rep("Negative Binomial GLM", 2)),
                p_value = c(perm_1tail_pvalue_M_0, perm_1tail_pvalue_F_0, nbGLM_1tail_pvalue_M_0, nbGLM_1tail_pvalue_F_0))
RRS_0 %>% 
  spread(Sex, p_value)

# (RRS_0 <- tibble(perm_1tail_pvalue_F_0, perm_1tail_pvalue_M_0, nbGLM_1tail_pvalue_F_0, nbGLM_1tail_pvalue_M_0))
```

There is a significant difference in RRS for females, but not males. 

#### Calculate CI's for RRS

We will calculate 95% CI's for RRS based on Kalinowski and Taper 2005 [link] (http://www.nrcresearchpress.com/doi/10.1139/f04-239). This is a maximum likelihood estimate based off of Hinrichsen 2003. This method has been used in other peer reviewed papers such as Ford et al. 2016 [link] (https://www.researchgate.net/publication/309147116_Broodstock_History_Strongly_Influences_Natural_Spawning_Success_in_Hatchery_Steelhead_Oncorhynchus_mykiss#pf14).

```{r RRS_confidence_intervals_function}
rrs_ci_kalinowski <- function(n_h_off, n_w_off, n_h_par, n_w_par, alpha){
  chi_alpha <- qchisq(p = (1 - alpha), df = 1)
  n_off <- sum(c(n_h_off, n_w_off))
  n_par <- sum(c(n_h_par, n_w_par))
  
  rs_h <- n_h_off / n_h_par
  rs_w <- n_w_off / n_w_par
  
  p_h_par <- n_h_par / n_par
  p_w_par <- n_w_par / n_par
  
  rrs_h <- rs_h / rs_w
  rrs_w <- rs_w / rs_w
  rrs_avg <- (rrs_h * p_h_par) + (rrs_w * p_w_par)
  
  rrs_ml <- (n_h_off * log(p_h_par * rrs_h / rrs_avg)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg))
  
  xi_dist <- bind_rows(
    lapply(seq(0.01, 5, by = 0.01), function(rrs_h_xi) {
      rrs_avg_xi <- (rrs_h_xi * p_h_par) + (rrs_w * p_w_par)
      tibble(rrs_crit = rrs_h_xi,
             logl = (n_h_off * log(p_h_par * rrs_h_xi / rrs_avg_xi)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg_xi)) - (rrs_ml - chi_alpha / 2)
      )
    } )
  )
  
  rrs_min <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit < rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  rrs_max <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit > rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  xi_plot <- xi_dist %>% 
    ggplot(aes(x = rrs_crit, y = logl)) +
    geom_line() +
    geom_hline(yintercept = 0, colour = "red", lwd = 2) +
    geom_vline(xintercept = c(rrs_h, rrs_min, rrs_max), colour = "blue") +
    ylim(c(-5, 5)) +
    xlim(c(0, 2)) +
    ylab("Log Likelihood - Chi Sq Value") +
    annotate("text", x = rrs_h + 0.1, y = xi_dist %>% filter(rrs_crit == xi_dist$rrs_crit[which.min(abs(xi_dist$rrs_crit  - rrs_h))]) %>% pull(logl) + 0.4, label = round(rrs_h, 2)) +
    annotate("text", x = rrs_min - 0.1, y = xi_dist %>% filter(rrs_crit == rrs_min) %>% pull(logl) + 0.4, label = rrs_min) +
    annotate("text", x = rrs_max + 0.1, y = xi_dist %>% filter(rrs_crit == rrs_max) %>% pull(logl) + 0.4, label = rrs_max)
  
  print(xi_plot)
  return(c(rrs_min, rrs_h, rrs_max))
}
```

##### Odd 2013/2015

Calculate 95% CIs for Stockdale 13-15 females. RRS = 0.02 (0.01-0.09)

```{r RRS_confidence_intervals_female_13_15}
# Females
# Get parent sample sizes
paired_13_15_filter_parents %>% 
  filter(SEX == "F") %>% 
  count(SEX, origin) %>% 
  spread(origin, nn)

# Get offspring sample sizes
paired_13_15_filter_parents %>% 
  filter(SEX == "F") %>% 
  group_by(SEX, origin) %>% 
  summarise(n_off = sum(n)) %>% 
  spread(origin, n_off)

rrs_ci_kalinowski(n_h_off = 1, n_w_off = 25, n_h_par = 281, n_w_par = 131, alpha = 0.05)
```

Calculate 95% CIs for Hogan 13-15 males. RRS = 0.12 (0.02-0.4)

```{r RRS_confidence_intervals_male_13_15}
# Males
# Get parent sample sizes
paired_13_15_filter_parents %>% 
  filter(SEX == "M") %>% 
  count(SEX, origin) %>% 
  spread(origin, nn)

# Get offspring sample sizes
paired_13_15_filter_parents %>% 
  filter(SEX == "M") %>% 
  group_by(SEX, origin) %>% 
  summarise(n_off = sum(n)) %>% 
  spread(origin, n_off)

rrs_ci_kalinowski(n_h_off = 2, n_w_off = 20, n_h_par = 162, n_w_par = 191, alpha = 0.05)
```

#### RS vs. field data plots

```{r Examine reproductive success by sample date odd with zeros}
paired_13_15_filter_parents %>% 
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(aes(shape = SEX), height = 0) +
  geom_smooth() +
  labs(title = "Reproductive Success by Parental Sample Date",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Sample Date") +
  ylab("Reproductive Success")

paired_13_15_filter_parents %>% 
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(height = 0) +
  geom_smooth() +
  facet_grid(~SEX)+
  labs(title = "Reproductive Success by Parental Sample Date",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Sample Date") +
  ylab("Reproductive Success")
```

```{r Examine reproductive success by parent length odd with zeros}
paired_13_15_filter_parents %>% 
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(aes(shape = SEX), height = 0) +
  geom_smooth() +
  labs(title = "Reproductive Success by Parental Length (mm)",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Reproductive Success")

paired_13_15_filter_parents %>% 
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(height = 0) +
  geom_smooth() +
  facet_grid(~SEX) +
  labs(title = "Reproductive Success by Parental Length (mm)",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Reproductive Success")
```

## Associations between parents and offspring
Here we briefly explore the heritability of:  

  * Body size
  * Body size x origin
  * Return timing
  * Return timing x origin

### Body size

Because body size is heritable, we compared length between parents and offspring.  

```{r Examine body length Odd}
parents_paired_13_15 %>% 
  filter(`Length Mm.par` != "NA") %>% 
  ggplot(aes(x = `Length Mm.par`, y = `Length Mm.off`, colour = origin, shape = SEX.off))+
  geom_point(size = 2)+
  labs(title = "Offspring Length vs. Parent Length",
       colour = "Parent: Origin",
       shape = "Offspring: Sex") +
  xlab("Parent Length (mm)") +
  ylab("Offspring Length (mm)")
```

Offspring lengths show a great deal of variation. Male offspring tend to be larger than female offspring, as expected.  

```{r Examine body length father Odd}
parents_paired_13_15 %>% 
  filter(`Length Mm.par` != "NA") %>% 
  filter(SEX.par == "M") %>% 
  ggplot(aes(x = `Length Mm.par`, y = `Length Mm.off`, colour = SEX.off))+
  geom_smooth(method = 'lm')+
  geom_point(size = 2)+
  labs(title = "Offspring Length vs. Father Length",
       colour = "Offspring Sex") +
  xlab("Father Length (mm)") +
  ylab("Offspring Length (mm)")
```

```{r Examine body length mother Odd}
parents_paired_13_15 %>% 
  filter(`Length Mm.par` != "NA") %>% 
  filter(SEX.par == "F") %>% 
  ggplot(aes(x = `Length Mm.par`, y = `Length Mm.off`, colour = SEX.off))+
  geom_smooth(method = 'lm')+
  geom_point(size = 2)+
  labs(title = "Offspring Length vs. Mother Length",
       colour = "Offspring Sex") +
  xlab("Mother Length (mm)") +
  ylab("Offspring Length (mm)")
```

```{r Examine body length faceted by sex odd}
parents_paired_13_15 %>% 
  filter(`Length Mm.par` != "NA") %>% 
  ggplot(aes(x = `Length Mm.par`, y = `Length Mm.off`, colour =  SEX.off))+
  facet_grid(~ SEX.par) + 
  geom_smooth(method = 'lm')+
  geom_point(size = 2)+
  labs(title = "Offspring Length vs. Parent Length",
       colour = "Offspring Sex") +
  xlab("Parent Length (mm)") +
  ylab("Offspring Length (mm)")
```

```{r Fit a linear model between parent and offspring length Odd}
fit.lm.length <- lm(`Length Mm.off` ~ `Length Mm.par` + SEX.off, parents_paired_13_15)  
summary(fit.lm.length)
```

There is a statistically significant relationship between parent and offspring length, and there is a significant effect of sex (p=0.008), which is what we observe in the scatterplot above, with males being larger on average than females. 

### Body size x origin

Let's examine the relationship between length and origin in the parent generation to see if there is a size difference between natural and hatchery origin fish. 

```{r Fit a linear model between parent length and origin Odd}
fit.lm.length.origin <- lm(`Length Mm` ~ origin + SEX, paired_13_15_filter %>% filter(`Sample Year` == "2013"))
summary(fit.lm.length.origin)
```

Length does not significantly differ due to origin, but differs by sex (p=0.002) in the parental generation (2013).

```{r Plot relationship between length and origin Odd}
paired_13_15_filter %>% 
  filter(`Sample Year` == "2013") %>% 
  ggplot(aes(x = SEX, y = `Length Mm`, fill = origin)) +
  geom_violin() +
  theme(axis.text = element_text(size = 10)) +
  ggtitle("Distribution of Length by Origin (2013)")
```

Below is a table of average length by origin and sex for 2013 (the parental generation).

```{r Calculate mean lengths of hatchery and natural origin fish}
paired_13_15_filter %>% 
  filter(`Sample Year` == 2013) %>% 
  group_by(origin, SEX) %>% 
  summarise(avg_length = round(mean(`Length Mm`, na.rm = TRUE))) %>% 
  spread(SEX, avg_length)
```

Hatchery males and females are smaller than natural males and females. Both hatchery and natural females are larger than males, but just by a few mm. This is probably not a biologically significant difference.  

### Sample date

Run timing is also highly heritable, so we compared sample date in parents and their offspring. 

```{r make Julian date column}
parents_paired_13_15 <- 
  parents_paired_13_15 %>% 
  mutate(Julian.off = yday(`Sample Date.off`), Julian.par = yday(`Sample Date.par`))

write.csv(parents_paired_13_15, "stock_parents_paired_13_15.csv")
```


```{r Examine sample date Odd}
parents_paired_13_15 %>% 
  ggplot(aes(x = `Sample Date.par`, y = `Sample Date.off`, colour = origin, shape = SEX.off))+
  geom_jitter(height = 0, size = 2) +
    labs(title = "Offspring Date vs. Parent Date",
       colour = "Parent: Origin",
       shape = "Offspring: Sex") +
  xlab("Parent Sample Date") +
  ylab("Offspring Sample Date") 
```

There appears to be a positive association between parent and offspring sample dates. Note that the hatchery parents were collected at the end of the season.  This plot also shows the huge gaps in sampling in 2013. 

```{r Fit lm for parent versus offspring sample date}
lm.date.odd <- lm(parents_paired_13_15$Julian.off ~ parents_paired_13_15$Julian.par + parents_paired_13_15$SEX.off)
summary(lm.date.odd)
```

Offspring sample date has a significant association with parental sample date (p < 0.001), but not offspring sex (p = 0.249). 

### Sample date x origin

Let's compare sample dates across all of the individuals collected in 2013 and 2015 to examine differences in run timing between hatchery and natural origin fish in the parental generation (2013) and look for completeness of sampling throughout the run. 

```{r Compare sample dates by origin for 2013}
paired_13_15_filter <- 
  paired_13_15_filter %>% 
  mutate(Julian = yday(`Sample Date`))

write.csv(paired_13_15_filter, "stock_paired_13_15_filter.csv")


paired_13_15_filter %>% 
  ggplot(aes(x=`Julian`, fill=origin)) +
  geom_histogram(position=position_dodge(), binwidth = 1) +
  facet_grid(`Sample Year` ~ .) +
  labs(title="Number of Samples per Year by Date and Origin")
```

From this histogram, we can see that in the parental generation (2013), natural origin fish tended to have earlier sampling dates than hatchery origin fish, which confirms what we know about differences in run timing. We can also see how sparse the field sampling was in 2013. Offspring in 2015 were sampled throughout the run.

### Length by sample date

We also examined length by sample date for the parental (2013) generation to see if size differences are associated with run timing. 

```{r Plot length by sample date odd}
paired_13_15_filter %>% 
  filter(`Sample Year` == "2013") %>% 
  ggplot(aes(x = `Sample Date`, y = `Length Mm`, color = origin, shape = SEX)) +
  geom_jitter(height = 0) +
  theme(axis.text = element_text(size = 10)) +
  ggtitle("Distribution of Length by Sample Date (2013)")
```

This is a plot of parental sample date by length. Note that it is a jitter plot. 

```{r Fit LM for parental sample date and length}
lm.date.length.odd <- lm(`Length Mm.par`~ `Sample Date.par`, parents_paired_13_15)
summary(lm.date.length.odd)
```

There is a significant relationship between parental sample date and length (p = 0.001). 

```{r Fit LM for parental sample date and length + sex + origin}
lm.date.length.odd <- lm(`Length Mm.par`~ `Sample Date.par` + SEX.par + origin, parents_paired_13_15)
summary(lm.date.length.odd)
```

Sample date and length are significantly associated. Length is not significantly associated with sex or origin. 

```{r Calculate mean length by sample date odd}
parents_paired_13_15 %>% 
  filter(`Sample Year.par` == 2013) %>% 
  group_by(origin, SEX.par, `Sample Date.par`) %>% 
  summarise(avg_length = round(mean(`Length Mm.par`, na.rm = TRUE))) %>% 
  spread(origin, avg_length)
```

These are mean lengths by sex and origin for each sample date. Bigger fish seem to return sooner than smaller fish.  

```{r Length by sample date pooled odd}
parents_paired_13_15 %>% 
  filter(`Sample Year.par` == 2013) %>% 
  group_by(`Sample Date.par`) %>% 
  summarise(avg_length = round(mean(`Length Mm.par`, na.rm = TRUE)))
```


These are mean lengths by sample date with all individuals pooled. Again, larger fish seem to return earlier. 

## Summary of odd lineage results
1. Exclusion probabilities were equal to 1, which means we can be confident in the ability of our marker set to correctly assign parents to offspring.  
2. 72 offspring were assigned to 54 parents (3 of which were hatchery-origin), for an assignment rate of **2.3%**.  
3. Family size varied from 0-2 for females and 0-4 for males.  
4. Hatchery fish appeared later in the run than natural fish. 
