---
title: "Stockdale Results for Science Panel"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(coin)
library(MASS)
library(scales)
library(tidyverse)
# library(leaflet)
library(lubridate)
library(PBSmapping)
library(grid)
library(ggthemes) # Load
```

Below are analyses of parentage results calculated separately for Odd (2013-2015) and Even (2014-2016) lineages using *FRANz*. 

# Odd lineage (2013/2015)

## *FRANz* parameters
*FRANz* was run at 15:35 on February 12, 2019. We used the following parameters:
FRANz.exe --Nmmax 5659 --Nfmax 5659 --femrepro 1:2 --malerepro 1:2 --typingerror 0.005 --updatefreqs --poutformat 2 --fullsibtest "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\stock_postQA_2013_2015_STOCK.dat"

The parameters are defined as follows:
--Nmmax and --Nfmax are the maximum numbers of candidate mothers and fathers. To obtain our values, we used an estimated escapement of 4,150 and divided by 2 (*note: above we forgot to divide by 2*). This escapement number came from the smaller of the two estimates (aerial survey AUC and stream walk AUC). This estimate is listed under 'pop est aerial' in the Excel file found here: "V:\Documents\5_Coastwide\Multispecies\AHRP\Field data\Sample Summary Extract and Genotype.xlsx"   
--femrepro and --malerepro specify the age range in which an individual can reproduce  
--typingerror refers to the overall genotyping error rate. Ours was ~0.005 (for Stockdale was actually 0.006, from QC)
--updatefreqs specifies that *FRANz* should update allele frequencies using MCMC sampling  
--poutformat specifies that all potential parents should be listed, not just the most likely  
--fullsibtest tests for full siblings among offspring   

All output files can be found here: "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\Stockdale\Odd_Run 2"  

The summary file provides information about the power of our marker suite:   
Cumulative exclusion probability when 1 to 7 fullsibs are genotyped  
  First Parent              : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Second Parent             : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Parent Pair               : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000
  
According to the FRANz manual, marker sets are not considered powerful if these cumulative exclusion probabilities are less than 0.95, which indicates that the probability that a random pair of individuals in the population has a 5% chance of having a genotype pair compatible to an offspring genotype. Since all of our probabilities are 1, we can be **confident** in the power of our 298 amplicons to make parent assignments. 

Plot a histogram of *FRANz* parentage posterior probabilities to show robustness of assignments
```{r FRANz_posterior_plot_odd}
parentage_13_15 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Franz/Stockdale/Odd_Run 2/parentage.csv")

parentage_13_15  %>% 
  filter(!is.na(`Parent 1`)) %>% 
  ggplot(aes(x = Posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  ggtitle("Histogram of FRANz posterior probabilities for parentage assignments")
```

All parentage assigments have a posterior probability of 1, which is very robust.

## Sample sizes
Quick look at what our sample sizes are for potential parents (2013) and potential offspring (2015).

```{r odd_sample_sizes}
paired_13_15_filter <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Stockdale/stock_paired_13_15_filter.csv")

paired_13_15_filter %>% 
  count(`Sample Year`, SEX, origin) %>% 
  spread(origin, n, fill = 0)
```

## Proportion test
Let's use a Chi-Square test to determine whether the proportions of offspring assigned to hatchery- and natural-origin parents significantly differ from the proportions of hatchery- and natural-origin parents sampled. 

```{r proportion_odd_data}

# How many spawners from each origin?
paired_13_15_filter %>% 
  filter(`Sample Year` == 2013) %>% 
  count(origin)


```

```{r Chi square test odd}
O.odd <- c("H", "N")
Spawners.odd <- c(163, 811) #  number of hatchery- and natural-origin parents
Offspring.odd <- c(3, 69)  # number of offspring assigned to hatchery- and natural-origin parents

df.odd <- data.frame(Spawners.odd, Offspring.odd)  # made a data.frame because tibbles don't like rownames
row.names(df.odd) <- O.odd

df.odd
chisq.test(df.odd)
```

There is a highly significant difference in the proportions of offspring assigned to hatchery- and natural-origin parents relative to the proportions of potential parents sampled, indicating an under-representation of offspring assigning to hatchery-origin parents. We see that although **`r round(df.odd["H", "Spawners.odd"] / sum(df.odd[, "Spawners.odd"]), 2) * 100`%** of parents genotyped were of hatchery-origin, only **`r round(df.odd["H", "Offspring.odd"] / sum(df.odd[, "Offspring.odd"]), 2) * 100`%** of offspring assigned to hatchery-origin parents.

## Family size
Here, we calculate and plot family size for both natural and hatchery parents.  Family size ranges from 1-4. All assignments were to 1 parent (*single parent-offspring duos*) - there were no 2 parent assignments (*parent pair-offspring trios*). 

Of the 3,093 offspring collected in 2015, 72 were assigned to parents, for an overall offspring assignment rate of **2.3%**. 

#### Calculate RRS including 0's 
We'll calculate RRS including sampled fish from the parental generation to which offspring were not assigned.

```{r calculate RRS odd}
paired_13_15_filter_parents <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Stockdale/stock_paired_13_15_filter_parents.csv")

# Mean RS including 0's
rs_13_15_0 <- paired_13_15_filter_parents %>%
  group_by(origin, SEX) %>% 
  summarise(RS = mean(n, na.rm = TRUE))

rs_13_15_0 %>%
  mutate(RS = round(RS, 2)) %>% 
  spread(SEX, RS)

# Plot histogram of proportion of parents with a given family size
paired_13_15_filter_parents %>% 
  count(SEX, origin, n) %>% 
  group_by(SEX, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  facet_grid(~ SEX) +
  labs(title="Distribution of Family Size by Females (F) and Males (M)\nProportion of Families",
       fill = "Parent: Origin") +
  xlab("Number of Offspring")+
  ylab("Proportion of Parents") 

# Save RS values
F_h_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Hatchery" & SEX == "F") %>% 
  pull(RS)
M_h_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Hatchery" & SEX == "M") %>% 
  pull(RS)
F_n_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Natural" & SEX == "F") %>% 
  pull(RS)
M_n_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Natural" & SEX == "M") %>% 
  pull(RS)
```

```{r Calculate relative reproductive success including 0s odd}
RRS_m_0_13_15 <- M_h_0_13_15/M_n_0_13_15
RRS_f_0_13_15 <- F_h_0_13_15/F_n_0_13_15

RRS_tibble_0_13_15 <- tibble(RRS_m_0_13_15, RRS_f_0_13_15)
round(RRS_tibble_0_13_15, 2)
```

**RRS for males = `r round(RRS_m_0_13_15, 2)`. RRS for females = `r round(RRS_f_0_13_15, 2)`.**

Let's test for significant differences in RRS between male and female hatchery and natural origin fish. We'll perform two different statistical tests:  

  1. Nonparametric permutation test
  2. Parametric negative binomial GLM  
  
```{r RRS statistical testing including 0s odd}
# The format of the data to test was very simple, it was a data.frame called “mydata” with 2 columns
# 1.	nOff = number of offspring per family
# 2.	Origin = “H” or “W” for if the parent was hatchery or natural (wild)
# 

paired_13_15 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Franz/stockdale_postQA_OceanAK_paired_2013_2015_STOCK.csv")

paired_13_15_filter <- paired_13_15 %>% 
  dplyr::select(franz_id, SILLY, `Fish ID`, `DNA Tray Code`, `DNA Tray Well Code`, `Sample Year`, `Sample Date`, SEX, `Length Mm`, `Otolith Mark Present`, `Otolith Mark ID`) %>% 
  mutate(origin = case_when(`Otolith Mark Present` == "NO" ~ "Natural",
                            `Otolith Mark Present` == "YES" ~ "Hatchery")) %>%  # add origin variable
  mutate(origin = factor(origin, c("Natural", "Hatchery")))  # make factor to ensure hatchery != red

paired_13_15_filter_parents <- paired_13_15_filter %>% 
  filter(`Sample Year` == "2013")

parents_paired_13_15 <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Stockdale/stock_parents_paired_13_15.csv")

parents_paired_13_15_count <- parents_paired_13_15 %>% 
  count(Parent_ID, origin, SEX.par, `Sample Year.par`, `Length Mm.par`, `Sample Date.par`) %>% 
  mutate(lineage = dplyr::case_when(`Sample Year.par` %in% c("2013", "2015") ~ "odd",
                                    `Sample Year.par` %in% c("2014", "2016") ~ "even"))

# Join with all parents that produced no progeny
paired_13_15_filter_parents <- left_join(paired_13_15_filter_parents, parents_paired_13_15_count, by = c("franz_id" = "Parent_ID")) %>% 
  mutate(n = replace_na(n, 0)) %>% 
  rename(origin = origin.x) %>% 
  select(-origin.y, -SEX.par, -`Sample Year.par`, -`Length Mm.par`, -`Sample Date.par`)

mydata_M_0 <- paired_13_15_filter_parents %>%
  filter(SEX == "M") 
  

mydata_F_0 <- paired_13_15_filter_parents %>%
  filter(SEX == "F")
  

perm_1tail_pvalue_M_0 <- round(x = as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_M_0, distribution = approximate(B = 10000), alternative = "greater"))), digits = 4)

perm_1tail_pvalue_F_0 <- round(x = as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_F_0, distribution = approximate(B = 10000), alternative = "greater"))), digits = 4)

#And to do a 1-tailed negative binomial GLM

fit_M_0 <- glm.nb(n ~ origin, data = mydata_M_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_M_0 <- round(summary(fit_M_0)$coefficients[2, 4], 4)

fit_F_0 <- glm.nb(n ~ origin, data = mydata_F_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_F_0 <- round(summary(fit_F_0)$coefficients[2, 4], 4)

RRS_0 <- tibble(Sex = c("M", "F", "M", "F"),
                Test = c(rep("Permutation", 2), rep("Negative Binomial GLM", 2)),
                p_value = c(perm_1tail_pvalue_M_0, perm_1tail_pvalue_F_0, nbGLM_1tail_pvalue_M_0, nbGLM_1tail_pvalue_F_0))
RRS_0 %>% 
  spread(Sex, p_value)

```

There is a significant difference in RRS for females, but not males. 

#### Calculate CI's for RRS

We will calculate 95% CI's for RRS based on Kalinowski and Taper 2005 [link] (http://www.nrcresearchpress.com/doi/10.1139/f04-239). This is a maximum likelihood estimate based off of Hinrichsen 2003. This method has been used in other peer reviewed papers such as Ford et al. 2016 [link] (https://www.researchgate.net/publication/309147116_Broodstock_History_Strongly_Influences_Natural_Spawning_Success_in_Hatchery_Steelhead_Oncorhynchus_mykiss#pf14).

```{r RRS_confidence_intervals_function}
rrs_ci_kalinowski <- function(n_h_off, n_w_off, n_h_par, n_w_par, alpha){
  chi_alpha <- qchisq(p = (1 - alpha), df = 1)
  n_off <- sum(c(n_h_off, n_w_off))
  n_par <- sum(c(n_h_par, n_w_par))
  
  rs_h <- n_h_off / n_h_par
  rs_w <- n_w_off / n_w_par
  
  p_h_par <- n_h_par / n_par
  p_w_par <- n_w_par / n_par
  
  rrs_h <- rs_h / rs_w
  rrs_w <- rs_w / rs_w
  rrs_avg <- (rrs_h * p_h_par) + (rrs_w * p_w_par)
  
  rrs_ml <- (n_h_off * log(p_h_par * rrs_h / rrs_avg)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg))
  
  xi_dist <- bind_rows(
    lapply(seq(0.01, 5, by = 0.01), function(rrs_h_xi) {
      rrs_avg_xi <- (rrs_h_xi * p_h_par) + (rrs_w * p_w_par)
      tibble(rrs_crit = rrs_h_xi,
             logl = (n_h_off * log(p_h_par * rrs_h_xi / rrs_avg_xi)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg_xi)) - (rrs_ml - chi_alpha / 2)
      )
    } )
  )
  
  rrs_min <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit < rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  rrs_max <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit > rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  xi_plot <- xi_dist %>% 
    ggplot(aes(x = rrs_crit, y = logl)) +
    geom_line() +
    geom_hline(yintercept = 0, colour = "red", lwd = 2) +
    geom_vline(xintercept = c(rrs_h, rrs_min, rrs_max), colour = "blue") +
    ylim(c(-5, 5)) +
    xlim(c(0, 2)) +
    ylab("Log Likelihood - Chi Sq Value") +
    annotate("text", x = rrs_h + 0.1, y = xi_dist %>% filter(rrs_crit == xi_dist$rrs_crit[which.min(abs(xi_dist$rrs_crit  - rrs_h))]) %>% pull(logl) + 0.4, label = round(rrs_h, 2)) +
    annotate("text", x = rrs_min - 0.1, y = xi_dist %>% filter(rrs_crit == rrs_min) %>% pull(logl) + 0.4, label = rrs_min) +
    annotate("text", x = rrs_max + 0.1, y = xi_dist %>% filter(rrs_crit == rrs_max) %>% pull(logl) + 0.4, label = rrs_max)
  
  print(xi_plot)
  return(c(rrs_min, rrs_h, rrs_max))
}
```

##### Odd 2013/2015

Calculate 95% CIs for Stockdale 13-15 females. RRS = 0.13 (0.01-0.59)

```{r RRS_confidence_intervals_female_13_15}
# Females
# Get parent sample sizes
paired_13_15_filter_parents %>% 
  filter(SEX == "F") %>% 
  count(SEX, origin) %>% 
  spread(origin, nn)

# Get offspring sample sizes
paired_13_15_filter_parents %>% 
  filter(SEX == "F") %>% 
  group_by(SEX, origin) %>% 
  summarise(n_off = sum(n)) %>% 
  spread(origin, n_off)

rrs_ci_kalinowski(n_h_off = 1, n_w_off = 33, n_h_par = 94, n_w_par = 393, alpha = 0.05)
```

Calculate 95% CIs for Stockdale 13-15 males. RRS = 0.34 (0.05-1.10)

```{r RRS_confidence_intervals_male_13_15}
# Males
# Get parent sample sizes
paired_13_15_filter_parents %>% 
  filter(SEX == "M") %>% 
  count(SEX, origin) %>% 
  spread(origin, nn)

# Get offspring sample sizes
paired_13_15_filter_parents %>% 
  filter(SEX == "M") %>% 
  group_by(SEX, origin) %>% 
  summarise(n_off = sum(n)) %>% 
  spread(origin, n_off)

rrs_ci_kalinowski(n_h_off = 2, n_w_off = 36, n_h_par = 69, n_w_par = 418, alpha = 0.05)
```

#### RS vs. field data plots

```{r Examine reproductive success by sample date odd with zeros}
paired_13_15_filter_parents %>% 
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(aes(shape = SEX), height = 0) +
  geom_smooth() +
  labs(title = "Reproductive Success by Parental Sample Date",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Sample Date") +
  ylab("Reproductive Success")

paired_13_15_filter_parents %>% 
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(height = 0) +
  geom_smooth() +
  facet_grid(~SEX)+
  labs(title = "Reproductive Success by Parental Sample Date",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Sample Date") +
  ylab("Reproductive Success")
```


```{r Examine reproductive success by parent length odd with zeros}
paired_13_15_filter_parents %>% 
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(aes(shape = SEX), height = 0) +
  geom_smooth() +
  labs(title = "Reproductive Success by Parental Length (mm)",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Reproductive Success")

paired_13_15_filter_parents %>% 
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(height = 0) +
  geom_smooth() +
  facet_grid(~SEX) +
  labs(title = "Reproductive Success by Parental Length (mm)",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Reproductive Success")
```

### Sample date x origin

Let's compare sample dates across all of the individuals collected in 2013 and 2015 to examine differences in run timing between hatchery and natural origin fish in the parental generation (2013) and look for completeness of sampling throughout the run. 

```{r Compare sample dates by origin for 2013}

paired_13_15_filter %>% 
  ggplot(aes(x=`Julian`, fill=origin)) +
  geom_histogram(position=position_dodge(), binwidth = 1) +
  facet_grid(`Sample Year` ~ .) +
  labs(title="Number of Samples per Year by Date and Origin")
```

From this histogram, we can see that in the parental generation (2013), natural origin fish tended to have earlier sampling dates than hatchery origin fish, which confirms what we know about differences in run timing. We can also see how sparse the field sampling was in 2013. Offspring in 2015 were sampled throughout the run.

## Summary of odd lineage results
1. Exclusion probabilities were equal to 1, which means we can be confident in the ability of our marker set to correctly assign parents to offspring.  
2. 72 offspring were assigned to 54 parents (3 of which were hatchery-origin), for an assignment rate of **2.3%**.  
3. Family size varied from 0-2 for females and 0-4 for males.  
4. Hatchery fish appeared later in the run than natural fish. 


# Even lineage (2014/2016)

## *FRANz* parameters
*FRANz* was run at 11:00 on February 13, 2019. We used the following parameters:
FRANz.exe --Nmmax 2019 --Nfmax 2019 --femrepro 1:2 --malerepro 1:2 --typingerror 0.006 --updatefreqs --poutformat 2 --fullsibtest "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\stock_postQA_2014_2016_STOCK.dat"

The parameters are defined as follows:
--Nmmax and --Nfmax are the maximum numbers of candidate mothers and fathers. To obtain our values, we used an estimated escapement of 4,150 and divided by 2 (*note: above we forgot to divide by 2 AND this is the odd-year parameter, not the even-year*). This escapement number came from the smaller of the two estimates (aerial survey AUC and stream walk AUC). This estimate is listed under 'pop est aerial' in the Excel file found here: "V:\Documents\5_Coastwide\Multispecies\AHRP\Field data\Sample Summary Extract and Genotype.xlsx"   
--femrepro and --malerepro specify the age range in which an individual can reproduce  
--typingerror refers to the overall genotyping error rate. Ours was ~0.005 (for Stockdale was actually 0.006, from QC)
--updatefreqs specifies that *FRANz* should update allele frequencies using MCMC sampling  
--poutformat specifies that all potential parents should be listed, not just the most likely  
--fullsibtest tests for full siblings among offspring   

All output files can be found here: "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\Stockdale\Even_Run 2"  

The summary file provides information about the power of our marker suite:   
Cumulative exclusion probability when 1 to 7 fullsibs are genotyped  
  First Parent              : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Second Parent             : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Parent Pair               : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000
  
According to the FRANz manual, marker sets are not considered powerful if these cumulative exclusion probabilities are less than 0.95, which indicates that the probability that a random pair of individuals in the population has a 5% chance of having a genotype pair compatible to an offspring genotype. Since all of our probabilities are 1, we can be **confident** in the power of our 298 amplicons to make parent assignments.  

```{r Setup Even, message=FALSE, warning=FALSE}
parentage_14_16 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Franz/Stockdale/Even_Run 2/parentage.csv")
```

Plot a histogram of *FRANz* parentage posterior probabilities to show robustness of assignments
```{r FRANz_posterior_plot_even}
parentage_14_16  %>% 
  filter(!is.na(`Parent 1`)) %>% 
  ggplot(aes(x = Posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  ggtitle("Histogram of FRANz posterior probabilities for parentage assignments")
```

All parentage assigments have a posterior probability of 1, which is very robust.

## Sample sizes
Quick look at what our sample sizes are for potential parents (2014) and potential offspring (2016).

```{r even_sample_sizes}

paired_14_16_filter <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Stockdale/stock_paired_14_16_filter.csv")

paired_14_16_filter %>% 
  count(`Sample Year`, SEX, origin) %>% 
  spread(origin, n, fill = 0)
```

**Note** that *FRANz* assigned 1054 offspring to 405 parents. 865 of these offsrping have natural-origin parents and 373 offspring have hatchery-origin parents. The hatchery origin fish are from AFK12B, CCH12, and WNH12PINKB.

## Proportion test
Let's use a Chi-Square test to determine whether the proportions of offspring assigned to hatchery- and natural-origin parents significantly differ from the proportions of hatchery- and natural-origin parents sampled. 

```{r proportion_even_data}
# How many spawners from each origin?
paired_14_16_filter %>% 
  filter(`Sample Year` == 2014) %>% 
  count(origin)


```

```{r Chi square test even}
O.even <- c("H", "N")
Spawners.even <- c(436, 358) #  number of hatchery- and natural-origin parents
Offspring.even <- c(373, 865)  # number of offspring assigned to hatchery- and natural-origin parents

df.even <- data.frame(Spawners.even, Offspring.even)  # made a data.frame because tibbles don't like rownames
row.names(df.even) <- O.even

df.even
chisq.test(df.even)
```

There is a highly significant difference in the proportions of offspring assigned to hatchery- and natural-origin parents relative to the proportions of potential parents sampled, indicating an under-representation of offspring assigning to hatchery-origin parents. We see that although **`r round(df.even["H", "Spawners.even"] / sum(df.even[, "Spawners.even"]), 2) * 100`%** of parents genotyped were of hatchery-origin, only **`r round(df.even["H", "Offspring.even"] / sum(df.even[, "Offspring.even"]), 2) * 100`%** of offspring assigned to hatchery-origin parents.

## Family size
Here, we calculate and plot family size for both natural and hatchery parents.  Family size ranges from 1-19. Most assignments were to 1 parent (*single parent-offspring duos*) - there were 183 2 parent assignments (*parent pair-offspring trios*). 

Of the 5,199 offspring collected in 2016, 1,054 were assigned to parents, for an overall offspring assignment rate of **20.3%**. 

### Single parent-offspring pairs
This plot shows the distribution of family size by sex (number of offspring assigned to a female parent or a male parent) for both hatchery and natural parents.  

When we plot the sexes separately, we can see that most of the assigned parents are males. Female parents produced 1-12 offspring, while male parents produced 1-19.  

Below, we'll calculate reproductive success of natural and hatchery females and males. 

#### Calculate RRS again including 0's 
We'll calculate RRS again including sampled fish from the parental generation to which offspring were not assigned.

```{r Re-calculate RRS even}

paired_14_16_filter_parents <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Stockdale/stock_paired_14_16_filter_parents.csv")

# Mean RS including 0's
rs_14_16_0 <- paired_14_16_filter_parents %>%
  group_by(origin, SEX) %>% 
  summarise(RS = mean(n, na.rm = TRUE))

rs_14_16_0 %>%
  mutate(RS = round(RS, 2)) %>% 
  spread(SEX, RS)

# Plot histogram of number of parents with a given family size
#paired_14_16_filter_parents %>% 
#  ggplot(aes(x = n, fill = origin)) +
#  geom_bar(position = position_dodge2(preserve = "single")) +
#  facet_grid( ~ SEX) +
#  xlab("Number of Offspring") +
#  ylab("Number of Parents") +
#  labs(title = "Distribution of Family Size by Females (F) and Males (M)\nNumber of Families",
#       fill = "Parent: Origin")

# Plot histogram of proportion of parents with a given family size
paired_14_16_filter_parents %>% 
  count(SEX, origin, n) %>% 
  group_by(SEX, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  geom_col(position = position_dodge2(preserve="single")) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  facet_grid(~ SEX) +
  labs(title="Distribution of Family Size by Females (F) and Males (M)\nProportion of Families",
       fill = "Parent: Origin") +
  xlab("Number of Offspring")+
  ylab("Proportion of Parents") 

# Save RS values
F_h_0_14_16 <- rs_14_16_0 %>% 
  filter(origin == "Hatchery" & SEX == "F") %>% 
  pull(RS)
M_h_0_14_16 <- rs_14_16_0 %>% 
  filter(origin == "Hatchery" & SEX == "M") %>% 
  pull(RS)
F_n_0_14_16 <- rs_14_16_0 %>% 
  filter(origin == "Natural" & SEX == "F") %>% 
  pull(RS)
M_n_0_14_16 <- rs_14_16_0 %>% 
  filter(origin == "Natural" & SEX == "M") %>% 
  pull(RS)
```

```{r Calculate relative reproductive success including 0s even}
RRS_m_0_14_16 <- M_h_0_14_16/M_n_0_14_16
RRS_f_0_14_16 <- F_h_0_14_16/F_n_0_14_16

RRS_tibble_0_14_16 <- tibble(RRS_m_0_14_16, RRS_f_0_14_16)
round(RRS_tibble_0_14_16, 2)
```

**RRS for males = `r round(RRS_m_0_14_16, 2)`. RRS for females = `r round(RRS_f_0_14_16, 2)`.**

Let's test for significant differences in RRS between male and female hatchery and natural origin fish. We'll perform two different statistical tests:  

  1. Nonparametric permutation test
  2. Parametric negative binomial GLM  

```{r RRS statistical testing including 0s even}
# The format of the data to test was very simple, it was a data.frame called “mydata” with 2 columns
# 1.	nOff = number of offspring per family
# 2.	Origin = “H” or “W” for if the parent was hatchery or natural (wild)
# 

paired_14_16 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Franz/stockdale_postQA_OceanAK_paired_2014_2016_STOCK.csv")

paired_14_16_filter <- paired_14_16 %>% 
  dplyr::select(franz_id, SILLY, `Fish ID`, `DNA Tray Code`, `DNA Tray Well Code`, `Sample Year`, `Sample Date`, SEX, `Length Mm`, `Otolith Mark Present`, `Otolith Mark ID`) %>% 
  mutate(origin = case_when(`Otolith Mark Present` == "NO" ~ "Natural",
                            `Otolith Mark Present` == "YES" ~ "Hatchery")) %>%  # add origin variable
  mutate(origin = factor(origin, c("Natural", "Hatchery")))  # make factor to ensure hatchery != red

paired_14_16_filter_parents <- paired_14_16_filter %>% 
  filter(`Sample Year` == "2014")

parents_paired_14_16 <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Stockdale/stock_parents_paired_14_16.csv")

parents_paired_14_16_count <- parents_paired_14_16 %>% 
  count(Parent_ID, origin, SEX.par, `Sample Year.par`, `Length Mm.par`, `Sample Date.par`) %>% 
  mutate(lineage = dplyr::case_when(`Sample Year.par` %in% c("2013", "2015") ~ "odd",
                                    `Sample Year.par` %in% c("2014", "2016") ~ "even"))

# Join with all parents that produced no progeny
paired_14_16_filter_parents <- left_join(paired_14_16_filter_parents, parents_paired_14_16_count, by = c("franz_id" = "Parent_ID")) %>% 
  mutate(n = replace_na(n, 0)) %>% 
  rename(origin = origin.x) %>% 
  select(-origin.y, -SEX.par, -`Sample Year.par`, -`Length Mm.par`, -`Sample Date.par`)

mydata_M_0 <- paired_14_16_filter_parents %>%
  filter(SEX == "M") %>% 
 mutate(origin = as.factor(origin))

mydata_F_0 <- paired_14_16_filter_parents %>%
  filter(SEX == "F") %>% 
   mutate(origin = as.factor(origin))

perm_1tail_pvalue_M_0 <- round(x = as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_M_0, distribution = approximate(B = 10000), alternative = "greater"))), digits = 4)

perm_1tail_pvalue_F_0 <- round(x = as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_F_0, distribution = approximate(B = 10000), alternative = "greater"))), digits = 4)

#And to do a 1-tailed negative binomial GLM

fit_M_0 <- glm.nb(n ~ origin, data = mydata_M_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_M_0 <- round(summary(fit_M_0)$coefficients[2, 4], 4)

fit_F_0 <- glm.nb(n ~ origin, data = mydata_F_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_F_0 <- round(summary(fit_F_0)$coefficients[2, 4], 4)

RRS_0 <- tibble(Sex = c("M", "F", "M", "F"),
                Test = c(rep("Permutation", 2), rep("Negative Binomial GLM", 2)),
                p_value = c(perm_1tail_pvalue_M_0, perm_1tail_pvalue_F_0, nbGLM_1tail_pvalue_M_0, nbGLM_1tail_pvalue_F_0))
RRS_0 %>% 
  spread(Sex, p_value)

```

There is a significant difference in RRS for both males and females. 

#### Calculate CI's for RRS

We will calculate 95% CI's for RRS based on Kalinowski and Taper 2005 [link] (http://www.nrcresearchpress.com/doi/10.1139/f04-239). This is a maximum likelihood estimate based off of Hinrichsen 2003. This method has been used in other peer reviewed papers such as Ford et al. 2016 [link] (https://www.researchgate.net/publication/309147116_Broodstock_History_Strongly_Influences_Natural_Spawning_Success_in_Hatchery_Steelhead_Oncorhynchus_mykiss#pf14).

```{r RRS_confidence_intervals_function_even}
rrs_ci_kalinowski <- function(n_h_off, n_w_off, n_h_par, n_w_par, alpha){
  chi_alpha <- qchisq(p = (1 - alpha), df = 1)
  n_off <- sum(c(n_h_off, n_w_off))
  n_par <- sum(c(n_h_par, n_w_par))
  
  rs_h <- n_h_off / n_h_par
  rs_w <- n_w_off / n_w_par
  
  p_h_par <- n_h_par / n_par
  p_w_par <- n_w_par / n_par
  
  rrs_h <- rs_h / rs_w
  rrs_w <- rs_w / rs_w
  rrs_avg <- (rrs_h * p_h_par) + (rrs_w * p_w_par)
  
  rrs_ml <- (n_h_off * log(p_h_par * rrs_h / rrs_avg)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg))
  
  xi_dist <- bind_rows(
    lapply(seq(0.01, 5, by = 0.01), function(rrs_h_xi) {
      rrs_avg_xi <- (rrs_h_xi * p_h_par) + (rrs_w * p_w_par)
      tibble(rrs_crit = rrs_h_xi,
             logl = (n_h_off * log(p_h_par * rrs_h_xi / rrs_avg_xi)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg_xi)) - (rrs_ml - chi_alpha / 2)
      )
    } )
  )
  
  rrs_min <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit < rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  rrs_max <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit > rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  xi_plot <- xi_dist %>% 
    ggplot(aes(x = rrs_crit, y = logl)) +
    geom_line() +
    geom_hline(yintercept = 0, colour = "red", lwd = 2) +
    geom_vline(xintercept = c(rrs_h, rrs_min, rrs_max), colour = "blue") +
    ylim(c(-5, 5)) +
    xlim(c(0, 2)) +
    ylab("Log Likelihood - Chi Sq Value") +
    annotate("text", x = rrs_h + 0.1, y = xi_dist %>% filter(rrs_crit == xi_dist$rrs_crit[which.min(abs(xi_dist$rrs_crit  - rrs_h))]) %>% pull(logl) + 0.4, label = round(rrs_h, 2)) +
    annotate("text", x = rrs_min - 0.1, y = xi_dist %>% filter(rrs_crit == rrs_min) %>% pull(logl) + 0.4, label = rrs_min) +
    annotate("text", x = rrs_max + 0.1, y = xi_dist %>% filter(rrs_crit == rrs_max) %>% pull(logl) + 0.4, label = rrs_max)
  
  print(xi_plot)
  return(c(rrs_min, rrs_h, rrs_max))
}
```

##### Even 2014/2016

Calculate 95% CIs for Stockdale 14-16 females. RRS = 0.42 (0.35-0.50)

```{r RRS_confidence_intervals_female_14_16}
# Females
# Get parent sample sizes
paired_14_16_filter_parents %>% 
  filter(SEX == "F") %>% 
  count(SEX, origin) %>% 
  spread(origin, nn)

# Get offspring sample sizes
paired_14_16_filter_parents %>% 
  filter(SEX == "F") %>% 
  group_by(SEX, origin) %>% 
  summarise(n_off = sum(n)) %>% 
  spread(origin, n_off)

rrs_ci_kalinowski(n_h_off = 196, n_w_off = 448, n_h_par = 230, n_w_par = 221, alpha = 0.05)
```

Calculate 95% CIs for Hogan 14-16 males. RRS = 0.28 (0.24-0.34)

```{r RRS_confidence_intervals_male_14_16}
# Males
# Get parent sample sizes
paired_14_16_filter_parents %>% 
  filter(SEX == "M") %>% 
  count(SEX, origin) %>% 
  spread(origin, nn)

# Get offspring sample sizes
paired_14_16_filter_parents %>% 
  filter(SEX == "M") %>% 
  group_by(SEX, origin) %>% 
  summarise(n_off = sum(n)) %>% 
  spread(origin, n_off)

rrs_ci_kalinowski(n_h_off = 177, n_w_off = 417, n_h_par = 206, n_w_par = 137, alpha = 0.05)
```

#### RS vs. field data plots

Natural-origin males and females both do better early in the season (density-dependence or increased fishing pressure later in the season?).
```{r Examine reproductive success by sample date even with zeros}
paired_14_16_filter_parents %>% 
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(aes(shape = SEX), height = 0) +
  geom_smooth() +
  labs(title = "Reproductive Success by Parental Sample Date",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Sample Date") +
  ylab("Reproductive Success")

paired_14_16_filter_parents %>% 
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(height = 0) +
  geom_smooth() +
  facet_grid(SEX ~ .)+
  labs(title = "Reproductive Success by Parental Sample Date",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Sample Date") +
  ylab("Reproductive Success")
```

Larger males are much more successful (but only natural-origin males).
```{r Examine reproductive success by parent length even with zeros}
paired_14_16_filter_parents %>% 
  filter(`Length Mm` > 300) %>% 
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(aes(shape = SEX), height = 0) +
  geom_smooth() +
  labs(title = "Reproductive Success by Parental Length (mm)",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Reproductive Success")

paired_14_16_filter_parents %>% 
  filter(`Length Mm` > 300) %>% 
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(height = 0) +
  geom_smooth() +
  facet_grid(SEX ~ .) +
  labs(title = "Reproductive Success by Parental Length (mm)",  colour = "Parent: Origin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Reproductive Success")
```

### Parent pair-offspring trios

Because we have 183 individuals who were assigned to 2 parents, we are able to examine instances of crosses between hatchery fish, between natural fish, and between natural and hatchery fish. **Note** that we have divided natural-hatchery crosses into "HN", which indicates a hatchery origin mother and natural origin father and "NH", which indicates a natural origin mother and hatchery origin father. 

```{r Create cross as new grouping variable even, message=FALSE}
# Which offspring have 2 parents?
offspring_trio_14_16 <- parents_paired_14_16 %>% 
  filter(Parent == "Parent 2") %>% 
  pull(Offspring)

parents_paired_14_16_cross <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Stockdale/stock_parents_paired_14_16_cross.csv")

parents_paired_14_16_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  count(cross, n) %>% 
  complete(cross, n, fill = list(nn = 0)) %>% 
  ggplot(aes(x = n, y = nn, fill = cross))+
  geom_col(position = position_dodge2(preserve = "single"), width = 0.5) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = 1:5) +
  # ylim(0, 40) +
  # xlim(0, 5) +
  labs(title = "Distribution of Family Size by Cross") +
  xlab("Number of Offspring")+
  ylab("Number of Families") 
```

Offspring from all four types of crosses are represented in our dataset. HH crosses produced 1-2 offspring, HN crosses produced 1-5 offspring, NH crosses produced 1-3 offspring, and NN crosses produced 1-5 offspring.

Calculate RS of cross types
```{r cross_rs_even}
parents_paired_14_16_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  group_by(cross) %>% 
  summarise(rs = mean(n))
```

Is this significant?
```{r RRS_confidence_intervals_14_16_setup}
# How many matings by cross type?
parents_paired_14_16_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  count(cross)

# Get offspring sample sizes
parents_paired_14_16_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  group_by(cross) %>% 
  summarise(nn = sum(n))
```

```{r RRS_confidence_intervals_14_16}
# HH vs. NN
rrs_ci_kalinowski(n_h_off = 16, n_w_off = 109, n_h_par = 15, n_w_par = 62, alpha = 0.05) # Significant!

# HN vs. NN
rrs_ci_kalinowski(n_h_off = 31, n_w_off = 109, n_h_par = 20, n_w_par = 62, alpha = 0.05) # NS

# NH vs. NN
rrs_ci_kalinowski(n_h_off = 27, n_w_off = 109, n_h_par = 18, n_w_par = 62, alpha = 0.05) # NS
```

Investigate the potential for multiple matings
```{r multiple_matings_even}
parents_paired_14_16_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  group_by(`Fish ID`) %>% 
  summarise(n_mates = n_distinct(`Fish ID.par1`)) %>% 
  count(n_mates)

parents_paired_14_16_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  group_by(`Fish ID.par1`) %>% 
  summarise(n_mates = n_distinct(`Fish ID`)) %>% 
  count(n_mates)

```


### Sample date x origin

Let's compare sample dates across all of the individuals collected in 2014 and 2016 to examine differences in run timing between hatchery and natural origin fish in the parental generation (2014) and look for completeness of sampling throughout the run. 

```{r Compare sample dates by origin for 2014}


paired_14_16_filter %>% 
  ggplot(aes(x=`Sample Date`, fill=origin)) +
  geom_histogram(position=position_dodge(), binwidth = 1) +
  facet_grid(`Sample Year` ~ .) +
  labs(title="Number of Samples per Year by Date and Origin")
```

From this histogram, we can see that in the parental generation (2014), natural origin fish tended to have earlier sampling dates than hatchery origin fish, which confirms what we know about differences in run timing. We can also see how sparse the field sampling was in 2014. Offspring in 2016 were sampled throughout the run.

## Summary of even lineage results
1. Exclusion probabilities were equal to 1, which means we can be confident in the ability of our marker set to correctly assign parents to offspring.  
2. 1,059 offspring were assigned to 405 parents, for an assignment rate of **20.3%**. 
3. 183 offspring assigned to 2 parents (triads).
4. RRS was significant for both males (0.28) and females (0.42).
5. RRS was significant for HH vs. NN crosses (0.61).
6. Hybrids (HN and NH) were intermediate between HH and NN crosses.
7. RS significantly varied with parent sample date (early fish do best, middle fish worst, late returning okay).
8. Rs significantly varied with parent size (large males had more offspring than smaller males).
9. Family size varied from 0-12 for females and 0-19 for males.  
10. Hatchery fish appeared later in the run than natural fish. 

