---
title: "Stockdale Sampling Questions Revisited"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

# Pro Tips {#protip}

Anything in [red](#protip) is a hyperlink that will either take you to the relevant location in this R Notebook .html file, or will link to an external reference (i.e. a publication we cite). If you just want to see the text and figures and **do not** want to see any of the `code` used to generate this R Notebook, click on the grey box in the upper right corner that says "Code" and select "Hide All Code". Happy data exploring!

# Introduction

The purpose of these analyses is to address Science Panel questions regarding the representativeness of sampling relative to the escapement for Stockdale Creek (Stockdale). The implicit question here is whether our samples are biased in any way that would affect how fitness results are extrapolated to represent the population. Specifically, we will investigate three areas:

  1) What proportion of the Stockdale spawning population in 2014 are represented by our samples?
  2) Provide a single RRS estimate for Stockdale 2014/2016, weighted by sex.
  3) Test samples for differences in reproductive success (RS) stratified by sample date, size, and location.

For the third area, we will do some naive tests, considering each variable separately and binning them into somewhat arbitrary categories. In another, soon to follow notebook, we will consider all information more holistically by model fitting to generalized linear models (GLM’s) as is commonly done in the literature (see [Bernston et al. 2011](https://afspubs.onlinelibrary.wiley.com/doi/pdf/10.1080/00028487.2011.584489), [Ford et al. 2012](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1755-263X.2012.00261.x), and [Janowitz-Koch et al. 2018](https://onlinelibrary.wiley.com/doi/full/10.1111/eva.12725)).

**Note** that we are only doing these analyses for the Stockdale even-year lineage (2014-2016) because there are not enough parent-offspring assignments from the odd lineage (2013-2015) to draw robust conclusions. Additionally, we are investigating Stockdale as opposed to Hogan Bay (Hogan) because Stockdale has a larger sample size of parent-offspring pairs and Stockdale is a longer system (~1500m of spawning habitat), so spawning location is likely more important in Stockdale than in Hogan (only 600m of spawning habitat).

**Methods Note:** All pedigree analyses and RRS results to date have been based on output from the computer program [*FRANz*](https://academic.oup.com/bioinformatics/article/25/16/2134/203763). In this notebook we will also introduce pedigree results from another computer program, [*Colony*](https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1755-0998.2009.02787.x). Both pedigree analysis programs are used extensively in the literature and have been shown to produce highly consistent results, so long as the genetic markerset is informative enough. We introduce results from *Colony* here because both the hypergeometric model trans-generational mark-recapture (tGMR) and trans-generational rarefaction curve (tGRC) methods from [Rawding et al. 2014](https://afspubs.onlinelibrary.wiley.com/doi/pdf/10.1080/00028487.2013.829122) require an estimate of the number of *unsampled*, *inferred* parents (i.e. how many unique parents would it have taken to create all of the offspring in our sample)

```{r setup, include=FALSE}
library(coin)
library(MASS)
library(scales)
library(tidyverse)
library(leaflet)
library(lubridate)
library(ggExtra)
library(gridExtra)
library(MuMIn)
library(GGally)
library(gganimate)

knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.width = 10)
```

# What Was the Sample Proportion of Stockdale 2014 Escapement?

In order to assess what proportion of the 2014 escapement is represented by our samples, we need our sample size and an estimate of the escapement. Sample size is the number of fish sampled in 2014 that were successfully genotyped and passed our genotype QA process. Here we provide four different estimates of escapement, all of which have slightly different assumptions.

## Sample Sizes

Here is the sample size of potential parents by origin and sex. These are fish sampled in 2014 that were successfully genotyped and passed our genotype QA process. The figure shows how many samples of each origin were collected by sample date for 2014 (parents) and 2016 (offspring). Total sample size for 2014 parents is 794 fish.
```{r}
paired_14_16_filter <- read_csv("stock_paired_14_16_filter.csv") %>% 
  mutate(Origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(Sex = case_when(SEX == "M" ~ "Male", SEX == "F" ~ "Female"))

paired_14_16_filter %>% 
  filter(`Sample Year` == 2014) %>% 
  count(`Sample Year`, Origin, Sex) %>% 
  spread(Sex, n)

paired_14_16_filter %>% 
  ggplot(aes(x = `Julian`)) +
  theme_bw() +
  geom_bar(aes(fill = Origin), colour = "black") +
  facet_grid(`Sample Year` ~ .) +
  geom_hline(yintercept = 0, colour = "black") +
  ylab("Count") +
  xlab("Julian Date") +
  labs(title= "Number of Samples per Year\nby Date and Origin",
       fill = "Origin") +
  theme(text = element_text(size = 20))
```

## Escapement Estimates

### Aerial Survey AUC (Area Under the Curve)

Below is the very sparse aerial survey data for Stockdale in 2014.
```{r}
auc <- readxl::read_xlsx(path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Field data/StreamSurveys 10-28-14 Ped only to Kyle 11-7-14_Updated_SM_2-25-2015.xlsm", sheet = "Raw_Data", skip = 4)

(stockdale_aerial <- auc %>% 
  filter(`ADF&G Stream Name` == "Stockdale Creek", Year == 2014, `Count Type` == "Aerial") %>% 
  select(Distict, `AWC Stream Number`, `ADF&G Stream Number`, `ADF&G Stream Name`, Year, `Survey Date`, `Survey Live Stream Count`, `Count Type`))

stockdale_aerial %>% 
  mutate(Julian = yday(`Survey Date`)) %>% 
  ggplot(aes(x = `Julian`, y = `Survey Live Stream Count`)) +
  theme_bw() +
  geom_col(colour = "black") +
  facet_grid(Year ~ .) +
  geom_hline(yintercept = 0, colour = "black") +
  ylab("Live Count") +
  xlab("Julian Date") +
  labs(title= "Aerial Survey 2014") +
  theme(text = element_text(size = 20))
```

Applying these three surveys to an Area Under the Curve (AUC) model with a stream life of 13.7 days and an observer efficiency of 43.6% [Fried et al. 1998] we get an estimated escapement of **2,869** fish.

### Stream Walk AUC

Below is the more comprehensive foot survey data for Stockdale in 2014.
```{r}
(stockdale_foot <- auc %>% 
  filter(`ADF&G Stream Name` == "Stockdale Creek", Year == 2014, `Count Type` == "Stream Walks") %>% 
  select(Distict, `AWC Stream Number`, `ADF&G Stream Number`, `ADF&G Stream Name`, Year, `Survey Date`, `Survey Live Stream Count`, `Count Type`))

stockdale_foot %>% 
  mutate(Julian = yday(`Survey Date`)) %>% 
  ggplot(aes(x = `Julian`, y = `Survey Live Stream Count`)) +
  theme_bw() +
  geom_col(colour = "black") +
  facet_grid(Year ~ .) +
  geom_hline(yintercept = 0, colour = "black") +
  ylab("Live Count") +
  xlab("Julian Date") +
  labs(title= "Foot Survey 2014") +
  theme(text = element_text(size = 20))
```

Applying these fourteen surveys to an Area Under the Curve (AUC) model with a stream life of 13.7 days and an observer efficiency of 70.3% [Fried et al. 1998] we get an estimated escapement of **4,038** fish.

### tGMR: Binomial Model

Another method of independently estimating the 2014 escapement to Stockdale Creek is to use trans-generational mark-recapture as detailed in [Rawding et al. 2014](https://afspubs.onlinelibrary.wiley.com/doi/pdf/10.1080/00028487.2013.829122). Briefly, marks are the genotyped potential parents, and the second event is genotyped potential offspring, with offspring assigning to parents representing recaptures. This model takes two forms, binomial (with replacement) and hypergeometric (without replacement). Here we provide the binomial estimate where:  

  * n1 = number of genotyped potential parents (2014)
  * n2 = number of genotypes from potential offspring (2016) = number of offspring x 2 (one copy from sire, one copy from dam)
  * m2 = number of recaptured parental genotypes (i.e. offspring that assign to parents)

Here is the sample size of individuals for potential parents (2014) and potential offspring (2016)
```{r}
read_csv("../Franz/stockdale_postQA_OceanAK_paired_2014_2016_STOCK.csv") %>% 
  count(`Sample Year`)
```

So that means that:  

  * n1 = 794
  * n2 = 5,199 x 2 = 10,398

And how many "recaptured genotypes" did we get in 2016?
```{r}
offspring_dat <- read_csv("../Franz/stockdale_postQA_OceanAK_paired_2014_2016_STOCK.csv") %>% 
  filter(`Sample Year` == 2016)

(offspring_franz <- read_csv("../../Franz/Stockdale/Even_Run 2/parentage.csv") %>% 
    filter(Posterior >= 0.9) %>% 
    left_join(offspring_dat, by = c("Offspring" = "franz_id")) %>% 
    select(Offspring, `Parent 1`, `Parent 2`) %>% 
    gather(Parent, Parent_ID, -Offspring) %>% 
    filter(!is.na(Parent_ID)))
```

We had 1,238 parent-offspring pair assignments, so that makes:  

  * m2 = 1,238

And our final calculation for Nc is:
```{r}
round(794 * 5199 * 2 / 1238)
```

Giving us an estimated escapement of **6,669** fish

### tGMR: Hypergeometric Model

**Note:** the previous version of this R Notebook that was sent out by Chris on Wednesday, August 28, 2019 did not include this method as it requires an estimate of the total number of parents it would have taken to create all of the offspring in our sample (i.e. *inferred* parents). Having an estimate of the number of unsampled, *inferred* parents is also required for the rarefaction curve method (tGRC). *Inferred* parents come from the maximum likelihood configuration pedigree from [*Colony*](https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1755-0998.2009.02787.x).

Here we provide the hypergeometric tGMR model (without replacement) estimates where:  

  * n1 = number of genotyped potential parents (2014)
  * n2 = number of unique adults estimated to have contributed to genotyped offspring (assigned + inferred parents from *Colony*)
  * m2 = number of genotyped parents assigned at least 1 offspring

Here is the sample size of individuals for potential parents (2014)
```{r}
read_csv("../Franz/stockdale_postQA_OceanAK_paired_2014_2016_STOCK.csv") %>% 
  filter(`Sample Year` == 2014) %>% 
  count(`Sample Year`)
```
And the number of unique parents inferred to have produced all offspring in our sample
```{r}
colony_best_config <- read_delim("../../Colony/Stockdale/Output/Stockdale_2014_2016.BestConfig", delim = " ", trim_ws = TRUE)

# colony_best_config <- read_delim("C:/Users/krshedd/Program Files/Colony/Example1/example1.BestConfig", delim = " ", trim_ws = TRUE)

# add
n_distinct(colony_best_config$FatherID) + n_distinct(colony_best_config$MotherID)

# split sire/dam
colony_best_config %>% 
  summarise(unique_sire = n_distinct(FatherID),
            unique_dam = n_distinct(MotherID))

# how many known vs. inferred sires
colony_best_config %>% 
  mutate(Father_Known = case_when(grepl(pattern = "\\*", x = FatherID) ~ FALSE,
                                  TRUE ~ TRUE)) %>% 
  group_by(Father_Known) %>% 
  summarise(n_Sire = n_distinct(FatherID))

# how many known vs. inferred dams
colony_best_config %>% 
  mutate(Mother_Known = case_when(grepl(pattern = "\\*", x = MotherID) ~ FALSE,
                                  TRUE ~ TRUE)) %>% 
  group_by(Mother_Known) %>% 
  summarise(n_Dam = n_distinct(MotherID))
```

And here are the number of parents that produced at least one offspring
```{r}
riverdist_parentage_stock <- read_csv("riverdist_parentage_stock.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

riverdist_parentage_stock %>% 
  filter(Success == "Successful") %>% 
  count(`Sample Year`)
```

So that gives us:  

  * n1 = 794
  * n2 = ???
  * m2 = 404

And our final calculation for Nc is:
```{r}
round(794 * XXX / 404)
```

Giving us an estimated escapement of **X,XXX** fish

### tGRC: Rarefaction Curve

**Note:** the previous version of this R Notebook that was sent out by Chris on Wednesday, August 28, 2019 did not apply this method correctly. In that version, the rarfaction curve only considered *assigned* parents and not *inferred* parents. This was a mistake, and is now corrected to consider both *assigned* **and** *inferred* parents. *Inferred* parents come from the maximum likelihood configuration pedigree from [*Colony*](https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1755-0998.2009.02787.x) (i.e. how many unique parents would it have taken to create all of the offspring in our sample).

In addition to tGMR, [Rawding et al. 2014](https://afspubs.onlinelibrary.wiley.com/doi/pdf/10.1080/00028487.2013.829122) provides a method to estimate the number of spawners using a rarefaction curve approach. Briefly, as each offspring is added to the sample, fewer new unique parental genotypes should be observed with the asymptote being equal to the number of spawners.

Here we will subsample the offspring (not their genotypes, *note* this is a subtle departure from Rawding et al. 2014 for simplicity) in levels of 10% to see how many unique parents (both assigned **and** inferred from *Colony*) we end up with.

```{r}
colony_best_config <- read_delim("../../Colony/Stockdale/Output/Stockdale_2014_2016.BestConfig", delim = " ", trim_ws = TRUE)

n_off <- n_distinct(colony_best_config$OffspringID)
n_sample <- round(seq(from = 0.1, to = 1, by = 0.1) * n_off)

rare_dat <- replicate(n = 1000, {
  lapply(n_sample, function(i) {
    tmp <- sample_n(colony_best_config, i)
    tibble(n_off = i, n_par = n_distinct(tmp$FatherID) + n_distinct(tmp$MotherID))
  }) %>% 
    bind_rows()
}, simplify = FALSE) %>% 
  bind_rows()
```

Apply Beverton-Holt model and plot
```{r}
library(FSA)

rare_dat <- rare_dat %>% 
  mutate(log_n_par = log(n_par))

bh1s <- srStarts(n_par ~ n_off, data = rare_dat, type = "BevertonHolt", param = 1)

bh1 <- log_n_par ~ log((a * n_off) / (1 + b * n_off))

bh1nls <- nls(bh1, data = rare_dat, start = bh1s)

bh0 <- log_n_par ~ log(a * n_off) # declare model
bh0s <- bh1s[1] # use the same starting value as above for a
bh0nls <- nls(bh0, data = rare_dat, start = bh0s)

xmax <- 12000
ymax <- max(rare_dat$n_par) * 1.2

plot(n_par ~ n_off, data = rare_dat, xlim = c(0, xmax), ylim = c(0, ymax), col = rgb(0, 0, 0, alpha = 0.01), pch = 16,
     xlab = "Sample Number of Offspring",
     ylab = "Number of Breeders", 
     main = "Stockdale 2014/2016 Genetic Rarefaction Curve")
curve((coef(bh1nls)[1] * x) / (1 + coef(bh1nls)[2] * x), from = 0, to = xmax, col = "red", lwd = 2, add = TRUE)
legend("topleft", legend = c("density dependent"), col = c("red"), lwd = 2, cex = 0.6)
```

Based on the parentage assignments, it looks like there were **~450** successful spawners.

## Summary

We have four separate estimates of the 2014 Stockdale escapement, which give us four different sampling proportions (parents sample size = 794):  

  1) Aerial Survey AUC: **2,869** fish, p = 28%
  2) Foot Survey AUC: **4,038** fish, p = 20%
  3) tGMR Binomial Model: **6,669** fish, p = 12%
  4) tGMR Hypergeometric Model: **X,XXX** fish, p = XX%
  4) tGRC Rarefaction Curve: **X,XXX** successful spawners, p = XX%

While these are four very different estimates, I think that we can be confident in saying that the Foot Survey AUC and tGMR Binomial Model estimates are likely the two most reliable estimates for the escapement. The aerial survey AUC estimate was only based on three flights, and the tGRC is biased downward in part due to the large number of parents that had no offspring assign back to them (only estimates successful spawners, not full escapement).

Based on this data, it is reasonable to conclude that our sample of 794 fish in 2014 likely represents between 12-20% of the escapement. Another very simple metric that we would expect to be correlated to the parental sampling rate would be the offspring assignment rate (i.e. what percentage of our 5,199 offspring assigned to parents?). Of our 5,199 offspring, 1,054 were assigned to at least one parent for an assignment rate of ~20%.

Sampling this proportion of the escapement (12-20%) provides opportunities for bias if the sampling is not representative of the escapement and other variables that covary with reproductive success.

# Single RRS Estimate for Stockdale 2014

While providing separate estimates of RRS by sex is standard throughout the literature (see [Williamson et al. 2010](https://www.nrcresearchpress.com/doi/abs/10.1139/F10-099), [Ford et al. 2012](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1755-263X.2012.00261.x), [Anderson et al. 2012](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1752-4571.2012.00271.x), [Berejikian etl al. 2009](https://www.nrcresearchpress.com/doi/abs/10.1139/f09-041), [Christie et al. 2014](https://onlinelibrary.wiley.com/doi/full/10.1111/eva.12183), etc.), especially in cases where most offspring assignments are to a single parent (as opposed to both parents), the Science Panel has expressed an interest in a single estimate of RRS, stratified by sex. In order to provide a stratified estimate, we need to know what we are stratifying by. I see two options, both of which have their own assumptions and drawbacks:  

  1) Assume that the sex ratio of the sample is representative of the population
  2) Sex-specific tGMR estimates as done in [Rawding et al. 2014](https://afspubs.onlinelibrary.wiley.com/doi/pdf/10.1080/00028487.2013.829122). 

I'll present both here for completeness, however, it is worth noting that using sex-specific tGMR estimates to stratify RRS data is potentially a bit circular (i.e. using the same parentage data to infer the population of males vs. females to stratify an estimate based on the parentage data).

## Sample Sex Ratio

The number of adults sampled in 2014 was 451 females and 343 males, resulting in a sex ratio of 1.32 (females/males).
```{r}
paired_14_16_filter_parents <- read_csv("stock_paired_14_16_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

paired_14_16_filter_parents %>% 
  count(`Sample Year`, sex)
```

If we assume that our sample sex ratio is representative of the sex ratio for the spawning population, then we can simply consider all parents and offspring together. **Note** that offspring with two parents get double counted since they assign to both a dam and a sire.

Here are sample sizes of parents by origin and the number of offspring that they produced (*both sexes combined*)
```{r}
paired_14_16_filter_parents %>% 
  group_by(origin) %>% 
  summarise(parents = n(), offspring = sum(n))
```

And here is our estimate of RRS with 95% confidence intervals following the methods of [Kalinowski and Taper 2005](http://www.nrcresearchpress.com/doi/10.1139/f04-239)
```{r}
rrs_ci_kalinowski <- function(n_h_off, n_w_off, n_h_par, n_w_par, alpha){
  chi_alpha <- qchisq(p = (1 - alpha), df = 1)
  n_off <- sum(c(n_h_off, n_w_off))
  n_par <- sum(c(n_h_par, n_w_par))
  
  rs_h <- n_h_off / n_h_par
  rs_w <- n_w_off / n_w_par
  
  p_h_par <- n_h_par / n_par
  p_w_par <- n_w_par / n_par
  
  rrs_h <- rs_h / rs_w
  rrs_w <- rs_w / rs_w
  rrs_avg <- (rrs_h * p_h_par) + (rrs_w * p_w_par)
  
  rrs_ml <- (n_h_off * log(p_h_par * rrs_h / rrs_avg)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg))
  
  xi_dist <- bind_rows(
    lapply(seq(0.01, 5, by = 0.01), function(rrs_h_xi) {
      rrs_avg_xi <- (rrs_h_xi * p_h_par) + (rrs_w * p_w_par)
      tibble(rrs_crit = rrs_h_xi,
             logl = (n_h_off * log(p_h_par * rrs_h_xi / rrs_avg_xi)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg_xi)) - (rrs_ml - chi_alpha / 2)
      )
    } )
  )
  
  rrs_min <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit < rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  rrs_max <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit > rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  xi_plot <- xi_dist %>% 
    ggplot(aes(x = rrs_crit, y = logl)) +
    geom_line() +
    geom_hline(yintercept = 0, colour = "red", lwd = 2) +
    geom_vline(xintercept = c(rrs_h, rrs_min, rrs_max), colour = "blue") +
    ylim(c(-5, 5)) +
    xlim(c(0, 2)) +
    ylab("Log Likelihood - Chi Sq Value") +
    annotate("text", x = rrs_h + 0.1, y = xi_dist %>% filter(rrs_crit == xi_dist$rrs_crit[which.min(abs(xi_dist$rrs_crit  - rrs_h))]) %>% pull(logl) + 0.4, label = round(rrs_h, 2)) +
    annotate("text", x = rrs_min - 0.1, y = xi_dist %>% filter(rrs_crit == rrs_min) %>% pull(logl) + 0.4, label = rrs_min) +
    annotate("text", x = rrs_max + 0.1, y = xi_dist %>% filter(rrs_crit == rrs_max) %>% pull(logl) + 0.4, label = rrs_max)
  
  print(xi_plot)
  return(c(rrs_min, rrs_h, rrs_max))
}

rrs_ci_kalinowski(n_h_off = 373, n_w_off = 865, n_h_par = 436, n_w_par = 358, alpha = 0.05)
```

**RRS hatchery vs. natural = 0.35 (0.31-0.40)**

## Sex-Specific tGRM Binomial Model

Here is the sample size of individuals for potential parents (2014) and recaptured offspring assigned to parents (2016)
```{r}
paired_14_16_filter_parents %>% 
  group_by(SEX) %>% 
  summarise(parents = n(), offspring = sum(n))
```

### Male

  * n1 = 343 male potential parents
  * n2 = 5,199 x 1 (1 copy from sire) = 5,199
  * m2 = 594 (offspring that assigned to males)

And our final calculation for male Nc is:
```{r}
round(343 * 5199 * 1 / 594)
```

Giving us an estimated escapement of **3,002** fish

### Female

  * n1 = 451 female potential parents
  * n2 = 5,199 x 1 (1 copy from dam) = 5,199
  * m2 = 644 (offspring that assigned to females)

And our final calculation for male Nc is:
```{r}
round(451 * 5199 * 1 / 644)
```

Giving us an estimated escapement of **3,641** fish

### Stratified Estimate

We found male RRS ~ 0.28 and female RRS ~ 0.42, if we stratify those by our population estimates, we get a combined RRS.
```{r}
round(((3002 * 0.28) + (3641 * 0.42)) / (3002 + 3641), 2)
```

**RRS hatchery vs. natural = 0.36**

# Test Covariates of Reproductive Success: Sample Date, Size, and Location

Here we will do some basic testing to see if there are RRS differences when arbitrarily splitting our covariates into bins (i.e. early vs. late, small vs. large, intertidal vs. upstream). Additionally, we will investigate whether differences in RRS between hatchery- and natural-origin fish hold, despite these covariates of RS. These results should be considered "preliminary" analyses, because a more rigorous analysis (see [Generalized Linear Models](#GLM)) will 1) not arbitrarily break continuous variables (i.e. sample date and size) into categorical bins, and 2) consider all variables and their interactions holistically in a single model rather than considering each on their own. In another, soon to follow notebook, we will consider all information more holistically by model fitting to generalized linear models (GLM's), as is commonly done in the literature (see [Bernston et al. 2011](https://afspubs.onlinelibrary.wiley.com/doi/pdf/10.1080/00028487.2011.584489), [Ford et al. 2012](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1755-263X.2012.00261.x), and [Janowitz-Koch et al. 2018](https://onlinelibrary.wiley.com/doi/full/10.1111/eva.12725)).

## Sample Date {#RS_vs_date}

First let us look at the influence of sample date on RS for hatchery- and natural-origin spawners.
```{r}
paired_14_16_filter_parents <- read_csv("stock_paired_14_16_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

paired_14_16_filter_parents %>%
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(height = 0, size = 2) +
  geom_smooth() +
  facet_grid(sex ~ .)+
  labs(title = "",  colour = "Parent\nOrigin",
       shape = "Parent: Sex") +
  xlab("Parent Sample Date") +
  ylab("Number of Offspring") +
  theme_bw() +
  theme(text = element_text(size = 20))
```

We can see that there are some temporal patterns in RS. In order to do a naive test, we can split parents sampled in August (early) vs. parents sampled in September (late). What are our sample sizes by month (*both origins and sexes combined*)?
```{r}
paired_14_16_filter_parents %>% 
  filter(`Sample Year` == 2014) %>% 
  mutate(Month = month(`Sample Date`, label = TRUE)) %>% 
  count(`Sample Year`, Month)
```

How many offspring were produced by these parents?
```{r}
paired_14_16_filter_parents %>% 
  filter(`Sample Year` == 2014) %>% 
  mutate(Month = month(`Sample Date`, label = TRUE)) %>% 
  group_by(Month) %>% 
  summarise(offspring = sum(n, na.rm = TRUE))
```

### Early (August) vs. Late (September)

Based on this naive test, early-run fish produced more offspring on average. Now we can calculate the RRS of late-run (September) parents relative to early-run (August) parents (*both origins and sexes combined*)
```{r}
rrs_ci_kalinowski(n_h_off = 407, n_w_off = 831, n_h_par = 398, n_w_par = 396, alpha = 0.05)
```

**RRS of late vs. early = 0.49 (0.43-0.55)**

Run timing (i.e. parent sample date) has an effect on RS with late-run (September) parents producing half as many offspring on average as early-run (August) parents. Thus, if sampling timing of parents is not fully representative of escapement timing, there is a potential for our estimates of RS (and RRS) for hatchery- vs. natural-origin to be biased. 

However, given that we now know (parent sample month has an influence on RS), what if we compare hatchery- vs. natural-origin RRS within each month?

### Hatchery vs. Natural Within Month

What are our sample sizes by month and origin (*both sexes combined*)?
```{r}
paired_14_16_filter_parents %>% 
  filter(`Sample Year` == 2014) %>% 
  mutate(Month = month(`Sample Date`, label = TRUE)) %>% 
  count(`Sample Year`, Month, origin) %>% 
  spread(origin, n)
```

How many offspring were produced by these parents?
```{r}
paired_14_16_filter_parents %>% 
  filter(`Sample Year` == 2014) %>% 
  mutate(Month = month(`Sample Date`, label = TRUE)) %>% 
  group_by(Month, origin) %>% 
  summarise(offspring = sum(n, na.rm = TRUE)) %>% 
  spread(origin, offspring)
```

Now we can calculate month-specific RRS (*both sexes combined*)

#### August (Early)

```{r}
rrs_ci_kalinowski(n_h_off = 197, n_w_off = 634, n_h_par = 172, n_w_par = 224, alpha = 0.05)
```

**RRS hatchery vs. natural = 0.40 (0.34-0.47)**

#### September (Late)

```{r}
rrs_ci_kalinowski(n_h_off = 176, n_w_off = 231, n_h_par = 264, n_w_par = 134, alpha = 0.05)
```

**RRS hatchery vs. natural = 0.39 (0.32-0.47)**

### Summary

When we compare RRS for hatchery- and natural-origin fish by sample month of the parents (August - early, September - late), we find that hatchery-origin fish produce an average of ~39-40% as many offspring as natural-origin parents. Sample date (timing) matters (see [plot](#RS_vs_date)), but we still see differences in RS between hatchery- and natural-origin fish when we compare within early- and late-run groups.

## Size {#RS_vs_size}

Next let us look at the influence of body size on RS for hatchery- and natural-origin spawners.
```{r}
paired_14_16_filter_parents <- read_csv("stock_paired_14_16_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

paired_14_16_filter_parents %>% 
  filter(`Length Mm` > 300) %>%  # remove outlier
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(height = 0, size = 2) +
  geom_smooth() +
  facet_grid(sex ~ .) +
  labs(title = "",  colour = "Parent\nOrigin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Number of Offspring") +
  theme_bw() +
  theme(text = element_text(size = 20))
```

We can see that there are some body size patterns in RS, specifically for males. In order to do a naive test, we can split the fish using the average fish size; "large" fish are above average, and "small" fish are below average size. What are our sample sizes by large/small (425mm is our cutoff) and origin (*both origins and sexes combined*)?
```{r}
paired_14_16_filter_parents %>% 
  filter(`Length Mm` > 300) %>%  # remove outlier
  filter(`Sample Year` == 2014) %>% 
  mutate(Size = case_when(`Length Mm` <= 425 ~ "Small",
                          `Length Mm` > 425 ~ "Large")) %>% 
  mutate(Size = factor(x = Size, levels = c("Small", "Large"))) %>% 
  count(`Sample Year`, Size)
```

How many offspring were produced by these parents?
```{r}
paired_14_16_filter_parents %>% 
  filter(`Length Mm` > 300) %>%  # remove outlier
  filter(`Sample Year` == 2014) %>% 
  mutate(Size = case_when(`Length Mm` <= 425 ~ "Small",
                          `Length Mm` > 425 ~ "Large")) %>% 
  mutate(Size = factor(x = Size, levels = c("Small", "Large"))) %>% 
  group_by(Size) %>% 
  summarise(offspring = sum(n, na.rm = TRUE))
```

### Small (<= 425mm) vs. Large (> 425mm)

It doesn't look like there are large differences in average RS between small and large parents. Now we can calculate the RRS of small (<= 425mm) parents relative to large (> 425mm) parents (*both origins and sexes combined*)
```{r}
rrs_ci_kalinowski(n_h_off = 531, n_w_off = 598, n_h_par = 359, n_w_par = 367, alpha = 0.05)
```

**RRS small vs. large = 0.91 (0.81-1.02)**

Body size (i.e. parent length) does not appear to have a strong effect (not significant with alpha = 0.05), but on average, smaller fish produce slightly fewer offspring (91%) than larger fish. Thus, if sampling of parents based on size is not fully representative of escapement, there is relatively little potential for our estimates of RS (and RRS) for hatchery- vs. natural-origin to be biased. 

Even though we now know that parent body size has relatively little influence on RS, what if we compare hatchery- vs. natural-origin RRS within each size class for curiosity?

### Hatchery vs. Natural Within Size Class

What are our sample sizes by size class and origin (*both sexes combined*)?
```{r}
paired_14_16_filter_parents %>% 
  filter(`Length Mm` > 300) %>%  # remove outlier
  filter(`Sample Year` == 2014) %>% 
  mutate(Size = case_when(`Length Mm` <= 425 ~ "Small",
                          `Length Mm` > 425 ~ "Large")) %>% 
  mutate(Size = factor(x = Size, levels = c("Small", "Large"))) %>% 
  count(`Sample Year`, Size, origin) %>% 
  spread(origin, n)
```

How many offspring were produced by these parents?
```{r}
paired_14_16_filter_parents %>% 
  filter(`Length Mm` > 300) %>%  # remove outlier
  filter(`Sample Year` == 2014) %>% 
  mutate(Size = case_when(`Length Mm` <= 425 ~ "Small",
                          `Length Mm` > 425 ~ "Large")) %>% 
  mutate(Size = factor(x = Size, levels = c("Small", "Large"))) %>% 
  group_by(Size, origin) %>% 
  summarise(offspring = sum(n, na.rm = TRUE)) %>% 
  spread(origin, offspring)
```

Now we can calculate size-specific RRS (*both sexes combined*)

#### Small (<= 425mm)

```{r}
rrs_ci_kalinowski(n_h_off = 120, n_w_off = 441, n_h_par = 151, n_w_par = 208, alpha = 0.05)
```

**RRS hatchery vs. natural = 0.37 (0.31-0.46)**

#### Large (> 425mm)

```{r}
rrs_ci_kalinowski(n_h_off = 223, n_w_off = 375, n_h_par = 243, n_w_par = 124, alpha = 0.05)
```

**RRS hatchery vs. natural = 0.30 (0.26-0.36)**

### Summary

When we compare RRS for hatchery- and natural-origin fish by body size of the parents (<= 425mm - small, > 425mm - large), we find that hatchery-origin fish produce an average of ~30-37% as many offspring as natural-origin parents. Body size has some effect (see [plot](#RS_vs_size)), but we still see differences in RS between hatchery- and natural-origin fish when we compare within small and large size classes.

## Location {#RS_vs_location}

Next let us look at the influence of spawning location on RS for hatchery- and natural-origin spawners. **Note** Parent sample location is listed in meters from the lowest extent of samples (i.e. mouth of stream), the black line denotes the upper extent of high tide (i.e. below is intertidal and above is upstream).
```{r}
riverdist_parentage_stock <- read_csv("riverdist_parentage_stock.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

riverdist_parentage_stock %>%
  ggplot(aes(x = Distance, y = n, color = origin)) +
  geom_vline(aes(xintercept = tidedist), size = 2) +
  geom_jitter(height = 0, size = 2) +
  geom_smooth() +
  facet_grid(sex ~ .)+
  labs(title = "",  colour = "Parent\nOrigin",
       shape = "Parent: Sex") +
  xlab("Parent Sample Location (m from mouth)") +
  ylab("Number of Offspring") +
  theme_bw() +
  theme(text = element_text(size = 20))
```

We can clearly see that there are some spatial patterns in RS, specifically for males. In order to do a naive test, we can split the fish into upstream and intertidal spawners (*both origins and sexes combined*)?
```{r}
riverdist_parentage_stock %>% 
  count(`Sample Year`, Intertidal)
```

How many offspring were produced by these parents?
```{r}
riverdist_parentage_stock %>% 
  filter(`Sample Year` == 2014) %>% 
  group_by(Intertidal) %>% 
  summarise(offspring = sum(n, na.rm = TRUE))
```

### Upstream vs. Intertidal

Intertidal spawners produced more offspring on average than upstream spawners. Now we can calculate the RRS of upstream parents relative to intertidal parents (*both origins and sexes combined*)
```{r}
rrs_ci_kalinowski(n_h_off = 458, n_w_off = 775, n_h_par = 488, n_w_par = 305, alpha = 0.05)
```

**RRS upstream vs. intertidal = 0.37 (0.33-0.41)**

Spawning location (i.e. parent sample location in the intertidal vs. upstream) clearly has an effect on RS, with upstream fish producing 37% as many offspring on average as intertidal parents! Thus, if sampling location of parents is not fully representative of the spatial distribution of escapement, there is a potential for our estimates of RS (and RRS) for hatchery- vs. natural-origin to be biased. 

However, given that we now know (parent sample location has an influence on RS), what if we compare hatchery- vs. natural-origin RRS within each location?

### Hatchery vs. Natural Within Location

What are our sample sizes by location and origin (*both sexes combined*)?
```{r}
riverdist_parentage_stock %>% 
  count(`Sample Year`, Intertidal, origin) %>% 
  spread(origin, n)
```

How many offspring were produced by these parents?
```{r}
riverdist_parentage_stock %>% 
  filter(`Sample Year` == 2014) %>% 
  group_by(Intertidal, origin) %>% 
  summarise(offspring = sum(n, na.rm = TRUE)) %>% 
  spread(origin, offspring)
```

Now we can calculate location-specific RRS (*both sexes combined*)

### Intertidal

```{r}
rrs_ci_kalinowski(n_h_off = 205, n_w_off = 570, n_h_par = 117, n_w_par = 188, alpha = 0.05)
```

**RRS hatchery vs. natural = 0.58 (0.49-0.68)**

### Upstream

```{r}
rrs_ci_kalinowski(n_h_off = 168, n_w_off = 290, n_h_par = 319, n_w_par = 169, alpha = 0.05)
```

**RRS hatchery vs. natural = 0.31 (0.25-0.37)**

### Summary

When we compare RRS for hatchery- and natural-origin fish by sample location of the parents (intertidal vs. upstream), we find that hatchery-origin fish produce an average of ~31-58% as many offspring as natural-origin parents depending on where the parents were sampled. Parent sample location clearly matters (see [plot](#RS_vs_location)), but we still see differences in RS between hatchery- and natural-origin fish when we compare within intertidal and upstream groups.

# Generalized Linear Models {#GLM}

These *preliminary* data explorations reveal two conclusions: 1) there is room for our samples to be biased relative to the escapement based on the proportion of 2014 spawners sampled (~12-20%), and 2) there are variables (covariates) that might affect bias in hatchery vs. natural RRS, if sampling is not representating of these covariates; of the three variables we investigated, the two that were most highly correlated with RS were sample date (timing) and sample location within the stream.

The next steps are:  

  1) to move on to a more robust approach and consider all variables (covariates) of RS holistically. General linear models are commonly used in the literature (see [Bernston et al. 2011](https://afspubs.onlinelibrary.wiley.com/doi/pdf/10.1080/00028487.2011.584489), [Ford et al. 2012](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1755-263X.2012.00261.x), and [Janowitz-Koch et al. 2018](https://onlinelibrary.wiley.com/doi/full/10.1111/eva.12725)) to determine the magnitude of the effect of origin on RRS while considering other variables (i.e. sample date, fish size, sample location, etc.). In another notebook soon to follow we will consider all information more holistically by model fitting to GLM's; and 
  2) determine if our samples are temporally representative of escapement (i.e. did we have a consistent sampling rate throughout the season), and, if not, determine the most appropriate way to adjust our RRS estimates for this (e.g. randomly subsampling fish by sample date at an even sampling rate and getting bootstrapped estimates of hatchery vs. natural RRS). This will be a topic of discussion at the upcoming Science Panel meeting.