---
title: "Stockdale 2014-2016 Sample Sizes for John H"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

# Introduction

John H. Clark wants to know:  
  1) sample size of potential parents by sex and origin (2014)
  2) number of offspring assigned to parents by sex and origin (2016)
  3) number of potential parents with at least one offspring by sex and origin (2014)

I presume that he wants to know this for some sort of tgMR exercise?

```{r setup, include=FALSE}
library(tidyverse)
```

# Data

```{r read_data}
(paired_14_16_filter_parents <- read_csv("stock_paired_14_16_filter_parents.csv") %>% 
   mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
   mutate(sex = case_when(SEX == "M" ~ "Male",
                          SEX == "F" ~ "Female")) %>% 
   mutate(sex = factor(x = sex, levels = c("Male", "Female"))))
```

# Sample Sizes

## Parents by origin and sex

```{r n_parents}
paired_14_16_filter_parents %>% 
  count(origin, sex)
```

## Offspring assigned to parents by origin and sex

```{r n_offspring_assigned}
paired_14_16_filter_parents %>% 
  group_by(origin, sex) %>% 
  summarise(n = sum(n))
```

## Parents with at least one offspring

```{r n_parents_successful}
paired_14_16_filter_parents %>% 
  filter(n > 0) %>% 
  count(origin, sex)
```

# Rarefaction

Based on [Rawding et al. 2014](https://afspubs.onlinelibrary.wiley.com/doi/pdf/10.1080/00028487.2013.829122).

## Data

Get offspring data

```{r}
offspring_dat <- read_csv("../Franz/stockdale_postQA_OceanAK_paired_2014_2016_STOCK.csv") %>% 
  filter(`Sample Year` == 2016)

(offspring_franz <- read_csv("../../Franz/Stockdale/Even_Run 2/parentage.csv") %>% 
    filter(Posterior >= 0.5) %>% 
    left_join(offspring_dat, by = c("Offspring" = "franz_id")))
```

## Analysis

Subsample the offspring (not their genotypes, this is a departure from Rawding for simplicity) in levels of 10% and see how many unique parents we end up with.
```{r rarefaction_loop}
n_off <- n_distinct(offspring_franz$Offspring)
n_sample <- round(seq(from = 0.1, to = 1, by = 0.1) * n_off)

rare_dat <- replicate(n = 1000, {
  lapply(n_sample, function(i) {
    tmp <- sample_n(offspring_franz, i) %>% 
      select(Offspring, `Parent 1`, `Parent 2`) %>% 
      gather(Parent, ID, -Offspring) %>% 
      arrange(Offspring)
    tibble(n_off = i, n_par = n_distinct(tmp$ID))
  }) %>% 
    bind_rows()
}, simplify = FALSE) %>% 
  bind_rows()
```

Make a plot of the curve.
```{r rarefaction_plot}
rare_dat %>% 
  ggplot(aes(x = n_off, y = n_par)) +
  geom_point(alpha = 0.01) +
  xlab("Sample Number of Offspring") +
  ylab("Number of Breeders") +
  ggtitle("Stockdale 2014/2016 Genetic Rarefaction Curve") +
  theme_bw()
```

Apply the Beverton-Holt equation.
```{r}
library(FSA)

rare_dat <- rare_dat %>% 
  mutate(log_n_par = log(n_par))

bh1s <- srStarts(n_par ~ n_off, data = rare_dat, type = "BevertonHolt", param = 1)

bh1 <- log_n_par ~ log((a * n_off) / (1 + b * n_off))

bh1nls <- nls(bh1, data = rare_dat, start = bh1s)

bh0 <- log_n_par ~ log(a * n_off) # declare model
bh0s <- bh1s[1] # use the same starting value as above for a
bh0nls <- nls(bh0, data = rare_dat, start = bh0s)

anova(bh0nls, bh1nls)

AIC(bh0nls, bh1nls)

xmax <- 12000
ymax <- 1000

plot(n_par ~ n_off, data = rare_dat, xlim = c(0, xmax), ylim = c(0, ymax), col = rgb(0, 0, 0, alpha = 0.01), pch = 16,
     xlab = "Sample Number of Offspring",
     ylab = "Number of Breeders", 
     main = "Stockdale 2014/2016 Genetic Rarefaction Curve")
curve((coef(bh1nls)[1] * x) / (1 + coef(bh1nls)[2] * x), from = 0, to = xmax, col = "red", lwd = 2, add = TRUE)
# curve(coef(bh0nls)[1] * x, from = 0, to = 12000, col = "blue", lwd = 2, add = TRUE)
legend("topleft", legend = c("density dependent"), col = c("red"), lwd = 2, cex = 0.6)
```


# tGMR

## Binomial

n1 = number of carcasses genotyped (potential parents)
n2 = genotypes from offspring (potential offspring x 2, 1 for dam, 1 for sire)
m2 = number of recaptured genotypes

### n1 and n2

```{r}
read_csv("../Franz/stockdale_postQA_OceanAK_paired_2014_2016_STOCK.csv") %>% 
  count(`Sample Year`)
```

n1 = 794
n2 = 5199 x 2 = 10398

### m2

```{r}
sum(!is.na(offspring_franz$`Parent 1`))
sum(!is.na(offspring_franz$`Parent 2`))
```

m2 = 1058 (parent 1) + 183 (parent 2) = 1241

```{r}
offspring_franz %>% 
  select(Offspring, `Parent 1`, `Parent 2`) %>% 
  gather(Parent, ID, -Offspring) %>% 
  filter(!is.na(ID))
```


### Nc

```{r}
round(794 * 5199 * 2 / 1241)
```

## Binomial strata

### n1 and n2

View 2014 Stockdale by sample date. Let's go with August vs. September.
```{r}
paired_2014_2016_dat <- read_csv("../Franz/stockdale_postQA_OceanAK_paired_2014_2016_STOCK.csv")

paired_2014_2016_dat %>% 
  filter(`Sample Year` == 2014) %>% 
  ggplot(aes(x = `Sample Date`, fill = `Otolith Mark Present`)) +
  geom_bar()
```

```{r}
paired_2014_2016_dat %>% 
  mutate(month = month(`Sample Date`, label = TRUE)) %>% 
  count(`Sample Year`, month)
```

```{r}
offspring_franz %>% 
  select(Offspring, `Parent 1`, `Parent 2`) %>% 
  gather(Parent, ID, -Offspring) %>% 
  filter(!is.na(ID)) %>% 
  left_join(paired_2014_2016_dat, by = c("ID" = "franz_id")) %>% 
  mutate(month = month(`Sample Date`, label = TRUE)) %>% 
  count(`Sample Year`, month)
```

