---
title: "Stockdale 2014 Area Under the Curve"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(readxl)
```

These are **not** adjusted for stream life or observer efficiency yet!
```{r}
(auc <- readxl::read_xlsx(path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Field data/StreamSurveys 10-28-14 Ped only to Kyle 11-7-14_Updated_SM_2-25-2015.xlsm", sheet = "AUC", skip = 2))
```

Now I adjust daily live AUC counts by stream life and observer efficiency and plot with raw dead counts.
```{r}
stream_life <- 13.7  # also a variable in the 
observer_efficiency <- 0.703

auc %>% 
  filter(Strname == "Stockdale C", Year == 2014) %>% 
  mutate(live_auc = `Live Count AUC` / `Salmon Stream Life` / observer_efficiency) %>% 
  rename(dead_raw = `Dead count`) %>% 
  select(Survdate, live_auc, dead_raw) %>% 
  gather(type, count, -Survdate) %>% 
  ggplot(aes(x = Survdate, y = count, colour = type)) +
  geom_line() + 
  theme_bw()
```

Read in OceanAK data for to get all samples collected.
```{r}
og_names <- suppressMessages(names(read_csv(file = "../OceanAK/PedigreeData_AHRP - Salmon Biological Data 2_PWS_2013-2018_no_otoliths.csv", progress = FALSE)))
oceanak <- read_csv(file = "../OceanAK/Salmon Biological Data 2 Export from DFGCFRESP.csv")
names(oceanak) <- og_names

(oceanak_mod <- oceanak  %>% 
    unite(SillySource, `Silly Code`, `Fish ID`, sep = "_", remove = FALSE) %>% 
    unite(TrayCodeID, `DNA Tray Code`, `DNA Tray Well Code`, sep = "_", remove = FALSE) %>% 
    mutate(otolith_read = !is.na(`Otolith Mark Status Code`) & `Otolith Mark Status Code` != "n") %>% 
    mutate(stream = factor(x = `Location Code`, levels = c("Gilmour Creek", "Paddy Creek", "Erb Creek", "Hogan Creek", "Stockdale Creek"))) %>% 
    rename(year = `Sample Year`) %>% 
    mutate(origin = case_when(`Otolith Mark Present` == "NO" ~ "natural",
                              `Otolith Mark Present` == "YES" ~ "hatchery")) %>% 
    mutate(origin = factor(origin, levels = c("natural", "hatchery"))) %>% 
    mutate(date = dmy(`Sample Date`))
)
```

Samples collected by day.
```{r}
(samples <- oceanak_mod %>% 
   filter(stream == "Stockdale Creek", year == 2014) %>% 
   count(date) %>% 
   rename(samples = n) %>% 
   bind_rows(tibble(date = as.Date("2014-09-12"), samples = 0))
)
```

Read in samples that were analyzed (we randomly selected a subset to genotype)
```{r}
(paired_data <- read_csv("Franz/stockdale_postQA_OceanAK_paired_2014_2016_STOCK.csv"))
```

Samples analyzed by day (make sure to add in last day on 09-12)
```{r}
(analyzed <- paired_data %>% 
   filter(`Sample Year` == 2014) %>% 
   mutate(date = as.Date(`Sample Date`)) %>% 
   count(date) %>% 
   rename(analyzed = n) %>% 
   bind_rows(tibble(date = as.Date("2014-09-12"), analyzed = 0))
)
```

Plot daily AUC, raw dead counts, samples collected, and samples analyzed.
```{r}
auc %>% 
  filter(Strname == "Stockdale C", Year == 2014) %>% 
  mutate(live_auc = `Live Count AUC` / `Salmon Stream Life` / observer_efficiency) %>% 
  rename(dead_raw = `Dead count`, date = Survdate) %>% 
  mutate(date = as.Date(date)) %>% 
  select(date, live_auc, dead_raw) %>% 
  full_join(samples) %>% 
  full_join(analyzed) %>% 
  gather(type, count, -date) %>% 
  drop_na(count) %>% 
  mutate(type = factor(x = type, levels = c("live_auc", "dead_raw", "samples", "analyzed"))) %>% 
  ggplot(aes(x = date, y = count, colour = type)) +
  geom_line(lwd = 2) + 
  theme_bw() +
  ggtitle("Stockdale 2014")
```

Now, what if we shift the daily live AUC counts over by stream life? Shouldn't that represent how many "new dead" accumulate each day?
```{r}
auc %>% 
  filter(Strname == "Stockdale C", Year == 2014) %>% 
  mutate(live_auc = `Live Count AUC` / `Salmon Stream Life` / observer_efficiency) %>% 
  rename(dead_raw = `Dead count`, date = Survdate) %>% 
  mutate(date = as.Date(date)) %>% 
  select(date, live_auc, dead_raw) %>% 
  full_join(samples) %>% 
  full_join(analyzed) %>% 
  gather(type, count, -date) %>% 
  drop_na(count) %>% 
  mutate(date = case_when(type == "live_auc" ~ date + 14,
                          TRUE ~ date)) %>% 
  filter(type %in% c("live_auc", "analyzed")) %>% 
  mutate(type = case_when(type == "live_auc" ~ "adjusted_live_auc",
                          TRUE ~ type)) %>% 
  mutate(type = factor(x = type, levels = c("adjusted_live_auc", "analyzed"))) %>% 
  ggplot(aes(x = date, y = count, colour = type)) +
  geom_line(lwd = 2) + 
  theme_bw() +
  ggtitle("Stockdale 2014")
```

