---
title: "Stockdale 2014-2016 Sample Sizes for John H"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

# Introduction

John H. Clark wants to know:  
  1) sample size of potential parents by sex and origin (2014)
  2) number of offspring assigned to parents by sex and origin (2016)
  3) number of potential parents with at least one offspring by sex and origin (2014)

I presume that he wants to know this for some sort of tgMR exercise?

```{r setup, include=FALSE}
library(tidyverse)
```

# Data

```{r read_data}
(paired_14_16_filter_parents <- read_csv("stock_paired_14_16_filter_parents.csv") %>% 
   mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
   mutate(sex = case_when(SEX == "M" ~ "Male",
                          SEX == "F" ~ "Female")) %>% 
   mutate(sex = factor(x = sex, levels = c("Male", "Female"))))
```

# Sample Sizes

## Parents by origin and sex

```{r n_parents}
paired_14_16_filter_parents %>% 
  count(origin, sex)
```

## Offspring assigned to parents by origin and sex

```{r n_offspring_assigned}
paired_14_16_filter_parents %>% 
  group_by(origin, sex) %>% 
  summarise(n = sum(n))
```

## Parents with at least one offspring

```{r n_parents_successful}
paired_14_16_filter_parents %>% 
  filter(n > 0) %>% 
  count(origin, sex)
```

