---
title: "Stockdale Figures"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---


Create Figures for Brown Bag presenation (same idea as Hogan for AMSS and NPRB Publication)

```{r setup, include=FALSE}
library(coin)
library(MASS)
library(scales)
library(tidyverse)
# library(leaflet)
library(lubridate)
library(PBSmapping)
library(grid)
library(ggthemes) 
library(ggrepel) # Load
```


# Figure 1: Map of study area

Try PBSmapping, clunky, but great basemap for AK coastlines. Got the idea from Ben Williams GitHub [link](https://ben-williams.github.io/updated_ggplot_figures.html). Still used the idea from Quantitative Palaeoecology [link](https://quantpalaeo.wordpress.com/2016/06/05/ggplot2-maps-with-inset/)

```{r map of study area, message=FALSE, warning=FALSE}
# Load data
dat <- read_csv("../map_data.csv")

dat <- dat %>%
  mutate(Legend = case_when(type == "Hogan" ~ "Fitness Stream",
                            type == "Fitness Stream" ~ "Fitness Stream",
                            type == "Hatchery" & name != "VFDA" ~ "PWSAC",
                            name == "VFDA" ~ "VFDA")) %>% 
  mutate(Legend = factor(x = Legend, levels = c("Fitness Stream", "VFDA", "PWSAC"))) %>% 
  mutate(abbr = case_when(name == "AFK" ~ "AFK",
                          name == "WHN" ~ "WNH",
                          name == "Cannery Creek" ~ "CCH",
                          name == "VFDA" ~ "SGH",
                          TRUE ~ name))

x_lim <- range(dat$long) + c(-0.25, 0.25)
y_lim <- range(dat$lat) + c(-0.25, 0.25)

dat <- dat %>%
  filter(type != "City")

# Plot map with PBS mapping
data("nepacLLhigh")
ak <- nepacLLhigh %>% 
  dplyr::select(group = PID, POS = POS,long = X, lat = Y)

maptheme <- theme(
  panel.grid = element_blank(),
  panel.border = element_rect(fill = NA, colour = "black"),
  panel.background = element_blank()
)

# PWS map
pws_map <- ggplot() + 
  geom_polygon(data = ak, aes(long, lat, group = group), fill = "grey90", color = "black", lwd = 0.3) +
  geom_point(data = dat, aes(x = long, y = lat, fill = Legend), shape = 21, colour = "black", stroke = 2, size = 5) +
  geom_label_repel(data = filter(dat, type %in% c("Hatchery", "Fitness Stream", "Hogan")), aes(long, lat, label = abbr), hjust = 0, nudge_x = 0.1) +
  geom_segment(aes(x = -148.7, xend = -148.7, y = 59.85, yend = 60.02), lineend = "butt", linejoin = "round", size = 2, arrow = arrow(length = unit(0.2, "inches"))) +
  theme(panel.background = element_rect(fill = 'white')) +
  scale_y_continuous(breaks = c(60.0, 60.5, 61.0)) +
  xlab(expression(paste(Longitude^o, ~'W'))) +
  ylab(expression(paste(Latitude^o, ~'N'))) +
  coord_map(xlim = x_lim, ylim = y_lim) +
  maptheme

# north2(pws_map, symbol = 12, 0.18, 0.20)  # couldn't get this to work with the inset

# Inset map
ak_map <- ggplot() + 
  geom_polygon(data = ak, aes(long, lat, group = group), fill = "grey90", color = "black", lwd = 0.3) +
  geom_rect(data = data.frame(), aes(xmin = x_lim[1], xmax = x_lim[2], ymin = y_lim[1], ymax = y_lim[2]), colour = "red", fill = NA, lwd = 1.3) +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  coord_map(xlim = c(-170, -130), ylim = c(53, 65))

# Create map with inset and North arrow
grid.newpage()
vp_b <- viewport(width = 1, height = 1, x = 0.5, y = 0.5)  # the larger map
vp_a <- viewport(width = 0.23, height = 0.35, x = 0.605, y = 0.24)  # the inset in lower left
print(pws_map, vp = vp_b)
print(ak_map + maptheme, vp = vp_a)
```

This is the code used to actually save the map as a .png file
```{r save_map}
png(filename = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/PWS_fitness_stream_map.png", width = 6.5, height = 6.5, units = "in", res = 300)
grid.newpage()
vp_b <- viewport(width = 1, height = 1, x = 0.5, y = 0.5)  # the larger map
vp_a <- viewport(width = 0.23, height = 0.35, x = 0.65, y = 0.3)  # the inset in lower left
print(pws_map, vp = vp_b)
print(ak_map + maptheme, vp = vp_a)
dev.off()
```

# Figure 2: Sample sizes throughout run

Create blank bar plot of even-year sampling, for use as builder slide.
```{r blank_samp_png}
paired_14_16_filter <- read_csv("stock_paired_14_16_filter.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

#Stacked
paired_14_16_filter %>% 
  count(`Sample Year`, Julian, origin) %>% 
  mutate(n = 0) %>%  # make 0 for blank plot
  ggplot(aes(x = `Julian`, y = n, fill = origin)) +
  theme_bw() +
  ylim(c(0, 580)) +
  geom_col(colour = "black") +
  geom_hline(yintercept = 0, colour = "black") +
  facet_grid(`Sample Year` ~ .) +
  ylab("Count") +
  xlab("Julian Date") +
  labs(title="",  # Number of Samples per Year by Date and Origin
       fill = "Origin") +
  theme(text = element_text(size = 20))  # strip.text = element_text(size = 12)
ggsave("samples_blank.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")
```

Create bar plot of even-year sampling with only 2014 data, for use as builder slide.
```{r even_2014_samp_png}
paired_14_16_filter <- read_csv("stock_paired_14_16_filter.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

#Stacked
paired_14_16_filter %>% 
  count(`Sample Year`, Julian, origin) %>% 
  mutate(nn = case_when(`Sample Year` == 2014 ~ n,
                        `Sample Year` == 2016 ~ as.integer(0))) %>%  # make 0 for blank plot
  ggplot(aes(x = `Julian`, y = nn, fill = origin)) +
  theme_bw() +
  ylim(c(0, 580)) +
  geom_col(colour = "black") +
  geom_hline(yintercept = 0, colour = "black") +
  facet_grid(`Sample Year` ~ .) +
  ylab("Count") +
  xlab("Julian Date") +
  labs(title="",  # Number of Samples per Year by Date and Origin
       fill = "Origin") +
  theme(text = element_text(size = 20))  # strip.text = element_text(size = 12)
ggsave("samples_even_2014.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")
```

Create bar plot of even-year sampling.
```{r even_samp_png}
paired_14_16_filter <- read_csv("stock_paired_14_16_filter.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

#Stacked
paired_14_16_filter %>% 
  count(`Sample Year`, Julian, origin) %>% 
  ggplot(aes(x = `Julian`, y = n, fill = origin)) +
  theme_bw() +
  ylim(c(0, 580)) +
  geom_col(colour = "black") +
  geom_hline(yintercept = 0, colour = "black") +
  facet_grid(`Sample Year` ~ .) +
  ylab("Count") +
  xlab("Julian Date") +
  labs(title="",  # Number of Samples per Year by Date and Origin
       fill = "Origin") +
  theme(text = element_text(size = 20))  # strip.text = element_text(size = 12)
ggsave("samples_even.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")
```

Create bar plot of odd-year sampling.
```{r odd_samp_png}
paired_13_15_filter <- read_csv("stock_paired_13_15_filter.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

#Stacked
paired_13_15_filter %>% 
  ggplot(aes(x = `Julian`)) +
  theme_bw() +
  geom_bar(aes(fill = origin), colour = "black") +
  facet_grid(`Sample Year` ~ .) +
  geom_hline(yintercept = 0, colour = "black") +
  ylab("Count") +
  xlab("Julian Date") +
  labs(title="",  # Number of Samples per Year by Date and Origin
       fill = "Origin") +
  theme(text = element_text(size = 20))  # strip.text = element_text(size = 12)
ggsave("samples_odd.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")
```


# Figure 3: Reproductive success

Create blank even-year figure as .png.
```{r rs_even_blank_png}
paired_14_16_filter_parents <- read_csv("stock_paired_14_16_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

paired_14_16_filter_parents %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female")) %>% 
  count(sex, origin, n) %>% 
  group_by(sex, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  mutate(p = 0) %>% 
  ggplot(aes(x=n, y=p, fill = origin)) +
  theme_bw() +
  geom_col(position = position_dodge2(preserve="single"), width=0.8, colour = "black") +
  geom_hline(yintercept = 0, colour = "black") +
  ylim(c(0, 0.77)) +
  # scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  xlab("Number of Offspring") +
  ylab("Proportion of Parents") +
  labs(title="",
       fill = "Parent\nOrigin") +
  facet_grid(. ~ sex) +
  theme(text = element_text(size = 20))
  ggsave("RS_Blank_Facet.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")
```

Create even-year figure with just hatchery females as .png.
```{r rs_even_female_hatchery_png}
paired_14_16_filter_parents <- read_csv("stock_paired_14_16_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

paired_14_16_filter_parents %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female")) %>% 
  count(sex, origin, n) %>% 
  group_by(sex, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  mutate(p = case_when(sex == "Male" ~ 0,
                       origin == "Natural" & sex == "Female" ~ 0,
                       origin == "Hatchery" & sex == "Female" ~ p)) %>% 
  ggplot(aes(x=n, y=p, fill = origin)) +
  theme_bw() +
  geom_col(position = position_dodge2(preserve="single"), width=0.8, colour = "black") +
  geom_hline(yintercept = 0, colour = "black") +
  ylim(c(0, 0.77)) +
  # scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  xlab("Number of Offspring") +
  ylab("Proportion of Parents") +
  labs(title="",
       fill = "Parent\nOrigin") +
  facet_grid(. ~ sex) +
  theme(text = element_text(size = 20))
  ggsave("RS_Even_Female_Hatchery_Facet.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")
```

Create even-year figure with just females as .png.
```{r rs_even_female_png}
paired_14_16_filter_parents <- read_csv("stock_paired_14_16_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

paired_14_16_filter_parents %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female")) %>% 
  count(sex, origin, n) %>% 
  group_by(sex, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  mutate(p = case_when(sex == "Male" ~ 0,
                       sex == "Female" ~ p)) %>% 
  ggplot(aes(x=n, y=p, fill = origin)) +
  theme_bw() +
  geom_col(position = position_dodge2(preserve="single"), width=0.8, colour = "black") +
  geom_hline(yintercept = 0, colour = "black") +
  ylim(c(0, 0.77)) +
  # scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  xlab("Number of Offspring") +
  ylab("Proportion of Parents") +
  labs(title="",
       fill = "Parent\nOrigin") +
  facet_grid(. ~ sex) +
  theme(text = element_text(size = 20))
  ggsave("RS_Even_Female_Facet.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")
```

Create even-year figure as .png.
```{r rs_even_png}
paired_14_16_filter_parents <- read_csv("stock_paired_14_16_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

paired_14_16_filter_parents %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female")) %>% 
  count(sex, origin, n) %>% 
  group_by(sex, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x=n, y=p, fill = origin)) +
  theme_bw() +
  geom_col(position = position_dodge2(preserve="single"), width=0.8, colour = "black") +
  geom_hline(yintercept = 0, colour = "black") +
  ylim(c(0, 0.77)) +
  # scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  xlab("Number of Offspring") +
  ylab("Proportion of Parents") +
  labs(title="",
       fill = "Parent\nOrigin") +
  facet_grid(. ~ sex) +
  theme(text = element_text(size = 20))
  ggsave("RS_Even_Facet.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")
```

Create odd-year figure as .png.
```{r rs_odd_png}
paired_13_15_filter_parents <- read_csv("stock_paired_13_15_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery")))

paired_13_15_filter_parents %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female")) %>% 
  count(sex, origin, n) %>% 
  group_by(sex, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x=n, y=p, fill = origin)) +
  theme_bw() +
  geom_col(position = position_dodge2(preserve="single"), width=0.8, colour = "black") +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  xlab("Number of Offspring") +
  ylab("Proportion of Parents") +
  labs(title="",
       fill = "Parent\nOrigin") +
  facet_grid(. ~ sex) +
  theme(text = element_text(size = 20))
  ggsave("RS_Odd_Facet.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")
```

# Figure 4: RS by length and time

```{r RS by length and time Even}
paired_14_16_filter_parents <- read_csv("stock_paired_14_16_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

paired_14_16_filter_parents %>%
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(height = 0, size = 2) +
  geom_smooth() +
  facet_grid(sex ~ .)+
  labs(title = "",  colour = "Parent\nOrigin",
       shape = "Parent: Sex") +
  xlab("Parent Sample Date") +
  ylab("Number of Offspring") +
  theme_bw() +
  theme(text = element_text(size = 20))
  ggsave("RS_Even_Date.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")

paired_14_16_filter_parents %>% 
  filter(`Length Mm` > 300) %>%  # remove outlier
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(height = 0, size = 2) +
  geom_smooth() +
  facet_grid(sex ~ .) +
  labs(title = "",  colour = "Parent\nOrigin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Number of Offspring") +
  theme_bw() +
  theme(text = element_text(size = 20))
  ggsave("RS_Even_Length.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")
```


# Figure 5: RS by distance

```{r RS by distance Even}
riverdist_parentage_stock <- read_csv("riverdist_parentage_stock.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

riverdist_parentage_stock %>%
  ggplot(aes(x = Distance, y = n, color = origin)) +
  geom_vline(aes(xintercept = tidedist), size = 2) +
  geom_jitter(height = 0, size = 2) +
  geom_smooth() +
  facet_grid(sex ~ .)+
  labs(title = "",  colour = "Parent\nOrigin",
       shape = "Parent: Sex") +
  xlab("Parent Sample Location (m from mouth)") +
  ylab("Number of Offspring") +
  theme_bw() +
  theme(text = element_text(size = 20))
ggsave("RS_Even_Distance.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/", width = 7.29, height = 4.5, units = "in")
```

# Sampling representativeness

```{r}
(survey <- read_csv("stockdale_stream_survey_2014_2016.csv") %>% 
   select(date, alive, dead, samples))

survey %>% 
  gather(type, count, -date) %>% 
  mutate(date = mdy(date)) %>% 
  mutate(julian = yday(date)) %>% 
  mutate(year = year(date)) %>% 
  drop_na() %>% 
  ggplot(aes(x = julian, y = count, colour = type)) +
  geom_line(lwd = 2) +
  facet_grid(year ~ .) +
  labs(title = "",  colour = "Count Type") +
  xlab("Parent Sample Date (Julian)") +
  ylab("Count") +
  theme_bw() +
  theme(text = element_text(size = 20))
ggsave("Stockdale Sampling.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/", width = 7.29, height = 4.5, units = "in")
```


# Figure S2: RS for triads  

```{r RS for triads}
parents_paired_14_16_cross <- read_csv("stock_parents_paired_14_16_cross.csv")

parents_paired_14_16_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  count(cross, n) %>% 
  complete(cross, n, fill = list(nn = 0)) %>% 
  group_by(cross) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = cross))+
  geom_col(position = position_dodge2(preserve = "single"), width = 0.5) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = 1:5) +
  # ylim(0, 40) +
  # xlim(0, 5) +
  labs(title = "",
       fill = "Cross Type") +
  xlab("Number of Offspring")+
  ylab("Proportion of Families") +
  theme_bw() +
  theme(text = element_text(size = 20))
ggsave("RS_cross.png", path = "V:/Documents/5_Coastwide/Multispecies/AHRP/Powerpoints/Brown Bags/Figures/")
```
```{r}
paired_14_16_filter_parents %>% 
  group_by(origin, `Otolith Mark ID`) %>% 
  summarise(n = sum(n))

paired_14_16_filter %>% 
  filter(`Sample Year` == 2014) %>% 
  count(origin, `Otolith Mark ID`)
```

```{r}
paired_14_16_filter_parents %>% filter(n > 0) %>% 
```

