---
title: "Stockdale + Hogan Even (2014/16) Parentage"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(coin)
library(MASS)
library(scales)
library(tidyverse)
library(leaflet)
library(lubridate)
library(ggExtra)
```

Below are analyses of parentage results for all Hogan and Stockdale Even (2014-2016) lineage samples considered together as potential parents and offspring using *FRANz*.

# Even lineage (2014/2016)

## *FRANz* parameters
*FRANz* was run at 11:16 on September 20, 2019. We used the following parameters:
FRANz.exe --Nmmax 6519 --Nfmax 6519 --femrepro 1:2 --malerepro 1:2 --typingerror 0.006 --updatefreqs --poutformat 2  "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\2014_2016_stockdale_hogan\stockdale_hogan_postQA_2014_2016_STOCK.dat"

The parameters are defined as follows:
--Nmmax and --Nfmax are the maximum numbers of candidate mothers and fathers. To obtain our values, I added the values we used for Hogan and Stockdale together 
--femrepro and --malerepro specify the age range in which an individual can reproduce  
--typingerror refers to the overall genotyping error rate. Ours was ~0.005 (for Stockdale was actually 0.006, from QC)
--updatefreqs specifies that *FRANz* should update allele frequencies using MCMC sampling  
--poutformat specifies that all potential parents should be listed, not just the most likely  

All output files can be found here: "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\Stockdale\2014_2016_stockdale_hogan"  

The summary file provides information about the power of our marker suite:   
Cumulative exclusion probability when 1 to 7 fullsibs are genotyped  
  First Parent              : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Second Parent             : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Parent Pair               : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000
  
According to the FRANz manual, marker sets are not considered powerful if these cumulative exclusion probabilities are less than 0.95, which indicates that the probability that a random pair of individuals in the population has a 5% chance of having a genotype pair compatible to an offspring genotype. Since all of our probabilities are 1, we can be **confident** in the power of our 298 amplicons to make parent assignments.  

## Import data
The first step is to read in .csv files for parentage assignments produced by *FRANz* as well as paired genotype and OceanAK data. 

```{r Setup Even, message=FALSE, warning=FALSE}
parentage_14_16 <- read_csv("../../Franz/2014_2016_stockdale_hogan/parentage.csv")
paired_14_16 <- bind_rows(
  read_csv("../Franz/stockdale_postQA_OceanAK_paired_2014_2016_STOCK.csv"),
  read_csv("../Franz/hogan_postQA_OceanAK_paired_2014_2016_HOGAN.csv"))
```

Plot a histogram of *FRANz* parentage posterior probabilities to show robustness of assignments
```{r FRANz_posterior_plot_even}
parentage_14_16  %>% 
  filter(!is.na(`Parent 1`)) %>% 
  ggplot(aes(x = Posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  ggtitle("Histogram of FRANz posterior probabilities for parentage assignments")
```

How many mismatches?
```{r}
parentage_14_16  %>% 
  filter(!is.na(`Parent 1`)) %>% 
  filter(Posterior > 0.9) %>% 
  count(Mismatches)
```


All parentage assigments have a posterior probability of 1, which is very robust.  

## Filter paired data
The file containing paired genotype and OceanAK data has a lot of information that we do not need at this time (e.g. genotypes for each marker). We therefore separate out only the columns that contain identifying information for each individual.

```{r Filter paired data Even}
# What are the non-genotype columns
grep(pattern = "RAD", x = colnames(paired_14_16), value = TRUE, invert = TRUE)

# Filter for non-genotype columns
paired_14_16_filter <- paired_14_16 %>% 
  dplyr::select(franz_id, SILLY, `Fish ID`, `DNA Tray Code`, `DNA Tray Well Code`, `Sample Year`, `Sample Date`, SEX, `Length Mm`, `Otolith Mark Present`, `Otolith Mark ID`) %>% 
  mutate(origin = case_when(`Otolith Mark Present` == "NO" ~ "Natural",
                            `Otolith Mark Present` == "YES" ~ "Hatchery")) %>%  # add origin variable
  mutate(origin = factor(origin, c("Natural", "Hatchery"))) %>%   # make factor to ensure hatchery != red
  mutate(stream = case_when(str_sub(string = SILLY, start = 1, end = 2) == "PS" ~ "Stockdale",
                            str_sub(string = SILLY, start = 1, end = 2) == "PH" ~ "Hogan"))
```

## Filter for "extra" duplicates
Filter for the "extra" duplicates that made it in to the parentage analysis that we need to remove before proceeding with RRS.

```{r dups_even}
duplicates_to_remove <-
  bind_rows (
    dget("../../Objects/Hogan/duplicates_to_remove.txt"),
    dget("../../Objects/Stockdale/duplicates_to_remove.txt")
  )
    
# Join `duplicates_to_remove` with the paired data, filter for only the "extra" duplicates that were NOT removed
(duplicates_to_remove_even_extra <- duplicates_to_remove %>% 
  dplyr::filter(SILLY_CODE %in% c("PHOGAN14", "PHOGAN16", "PSTOCK14", "PSTOCK16")) %>% 
  dplyr::mutate(`DNA_TRAY_CODE` = as.numeric(`DNA_TRAY_CODE`)) %>% 
  dplyr:: mutate(`DNA_TRAY_WELL_CODE` = as.numeric(`DNA_TRAY_WELL_CODE`)) %>% 
  dplyr::left_join(paired_14_16_filter, by = c("DNA_TRAY_CODE" = "DNA Tray Code", "DNA_TRAY_WELL_CODE" = "DNA Tray Well Code")) %>% 
  tidyr::drop_na(franz_id))

dups_even_franz_id <- duplicates_to_remove_even_extra %>% 
  pull(franz_id)

# Remove these "extra" duplicates from `paired_14_16_filter`
paired_14_16_filter <- paired_14_16_filter %>% 
  filter(!franz_id %in% dups_even_franz_id)
```

## Filter parentage data
```{r Filter parentage data Even}
# Filter for posterior > 0.9 + make tidy relative to assigment (1 row per parentage assignment)
parentage_14_16_filter <- parentage_14_16 %>% 
  filter(Posterior > 0.9) %>% 
  dplyr::select(Offspring, `Parent 1`, `Parent 2`) %>% 
  gather(Parent, Parent_ID, - Offspring) %>% 
  filter(!is.na(Parent_ID))
```

## Sample sizes
Quick look at what our sample sizes are for potential parents (2014) and potential offspring (2016).

```{r even_sample_sizes}
paired_14_16_filter %>% 
  count(`Sample Year`, stream, SEX, origin) %>% 
  spread(origin, n, fill = 0)
```


## Join parentage and individual data
Here, we joined parentage data from *FRANz* with individual data so that we can match data to each offspring and parent. **Note** that any column ending in `.off` refers to offspring and `.par` refers to parent.  

```{r Join parentage and paired objects Even}
# Creating a single, "tidy" object where each row is a parent-offspring relationship
# Offspring with 2 parents have 2 rows
# .off data is offspring, .par data is parents
(parents_paired_14_16 <- parentage_14_16_filter %>% 
   left_join(paired_14_16_filter, by = c("Offspring" = "franz_id")) %>% 
   left_join(paired_14_16_filter, by = c("Parent_ID" = "franz_id"), suffix = c(".off", ".par")) %>% 
   mutate(origin = case_when(`Otolith Mark Present.par` == "NO" ~ "Natural",
                             `Otolith Mark Present.par` == "YES" ~ "Hatchery")) %>% 
   mutate(origin = factor(origin, c("Natural", "Hatchery"))))  # added hatchery/natural variable

# How many single-parent offspring pair assignments?
parents_paired_14_16 %>% 
  count(Parent)

# How many unique parents had offspring assigned?
n_distinct(parents_paired_14_16$Parent_ID)

# For each offspring assigment, what was the parent's origin?
parents_paired_14_16 %>% 
  count(origin, `Otolith Mark ID.par`) %>% 
  spread(origin, n)
```

**Note** that *FRANz* assigned 1,744 offspring to 596 parents. 1,093 of these offsrping have natural-origin parents and 651 offspring have hatchery-origin parents. The hatchery origin fish are from AFK12B, CCH12, and WNH12PINKB.

How many offspring had parents from each stream?
```{r}
parents_paired_14_16 %>% 
  count(stream.par, origin.par, stream.off) %>% 
  spread(stream.off, n, fill = 0)
```

Filter for stray parent-offspring relationships
```{r}
parents_paired_14_16 %>% 
  filter(stream.par != stream.off)
```

Filter for the parent pair that made a stray
```{r}
parents_paired_14_16 %>% 
  filter(Parent_ID %in% c("PS14_01076", "PS14_01508"))
```

Get parents that produced strays
```{r}
stray_parents <- parents_paired_14_16 %>% 
  filter(stream.par != stream.off) %>% 
  select(Parent_ID) %>% 
  distinct() %>% 
  pull
```

```{r}
parents_paired_14_16 %>% 
  filter(Parent_ID %in% stray_parents) %>% 
  group_by(Parent_ID, stream.par, origin.par) %>% 
  count(stream.off) %>% 
  spread(stream.off, n, fill = 0)
```




## Proportion test
Let's use a Chi-Square test to determine whether the proportions of offspring assigned to hatchery- and natural-origin parents significantly differ from the proportions of hatchery- and natural-origin parents sampled. 

```{r proportion_even_data}
# How many spawners from each origin?
paired_14_16_filter %>% 
  filter(`Sample Year` == 2014) %>% 
  count(origin)

# How many offspring from each origin?
parents_paired_14_16 %>% 
  count(origin)
```

```{r Chi square test even}
O.even <- c("H", "N")
Spawners.even <- c(873, 572) #  number of hatchery- and natural-origin parents
Offspring.even <- c(651, 1093)  # number of offspring assigned to hatchery- and natural-origin parents

df.even <- data.frame(Spawners.even, Offspring.even)  # made a data.frame because tibbles don't like rownames
row.names(df.even) <- O.even

df.even
chisq.test(df.even)
```