---
title: "Hogan 2013-2016 Results for Science Panel"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(coin)
library(MASS)
library(scales)
library(tidyverse)
library(lubridate)
# library(leaflet)
library(lubridate)
library(PBSmapping)
library(grid)
library(ggthemes) # Load

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, out.width = "100%", fig.width = 10)
```

# Introduction
This is a summary of parentage analysis and relative reproductive success (RRS) results from Hogan Bay 2013-2016 for the Science Panel. The even year lineage had many more parentage assignments than the odd year lineage due to the higher sampling rates in 2014 compared to 2013 and the smaller escapements in 2014/2016 than in 2013/2015. While this does not mean that the odd lineage results are not valid, they are certainly not as robust as the even lineage results. These results will be used to inform additional analyses  (e.g. investigating the impact of spawning location on reproductive success). We intend to present these results at both the March 7th AHRP Informational Meeting and at the Alaska Chapter of the American Fisheries Society meeting in Sitka March 19-22. These analyses were funded by the North Pacific Research Board. 

# Odd lineage (2013/2015)

## Sample sizes
Below is a comparison of sample dates for genotyped individuals collected in 2013 and 2015. These plots allow us to examine differences in run timing between hatchery- and natural-origin fish in the parent generation (2013) and look for completeness of sampling throughout the run for both years. Note that only natural-origin individuals collected in 2015 were genotyped as potential offspring.  

From this histogram, we can see that in the parent generation (2013), natural-origin fish tended to have earlier sampling dates than hatchery-origin fish, which confirms what we know about differences in run timing. We can also see how sparse the field sampling was in 2013 compared to 2015. Offspring in 2015 were sampled throughout the run.

```{r odd_sample_sizes_plot}
paired_13_15_filter <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Hogan/304/paired_13_15_filter.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(Sex = case_when(SEX == "M" ~ "Male", SEX == "F" ~ "Female"))

paired_13_15_filter %>% 
  ggplot(aes(x = `Julian`)) +
  theme_bw() +
  geom_bar(aes(fill = origin), colour = "black") +
  facet_grid(`Sample Year` ~ .) +
  geom_hline(yintercept = 0, colour = "black") +
  ylab("Count") +
  xlab("Julian Date") +
  labs(title= "Number of Samples per Year\nby Date and Origin",
       fill = "Origin") +
  theme(text = element_text(size = 20))
```

Here are our sample sizes are for potential parents (2013) and potential offspring (2015) by sex and origin.

```{r odd_sample_sizes}
paired_13_15_filter %>% 
  count(`Sample Year`, Sex, origin) %>% 
  spread(origin, n, fill = 0)
```

Below are analyses of parentage results calculated separately for Odd (2013-2015) lineage using *FRANz*. 

## *FRANz* parameters
*FRANz* was run on November 6, 2018. We used the following parameters:
FRANz.exe --Nmmax 23500 --Nfmax 23500 --femrepro 1:2 --malerepro 1:2 --typingerror 0.005 --updatefreqs --poutformat 2 --fullsibtest "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\hogan_postQA_2013_2015_HOGAN.dat"

The parameters are defined as follows:  
--Nmmax and --Nfmax are the maximum numbers of candidate mothers and fathers. To obtain our values, we used an estimated escapement of 4,150 and divided by 2. This escapement number came from the smaller of the two estimates (aerial survey AUC and stream walk AUC). This estimate is listed under 'pop est aerial' in the Excel file found here: "V:\Documents\5_Coastwide\Multispecies\AHRP\Field data\Sample Summary Extract and Genotype.xlsx"  
--femrepro and --malerepro specify the age range in which an individual can reproduce  
--typingerror refers to the overall genotyping error rate. Ours was ~0.005  
--updatefreqs specifies that *FRANz* should update allele frequencies using MCMC sampling  
--poutformat specifies that all potential parents should be listed, not just the most likely  
--fullsibtest tests for full siblings among offspring  

All output files can be found here: "v:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\hogan_postQA_2013_2015_HOGAN_Run1"  

The summary file provides information about the power of our marker suite:   
Cumulative exclusion probability when 1 to 7 fullsibs are genotyped  
  First Parent              : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Second Parent             : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Parent Pair               : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000
  
According to the *FRANz* manual, marker sets are not considered powerful if these cumulative exclusion probabilities are less than 0.95, which indicates that the probability that a random pair of individuals in the population has a 5% chance of having a genotype pair compatible to an offspring genotype. Since all of our probabilities are 1, we can be **confident** in the power of our 298 amplicons to make parent assignments. 

Below is a histogram of *FRANz* parentage posterior probabilities to show robustness of assignments
```{r FRANz_posterior_plot_odd}
parentage_13_15 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Franz/304/hogan_postQA_2013_2015_HOGAN_Run1/parentage_304.csv")

parentage_13_15  %>% 
  filter(!is.na(`Parent 1`)) %>% 
  ggplot(aes(x = Posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  ylab("Count") +
  xlab("Posterior Probability") +
  ggtitle("Histogram of FRANz posterior probabilities for parentage assignments")
```

All parentage assigments have a posterior probability of 1.00, which means that no offspring assignments were split among multiple potential parents. 

Of the 1,919 offspring collected in 2015, 48 were assigned to parents, for an overall offspring assignment rate of **2.5%**. This assignment rate is in line with expectations based on the sampling rate of the parent year escapement in 2013. All assignments were to one parent (*single parent-offpsring duos*) - there were no two-parent assignments (*parents pair-offspring trios*).

## Proportion test
We used a Chi-Square test to determine whether the proportions of offspring assigned to hatchery- and natural-origin parents significantly differ from the proportions of hatchery- and natural-origin parents sampled. 

```{r Chi square test odd}
paired_13_15_filter_parents <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Hogan/paired_13_15_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

O.odd <- levels(paired_13_15_filter$origin)
Spawners.odd <- paired_13_15_filter %>% 
  filter(`Sample Year` == 2013) %>% 
  count(origin) %>% 
  pull(n) #  number of hatchery- and natural-origin parents
Offspring.odd <- paired_13_15_filter_parents %>% 
  group_by(origin) %>% 
  summarise(n = sum(n)) %>% 
  pull(n)  # number of offspring assigned to hatchery- and natural-origin parents

df.odd <- data.frame(Spawners.odd, Offspring.odd)  # made a data.frame because tibbles don't like rownames
row.names(df.odd) <- O.odd

df.odd
chisq.test(df.odd)
```

There is a significant difference in the proportions of offspring assigned to hatchery- and natural-origin parents relative to the proportions of potential parents sampled, indicating an under-representation of offspring assigning to hatchery-origin parents. We see that although **`r round(df.odd["H", "Spawners.odd"] / sum(df.odd[, "Spawners.odd"]), 2) * 100`%** of parents genotyped were of hatchery-origin, only **`r round(df.odd["H", "Offspring.odd"] / sum(df.odd[, "Offspring.odd"]), 2) * 100`%** of offspring assigned to hatchery-origin parents.

## Reproductive success
Here, we calculate and plot reproductive success (RS; number of offspring assigned to parents) for both natural- and hatchery-origin parents. The number of offspring assigned ranges from 0-4. 

### Calculate mean RS
We calculated the mean RS by origin and sex including sampled fish from the parental generation to which offspring were not assigned.

```{r calculate RS odd}
# Mean RS including 0's
rs_13_15_0 <- paired_13_15_filter_parents %>%
  group_by(origin, sex) %>% 
  summarise(RS = mean(n, na.rm = TRUE))

rs_13_15_0 %>%
  mutate(RS = round(RS, 2)) %>% 
  spread(sex, RS)
```

### Distribution of RS
```{r plot RS odd}
# Plot histogram of proportion of parents with a given family size
paired_13_15_filter_parents %>% 
  count(sex, origin, n) %>%
  group_by(sex, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  theme_bw() +
  geom_col(position = position_dodge2(preserve="single"), width=0.8, colour = "black") +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  facet_grid(~ sex) +
  labs(title="Distribution of Reproductive Success",
       fill = "Parent\nOrigin") +
  xlab("Number of Offspring")+
  ylab("Proportion of Parents") +
  theme(text = element_text(size = 20))

# Save RS values
F_h_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Hatchery" & sex == "Female") %>% 
  pull(RS)
M_h_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Hatchery" & sex == "Male") %>% 
  pull(RS)
F_n_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Natural" & sex == "Female") %>% 
  pull(RS)
M_n_0_13_15 <- rs_13_15_0 %>% 
  filter(origin == "Natural" & sex == "Male") %>% 
  pull(RS)
```

Most individuals were assigned 0 offspring. This doesn't mean that their fitness is necessarily 0 - rather, it means that we did not sample offspring that were assigned to them. We need to bear in mind that our RS estimates are based on the samples that we genotype. 

### Calculate and test RRS
```{r Calculate relative reproductive success including 0s odd}
Male_RRS_odd <- M_h_0_13_15/M_n_0_13_15
Female_RRS_odd <- F_h_0_13_15/F_n_0_13_15

# RRS_tibble_0_13_15 <- tibble(Male_RRS_odd, Female_RRS_odd)
# round(RRS_tibble_0_13_15, 2)
```

We calculate RRS as mean Hatchery RS divided by mean Natural RS. 

**RRS for females = `r round(Female_RRS_odd, 2)`**

**RRS for males = `r round(Male_RRS_odd, 2)`.**

#### Statistical testing
We tested for statistically significant differences in RS between male and female hatchery- and natural-origin fish using two different statistical tests:  

  1. Nonparametric permutation test
  2. Parametric negative binomial GLM  
  
```{r RRS statistical testing including 0s odd}
# The format of the data to test was very simple, it was a data.frame called “mydata” with 2 columns
# 1.	nOff = number of offspring per family
# 2.	Origin = “H” or “W” for if the parent was hatchery or natural (wild)
# 
mydata_M_0 <- paired_13_15_filter_parents %>%
  filter(SEX == "M") 
  

mydata_F_0 <- paired_13_15_filter_parents %>%
  filter(SEX == "F")
  

perm_1tail_pvalue_M_0 <- round(x = as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_M_0, distribution = approximate(B = 10000), alternative = "greater"))), digits = 4)

perm_1tail_pvalue_F_0 <- round(x = as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_F_0, distribution = approximate(B = 10000), alternative = "greater"))), digits = 4)

#And to do a 1-tailed negative binomial GLM

fit_M_0 <- glm.nb(n ~ origin, data = mydata_M_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_M_0 <- round(summary(fit_M_0)$coefficients[2, 4], 4)

fit_F_0 <- glm.nb(n ~ origin, data = mydata_F_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_F_0 <- round(summary(fit_F_0)$coefficients[2, 4], 4)

RRS_0 <- tibble(Sex = c("Male", "Female", "Male", "Female"),
                Test = c(rep("Permutation", 2), rep("Negative Binomial GLM", 2)),
                p_value = c(perm_1tail_pvalue_M_0, perm_1tail_pvalue_F_0, nbGLM_1tail_pvalue_M_0, nbGLM_1tail_pvalue_F_0))
RRS_0 %>% 
  spread(Sex, p_value)  
```

The above table shows the *p-values* for each test and sex. We use a *p-value* cut-off of 0.05. Both male and female hatchery-origin fish produce significantly fewer offspring than their natural-origin counterparts.  

#### Confidence intervals

We calculated 95% CI's for RRS based on [Kalinowski and Taper 2005](http://www.nrcresearchpress.com/doi/10.1139/f04-239). This is a maximum likelihood estimate based off of Hinrichsen 2003. This method has been used in other peer reviewed papers such as [Ford et al. 2016](https://www.researchgate.net/publication/309147116_Broodstock_History_Strongly_Influences_Natural_Spawning_Success_in_Hatchery_Steelhead_Oncorhynchus_mykiss#pf14).

```{r RRS_confidence_intervals_function}
rrs_ci_kalinowski <- function(n_h_off, n_w_off, n_h_par, n_w_par, alpha){
  chi_alpha <- qchisq(p = (1 - alpha), df = 1)
  n_off <- sum(c(n_h_off, n_w_off))
  n_par <- sum(c(n_h_par, n_w_par))
  
  rs_h <- n_h_off / n_h_par
  rs_w <- n_w_off / n_w_par
  
  p_h_par <- n_h_par / n_par
  p_w_par <- n_w_par / n_par
  
  rrs_h <- rs_h / rs_w
  rrs_w <- rs_w / rs_w
  rrs_avg <- (rrs_h * p_h_par) + (rrs_w * p_w_par)
  
  rrs_ml <- (n_h_off * log(p_h_par * rrs_h / rrs_avg)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg))
  
  xi_dist <- bind_rows(
    lapply(seq(0.01, 5, by = 0.01), function(rrs_h_xi) {
      rrs_avg_xi <- (rrs_h_xi * p_h_par) + (rrs_w * p_w_par)
      tibble(rrs_crit = rrs_h_xi,
             logl = (n_h_off * log(p_h_par * rrs_h_xi / rrs_avg_xi)) + (n_w_off * log(p_w_par * rrs_w / rrs_avg_xi)) - (rrs_ml - chi_alpha / 2)
      )
    } )
  )
  
  rrs_min <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit < rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  rrs_max <- xi_dist %>% 
    mutate(abs_logl = abs(logl)) %>% 
    filter(rrs_crit > rrs_h) %>% 
    top_n(-1, abs_logl) %>% 
    pull(rrs_crit)
  
  xi_plot <- xi_dist %>% 
    ggplot(aes(x = rrs_crit, y = logl)) +
    geom_line() +
    geom_hline(yintercept = 0, colour = "red", lwd = 2) +
    geom_vline(xintercept = c(rrs_h, rrs_min, rrs_max), colour = "blue") +
    ylim(c(-5, 5)) +
    xlim(c(0, 2)) +
    ylab("Log Likelihood - Chi Sq Value") +
    annotate("text", x = rrs_h + 0.1, y = xi_dist %>% filter(rrs_crit == xi_dist$rrs_crit[which.min(abs(xi_dist$rrs_crit  - rrs_h))]) %>% pull(logl) + 0.4, label = round(rrs_h, 2)) +
    annotate("text", x = rrs_min - 0.1, y = xi_dist %>% filter(rrs_crit == rrs_min) %>% pull(logl) + 0.4, label = rrs_min) +
    annotate("text", x = rrs_max + 0.1, y = xi_dist %>% filter(rrs_crit == rrs_max) %>% pull(logl) + 0.4, label = rrs_max)
  
  # print(xi_plot)  # hide this for Science Panel notebook
  return(round(c(rrs_min, rrs_h, rrs_max), 2))
}
```

```{r RRS_confidence_intervals_13_15}
odd_female_rrs_CI <- rrs_ci_kalinowski(n_h_off = paired_13_15_filter_parents %>% 
                                         filter(SEX == "F", origin == "Hatchery") %>% 
                                         group_by(SEX, origin) %>% 
                                         summarise(n_off = sum(n)) %>% 
                                         pull(n_off),
                                       n_w_off = paired_13_15_filter_parents %>% 
                                         filter(SEX == "F", origin == "Natural") %>% 
                                         group_by(SEX, origin) %>% 
                                         summarise(n_off = sum(n)) %>% 
                                         pull(n_off), 
                                       n_h_par = paired_13_15_filter_parents %>% 
                                         filter(SEX == "F", origin == "Hatchery") %>% 
                                         count(SEX, origin) %>% 
                                         pull(nn), 
                                       n_w_par = paired_13_15_filter_parents %>% 
                                         filter(SEX == "F", origin == "Natural") %>% 
                                         count(SEX, origin) %>% 
                                         pull(nn),
                                       alpha = 0.05)

odd_male_rrs_CI <- rrs_ci_kalinowski(n_h_off = paired_13_15_filter_parents %>% 
                                       filter(SEX == "M", origin == "Hatchery") %>% 
                                       group_by(SEX, origin) %>% 
                                       summarise(n_off = sum(n)) %>% 
                                       pull(n_off),
                                     n_w_off = paired_13_15_filter_parents %>% 
                                       filter(SEX == "M", origin == "Natural") %>% 
                                       group_by(SEX, origin) %>% 
                                       summarise(n_off = sum(n)) %>% 
                                       pull(n_off), 
                                     n_h_par = paired_13_15_filter_parents %>% 
                                       filter(SEX == "M", origin == "Hatchery") %>% 
                                       count(SEX, origin) %>% 
                                       pull(nn), 
                                     n_w_par = paired_13_15_filter_parents %>% 
                                       filter(SEX == "M", origin == "Natural") %>% 
                                       count(SEX, origin) %>% 
                                       pull(nn),
                                     alpha = 0.05)
```

Hogan 2013/2015 **female** RRS with 95% CI = **`r paste0(odd_female_rrs_CI[2], " (", odd_female_rrs_CI[1], "-", odd_female_rrs_CI[3], ")")`**

Hogan 2013/2015 **male** RRS with 95% CI = **`r paste0(odd_male_rrs_CI[2], " (", odd_male_rrs_CI[1], "-", odd_male_rrs_CI[3], ")")`**

### Plots of RS vs. field data
We examined associations between two potential explanatory variables (sample date and body size) and RS. These plots include LOESS (locally weighted smoothing) lines with 95% confidence intervals to help illustrate trends. We use LOESS rather than linear models because our data violate assumptions of linear regression (normality, homoscedasticityc, etc.). 

```{r Examine reproductive success by sample date odd with zeros}
paired_13_15_filter_parents %>%
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(height = 0, size = 2) +
  geom_smooth() +
  facet_grid(sex ~ .)+
  labs(title = "Reproductive Success by Parent Sample Date",  colour = "Parent\nOrigin",
       shape = "Parent: Sex") +
  xlab("Parent Sample Date") +
  ylab("Number of Offspring") +
  theme_bw() +
  theme(text = element_text(size = 20))
```

The above plot show RS by parent sample date, separately for females (top) and males (bottom). We see that there were no hatchery-origin fish sampled in the first event. RS for natural-origin males appears consistent throughout the season, whereas female natural-origin fish appear to be more productive early in the season than later on. Note that points are jittered by sample date (x-axis) to avoid overplotting. Not enough offspring were assigned to hatchery-origin parents to draw conclusions about RS and sample date. 

```{r Examine reproductive success by parent length odd with zeros}
paired_13_15_filter_parents %>% 
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(height = 0, size = 2) +
  geom_smooth() +
  facet_grid(sex ~ .) +
  labs(title = "Reproductive Success by Parent Length",  colour = "Parent\nOrigin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Number of Offspring") +
  theme_bw() +
  theme(text = element_text(size = 20))
```

The above plot show RS by parent body length, separately for females (top) and males (bottom). It is difficult to draw conclusions about this relationship for hatchery-origin parents since there were so few assignments, which makes it difficult to draw robust conclusions.

## Summary of odd lineage results
1. Exclusion probabilities were equal to 1.00, which means we can be confident in the ability of our marker set to correctly assign parents to offspring.  
2. 48 offspring were assigned to 20 parents (3 of which were hatchery-origin), for an assignment rate of **2.5%**. This is in line with our expectation given the sampling rate of the parent year escapement in 2013.  
3. Given the small sample size of parent-offspring relationships for the odd lineage, these results should be considered valid, but not particularly robust.
4. Female RRS was 0.02 (0.01-0.09). Male RRS was 0.12 (0.02-0.40). Both male and female hatchery-origin fish produced significantly fewer offspring than natural-origin fish.  
5. The number of offpsring assigned to parents varied from 0-16 for females and 0-4 for males.  
6. It is difficult to draw clear conclusions about relationships between RS and either sample date or parental size due to the low offspring assignment rate.   


# Even lineage (2014/2016)

## Sample sizes
Below is a comparison of sample dates for genotyped individuals collected in 2014 and 2016. These plots allow us to examine differences in run timing between hatchery- and natural-origin fish in the parent generation (2014) and look for completeness of sampling throughout the run for both years. Note that only natural-origin individuals collected in 2016 were genotyped as potential offspring.  

From this histogram, we can see that in the parent generation (2014), natural-origin fish tended to have earlier sampling dates than hatchery-origin fish, which confirms what we know about differences in run timing. We can also see that field sampling in 2014 was more robust than 2013 (more stream visits). Offspring in 2016 were sampled throughout the run.

```{r even_sample_sizes_plot}
paired_14_16_filter <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Hogan/paired_14_16_filter.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(Sex = case_when(SEX == "M" ~ "Male", SEX == "F" ~ "Female"))

paired_14_16_filter %>% 
  ggplot(aes(x = `Julian`)) +
  theme_bw() +
  geom_bar(aes(fill = origin), colour = "black") +
  facet_grid(`Sample Year` ~ .) +
  geom_hline(yintercept = 0, colour = "black") +
  ylab("Count") +
  xlab("Julian Date") +
  labs(title= "Number of Samples per Year\nby Date and Origin",
       fill = "Origin") +
  theme(text = element_text(size = 20))
```

Here are our sample sizes are for potential parents (2014) and potential offspring (2016) by sex and origin.

```{r even_sample_sizes}
paired_14_16_filter %>% 
  count(`Sample Year`, Sex, origin) %>% 
  spread(origin, n, fill = 0)
```

Below are analyses of parentage results calculated separately for Even (2014-2016) lineage using *FRANz*. 

## *FRANz* parameters
*FRANz* was run on November 6, 2018. We used the following parameters:
FRANz.exe --Nmmax 4500 --Nfmax 4500 --femrepro 1:2 --malerepro 1:2 --typingerror 0.005 --updatefreqs --poutformat 2 --fullsibtest "V:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\hogan_postQA_2014_2016_HOGAN.dat"

The parameters are defined as follows:  
--Nmmax and --Nfmax are the maximum numbers of candidate mothers and fathers. To obtain our values, we used an estimated escapement of 4,150 and divided by 2. This escapement number came from the smaller of the two estimates (aerial survey AUC and stream walk AUC). This estimate is listed under 'pop est aerial' in the Excel file found here: "V:\Documents\5_Coastwide\Multispecies\AHRP\Field data\Sample Summary Extract and Genotype.xlsx"  
--femrepro and --malerepro specify the age range in which an individual can reproduce  
--typingerror refers to the overall genotyping error rate. Ours was ~0.005  
--updatefreqs specifies that *FRANz* should update allele frequencies using MCMC sampling  
--poutformat specifies that all potential parents should be listed, not just the most likely  
--fullsibtest tests for full siblings among offspring  

All output files can be found here: "v:\Analysis\5_Coastwide\Multispecies\Alaska Hatchery Research Program\PWS Pink\Franz\hogan_postQA_2014_2016_HOGAN_Run1"  

The summary file provides information about the power of our marker suite:   
Cumulative exclusion probability when 1 to 7 fullsibs are genotyped  
  First Parent              : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Second Parent             : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000   
  Parent Pair               : 1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000  1.0000000
  
According to the FRANz manual, marker sets are not considered powerful if these cumulative exclusion probabilities are less than 0.95, which indicates that the probability that a random pair of individuals in the population has a 5% chance of having a genotype pair compatible to an offspring genotype. Since all of our probabilities are 1, we can be **confident** in the power of our 298 amplicons to make parent assignments. 

Below is a histogram of *FRANz* parentage posterior probabilities to show robustness of assignments
```{r FRANz_posterior_plot_even}
parentage_14_16 <- read_csv("v:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/Franz/2014_2016_Hogan_Run1_298/parentage.csv")

parentage_14_16  %>% 
  filter(!is.na(`Parent 1`)) %>% 
  ggplot(aes(x = Posterior)) +
  geom_histogram(breaks = seq(0, 1, 0.01)) +
  ylab("Count") +
  xlab("Posterior Probability") +
  ggtitle("Histogram of FRANz posterior probabilities for parentage assignments")
```

All parentage assigments have a posterior probability of 1.00, which is very robust. No assignments were split among multiple potential parents. 

Of the 4,020 offspring collected in 2016, 451 were assigned to at least one parent, for an an overall offspring assignment rate of **11.2%**. This is in line with expectations based on parent sampling depth

**Note** that *FRANz* assigned 451 offspring to 184 parents in single parent assignments (*single parent-offspring duos*). 208 of these assignments are to natural origin parents and 265 assignments were to hatchery origin parents. Hatchery origin parents are from AFK12B, CCH12, and WNH12PINKB, none are from Solomon Gulch (VFDA).  

Additionally, *FRANz* made two parent assignments (*parent pair-offspring trios*) for 22 offspring.

## Proportion test
We used a Chi-Square test to determine whether the proportions of offspring assigned to hatchery- and natural-origin parents significantly differ from the proportions of hatchery- and natural-origin parents sampled. 

```{r Chi square test even}
paired_14_16_filter_parents <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Hogan/paired_14_16_filter_parents.csv") %>% 
  mutate(origin = factor(x = origin, levels = c("Natural", "Hatchery"))) %>% 
  mutate(sex = case_when(SEX == "M" ~ "Male",
                         SEX == "F" ~ "Female"))

O.even <- levels(paired_14_16_filter$origin)
Spawners.even <- paired_14_16_filter %>% 
  filter(`Sample Year` == 2014) %>% 
  count(origin) %>% 
  pull(n) #  number of hatchery- and natural-origin parents
Offspring.even <- paired_14_16_filter_parents %>% 
  group_by(origin) %>% 
  summarise(n = sum(n)) %>% 
  pull(n)  # number of offspring assigned to hatchery- and natural-origin parents

df.even <- data.frame(Spawners.even, Offspring.even)  # made a data.frame because tibbles don't like rownames
row.names(df.even) <- O.even

df.even
chisq.test(df.even)
```

There is a significant difference in the proportions of offspring assigned to hatchery- and natural-origin parents relative to the proportions of potential parents sampled, indicating an under-representation of offspring assigning to hatchery-origin parents. We see that although **`r round(df.even["H", "Spawners.even"] / sum(df.even[, "Spawners.even"]), 2) * 100`%** of parents genotyped were of hatchery-origin, only **`r round(df.even["H", "Offspring.even"] / sum(df.even[, "Offspring.even"]), 2) * 100`%** of offspring assigned to hatchery-origin parents.

## Reproductive success
Here, we calculate and plot reproductive success (RS; number of offspring assigned to parents) for both natural and hatchery parents. The number of offspring assigned ranges from 0-13. 

### Calculate mean RS
We'll calculate the mean RS by origin and sex including sampled fish from the parental generation to which offspring were not assigned.

```{r calculate RS even}
# Mean RS including 0's
rs_14_16_0 <- paired_14_16_filter_parents %>%
  group_by(origin, sex) %>% 
  summarise(RS = mean(n, na.rm = TRUE))

rs_14_16_0 %>%
  mutate(RS = round(RS, 2)) %>% 
  spread(sex, RS)
```

### Distribution of RS
```{r plot RS even}
# Plot histogram of proportion of parents with a given family size
paired_14_16_filter_parents %>% 
  count(sex, origin, n) %>%
  group_by(sex, origin) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = origin)) +
  theme_bw() +
  geom_col(position = position_dodge2(preserve="single"), width=0.8, colour = "black") +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  facet_grid(~ sex) +
  labs(title="Distribution of Reproductive Success",
       fill = "Parent\nOrigin") +
  xlab("Number of Offspring")+
  ylab("Proportion of Parents") +
  theme(text = element_text(size = 20))

# Save RS values
F_h_0_14_16 <- rs_14_16_0 %>% 
  filter(origin == "Hatchery" & sex == "Female") %>% 
  pull(RS)
M_h_0_14_16 <- rs_14_16_0 %>% 
  filter(origin == "Hatchery" & sex == "Male") %>% 
  pull(RS)
F_n_0_14_16 <- rs_14_16_0 %>% 
  filter(origin == "Natural" & sex == "Female") %>% 
  pull(RS)
M_n_0_14_16 <- rs_14_16_0 %>% 
  filter(origin == "Natural" & sex == "Male") %>% 
  pull(RS)
```

Note the difference in the number of natural- versus hatchery-origin parents assigned 0 offspring, particularly for females. Greater sampling depth allowed for a more complete distribution of RS. Many fewer parents were assigned 0 offspring as compared to the odd-year lineage.  

### Calculate and test RRS
```{r Calculate relative reproductive success including 0s even}
Male_RRS_even <- M_h_0_14_16/M_n_0_14_16
Female_RRS_even <- F_h_0_14_16/F_n_0_14_16

# RRS_tibble_0_14_16 <- tibble(Male_RRS_even, Female_RRS_even)
# round(RRS_tibble_0_14_16, 2)
```

We calcualte RRS as mean Hatchery RS divided by mean Natural RS.

**RRS for females = `r round(Female_RRS_even, 2)`**

**RRS for males = `r round(Male_RRS_even, 2)`.**

#### Statistical testing
We tested for significant differences in RS between male and female hatchery- and natural-origin fish using two different statistical tests:  

  1. Nonparametric permutation test
  2. Parametric negative binomial GLM  
  
```{r RRS statistical testing including 0s even}
# The format of the data to test was very simple, it was a data.frame called “mydata” with 2 columns
# 1.	nOff = number of offspring per family
# 2.	Origin = “H” or “W” for if the parent was hatchery or natural (wild)
# 
mydata_M_0 <- paired_14_16_filter_parents %>%
  filter(SEX == "M") 
  

mydata_F_0 <- paired_14_16_filter_parents %>%
  filter(SEX == "F")
  

perm_1tail_pvalue_M_0 <- as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_M_0, distribution = approximate(B = 10000), alternative = "greater")))

perm_1tail_pvalue_F_0 <- as.numeric(coin::pvalue(oneway_test(n ~ origin, data = mydata_F_0, distribution = approximate(B = 10000), alternative = "greater")))

#And to do a 1-tailed negative binomial GLM

fit_M_0 <- glm.nb(n ~ origin, data = mydata_M_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_M_0 <- summary(fit_M_0)$coefficients[2, 4]

fit_F_0 <- glm.nb(n ~ origin, data = mydata_F_0, init.theta = 1, link = log)
nbGLM_1tail_pvalue_F_0 <- summary(fit_F_0)$coefficients[2, 4]

RRS_0 <- tibble(Sex = c("Male", "Female", "Male", "Female"),
                Test = c(rep("Permutation", 2), rep("Negative Binomial GLM", 2)),
                p_value = c(perm_1tail_pvalue_M_0, perm_1tail_pvalue_F_0, nbGLM_1tail_pvalue_M_0, nbGLM_1tail_pvalue_F_0))
RRS_0 %>% 
  spread(Sex, p_value)  
```

The above table shows the *p-values* for each test and sex. We use a *p-value* cut-off of 0.05. Female hatchery-origin fish produced significantly fewer offspring on average than their natural-origin counterparts, but there was no significant difference in RS for males.  

#### Confidence intervals

We will calculate 95% CI's for RRS based on [Kalinowski and Taper 2005](http://www.nrcresearchpress.com/doi/10.1139/f04-239). This is a maximum likelihood estimate based off of Hinrichsen 2003. This method has been used in other peer reviewed papers such as [Ford et al. 2016](https://www.researchgate.net/publication/309147116_Broodstock_History_Strongly_Influences_Natural_Spawning_Success_in_Hatchery_Steelhead_Oncorhynchus_mykiss#pf14).

```{r RRS_confidence_intervals_14_16}
even_female_rrs_CI <- rrs_ci_kalinowski(n_h_off = paired_14_16_filter_parents %>% 
                                         filter(SEX == "F", origin == "Hatchery") %>% 
                                         group_by(SEX, origin) %>% 
                                         summarise(n_off = sum(n)) %>% 
                                         pull(n_off),
                                       n_w_off = paired_14_16_filter_parents %>% 
                                         filter(SEX == "F", origin == "Natural") %>% 
                                         group_by(SEX, origin) %>% 
                                         summarise(n_off = sum(n)) %>% 
                                         pull(n_off), 
                                       n_h_par = paired_14_16_filter_parents %>% 
                                         filter(SEX == "F", origin == "Hatchery") %>% 
                                         count(SEX, origin) %>% 
                                         pull(nn), 
                                       n_w_par = paired_14_16_filter_parents %>% 
                                         filter(SEX == "F", origin == "Natural") %>% 
                                         count(SEX, origin) %>% 
                                         pull(nn),
                                       alpha = 0.05)

even_male_rrs_CI <- rrs_ci_kalinowski(n_h_off = paired_14_16_filter_parents %>% 
                                       filter(SEX == "M", origin == "Hatchery") %>% 
                                       group_by(SEX, origin) %>% 
                                       summarise(n_off = sum(n)) %>% 
                                       pull(n_off),
                                     n_w_off = paired_14_16_filter_parents %>% 
                                       filter(SEX == "M", origin == "Natural") %>% 
                                       group_by(SEX, origin) %>% 
                                       summarise(n_off = sum(n)) %>% 
                                       pull(n_off), 
                                     n_h_par = paired_14_16_filter_parents %>% 
                                       filter(SEX == "M", origin == "Hatchery") %>% 
                                       count(SEX, origin) %>% 
                                       pull(nn), 
                                     n_w_par = paired_14_16_filter_parents %>% 
                                       filter(SEX == "M", origin == "Natural") %>% 
                                       count(SEX, origin) %>% 
                                       pull(nn),
                                     alpha = 0.05)
```

Hogan 2014/2016 **female** RRS with 95% CI = **`r paste0(even_female_rrs_CI[2], " (", even_female_rrs_CI[1], "-", even_female_rrs_CI[3], ")")`**

Hogan 2014/2016 **male** RRS with 95% CI = **`r paste0(even_male_rrs_CI[2], " (", even_male_rrs_CI[1], "-", even_male_rrs_CI[3], ")")`**

### Plots of RS vs. field data
As in the odd-lineage above, we examined associations between two potential explanatory variables (sample date and body size) and RS. Since we have many more offspring assignments, we can more clearly visualize these relationships. These plots include LOESS lines with 95% confidence intervals to help illustrate trends. We use LOESS rather than linear models because our data violate assumptions of linear regression.

```{r Examine reproductive success by sample date even with zeros}
paired_14_16_filter_parents %>%
  ggplot(aes(x = `Sample Date`, y = n, color = origin)) +
  geom_jitter(height = 0, size = 2) +
  geom_smooth() +
  facet_grid(sex ~ .)+
  labs(title = "Reproductive Success by Parent Sample Date",  colour = "Parent\nOrigin",
       shape = "Parent: Sex") +
  xlab("Parent Sample Date") +
  ylab("Number of Offspring") +
  theme_bw() +
  theme(text = element_text(size = 20))
```

The above plot shows RS by parent sample date, separately for females (top) and males (bottom). We see that RS does vary due to sample date, with the lowest RS values in the middle of the season (perhaps due to increased fishing pressure or density dependence on the spawning grounds). There do appear to be some differences in RS between natural- and hatchery-origin fish early in the season for females and later in the season for males.

```{r Examine reproductive success by parent length even with zeros}
paired_14_16_filter_parents %>% 
  filter(`Length Mm` > 300) %>%  # remove outlier
  ggplot(aes(x = `Length Mm`, y = n, color = origin)) +
  geom_jitter(height = 0, size = 2) +
  geom_smooth() +
  facet_grid(sex ~ .) +
  labs(title = "Reproductive Success by Parent Length",  colour = "Parent\nOrigin",
       shape = "Parent: Sex") +
  xlab("Parental Length (mm)") +
  ylab("Number of Offspring") +
  theme_bw() +
  theme(text = element_text(size = 20))
```

The above plot show RS by parent body length, separately for females (top) and males (bottom). Larger males and females appear more successful than smaller individuals. 

## Parent pair-offspring trios
Because we have 22 individuals who were assigned to two parents (families), we are able to examine instances of crosses between hatchery fish, between natural fish, and between natural and hatchery fish. **Note** that we have divided natural-hatchery crosses into "HN", which indicates a hatchery-origin dam and natural-origin sire and "NH", which indicates a natural-origin dam and hatchery-origin sire. 

### Sample sizes
These are our sample sizes are for the number of families by cross type.

```{r even_cross_type}
parents_paired_14_16_cross <- read_csv("V:/Analysis/5_Coastwide/Multispecies/Alaska Hatchery Research Program/PWS Pink/GitHub-PWS-Pink-Parentage/Hogan/parents_paired_14_16_cross.csv") %>% 
  mutate(cross = factor(cross, c("NN", "NH", "HN", "HH")))

parents_paired_14_16_cross %>% 
  mutate(`Cross Type` = cross) %>% 
  count(`Cross Type`, `Fish ID`, `Fish ID.par1`) %>% 
  count(`Cross Type`) %>% 
  rename(`Number of Families` = nn)
```

### Distribution of RS

```{r Create cross as new grouping variable even}
parents_paired_14_16_cross %>% 
  count(cross, `Fish ID`, `Fish ID.par1`) %>% 
  count(cross, n) %>% 
  complete(cross, n, fill = list(nn = 0)) %>% 
  group_by(cross) %>% 
  mutate(p = nn / sum(nn)) %>% 
  ggplot(aes(x = n, y = p, fill = cross))+
  geom_col(position = position_dodge2(preserve = "single"), width = 0.5) +
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = pretty_breaks()) +
  scale_x_continuous(breaks = 1:5) +
  # ylim(0, 40) +
  # xlim(0, 5) +
  labs(title = "Distribution of Reproductive Success by Cross Type",
       fill = "Cross Type") +
  xlab("Number of Offspring")+
  ylab("Proportion of Families") +
  theme_bw() +
  theme(text = element_text(size = 20))
```

Offspring from all four types of crosses are represented in our dataset. HH crosses produced 1-2 offspring, HN and NH crosses produced 1-3 offspring, and NN crosses produced 1-2 offspring. Note that analyses of trios do not include matings with RS = 0 because it is not possible for us to collect this information (we only know a mating between two fish ocurred if we assign offspring to those two fish). 

### Calculate mean RS
Here we calculated the mean RS for each cross type. Note that sample sizes of families for each cross type are very small.

```{r cross_rs_even}
parents_paired_14_16_cross %>% 
  mutate(`Cross Type` = cross) %>% 
  count(`Cross Type`, `Fish ID`, `Fish ID.par1`) %>% 
  group_by(`Cross Type`) %>% 
  summarise(`Average RS` = round(mean(n), digits = 2))  
```

### Calculate and test RRS

#### Confidence intervals

```{r RRS_confidence_intervals_cross_14_16}
even_HH_NN_rrs_CI <- rrs_ci_kalinowski(n_h_off = parents_paired_14_16_cross %>% 
                                         filter(cross == "HH") %>% 
                                         count(cross) %>% 
                                         pull(n),
                                       n_w_off = parents_paired_14_16_cross %>% 
                                         filter(cross == "NN") %>% 
                                         count(cross) %>% 
                                         pull(n), 
                                       n_h_par = parents_paired_14_16_cross %>% 
                                         filter(cross == "HH") %>% 
                                         count(cross, `Fish ID`, `Fish ID.par1`) %>% 
                                         count(cross) %>% 
                                         pull(nn), 
                                       n_w_par = parents_paired_14_16_cross %>% 
                                         filter(cross == "NN") %>% 
                                         count(cross, `Fish ID`, `Fish ID.par1`) %>% 
                                         count(cross) %>% 
                                         pull(nn),
                                       alpha = 0.05)

even_HN_NN_rrs_CI <- rrs_ci_kalinowski(n_h_off = parents_paired_14_16_cross %>% 
                                         filter(cross == "HN") %>% 
                                         count(cross) %>% 
                                         pull(n),
                                       n_w_off = parents_paired_14_16_cross %>% 
                                         filter(cross == "NN") %>% 
                                         count(cross) %>% 
                                         pull(n), 
                                       n_h_par = parents_paired_14_16_cross %>% 
                                         filter(cross == "HN") %>% 
                                         count(cross, `Fish ID`, `Fish ID.par1`) %>% 
                                         count(cross) %>% 
                                         pull(nn), 
                                       n_w_par = parents_paired_14_16_cross %>% 
                                         filter(cross == "NN") %>% 
                                         count(cross, `Fish ID`, `Fish ID.par1`) %>% 
                                         count(cross) %>% 
                                         pull(nn),
                                       alpha = 0.05)

even_NH_NN_rrs_CI <- rrs_ci_kalinowski(n_h_off = parents_paired_14_16_cross %>% 
                                         filter(cross == "NH") %>% 
                                         count(cross) %>% 
                                         pull(n),
                                       n_w_off = parents_paired_14_16_cross %>% 
                                         filter(cross == "NN") %>% 
                                         count(cross) %>% 
                                         pull(n), 
                                       n_h_par = parents_paired_14_16_cross %>% 
                                         filter(cross == "NH") %>% 
                                         count(cross, `Fish ID`, `Fish ID.par1`) %>% 
                                         count(cross) %>% 
                                         pull(nn), 
                                       n_w_par = parents_paired_14_16_cross %>% 
                                         filter(cross == "NN") %>% 
                                         count(cross, `Fish ID`, `Fish ID.par1`) %>% 
                                         count(cross) %>% 
                                         pull(nn),
                                       alpha = 0.05)
```

Hogan 2014/2016 **HH/NN** RRS with 95% CI = **`r paste0(even_HH_NN_rrs_CI[2], " (", even_HH_NN_rrs_CI[1], "-", even_HH_NN_rrs_CI[3], ")")`**

Hogan 2014/2016 **HN/NN** RRS with 95% CI = **`r paste0(even_HN_NN_rrs_CI[2], " (", even_HN_NN_rrs_CI[1], "-", even_HN_NN_rrs_CI[3], ")")`**

Hogan 2014/2016 **NH/NN** RRS with 95% CI = **`r paste0(even_NH_NN_rrs_CI[2], " (", even_NH_NN_rrs_CI[1], "-", even_NH_NN_rrs_CI[3], ")")`**

**Note** that the confidence intervals for both cross type RRS are large due to the small numbers of offspring assigned to two parents and the small sample size of families. 

## Summary of even lineage results
1. Exclusion probabilities were equal to 1.00, which means we can be confident in the ability of our marker set to correctly assign parents to offspring.  
2. 451 offspring were assigned to 184 parents, for an assignment rate of **11.2%**. This is in line with our expectations based on the sampling rate of the parent escapement in 2014. 
3. Given the larger sample size of parent-offspring relationships, the even lineage results are much more robust than the odd lineage results.
4. The number of offspring assigned to parents varied from 0-10 for females and 0-13 for males.  
5. RRS was significant for females (0.47 [CI's 0.37-0.62]), but not for males (0.87 [CI's 0.67-1.12]).  
5. Hatchery and natural fish interbreed (both sexes).  
6. For the most part, natural-origin fish had higher RS earlier than later in the run.  
7. Larger male and female fish appear to have higher RS than smaller fish. 

