---
title: "AK AFS 2022"
author: "Kyle Shedd"
date: "17 February 2022"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list = ls())
library(coin)
library(MASS)
library(scales)
library(leaflet)
library(lubridate)
library(ggExtra)
library(colorspace)
library(ggmap)
library(maps)
library(mapdata)
library(PBSmapping)
library(grid)
library(ggsn)
library(tidyverse)

knitr::opts_chunk$set(out.width = "100%", fig.width = 10)
load("../Stockdale_Hogan_Gilmour_Paddy_Erb/2_Parentage_All_Streams_2014_2016_develop_20220212.RData")
```

# RS Distributions

## Erb 2014

Blank
```{r fig.height=6, fig.width=10}
paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Erb") %>% 
  dplyr::count(stream, sex, origin, n, name = "n_par") %>% 
  dplyr::group_by(stream, sex, origin) %>% 
  dplyr::mutate(p = n_par / sum(n_par)) %>% 
  ggplot2::ggplot(aes(x = n, y = p, fill = origin)) +
  # ggplot2::geom_col(position = position_dodge2(preserve="single"), colour = "black") +
  ggplot2::geom_blank() +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::scale_y_continuous(breaks = scales::pretty_breaks(), limits = c(0, 0.65)) +
  ggplot2::scale_x_continuous(breaks = scales::pretty_breaks(), limits = c(-1, 28)) +
  ggplot2::facet_grid(cols = ggplot2::vars(sex)) +
  ggplot2::labs(fill = "Parent: Origin",
                x = "Number of Offspring",
                y = "Proportion of Parents (RS)") +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(legend.position = c(0.85, 0.85), legend.background = element_rect(colour = "black"))

ggplot2::ggsave(filename = "RS_distribution_erb_2014_blank.png", width = 10, height = 6.5, units = "in")
```

Hatchery-origin females
```{r fig.height=6, fig.width=10}
paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Erb") %>% 
  dplyr::count(stream, sex, origin, n, name = "n_par") %>% 
  dplyr::group_by(stream, sex, origin) %>% 
  dplyr::mutate(p = n_par / sum(n_par)) %>% 
  dplyr::mutate(p = dplyr::case_when(sex == "Male" ~ 0,
                                     origin == "Natural" & sex == "Female" ~ 0,
                                     origin == "Hatchery" & sex == "Female" ~ p)) %>% 
  ggplot2::ggplot(aes(x = n, y = p, fill = origin)) +
  ggplot2::geom_col(position = position_dodge2(preserve="single"), colour = "black") +
  # ggplot2::geom_blank() +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::scale_y_continuous(breaks = scales::pretty_breaks(), limits = c(0, 0.65)) +
  ggplot2::scale_x_continuous(breaks = scales::pretty_breaks(), limits = c(-1, 28)) +
  ggplot2::scale_fill_discrete(drop = FALSE) +
  ggplot2::facet_grid(cols = ggplot2::vars(sex)) +
  ggplot2::labs(fill = "Parent: Origin",
                x = "Number of Offspring",
                y = "Proportion of Parents") +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(legend.position = c(0.85, 0.85), legend.background = element_rect(colour = "black"))

ggplot2::ggsave(filename = "RS_distribution_erb_2014_hatchery_female.png", width = 10, height = 6.5, units = "in")
```

All females
```{r fig.height=6, fig.width=10}
paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Erb") %>% 
  dplyr::count(stream, sex, origin, n, name = "n_par") %>% 
  dplyr::group_by(stream, sex, origin) %>% 
  dplyr::mutate(p = n_par / sum(n_par)) %>% 
  dplyr::mutate(p = dplyr::case_when(sex == "Male" ~ 0,
                                     sex == "Female" ~ p)) %>% 
  ggplot2::ggplot(aes(x = n, y = p, fill = origin)) +
  ggplot2::geom_col(position = position_dodge2(preserve="single"), colour = "black") +
  # ggplot2::geom_blank() +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::scale_y_continuous(breaks = scales::pretty_breaks(), limits = c(0, 0.65)) +
  ggplot2::scale_x_continuous(breaks = scales::pretty_breaks(), limits = c(-1, 28)) +
  ggplot2::scale_fill_discrete(drop = FALSE) +
  ggplot2::facet_grid(cols = ggplot2::vars(sex)) +
  ggplot2::labs(fill = "Parent: Origin",
                x = "Number of Offspring",
                y = "Proportion of Parents") +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(legend.position = c(0.85, 0.85), legend.background = element_rect(colour = "black"))

ggplot2::ggsave(filename = "RS_distribution_erb_2014_all_female.png", width = 10, height = 6.5, units = "in")
```

All~
```{r fig.height=6, fig.width=10}
paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Erb") %>% 
  dplyr::count(stream, sex, origin, n, name = "n_par") %>% 
  dplyr::group_by(stream, sex, origin) %>% 
  dplyr::mutate(p = n_par / sum(n_par)) %>% 
  # dplyr::mutate(p = dplyr::case_when(sex == "Male" ~ 0,
  #                                    sex == "Female" ~ p)) %>% 
  ggplot2::ggplot(aes(x = n, y = p, fill = origin)) +
  ggplot2::geom_col(position = position_dodge2(preserve="single"), colour = "black") +
  # ggplot2::geom_blank() +
  ggplot2::geom_hline(yintercept = 0, colour = "black") +
  ggplot2::scale_y_continuous(breaks = scales::pretty_breaks(), limits = c(0, 0.65)) +
  ggplot2::scale_x_continuous(breaks = scales::pretty_breaks(), limits = c(-1, 28)) +
  ggplot2::scale_fill_discrete(drop = FALSE) +
  ggplot2::facet_grid(cols = ggplot2::vars(sex)) +
  ggplot2::labs(fill = "Parent: Origin",
                x = "Number of Offspring",
                y = "Proportion of Parents") +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(legend.position = c(0.85, 0.85), legend.background = element_rect(colour = "black"))

ggplot2::ggsave(filename = "RS_distribution_erb_2014.png", width = 10, height = 6.5, units = "in")
```

## Density Plot

```{r}
paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Erb") %>% 
  ggplot2::ggplot(aes(x = n, fill = origin)) +
  ggplot2::geom_density(alpha = 0.5, n = 25, adjust = 5) +
  # ggplot2::geom_blank() +
  ggplot2::scale_y_continuous(breaks = scales::pretty_breaks()) +
  ggplot2::scale_x_continuous(breaks = scales::pretty_breaks()) +
  ggplot2::facet_grid(cols = ggplot2::vars(sex)) +
  ggplot2::labs(fill = "Parent: Origin",
                x = "Number of Offspring (RS)",
                y = "Proportion of Parents") +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(legend.position = c(0.85, 0.85), legend.background = element_rect(colour = "black"))

ggplot2::ggsave(filename = "RS_distribution_erb_2014_density.png", width = 10, height = 6.5, units = "in")
```

```{r fig.width=10, fig.height=6}
paired_14_16_filter_parents %>% 
  dplyr::mutate(stream = factor(
    x = stream,
    levels = c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale")
  )) %>% 
  ggplot2::ggplot(aes(x = n, fill = origin)) +
  ggplot2::geom_density(alpha = 0.5, n = 25, adjust = 5) +
  # ggplot2::geom_blank() +
  ggplot2::scale_y_continuous(breaks = scales::pretty_breaks()) +
  ggplot2::scale_x_continuous(breaks = scales::pretty_breaks()) +
  ggplot2::facet_grid(cols = ggplot2::vars(stream), rows = ggplot2::vars(sex)) +
  ggplot2::labs(fill = "Origin",
                x = "Number of Offspring (RS)",
                y = "Proportion of Parents") +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(legend.position = c(0.475, 0.375), legend.background = element_rect(colour = "black"),
                 legend.title = element_text(size = 16), legend.text = element_text(size = 12),
                 legend.margin = margin(2, 2, 2, 2))

ggplot2::ggsave(filename = "RS_distribution_all_streams_2014_density.png", width = 10, height = 6.5, units = "in")
```

# RRS to Date

## Dyad

Read in the data
```{r read_rrs_to_date}
(rrs_to_date_dyad <- readr::read_csv(file = "rrs_to_date_dyad.csv"))
```

```{r}
rrs_to_date_dyad %>%
  dplyr::mutate(stream = factor(
    x = stream,
    levels = c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale")
  ),
  year = as.character(year)) %>%
  dplyr::filter(year == "2014") %>% 
  summarise(min(rrs),
            max(rrs))
```


### Bar plot

```{r fig.height=6.5, fig.width=5}
rrs_to_date_dyad %>%
  dplyr::mutate(stream = factor(
    x = stream,
    levels = c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale")
  ),
  year = as.character(year)) %>%
  dplyr::filter(year == "2014") %>%
  ggplot2::ggplot(aes(
    x = year,
    y = rrs,
    ymin = lower_ci,
    ymax = upper_ci,
    fill = sex
  )) +
  ggplot2::geom_col(position = position_dodge(), colour = "black") +
  ggplot2::geom_errorbar(position = position_dodge(width = 0.9),
                         width = 0.5,
                         size = 0.75) +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_hline(yintercept = 1,
                      linetype = 2,
                      size = 1) +
  ggplot2::scale_fill_manual(values = c("orange", "skyblue")) +
  ggplot2::facet_grid(stream ~ lineage, scales = "free_x") +
  ggplot2::labs(x = "Run Year", y = "RRS (95% CI)", fill = "Sex") +
  ggplot2::theme_bw(base_size = 16)

ggplot2::ggsave(filename = "RRS_to_date_dyad_2014.png", width = 5, height = 6.5, units = "in")
```

## Triad

Read in the data
```{r read_rrs_to_date_triad}
(rrs_to_date_triad <- readr::read_csv(file = "rrs_to_date_triad.csv"))
```

### Bar plot

Get default `ggplot2` colors
```{r}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```

### Bar plot

```{r fig.height=6.5, fig.width=5}
rrs_to_date_triad %>% 
  dplyr::mutate(year = as.character(year)) %>% 
  ggplot2::ggplot(aes(x = year, y = rrs, ymin = lower_ci, ymax = upper_ci, fill = cross)) +
  ggplot2::geom_col(position = position_dodge(), colour = "black") +
  ggplot2::geom_errorbar(position = position_dodge(width = 0.9), width = 0.5, size = 0.75) +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_hline(yintercept = 1, linetype = 2, size = 1) +
  ggplot2::scale_fill_manual(values = gg_color_hue(n = 4)[4:1]) +
  ggplot2::facet_grid(stream ~ lineage, scales = "free_x") +
  ggplot2::labs(x = "Run Year", y = "RRS (95% CI)", fill = "Cross Type") +
  ggplot2::theme_bw(base_size = 16)
  # ggplot2::theme(legend.position = c(0.8, 0.8), legend.background = element_rect(colour = "black"))

ggplot2::ggsave(filename = "RRS_to_date_triad_AK_AFS.png", width = 5, height = 6.5, units = "in")
```

# Covariate Density Distributions

## Sample Date

In depth look at run timing
```{r fig.height=6, fig.width=10}
paired_14_16_filter_parents %>%
  dplyr::mutate(stream = factor(
    x = stream,
    levels = c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale")
  )) %>% 
  ggplot2::ggplot(ggplot2::aes(x = date, fill = origin)) +
  ggplot2::geom_density(alpha = 0.5, adjust = 2) +
  ggplot2::labs(x = "Parent Sample Date",
                y = "Density",
                fill = "Origin") +
  ggplot2::facet_grid(rows = ggplot2::vars(sex), cols = ggplot2::vars(stream)) +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(legend.position = c(0.475, 0.375), legend.background = element_rect(colour = "black"),
                 legend.title = element_text(size = 16), legend.text = element_text(size = 12),
                 legend.margin = margin(2, 2, 2, 2)) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5))


ggplot2::ggsave(filename = "density_date_2014.png", width = 10, height = 6.5, units = "in")
```

## Body Length

In depth look at body length
```{r fig.height=6, fig.width=10}
paired_14_16_filter_parents %>%
  dplyr::mutate(stream = factor(
    x = stream,
    levels = c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale")
  )) %>% 
  ggplot2::ggplot(ggplot2::aes(x = length, fill = origin)) +
  ggplot2::geom_density(alpha = 0.5, adjust = 2) +
  ggplot2::labs(x = "Parent Body Length (mm)",
                y = "Density",
                fill = "Origin") +
  ggplot2::facet_grid(rows = ggplot2::vars(sex), cols = ggplot2::vars(stream)) +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(legend.position = c(0.475, 0.375), legend.background = element_rect(colour = "black"),
                 legend.title = element_text(size = 16), legend.text = element_text(size = 12),
                 legend.margin = margin(2, 2, 2, 2)) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5))


ggplot2::ggsave(filename = "density_length_2014.png", width = 10, height = 6.5, units = "in")
```

## Intertidal

In depth look at body length
```{r fig.height=6, fig.width=10}
paired_14_16_filter_parents %>%
  dplyr::mutate(stream = factor(
    x = stream,
    levels = c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale")
  )) %>% 
  dplyr::filter(!is.na(intertidal)) %>% 
  dplyr::count(stream, sex, origin, intertidal, name = "n_par") %>% 
  dplyr::group_by(stream, sex, origin) %>% 
  dplyr::mutate(p = n_par / sum(n_par)) %>% 
  ggplot2::ggplot(ggplot2::aes(x = intertidal, y = p, fill = origin)) +
  ggplot2::geom_col(alpha = 0.5, position = position_dodge2()) +
  ggplot2::labs(x = "Parent Sample Location",
                y = "Proportion of Parents",
                fill = "Origin") +
  ggplot2::facet_grid(rows = ggplot2::vars(sex), cols = ggplot2::vars(stream)) +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(legend.position = c(0.56, 0.375), legend.background = element_rect(colour = "black"),
                 legend.title = element_text(size = 16), legend.text = element_text(size = 12),
                 legend.margin = margin(2, 2, 2, 2)) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1, vjust = 1))


ggplot2::ggsave(filename = "density_intertidal_2014.png", width = 10, height = 6.5, units = "in")
```

# RS vs. Covariates

## Sample Date

```{r fig.width=10, fig.height=6}
paired_14_16_filter_parents %>%
  dplyr::mutate(stream = factor(
    x = stream,
    levels = c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale")
  )) %>%
  ggplot2::ggplot(aes(x = date, y = n, fill = origin)) +
  ggplot2::geom_hline(yintercept = 0, color = "grey50") +
  ggplot2::geom_jitter(
    height = 0,
    size = 4,
    shape = 21,
    color = "black",
    alpha = 0.3
  ) +
  ggplot2::geom_smooth(
    method =  "lm",
    formula = 'y ~ x',
    color = "black",
    size = 1.5,
    alpha = 0.9
  ) +
  ggplot2::facet_grid(sex ~ stream, scales = "free") +
  ggplot2::labs(title = "",
                fill = "Origin",
                x = "Parent Sample Date",
                y = "Reproductive Success") +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(
    legend.position = c(0.30, 0.355),
    legend.background = element_rect(colour = "black"),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 12),
    legend.margin = margin(2, 2, 2, 2)
  ) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5))

ggplot2::ggsave(filename = "RS_x_date_2014.png", width = 10, height = 6.5, units = "in")
```

## Body Length

```{r fig.width=10, fig.height=6}
paired_14_16_filter_parents %>%
  dplyr::mutate(stream = factor(
    x = stream,
    levels = c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale")
  )) %>%
  dplyr::filter(length > 300) %>% 
  ggplot2::ggplot(aes(x = length, y = n, fill = origin)) +
  ggplot2::geom_hline(yintercept = 0, color = "grey50") +
  ggplot2::geom_jitter(
    height = 0,
    size = 4,
    shape = 21,
    color = "black",
    alpha = 0.3
  ) +
  ggplot2::geom_smooth(
    method =  "lm",
    formula = 'y ~ x',
    color = "black",
    size = 1.5,
    alpha = 0.9
  ) +
  ggplot2::facet_grid(sex ~ stream, scales = "free_y") +
  ggplot2::labs(title = "",
                fill = "Origin",
                x = "Parent Body Length (mm)",
                y = "Reproductive Success") +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(
    legend.position = c(0.30, 0.355),
    legend.background = element_rect(colour = "black"),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 12),
    legend.margin = margin(2, 2, 2, 2)
  ) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust = 1, vjust = 0.5))

ggplot2::ggsave(filename = "RS_x_length_2014.png", width = 10, height = 6.5, units = "in")
```

## Intertidal

```{r fig.width=10, fig.height=6}
paired_14_16_filter_parents %>%
  dplyr::mutate(stream = factor(
    x = stream,
    levels = c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale")
  )) %>%
  dplyr::filter(!is.na(intertidal)) %>% 
  ggplot2::ggplot(aes(x = intertidal, y = n, fill = origin)) +
  ggplot2::geom_hline(yintercept = 0, color = "grey50") +
  ggplot2::geom_boxplot() +
  ggplot2::facet_grid(sex ~ stream, scales = "free_y") +
  ggplot2::labs(title = "",
                fill = "Origin",
                x = "Parent Sample Location",
                y = "Reproductive Success") +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::theme(
    legend.position = c(0.30, 0.35),
    legend.background = element_rect(colour = "black"),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 12),
    legend.margin = margin(2, 2, 2, 2)
  ) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1, vjust = 1))

ggplot2::ggsave(filename = "RS_x_intertidal_2014.png", width = 10, height = 6.5, units = "in")
```

# GLM

## Scale Covariates

```{r}
(
  paired_14_16_filter_parents_scaled <-
    paired_14_16_filter_parents %>%
    dplyr::group_by(stream, sex) %>%
    dplyr::mutate(
      DOY_scaled = scale(x = DOY, center = TRUE, scale = TRUE),
      length_scaled = scale(x = length, center = TRUE, scale = TRUE),
      distance_mouth_scaled = scale(x = distance_mouth, center = TRUE, scale = TRUE),
      n_relative = n / mean(n)
    ) %>%
    dplyr::ungroup() %>% 
    dplyr::select(franz_id, n, n_relative, tidyselect::everything())
)
```

## Male

### Model

Just doing a very simple model, no interactions
```{r}
male_scaled <- paired_14_16_filter_parents_scaled %>%
  dplyr::filter(sex == "Male")

glm_male <- MASS::glm.nb(
  n ~ stream + DOY_scaled + length_scaled + intertidal + origin,
  data = male_scaled,
  link = log
)

summary(glm_male)
```

Compare effect sizes for basic model using `modelplot`
```{r fig.height=6.5, fig.width=5}
renamed_coeff <- c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale", "Date", "Length", "Upstream", "Hatchery")

p_male <- modelsummary::modelplot(models = glm_male)

p_male$data %>% 
  mutate(name = factor(x = renamed_coeff, levels = renamed_coeff)) %>% 
  ggplot(aes(y = name, x = estimate, xmin = conf.low, xmax = conf.high)) +
  geom_vline(xintercept = 0, size = 1) +
  geom_errorbar(width = 0, size = 2) +
  geom_point(shape = 21, colour = "black", fill = "grey50", size = 5) +
  scale_x_continuous(breaks = pretty_breaks(), limits = c(-1.25, 1.25)) +
  theme_bw(base_size = 20) +
  labs(x = "Coefficients with 95% CIs",
       y = "",
       title = "Male")

ggplot2::ggsave(filename = "GLM_coef_male_2014.png", width = 5, height = 6.5, units = "in")
```

Transform coefficients out of log link function with exponent.
```{r}
(male_coeffs <- cbind(coef = coef(glm_male), confint(glm_male)))
exp(male_coeffs)
```

## Female

### Model

Just doing a very simple model, no interactions
```{r}
female_scaled <- paired_14_16_filter_parents_scaled %>%
  dplyr::filter(sex == "Female")

glm_female <- MASS::glm.nb(
  n ~ stream + DOY_scaled + length_scaled + intertidal + origin,
  data = female_scaled,
  link = log
)

summary(glm_female)
```

Compare effect sizes for basic model using `modelplot`
```{r fig.height=6.5, fig.width=5}
renamed_coeff <- c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale", "Date", "Length", "Upstream", "Hatchery")

p_female <- modelsummary::modelplot(models = glm_female)

p_female$data %>% 
  mutate(name = factor(x = renamed_coeff, levels = renamed_coeff)) %>% 
  ggplot(aes(y = name, x = estimate, xmin = conf.low, xmax = conf.high, fill = sex)) +
  geom_vline(xintercept = 0, size = 1) +
  geom_errorbar(width = 0, size = 2) +
  geom_point(shape = 21, colour = "black", fill = "grey50", size = 5) +
  scale_x_continuous(breaks = pretty_breaks(), limits = c(-1.25, 1.25)) +
  theme_bw(base_size = 20) +
  labs(x = "Coefficients with 95% CIs",
       y = "",
       title = "Female")

ggplot2::ggsave(filename = "GLM_coef_female_2014.png", width = 5, height = 6.5, units = "in")
```

Transform coefficients out of log link function with exponent.
```{r}
(female_coeffs <- cbind(coef = coef(glm_female), confint(glm_female)))
exp(female_coeffs)
```

## Single Plot

```{r}
glm_coeffs_male <- p_male$data %>%
  dplyr::mutate(name = factor(x = renamed_coeff, levels = renamed_coeff),
                sex = "Male")

glm_coeffs_female <- p_female$data %>%
  dplyr::mutate(name = factor(x = renamed_coeff, levels = renamed_coeff),
                sex = "Female")

dplyr::bind_rows(glm_coeffs_male, glm_coeffs_female) %>%
  ggplot2::ggplot(aes(
    y = name,
    x = estimate,
    xmin = conf.low,
    xmax = conf.high,
    fill = sex
  )) +
  ggplot2::geom_vline(xintercept = 0, size = 1) +
  ggplot2::geom_errorbar(width = 0, size = 2) +
  ggplot2::geom_point(shape = 21,
                      colour = "black",
                      size = 5) +
  ggplot2::scale_fill_manual(values = c("orange", "skyblue")) +
  ggplot2::scale_x_continuous(breaks = pretty_breaks(), limits = c(-1.25, 1.25)) +
  ggplot2::theme_bw(base_size = 20) +
  ggplot2::labs(x = "Coefficients with 95% CIs",
                y = "") +
  ggplot2::facet_grid(cols = ggplot2::vars(sex)) +
  ggplot2::guides(fill = "none")

ggplot2::ggsave(
  filename = "GLM_coef_combined_2014.png",
  width = 10,
  height = 6.5,
  units = "in"
)
```


End