---
title: "Board of Fisheries Hatchery Committee Meeting 2022"
author: "Kyle Shedd"
date: "13 February 2022"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list = ls())
library(coin)
library(MASS)
library(scales)
library(leaflet)
library(lubridate)
library(ggExtra)
library(colorspace)
library(ggmap)
library(maps)
library(mapdata)
library(PBSmapping)
library(grid)
library(ggsn)
library(tidyverse)

knitr::opts_chunk$set(out.width = "100%", fig.width = 10)
```

# All Samples by Sample Date

```{r}
(oceanak <- readr::read_csv(file = "~/../Desktop/Local_PWS_pinks/OceanAK/AHRP Salmon Biological Data 20220124_151655_names.csv"))
```

Rename to fit *new* conventions ;)
```{r}
(
  oceanak <- oceanak %>%
    dplyr::select(
      `Silly Code`,
      `Fish ID`,
      `DNA Tray Code`,
      `DNA Tray Well Code`,
      `Location Code`,
      `Sample Year`,
      `Sample Date`,
      Sex,
      `Length Mm`,
      `Otolith Mark Present`,
      `Otolith Mark ID`
    ) %>%
    dplyr::rename(
      silly = "Silly Code",
      fish_id = "Fish ID",
      dna_tray_code = "DNA Tray Code",
      dna_tray_well_code = "DNA Tray Well Code",
      year = "Sample Year",
      length = "Length Mm",
      otolith_mark_present = "Otolith Mark Present",
      otolith_mark_id = "Otolith Mark ID"
    ) %>%
    dplyr::mutate(
      stream = stringr::str_remove(string = `Location Code`, pattern = " Creek"),
      date = lubridate::ymd(`Sample Date`),
      DOY = lubridate::yday(date),
      # day of year
      sex = dplyr::case_when(Sex == "M" ~ "Male",
                             Sex == "F" ~ "Female"),
      origin = dplyr::case_when(
        otolith_mark_present == "NO" ~ "Natural",
        otolith_mark_present == "YES" ~ "Hatchery"
      ),
      # add origin variable
      franz_id = base::paste0(
        stringr::str_sub(
          string = silly,
          start = 1,
          end = 2
        ),
        stringr::str_sub(
          string = silly,
          start = -2,
          end = -1
        ),
        "_",
        stringr::str_pad(
          string = fish_id,
          width = 5,
          side = "left",
          pad = "0"
        )
      ) # create franz_id
    ) %>%
    dplyr::mutate(origin = base::factor(origin, c(
      "Natural", "Hatchery"
    ))) %>%  # make factor to ensure hatchery != red
    tidyr::unite(
      col = "sample",
      c("dna_tray_code", "dna_tray_well_code"),
      sep = "_",
      remove = FALSE
    ) %>%
    dplyr::select(
      franz_id,
      stream,
      year,
      origin,
      sex,
      date,
      DOY,
      length,
      otolith_mark_present,
      otolith_mark_id,
      silly,
      fish_id,
      sample,
      dna_tray_code,
      dna_tray_well_code
    )
)
```

```{r fig.height=6.5, fig.width=10}
oceanak %>%
  ggplot2::ggplot(aes(x = DOY, fill = origin)) +
  ggplot2::geom_histogram(binwidth = 1) +
  ggplot2::scale_x_continuous(
    limits = c(210, 270),
    breaks = seq(213, 269, by = 14),
    labels = format(x = (
      as.Date("2012-12-31") + seq(213, 269, by = 14)
    ), "%b %d")
  ) +
  ggplot2::facet_grid(year ~ stream) +
  ggplot2::labs(x = "Date",
                y = "Number of Samples",
                fill = "Origin") +
  ggplot2::theme_bw(base_size = 18) +
  ggplot2::theme(axis.text.x = element_text(
    angle = 90,
    hjust = 1,
    vjust = 0.5
  ))

ggplot2::ggsave(plot = last_plot(), filename = "pink_samples_collected.png", width = 10, height = 6.5, units = "in")
```


# Map

```{r read_map_data}
(pws_map <- readr::read_csv(file = "../map_data.csv"))
```

## ggplot + PBSmapping

Try PBSmapping, clunky, but great basemap for AK coastlines. Got the idea from Ben Williams GitHub [link](https://ben-williams.github.io/updated_ggplot_figures.html). Still used the idea from the previous link on Quantitative Palaeoecology

```{r pbs_mapping}
x_lim <- range(pws_map$long) + c(-0.1, 0.1)
y_lim <- range(pws_map$lat) + c(-0.2, 0.2)

data("nepacLLhigh")
ak <- nepacLLhigh %>% 
  dplyr::select(group = PID, POS = POS,long = X, lat = Y)

maptheme <- theme(
  panel.grid = element_blank(),
  panel.border = element_rect(fill = NA, colour = "black"),
  panel.background = element_blank()
)

# PWS map
pws_map <- ggplot() + 
  geom_polygon(data = ak, aes(long, lat, group = group), fill = "grey90", color = "black", lwd = 0.3) +
  geom_point(data = pws_map, aes(x = long, y = lat, fill = legend, shape = legend, size = legend), stroke = 1.5) +  # shape = 21, colour = "black",
  scale_shape_manual(values = c(24, 21, 22, 22)) +
  scale_size_manual(values = c(3.5, 5, 5, 5)) +
  geom_label(data = filter(pws_map, type == "Hatchery"), aes(long, lat, label = name), hjust = 0, nudge_x = 0.1) +
  geom_label(data = filter(pws_map, type == "City"), aes(long, lat, label = name), hjust = 0.5, vjust = 0, nudge_y = 0.05) +
  geom_label(data = filter(pws_map, name == "Hogan"), aes(long, lat, label = name), hjust = 0, nudge_x = 0.1, nudge_y = -0.05) +
  geom_label(data = filter(pws_map, name == "Stockdale"), aes(long, lat, label = name), hjust = 0, nudge_x = 0.1, nudge_y = 0.02) +
  geom_label(data = filter(pws_map, name == "Gilmour"), aes(long, lat, label = name), hjust = 0, nudge_x = 0.1, nudge_y = -0.02) +
  geom_label(data = filter(pws_map, name == "Paddy"), aes(long, lat, label = name), hjust = 1, nudge_x = -0.1, nudge_y = 0.04) +
  geom_label(data = filter(pws_map, name == "Erb"), aes(long, lat, label = name), hjust = 1, nudge_x = -0.1, nudge_y = -0.04) +
  geom_segment(aes(x = -148.7, xend = -148.7, y = 59.85, yend = 60.02), lineend = "butt", linejoin = "round", size = 2, arrow = arrow(length = unit(0.2, "inches"))) +
  annotate("text", label = "Prince\nWilliam\nSound", x = -147, y = 60.6, size = 5, colour = "black") +
  theme(panel.background = element_rect(fill = 'white')) +
  scale_y_continuous(breaks = c(60.0, 60.5, 61.0)) +
  xlab(expression(paste(Longitude^o, ~'W'))) +
  ylab(expression(paste(Latitude^o, ~'N'))) +
  labs(fill = "Legend", shape = "Legend", size = "Legend") +
  coord_map(xlim = x_lim, ylim = y_lim) +
  maptheme

# north2(pws_map, symbol = 12, 0.18, 0.20)  # couldn't get this to work with the inset

# Inset map
ak_map <- ggplot() + 
  geom_polygon(data = ak, aes(long, lat, group = group), fill = "grey90", color = "black", lwd = 0.3) +
  borders("world", xlim = c(-170, -130, ylim = c(20, 50)), lwd = 1) +
  geom_rect(data = data.frame(), aes(xmin = x_lim[1], xmax = x_lim[2], ymin = y_lim[1], ymax = y_lim[2]), colour = "red", fill = NA, lwd = 1) +
  annotate("text", label = "Alaska", x = -150.5, y = 63.5, size = 5, colour = "black") +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  coord_map(xlim = c(-170, -130), ylim = c(53, 65))

grid.newpage()
vp_b <- viewport(width = 1, height = 1, x = 0.5, y = 0.5)  # the larger map
vp_a <- viewport(width = 0.23, height = 0.35, x = 0.585, y = 0.32)  # the inset in lower left
print(pws_map, vp = vp_b)
print(ak_map + maptheme, vp = vp_a)
```

Save
```{r save_png}
png(filename = "PWS_stream_map.png", width = 6.5, height = 6.5, units = "in", res = 300)
grid.newpage()
vp_b <- viewport(width = 1, height = 1, x = 0.5, y = 0.5)  # the larger map
vp_a <- viewport(width = 0.23, height = 0.40, x = 0.7, y = 0.26)  # the inset in lower left; x = 0.585, y = 0.32
print(pws_map, vp = vp_b)
print(ak_map + maptheme, vp = vp_a)
dev.off()
```

## Natural Straying

```{r pbs_mapping_zoom}
pws_map <- readr::read_csv(file = "../map_data.csv")

pws_map_streams <- pws_map %>% 
  dplyr::filter(type == "Fitness Stream")


x_lim <- range(pws_map_streams$long) + c(-0.1, 0.1)
y_lim <- range(pws_map_streams$lat) + c(-0.1, 0.1)

data("nepacLLhigh")
ak <- nepacLLhigh %>% 
  dplyr::select(group = PID, POS = POS,long = X, lat = Y)

maptheme <- theme(
  panel.grid = element_blank(),
  panel.border = element_rect(fill = NA, colour = "black"),
  panel.background = element_blank()
)

# PWS map
pws_map <- ggplot() + 
  geom_polygon(data = ak, aes(long, lat, group = group), fill = "grey90", color = "black", lwd = 0.3) +
  geom_point(data = pws_map, aes(x = long, y = lat, fill = legend, shape = legend, size = legend), stroke = 1.5) +  # shape = 21, colour = "black",
  scale_shape_manual(values = c(24, 21, 22, 22)) +
  scale_size_manual(values = c(3.5, 5, 5, 5)) +
  geom_label(data = filter(pws_map, type == "Hatchery"), aes(long, lat, label = name), hjust = 0, nudge_x = 0.04) +
  geom_label(data = filter(pws_map, type == "City"), aes(long, lat, label = name), hjust = 0.5, vjust = 0, nudge_y = 0.01) +
  geom_label(data = filter(pws_map, name == "Hogan"), aes(long, lat, label = name), hjust = 0, nudge_x = 0.04, nudge_y = -0.01) +
  geom_label(data = filter(pws_map, name == "Stockdale"), aes(long, lat, label = name), hjust = 0.5, nudge_x = 0.04, nudge_y = 0.03) +
  geom_label(data = filter(pws_map, name == "Gilmour"), aes(long, lat, label = name), hjust = 0, nudge_x = 0.04, nudge_y = -0.01) +
  geom_label(data = filter(pws_map, name == "Paddy"), aes(long, lat, label = name), hjust = 1, nudge_x = -0.04, nudge_y = 0.01) +
  geom_label(data = filter(pws_map, name == "Erb"), aes(long, lat, label = name), hjust = 1, nudge_x = -0.04, nudge_y = -0.01) +
  geom_segment(aes(x = -148.7, xend = -148.7, y = 59.85, yend = 60.02), lineend = "butt", linejoin = "round", size = 2, arrow = arrow(length = unit(0.2, "inches"))) +
  annotate("text", label = "Prince\nWilliam\nSound", x = -147, y = 60.6, size = 5, colour = "black") +
  theme(panel.background = element_rect(fill = 'white')) +
  scale_y_continuous(breaks = c(60.0, 60.1, 60.2, 60.3, 60.4, 60.5)) +
  xlab(expression(paste(Longitude^o, ~'W'))) +
  ylab(expression(paste(Latitude^o, ~'N'))) +
  labs(fill = "Legend", shape = "Legend", size = "Legend") +
  theme(legend.position = "none") +
  coord_map(xlim = x_lim, ylim = y_lim) +
  maptheme

# north2(pws_map, symbol = 12, 0.18, 0.20)  # couldn't get this to work with the inset

# Inset map
ak_map <- ggplot() + 
  geom_polygon(data = ak, aes(long, lat, group = group), fill = "grey90", color = "black", lwd = 0.3) +
  # borders("world", xlim = x_lim + c(-2, 2), ylim = y_lim + c(-3, 3), lwd = 1) +
  geom_rect(data = data.frame(), aes(xmin = x_lim[1], xmax = x_lim[2], ymin = y_lim[1], ymax = y_lim[2]), colour = "red", fill = NA, lwd = 1) +
  # annotate("text", label = "Alaska", x = -150.5, y = 63.5, size = 5, colour = "black") +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  coord_map(xlim = x_lim + c(-1, 1.5), ylim = y_lim + c(-0.5, 0.75))

grid.newpage()
vp_b <- viewport(width = 1, height = 1, x = 0.5, y = 0.5)  # the larger map
vp_a <- viewport(width = 0.23, height = 0.35, x = 0.585, y = 0.32)  # the inset in lower left
print(pws_map, vp = vp_b)
print(ak_map + maptheme, vp = vp_a)
```

Save
```{r save_zoom_png}
png(filename = "PWS_stream_map_zoom.png", width = 7.5, height = 4.5, units = "in", res = 300)
grid.newpage()
vp_b <- viewport(width = 1, height = 1, x = 0.5, y = 0.5)  # the larger map
vp_a <- viewport(width = 0.2, height = 0.35, x = 0.802, y = 0.825)  # the inset in lower left; x = 0.585, y = 0.32
print(pws_map, vp = vp_b)
print(ak_map + maptheme, vp = vp_a)
dev.off()
```


# RRS to Date

## Dyad

Read in the data
```{r read_rrs_to_date}
(rrs_to_date_dyad <- readr::read_csv(file = "rrs_to_date_dyad.csv"))
```

### Bar plot

```{r fig.height=7.5, fig.width=6.5}
rrs_to_date_dyad %>% 
  dplyr::mutate(year = as.character(year)) %>% 
  ggplot2::ggplot(aes(x = year, y = rrs, ymin = lower_ci, ymax = upper_ci, fill = sex)) +
  ggplot2::geom_col(position = position_dodge(), colour = "black") +
  ggplot2::geom_errorbar(position = position_dodge(width = 0.9), width = 0.5, size = 0.75) +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_hline(yintercept = 1, linetype = 2, size = 1) +
  ggplot2::scale_fill_manual(values = c("orange", "skyblue")) +
  ggplot2::facet_grid(stream ~ lineage, scales = "free_x") +
  ggplot2::labs(x = "Run Year", y = "RRS (95% CI)", fill = "Sex") +
  ggplot2::theme_bw(base_size = 16) +
  ggplot2::theme(legend.position = c(0.8, 0.8), legend.background = element_rect(colour = "black"))

ggplot2::ggsave(filename = "RRS_to_date_dyad.png", width = 6.5, height = 7.5, units = "in")
```

## Triad

Read in the data
```{r read_rrs_to_date_triad}
(rrs_to_date_triad <- readr::read_csv(file = "rrs_to_date_triad.csv"))
```

### Bar plot

Get default `ggplot2` colors
```{r}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```

### Bar plot

```{r fig.height=7.5, fig.width=6.5}
rrs_to_date_triad %>% 
  dplyr::mutate(year = as.character(year)) %>% 
  ggplot2::ggplot(aes(x = year, y = rrs, ymin = lower_ci, ymax = upper_ci, fill = cross)) +
  ggplot2::geom_col(position = position_dodge(), colour = "black") +
  ggplot2::geom_errorbar(position = position_dodge(width = 0.9), width = 0.5, size = 0.75) +
  ggplot2::geom_hline(yintercept = 0) +
  ggplot2::geom_hline(yintercept = 1, linetype = 2, size = 1) +
  ggplot2::scale_fill_manual(values = gg_color_hue(n = 4)[4:1]) +
  ggplot2::facet_grid(stream ~ lineage, scales = "free_x") +
  ggplot2::labs(x = "Run Year", y = "RRS (95% CI)", fill = "Cross Type") +
  ggplot2::theme_bw(base_size = 16)
  # ggplot2::theme(legend.position = c(0.8, 0.8), legend.background = element_rect(colour = "black"))

ggplot2::ggsave(filename = "RRS_to_date_triad.png", width = 6.5, height = 7.5, units = "in")
```

# Offspring Sample Date - Density Plot

## Even

### 2014/2016

```{r}
(
  all_streams_parents_paired_14_16 <-
    readr::read_csv(file = "../Stockdale_Hogan_Gilmour_Paddy_Erb/all_streams_parents_paired_14_16.csv")
)
```

Create a density plot of offspring sample date by parent-origin, faceted by offspring sample stream and offspring sex.
```{r fig.height=7.5}
all_streams_parents_paired_14_16 %>% 
  dplyr::mutate(origin_par = base::factor(x = origin_par, levels = c("Natural", "Hatchery"))) %>% 
  ggplot2::ggplot(aes(x = date_off, fill = origin_par)) +
  ggplot2::geom_density(alpha = 0.5) +
  ggplot2::facet_grid(stream_off ~ sex_off, scales = "free_y") +
  ggplot2::labs(title = "Density Plot of Offspring (2016) Sample Date by Parent Origin",
                fill = "Parent: Origin",
                x = "Offspring Sample Date",
                y = "Density") +
  ggplot2::theme_bw() 
  # ggplot2::theme(panel.grid.minor = element_blank()) +
  # ggplot2::scale_fill_manual(values = cbPalette2)

ggplot2::ggsave(filename = "offspring_date_density_2014_2016.png", width = 6.5, height = 7.5, units = "in")
```

### 2016/2018

Only showing data from Stockdale and Hogan so far due to GOD affecting the Paddy, Erb, and Gilmour results
```{r}
(
  stockdale_parents_paired_16_18 <-
    readr::read_csv(file = "../Stockdale/stock_parents_paired_16_18.csv") %>% 
    dplyr::select(-`Sample Date.off`, -`Sample Date.par`)
)

(
  hogan_parents_paired_16_18 <-
    readr::read_csv(file = "../Hogan/hogan_parents_paired_16_18.csv") %>% 
    dplyr::mutate(date.off = `Sample Date.off`) %>% 
    dplyr::select(-`Sample Date.off`, -`Sample Date.par`)
)
```

Join, rename to fit *new* conventions ;)
```{r}
stockdale_hogan_parents_paired_16_18 <-
  dplyr::bind_rows(stockdale_parents_paired_16_18, hogan_parents_paired_16_18) %>%
  dplyr::rename(origin_par = origin.par,
                date_off = date.off,
                sex_off = SEX.off) %>%
  dplyr::mutate(
    stream_off = dplyr::case_when(
      base::grepl(pattern = "STOCK", x = SILLY.off) ~ "Stockdale",
      base::grepl(pattern = "HOGAN", x = SILLY.off) ~ "Hogan"
    ),
    sex_off = dplyr::case_when(sex_off == "F" ~ "Female",
                               sex_off == "M" ~ "Male")
  )
```

Create a density plot of offspring sample date by parent-origin, faceted by offspring sample stream and offspring sex.
```{r fig.height=7.5}
stockdale_hogan_parents_paired_16_18 %>% 
  dplyr::filter(!is.na(sex_off)) %>% 
  dplyr::mutate(origin_par = base::factor(x = origin_par, levels = c("Natural", "Hatchery"))) %>% 
  ggplot2::ggplot(aes(x = date_off, fill = origin_par)) +
  ggplot2::geom_density(alpha = 0.5) +
  ggplot2::facet_grid(stream_off ~ sex_off, scales = "free_y") +
  ggplot2::labs(title = "Density Plot of Offspring (2018) Sample Date by Parent Origin",
                fill = "Parent: Origin",
                x = "Offspring Sample Date",
                y = "Density") +
  ggplot2::theme_bw() 
  # ggplot2::theme(panel.grid.minor = element_blank()) +
  # ggplot2::scale_fill_manual(values = cbPalette2)

ggplot2::ggsave(filename = "offspring_date_density_2016_2018.png", width = 6.5, height = 7.5, units = "in")
```

## Odd

### 2013/2015

Only have data from Stockdale and Hogan so far due to GOD
```{r}
(
  stockdale_parents_paired_13_15 <-
    readr::read_csv(file = "../Stockdale/stock_parents_paired_13_15.csv")
)

(
  hogan_parents_paired_13_15 <-
    readr::read_csv(file = "../Hogan/hogan_parents_paired_13_15.csv")
)
```

Join, rename to fit *new* conventions ;)
```{r}
stockdale_hogan_parents_paired_13_15 <-
  dplyr::bind_rows(stockdale_parents_paired_13_15, hogan_parents_paired_13_15) %>%
  dplyr::rename(origin_par = origin.par,
                date_off = date.off,
                sex_off = SEX.off) %>%
  dplyr::mutate(
    stream_off = dplyr::case_when(
      base::grepl(pattern = "STOCK", x = SILLY.off) ~ "Stockdale",
      base::grepl(pattern = "HOGAN", x = SILLY.off) ~ "Hogan"
    ),
    sex_off = dplyr::case_when(sex_off == "F" ~ "Female",
                               sex_off == "M" ~ "Male")
  )
```


Create a density plot of offspring sample date by parent-origin, faceted by offspring sample stream and offspring sex.
```{r fig.height=7.5}
stockdale_hogan_parents_paired_13_15 %>% 
  dplyr::mutate(origin_par = base::factor(x = origin_par, levels = c("Natural", "Hatchery"))) %>% 
  ggplot2::ggplot(aes(x = date_off, fill = origin_par)) +
  ggplot2::geom_density(alpha = 0.5) +
  ggplot2::facet_grid(stream_off ~ sex_off, scales = "free_y") +
  ggplot2::labs(title = "Density Plot of Offspring (2015) Sample Date by Parent Origin",
                fill = "Parent: Origin",
                x = "Offspring Sample Date",
                y = "Density") +
  ggplot2::theme_bw() 
  # ggplot2::theme(panel.grid.minor = element_blank()) +
  # ggplot2::scale_fill_manual(values = cbPalette2)

ggplot2::ggsave(filename = "offspring_date_density_2013_2015.png", width = 6.5, height = 7.5, units = "in")
```

### 2015/2017

Only have data from Stockdale and Hogan so far due to GOD
```{r}
(
  stockdale_parents_paired_15_17 <-
    readr::read_csv(file = "../Stockdale/stock_parents_paired_15_17.csv")
)

(
  hogan_parents_paired_15_17 <-
    readr::read_csv(file = "../Hogan/hogan_parents_paired_15_17.csv")
)
```

Join, rename to fit *new* conventions ;)
```{r}
stockdale_hogan_parents_paired_15_17 <-
  dplyr::bind_rows(stockdale_parents_paired_15_17, hogan_parents_paired_15_17) %>%
  dplyr::rename(origin_par = origin.par,
                date_off = date.off,
                sex_off = SEX.off) %>%
  dplyr::mutate(
    stream_off = dplyr::case_when(
      base::grepl(pattern = "STOCK", x = SILLY.off) ~ "Stockdale",
      base::grepl(pattern = "HOGAN", x = SILLY.off) ~ "Hogan"
    ),
    sex_off = dplyr::case_when(sex_off == "F" ~ "Female",
                               sex_off == "M" ~ "Male")
  )
```


Create a density plot of offspring sample date by parent-origin, faceted by offspring sample stream and offspring sex.
```{r fig.height=7.5}
stockdale_hogan_parents_paired_15_17 %>% 
  dplyr::mutate(origin_par = base::factor(x = origin_par, levels = c("Natural", "Hatchery"))) %>% 
  ggplot2::ggplot(aes(x = date_off, fill = origin_par)) +
  ggplot2::geom_density(alpha = 0.5) +
  ggplot2::facet_grid(stream_off ~ sex_off, scales = "free_y") +
  ggplot2::labs(title = "Density Plot of Offspring (2017) Sample Date by Parent Origin",
                fill = "Parent: Origin",
                x = "Offspring Sample Date",
                y = "Density") +
  ggplot2::theme_bw() 
  # ggplot2::theme(panel.grid.minor = element_blank()) +
  # ggplot2::scale_fill_manual(values = cbPalette2)

ggplot2::ggsave(filename = "offspring_date_density_2015_2017.png", width = 6.5, height = 7.5, units = "in")
```

### 2017/2019

Only have data from Stockdale and Hogan so far due to GOD
```{r}
(
  stockdale_parents_paired_17_19 <-
    readr::read_csv(file = "../Stockdale/stock_parents_paired_17_19.csv") %>% 
    dplyr::select(-`Sample Date.off`, -`Sample Date.par`)
)

(
  hogan_parents_paired_17_19 <-
    readr::read_csv(file = "../Hogan/hogan_parents_paired_17_19.csv") %>% 
    dplyr::mutate(date.off = `Sample Date.off`) %>% 
    dplyr::select(-`Sample Date.off`, -`Sample Date.par`)
)
```

Join, rename to fit *new* conventions ;)
```{r}
stockdale_hogan_parents_paired_17_19 <-
  dplyr::bind_rows(stockdale_parents_paired_17_19, hogan_parents_paired_17_19) %>%
  dplyr::rename(origin_par = origin.par,
                date_off = date.off,
                sex_off = SEX.off) %>%
  dplyr::mutate(
    stream_off = dplyr::case_when(
      base::grepl(pattern = "STOCK", x = SILLY.off) ~ "Stockdale",
      base::grepl(pattern = "HOGAN", x = SILLY.off) ~ "Hogan"
    ),
    sex_off = dplyr::case_when(sex_off == "F" ~ "Female",
                               sex_off == "M" ~ "Male")
  )
```


Create a density plot of offspring sample date by parent-origin, faceted by offspring sample stream and offspring sex.
```{r fig.height=7.5}
stockdale_hogan_parents_paired_17_19 %>% 
  dplyr::mutate(origin_par = base::factor(x = origin_par, levels = c("Natural", "Hatchery"))) %>% 
  ggplot2::ggplot(aes(x = date_off, fill = origin_par)) +
  ggplot2::geom_density(alpha = 0.5) +
  ggplot2::facet_grid(stream_off ~ sex_off, scales = "free_y") +
  ggplot2::labs(title = "Density Plot of Offspring (2019) Sample Date by Parent Origin",
                fill = "Parent: Origin",
                x = "Offspring Sample Date",
                y = "Density") +
  ggplot2::theme_bw() 
  # ggplot2::theme(panel.grid.minor = element_blank()) +
  # ggplot2::scale_fill_manual(values = cbPalette2)

ggplot2::ggsave(filename = "offspring_date_density_2017_2019.png", width = 6.5, height = 7.5, units = "in")
```










