---
title: "All Streams 2014/2016 GLMs"
subtitle: "Hogan + Stockdale + Gilmour + Paddy + Erb"
author: "Kyle Shedd"
date: "16 February 2022"
output:
  html_notebook:
    theme: united
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
rm(list = ls())
library(coin)
library(MASS)
library(scales)
library(tidyverse)
library(leaflet)
library(lubridate)
library(ggExtra)
library(gridExtra)
library(MuMIn)
library(GGally)
library(gganimate)
library(QuantPsyc)
library(hier.part)
source("../hier.part_functions.R")

knitr::opts_chunk$set(warning=FALSE, message=FALSE, fig.width = 10)
```

# Introduction

The purpose of these analyses are to follow up on the previous notebook and fully consider all of the RS data holistically by model fitting reproductive success (RS) to generalized linear models (GLM’s), as is commonly done in the literature (see [Anderson et al. 2013](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1752-4571.2012.00271.x), [Bernston et al. 2011](https://afspubs.onlinelibrary.wiley.com/doi/pdf/10.1080/00028487.2011.584489), [Ford et al. 2012](https://onlinelibrary.wiley.com/doi/full/10.1111/j.1755-263X.2012.00261.x), and [Janowitz-Koch et al. 2018](https://onlinelibrary.wiley.com/doi/full/10.1111/eva.12725)). First, we will explore the data a bit so we know what we are modeling, next we will check for multi-collinearity among variables, and finally we will construct a series of GLM's to model RS and use Akaike model selection to calculate variable weights.

# Load Parent Data

```{r}
(
  all_streams_paired_14_16_filter_parents <-
    readr::read_csv("all_streams_paired_14_16_filter_parents.csv") %>%
    dplyr::mutate(
      origin = factor(x = origin, levels = c("Natural", "Hatchery")),
      stream = factor(x = stream, levels = c("Erb", "Paddy", "Hogan", "Gilmour", "Stockdale")),
      intertidal = factor(x = intertidal, levels = c("Intertidal", "Upstream"))
    )
)
```

# Explore Parent Data

While we've previously shown figures relating RS to individual variables (e.g. Sample Date, Length of parents, etc.), we have now incorporated the spatial data (Sample Location of parents) and can now explore spatio-temporal trends in RS.  

Here we'll create an animation that walks through each sample date to show that fishes sample location (latitude and longitude) with the color of the point indicating the fish's origin (hatchery vs. natural), the shape the fish's sex (male/female), and the size of the point indicating how many offspring were assigned to it (reproductive success). This will show us two important things:  

  1) where high reproductive success occurred in space and time
  2) the degree of spatio-temporal overlap between hatchery- and natural-origin pink salmon in streams in 2014

I made the GIFs in a separate notebook.

Another way to look at this is to use river distance (distance in meters upstream from the lowest sampling extent) instead of latitude/longitude, so we can view it in 2D as opposed to an animation. Here is a simple plot of RS by parent sample date and distance (both origins and sexes combined) split out by stream and whether parents were successful or not (n_offspring > 0).
```{r, out.width="100%", fig.height=10}
all_streams_paired_14_16_filter_parents %>%
  dplyr::mutate(success = dplyr::case_when(n > 0 ~ "Successful",
                                           n == 0 ~ "Unsuccessful")) %>%
  ggplot2::ggplot(ggplot2::aes(
    x = date,
    y = distance_mouth,
    size = n,
    colour = origin,
    alpha = 0.05
  )) +
  ggplot2::geom_hline(ggplot2::aes(yintercept = high_tide), size = 1) +
  ggplot2::geom_jitter(width = 0.25) +
  ggplot2::theme_bw() +
  ggplot2::guides(
    alpha = "none",
    colour = ggplot2::guide_legend(order = 1),
    size = ggplot2::guide_legend(order = 3)
  ) +
  ggplot2::labs(x = "Parent Sample Date",
                y = "Parent Sample Location from Stream Mouth (m)",
                size = "RS",
                colour = "Parent: Origin",
                title = "RS by Sample Date and Location of Parents") +
  ggplot2::facet_grid(
    rows = ggplot2::vars(stream),
    cols = ggplot2::vars(success),
    scales = "free_y"
  )
```

Interesting, we can certainly see how hatchery- and natural-origin fish are separated in space and time, although there is some spatio-temporal overlap. Considerable varation by stream.

# Check for Multi-collinearity

Before modeling RS, we need to check for multi-collinearity among variables because there may be some degree of non-independence among variables. For visualization purposes, data are grouped by origin. The diagonal shows density plots for continuous variables (e.g. Sample Date) and side by side bar plots for categorical variables (e.g. Sex). The upper plots above the diagonal show correlation coefficients between continuous variables, grouped box plots between continuous and categorical variables, and stacked bar plots between categorical variables. The lower plots below the diagonal show scatter plots with loess curves between continuous variables, grouped histograms between continuous and categorical variables, and stacked bar plots between categorical variables. The variable *Success* is derived from *RS* (Successful for RS > 0, Unsuccessful for RS = 0), and thus would not be used in a model concurrent with *RS*. **Note** that these plots exclude individuals with missing data.


Where is high-tide?
**NOTE** some streams with multiple tributaries have separate high tide marks for each tributary.
```{r}
all_streams_paired_14_16_filter_parents %>% 
  dplyr::group_by(stream) %>% 
  dplyr::mutate(high_tide = base::round(high_tide)) %>% 
  dplyr::distinct(high_tide) %>% 
  dplyr::arrange(stream)
```

## Erb

```{r Erb_ggpairs, fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
reduced_Erb <- all_streams_paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Erb",
                length > 300) %>% 
  dplyr::mutate(success = dplyr::case_when(n > 0 ~ "Successful",
                                           n == 0 ~ "Unsuccessful")) %>%
  dplyr::select(stream, origin, n, success, sex, date, length, intertidal, distance_mouth) %>% 
  tidyr::drop_na()

GGally::ggpairs(
  data = reduced_Erb,
  mapping = ggplot2::aes(colour = origin, alpha = 0.5),
  columns = c("date", "length", "distance_mouth", "n", "success", "intertidal", "sex"),
  columnLabels = c("Sample Date", "Length (mm)", "Distance (m)", "RS", "Success", "Intertidal", "Sex"),
  progress = FALSE,
  lower = list(continuous = "smooth_loess", combo = wrap(GGally::ggally_facethist, position = "dodge2")),
  upper = list(continuous = wrap("cor", size = 4, hjust = 0.5), combo = "box"),
  diag = list(discrete = wrap(GGally::ggally_barDiag, position = "dodge2"))
) +
  ggplot2::theme_bw() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

## Paddy

```{r Paddy_ggpairs, fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
reduced_Paddy <- all_streams_paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Paddy",
                length > 300) %>% 
  dplyr::mutate(success = dplyr::case_when(n > 0 ~ "Successful",
                                           n == 0 ~ "Unsuccessful")) %>%
  dplyr::select(stream, origin, n, success, sex, date, length, intertidal, distance_mouth) %>% 
  tidyr::drop_na()

GGally::ggpairs(
  data = reduced_Paddy,
  mapping = ggplot2::aes(colour = origin, alpha = 0.5),
  columns = c("date", "length", "distance_mouth", "n", "success", "intertidal", "sex"),
  columnLabels = c("Sample Date", "Length (mm)", "Distance (m)", "RS", "Success", "Intertidal", "Sex"),
  progress = FALSE,
  lower = list(continuous = "smooth_loess", combo = wrap(GGally::ggally_facethist, position = "dodge2")),
  upper = list(continuous = wrap("cor", size = 4, hjust = 0.5), combo = "box"),
  diag = list(discrete = wrap(GGally::ggally_barDiag, position = "dodge2"))
) +
  ggplot2::theme_bw() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

## Hogan

```{r Hogan_ggpairs, fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
reduced_Hogan <- all_streams_paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Hogan",
                length > 300) %>% 
  dplyr::mutate(success = dplyr::case_when(n > 0 ~ "Successful",
                                           n == 0 ~ "Unsuccessful")) %>%
  dplyr::select(stream, origin, n, success, sex, date, length, intertidal, distance_mouth) %>% 
  tidyr::drop_na()

GGally::ggpairs(
  data = reduced_Hogan,
  mapping = ggplot2::aes(colour = origin, alpha = 0.5),
  columns = c("date", "length", "distance_mouth", "n", "success", "intertidal", "sex"),
  columnLabels = c("Sample Date", "Length (mm)", "Distance (m)", "RS", "Success", "Intertidal", "Sex"),
  progress = FALSE,
  lower = list(continuous = "smooth_loess", combo = wrap(GGally::ggally_facethist, position = "dodge2")),
  upper = list(continuous = wrap("cor", size = 4, hjust = 0.5), combo = "box"),
  diag = list(discrete = wrap(GGally::ggally_barDiag, position = "dodge2"))
) +
  ggplot2::theme_bw() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

## Gilmour

```{r Gilmour_ggpairs, fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
reduced_Gilmour <- all_streams_paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Gilmour",
                length > 300) %>% 
  dplyr::mutate(success = dplyr::case_when(n > 0 ~ "Successful",
                                           n == 0 ~ "Unsuccessful")) %>%
  dplyr::select(stream, origin, n, success, sex, date, length, intertidal, distance_mouth) %>% 
  tidyr::drop_na()

GGally::ggpairs(
  data = reduced_Gilmour,
  mapping = ggplot2::aes(colour = origin, alpha = 0.5),
  columns = c("date", "length", "distance_mouth", "n", "success", "intertidal", "sex"),
  columnLabels = c("Sample Date", "Length (mm)", "Distance (m)", "RS", "Success", "Intertidal", "Sex"),
  progress = FALSE,
  lower = list(continuous = "smooth_loess", combo = wrap(GGally::ggally_facethist, position = "dodge2")),
  upper = list(continuous = wrap("cor", size = 4, hjust = 0.5), combo = "box"),
  diag = list(discrete = wrap(GGally::ggally_barDiag, position = "dodge2"))
) +
  ggplot2::theme_bw() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

## Stockdale

```{r Stockdale_ggpairs, fig.width=12, fig.height=12, warning=FALSE, message=FALSE}
reduced_Stockdale <- all_streams_paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Stockdale",
                length > 300) %>% 
  dplyr::mutate(success = dplyr::case_when(n > 0 ~ "Successful",
                                           n == 0 ~ "Unsuccessful")) %>%
  dplyr::select(stream, origin, n, success, sex, date, length, intertidal, distance_mouth) %>% 
  tidyr::drop_na()

GGally::ggpairs(
  data = reduced_Stockdale,
  mapping = ggplot2::aes(colour = origin, alpha = 0.5),
  columns = c("date", "length", "distance_mouth", "n", "success", "intertidal", "sex"),
  columnLabels = c("Sample Date", "Length (mm)", "Distance (m)", "RS", "Success", "Intertidal", "Sex"),
  progress = FALSE,
  lower = list(continuous = "smooth_loess", combo = wrap(GGally::ggally_facethist, position = "dodge2")),
  upper = list(continuous = wrap("cor", size = 4, hjust = 0.5), combo = "box"),
  diag = list(discrete = wrap(GGally::ggally_barDiag, position = "dodge2"))
) +
  ggplot2::theme_bw() +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
```

# Density Distributions

## Sample Date

In depth look at run timing
```{r fig.height=10}
all_streams_paired_14_16_filter_parents %>%
  ggplot2::ggplot(ggplot2::aes(x = date, fill = origin)) +
  ggplot2::geom_density(alpha = 0.5, adjust = 2) +
  ggplot2::labs(x = "Parent Sample Date",
                y = "Density",
                fill = "Parent: Origin") +
  ggplot2::theme_bw() +
  ggplot2::facet_grid(rows = ggplot2::vars(stream), cols = ggplot2::vars(sex))
```

Interesting patterns of temporal overlap among streams. Big differences in run timing between natural-origin fish and hatchery strays for Erb, Paddy, and Hogan, but less so for Stockdale, and almost compelte overlap for Gilmour.

## Stream Distance

In depth look at spawning location

**NOTE** excluding the Paddy due to it's weird sample location profile (almost all samples from one location, some samples waaay upstream).
```{r fig.height=10}
all_streams_paired_14_16_filter_parents %>%
  dplyr::filter(stream != "Paddy") %>% 
  ggplot2::ggplot(ggplot2::aes(y = distance_mouth, fill = origin)) +
  ggplot2::geom_density(alpha = 0.5, adjust = 2) +
  ggplot2::geom_hline(ggplot2::aes(yintercept = high_tide)) +
  ggplot2::labs(x = "Density",
                y = "Parent Sample Location from Stream Mouth (m)",
                fill = "Parent: Origin") +
  ggplot2::theme_bw() +
  ggplot2::facet_grid(rows = ggplot2::vars(stream), cols = ggplot2::vars(sex), scales = "free_y")
```

In general, natural-origin fish were sampled lower in each stream, with a higher proportion of samples in the intertidal.

### Paddy

What's going on here?
```{r}
all_streams_paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Paddy",
                distance_mouth > 500)
```

Very few fish sampled way the heck up stream.

What is our sample size by origin and sex for upstream?
```{r}
all_streams_paired_14_16_filter_parents %>% 
  dplyr::filter(stream == "Paddy") %>% 
  dplyr::count(origin, sex, intertidal) %>% 
  dplyr::arrange(dplyr::desc(intertidal))
```

Ah, very few natural-origin fish present upstream...

## Body Length

In depth look at body length
```{r fig.height=10}
all_streams_paired_14_16_filter_parents %>%
  ggplot2::ggplot(ggplot2::aes(x = length, fill = origin)) +
  ggplot2::geom_density(alpha = 0.5) +
  ggplot2::labs(x = "Parent Body Length (mm)",
                y = "Density",
                fill = "Parent: Origin") +
  ggplot2::theme_bw() +
  ggplot2::facet_grid(rows = ggplot2::vars(stream), cols = ggplot2::vars(sex))
```

In general, hatchery-origin fish were larger than natural-origin fish across all streams, particularly for males.

# GLMs

## Scale Covariates

```{r}
(
  all_streams_paired_14_16_filter_parents_scaled <-
    all_streams_paired_14_16_filter_parents %>%
    dplyr::group_by(stream, sex) %>%
    dplyr::mutate(
      DOY_scaled = scale(x = DOY, center = TRUE, scale = TRUE),
      length_scaled = scale(x = length, center = TRUE, scale = TRUE),
      distance_mouth_scaled = scale(x = distance_mouth, center = TRUE, scale = TRUE),
      n_relative = n / mean(n)
    ) %>%
    dplyr::ungroup() %>% 
    dplyr::select(franz_id, n, n_relative, tidyselect::everything())
)
```

## Male

### Models

```{r}
male_scaled <- all_streams_paired_14_16_filter_parents_scaled %>%
  dplyr::filter(sex == "Male")

glm_male_1 <-
  MASS::glm.nb(
    n ~ stream + DOY_scaled:stream + length_scaled:stream + intertidal:stream + origin,
    data = male_scaled,
    link = log
  )

glm_male_2 <-
  MASS::glm.nb(
    n ~ stream + DOY_scaled:stream + length_scaled:stream + distance_mouth_scaled:stream + origin,
    data = male_scaled,
    link = log
  )

glm_male_3 <-
  MASS::glm.nb(
    n ~ stream + DOY_scaled + length_scaled + intertidal + origin,
    data = male_scaled,
    link = log
  )

glm_male_4 <-
  MASS::glm.nb(
    n ~ stream + DOY_scaled + length_scaled + distance_mouth_scaled + origin,
    data = male_scaled,
    link = log
  )
```

### Extract Akaike Weights {#male_glm}

Akaike weights are another way of determining which model has the best fit, by calculating their conditional probabilities. 
```{r male_glm_summary}
male_glm_model_summary <- bind_rows(lapply(c(1:4), function(i) {
  fit <- get(paste0("glm_male_", i))
  summary(fit)$coefficients %>%
    as_tibble(rownames = "variable") %>%
    mutate(name = paste0("glm_male_", i)) %>%
    mutate(model = as.character(fit$call)[2]) %>%
    mutate(AIC = fit$aic) %>%
    mutate(residual_deviance = fit$deviance) %>%
    mutate(residual_df = fit$df.residual) %>%
    mutate(null_deviance = fit$null.deviance) %>%
    mutate(null_df = fit$df.null)
}))

male_glm_aic <- bind_rows(lapply(paste0("glm_male_", 1:4), function(mod) {
  data.frame(
    row.names = mod,
    df = length(get(mod)$coefficients) + 1,
    AIC = AIC(get(mod))
  )
}))

rownames(male_glm_aic) <- paste0("glm_male_", 1:4)

male_glm_aic <- male_glm_aic %>%
  mutate(AIC_weight = as.numeric(Weights(male_glm_aic))) %>%
  mutate(name = rownames(male_glm_aic)) %>%
  arrange(AIC) %>%
  mutate(delta_AIC = AIC - min(AIC))

male_glm_model_summary <- male_glm_model_summary %>%
  left_join(male_glm_aic, by = c("name", "AIC")) %>%
  arrange(AIC)
# select(name, model, df, AIC, delta_AIC, AIC_weight, variable, Estimate, `Std. Error`, residual_deviance, residual_df, null_deviance, null_df)

male_glm_model_summary %>%
  group_by(name, df, residual_deviance, null_deviance) %>%
  summarise(
    AIC_weight = round(max(AIC_weight), 3),
    model = unique(model),
    deviance_explained = round(1 - (
      max(residual_deviance) / max(null_deviance)
    ), 3)
  ) %>%
  ungroup() %>%
  select(name, df, AIC_weight, deviance_explained, model) %>%
  arrange(desc(AIC_weight))

# write_csv(male_glm_model_summary, "glm_male_summary.csv")
```

### Top Model Summary

```{r}
summary(glm_male_1)
```

What are the coefficients and 95% CIs for each of the explanatory variables?

**NOTE** need to drop "streamPaddy:intertidalUpstream"
```{r}
male_est <-
  dplyr::as_tibble(coef(glm_male_1), rownames = "coefficients") %>%
  dplyr::rename(estimate = value)

male_est_CIs <-
  stats::confint(glm_male_1, parm = setdiff(names(coef(glm_male_1)), "streamPaddy:intertidalUpstream")) %>%
  dplyr::as_tibble(rownames = "coefficients")

(male_est <- male_est %>%
    dplyr::left_join(male_est_CIs, by = "coefficients"))
```

**Note** those coefficients are still in log space due to the logit link function. If we take the exponent, we can look at incident ratios.
```{r}
male_est %>% 
  dplyr::mutate(dplyr::across(where(is.numeric), exp)) %>% 
  dplyr::mutate(dplyr::across(where(is.numeric), round, 2))
```

## Female

### Models

```{r}
female_scaled <- all_streams_paired_14_16_filter_parents_scaled %>%
  dplyr::filter(sex == "Female")

glm_female_1 <-
  MASS::glm.nb(
    n ~ stream + DOY_scaled:stream + length_scaled:stream + intertidal:stream + origin,
    data = female_scaled,
    link = log
  )

glm_female_2 <-
  MASS::glm.nb(
    n ~ stream + DOY_scaled:stream + length_scaled:stream + distance_mouth_scaled:stream + origin,
    data = female_scaled,
    link = log
  )

glm_female_3 <-
  MASS::glm.nb(
    n ~ stream + DOY_scaled + length_scaled + intertidal + origin,
    data = female_scaled,
    link = log
  )

glm_female_4 <-
  MASS::glm.nb(
    n ~ stream + DOY_scaled + length_scaled + distance_mouth_scaled + origin,
    data = female_scaled,
    link = log
  )
```

### Extract Akaike Weights {#female_glm}

Akaike weights are another way of determining which model has the best fit, by calculating their conditional probabilities. 
```{r female_glm_summary}
female_glm_model_summary <- bind_rows(lapply(c(1:4), function(i) {
  fit <- get(paste0("glm_female_", i))
  summary(fit)$coefficients %>%
    as_tibble(rownames = "variable") %>%
    mutate(name = paste0("glm_female_", i)) %>%
    mutate(model = as.character(fit$call)[2]) %>%
    mutate(AIC = fit$aic) %>%
    mutate(residual_deviance = fit$deviance) %>%
    mutate(residual_df = fit$df.residual) %>%
    mutate(null_deviance = fit$null.deviance) %>%
    mutate(null_df = fit$df.null)
}))

female_glm_aic <- bind_rows(lapply(paste0("glm_female_", 1:4), function(mod) {
  data.frame(
    row.names = mod,
    df = length(get(mod)$coefficients) + 1,
    AIC = AIC(get(mod))
  )
}))

rownames(female_glm_aic) <- paste0("glm_female_", 1:4)

female_glm_aic <- female_glm_aic %>%
  mutate(AIC_weight = as.numeric(Weights(female_glm_aic))) %>%
  mutate(name = rownames(female_glm_aic)) %>%
  arrange(AIC) %>%
  mutate(delta_AIC = AIC - min(AIC))

female_glm_model_summary <- female_glm_model_summary %>%
  left_join(female_glm_aic, by = c("name", "AIC")) %>%
  arrange(AIC)
# select(name, model, df, AIC, delta_AIC, AIC_weight, variable, Estimate, `Std. Error`, residual_deviance, residual_df, null_deviance, null_df)

female_glm_model_summary %>%
  group_by(name, df, residual_deviance, null_deviance) %>%
  summarise(
    AIC_weight = round(max(AIC_weight), 3),
    model = unique(model),
    deviance_explained = round(1 - (
      max(residual_deviance) / max(null_deviance)
    ), 3)
  ) %>%
  ungroup() %>%
  select(name, df, AIC_weight, deviance_explained, model) %>%
  arrange(desc(AIC_weight))

# write_csv(female_glm_model_summary, "glm_female_summary.csv")
```

### Top Model Summary

```{r}
summary(glm_female_1)
```

What are the coefficients and 95% CIs for each of the explanatory variables?

**NOTE** need to drop "streamPaddy:intertidalUpstream"
```{r}
female_est <-
  dplyr::as_tibble(coef(glm_female_1), rownames = "coefficients") %>%
  dplyr::rename(estimate = value)

female_est_CIs <-
  stats::confint(glm_female_1) %>%
  dplyr::as_tibble(rownames = "coefficients")

(female_est <- female_est %>%
    dplyr::left_join(female_est_CIs, by = "coefficients"))
```

**Note** those coefficients are still in log space due to the logit link function. If we take the exponent, we can look at incident ratios.
```{r}
female_est %>% 
  dplyr::mutate(dplyr::across(where(is.numeric), exp)) %>% 
  dplyr::mutate(dplyr::across(where(is.numeric), round, 2))
```

End